### 2018-08-10 NYC_Twitter_SDM_models.Rmd
Upload files to repository
New park access analysis with data and code.

---
title: "NYC Spatial Models - Comparison of SDM Models With and Without Twitter Data"
author: "Nico Boyd"
date: "June 26, 2017"
output: html_document
---

For both obesity and physical activity, the first SDM model uses a size beta of .00001 and a distance beta of -5. The second SDM model uses the same size and distance betas with the addition of a tweet beta of .001. The third SDM model again uses the same size and distance betas but uses a negative tweet beta of -.001.

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
library(geojsonio)
library(tidyverse)
library(maptools)
library(rgdal)
library(rgeos)
library(spdep)
library(texreg)
library(leaflet)
```

```{r read_tracts, echo=FALSE}
tracts <- geojson_read("nyc_tracts.geojson", what = "sp")
#plot(tracts)
```

```{r read_parks, echo=FALSE}
parks <- geojson_read("nyc_parks.geojson", what = "sp")
#plot(parks)
```

```{r show_data, echo=FALSE}
tracts@data <- tbl_df(tracts@data) %>%
  select(GEOID, OBESITY, Park_Percent, Phys_Act, MENTAL, Income1, Income2, Income3, Income4, Income5, Income6, Income7, Income8, Income9, Income10, Pop_Density, FulltimeWork, CollegeDeg, Pct0to17, Pct18to29, Pct30to64, Pct65plus, Single_Percent, PctWhite, PctBlack, PctNative, PctAsian, PctPacific, PctOther, PctHispanic) %>%
    mutate_each(funs(ifelse(is.na(.), 0, .)), Park_Percent)

parks@data <- tbl_df(parks@data)

tracts@data
#parks@data
```

Calculate the logsum given a tracts and parks dataset. `tracts$park_ls1` uses only size and distance parameters. `tracts$park_ls2` uses size, distance and tweet parameters. `tracts$park_ls3` uses a size and distance parameter with a negative tweet parameter.
```{r accessibility}
source('size00001dist5.R', echo=TRUE)
tracts$park_ls1 <- calculate_park_logsums(tracts, parks, betas = c(.00001, -5))

source('size00001dist5tweet001.R', echo=TRUE)
tracts$park_ls2 <- calculate_park_logsums(tracts, parks, betas = c(.00001, -5, .001))

source('size00001dist5tweetneg001.R', echo=TRUE)
tracts$park_ls3 <- calculate_park_logsums(tracts, parks, betas = c(.00001, -5, -.001))

source('size00001dist15.R', echo=TRUE)
tracts$park_ls4 <- calculate_park_logsums(tracts, parks, betas = c(.00001, -15))

source('size00001dist15tweet001.R', echo=TRUE)
tracts$park_ls5 <- calculate_park_logsums(tracts, parks, betas = c(.00001, -15, .001))
```

Map the newly created Park Access variables. Here is a map of accessibility using only the size and distance parameters that are most likely when using obesity as the dependent variable.
```{r accessibility_leaflet_1, echo=FALSE}
# Create a continuous palette function
pal <- colorQuantile(
  palette = "Greens",
  domain = c(0, tracts$park_ls1)
)

pop_content <- function(shp){
  paste0(
    "Tract: ", shp$NAME, "<br/>",
    "% Park: ", round(shp$Park_Percent, 2), "<br/>",
    "Park Access: ", round(shp$park_ls1, 2), "<br/>"
  )
}

leaflet(tracts %>% spTransform(., CRS("+init=epsg:4326"))) %>%
  addProviderTiles("CartoDB.Positron") %>%
  addPolygons(weight = 1, popup = ~pop_content(tracts),
              color = ~pal(park_ls1))
```

And here is a map of accessibility using the size, distance and tweet parameters (using obesity as the DV). The tweet parameter here is positive.
```{r accessibility_leaflet_2, echo=FALSE}
# Create a continuous palette function
pal <- colorQuantile(
  palette = "Greens",
  domain = c(0, tracts$park_ls2)
)

pop_content <- function(shp){
  paste0(
    "Tract: ", shp$NAME, "<br/>",
    "% Park: ", round(shp$Park_Percent, 2), "<br/>",
    "Park Access: ", round(shp$park_ls2, 2), "<br/>"
  )
}

leaflet(tracts %>% spTransform(., CRS("+init=epsg:4326"))) %>%
  addProviderTiles("CartoDB.Positron") %>%
  addPolygons(weight = 1, popup = ~pop_content(tracts),
              color = ~pal(park_ls2))
```

Here is a map of accessibility using the betas that produce the most likely model when using physical activity as the dependent variable. Only size and distance parameters are used.
```{r accessibility_leaflet_3, echo=FALSE}
# Create a continuous palette function
pal <- colorQuantile(
  palette = "Greens",
  domain = c(0, tracts$park_ls4)
)

pop_content <- function(shp){
  paste0(
    "Tract: ", shp$NAME, "<br/>",
    "% Park: ", round(shp$Park_Percent, 2), "<br/>",
    "Park Access: ", round(shp$park_ls4, 2), "<br/>"
  )
}

leaflet(tracts %>% spTransform(., CRS("+init=epsg:4326"))) %>%
  addProviderTiles("CartoDB.Positron") %>%
  addPolygons(weight = 1, popup = ~pop_content(tracts),
              color = ~pal(park_ls4))
```

And here is a map of accessibility using size, distance and tweet parameters (with the combination of betas that is most likely for physical activity as the DV).
```{r accessibility_leaflet_4, echo=FALSE}
# Create a continuous palette function
pal <- colorQuantile(
  palette = "Greens",
  domain = c(0, tracts$park_ls5)
)

pop_content <- function(shp){
  paste0(
    "Tract: ", shp$NAME, "<br/>",
    "% Park: ", round(shp$Park_Percent, 2), "<br/>",
    "Park Access: ", round(shp$park_ls5, 2), "<br/>"
  )
}

leaflet(tracts %>% spTransform(., CRS("+init=epsg:4326"))) %>%
  addProviderTiles("CartoDB.Positron") %>%
  addPolygons(weight = 1, popup = ~pop_content(tracts),
              color = ~pal(park_ls5))
```

```{r accessibility_leaflet_X, echo=FALSE, eval=FALSE}
# Create a continuous palette function
pal <- colorQuantile(
  palette = "Greens",
  domain = c(0, tracts$park_ls3)
)

pop_content <- function(shp){
  paste0(
    "Tract: ", shp$NAME, "<br/>",
    "% Park: ", round(shp$Park_Percent, 2), "<br/>",
    "Park Access: ", round(shp$park_ls3, 2), "<br/>"
  )
}

leaflet(tracts %>% spTransform(., CRS("+init=epsg:4326"))) %>%
  addProviderTiles("CartoDB.Positron") %>%
  addPolygons(weight = 1, popup = ~pop_content(tracts),
              color = ~pal(park_ls3))
```

```{r obese_lm, echo=FALSE, message=FALSE, include=FALSE}
base_formula <- formula(~log(Pop_Density) + FulltimeWork + 
    CollegeDeg + Single_Percent + PctWhite + PctBlack + PctAsian + PctOther + PctHispanic)
base_formula_age <- formula(~log(Pop_Density) + FulltimeWork + 
    CollegeDeg + Single_Percent + Pct18to29 + Pct30to64 + Pct65plus + PctWhite + PctBlack + PctAsian + PctOther + PctHispanic)

parks_formula_1 <- update(base_formula, . ~ . + log(park_ls1))
parks_formula_1a <- update(base_formula_age, . ~ . + log(park_ls1))
parks_formula_2 <- update(base_formula, . ~ . + log(park_ls2))
parks_formula_2a <- update(base_formula_age, . ~ . + log(park_ls2))
parks_formula_3 <- update(base_formula, . ~ . + log(park_ls4))
parks_formula_3a <- update(base_formula_age, . ~ . + log(park_ls4))
parks_formula_4 <- update(base_formula, . ~ . + log(park_ls5))
parks_formula_4a <- update(base_formula_age, . ~ . + log(park_ls5))

#parks_formula_3 <- update(base_formula, . ~ . + park_ls3)
```

Create a spatial weights matrix:

```{r W, echo=FALSE}
neighbors <- poly2nb(tracts, queen = FALSE)
W <- nb2listw(neighbours = neighbors, zero.policy = TRUE)
print(W, zero.policy = TRUE)
```

Estimate SDM models for the `OBESITY` variable. Model 1 does not include park access. Models 2 & 3 each use a different park access variable with a different combination of parameters:

```{r spmodels obesity, echo=FALSE}
obese_sdm <- lagsarlm(update(base_formula, OBESITY ~ .), data = tracts@data, 
                      listw = W, zero.policy = TRUE, type = "mixed")
obese_sdm2 <- lagsarlm(update(parks_formula_1, OBESITY ~ .), data = tracts@data, 
                      listw = W, zero.policy = TRUE, type = "mixed")
obese_sdm3 <- lagsarlm(update(parks_formula_2, OBESITY ~ .), data = tracts@data, 
                      listw = W, zero.policy = TRUE, type = "mixed")
obese_sdem <- errorsarlm(update(base_formula, OBESITY ~ .), data = tracts@data, 
                      listw = W, zero.policy = TRUE, etype = "emixed",tol.solve = 1e-16)
obese_sdem2 <- errorsarlm(update(parks_formula_1, OBESITY ~ .), data = tracts@data, 
                      listw = W, zero.policy = TRUE, etype = "emixed",tol.solve = 1e-16)
obese_sdem3 <- errorsarlm(update(parks_formula_2, OBESITY ~ .), data = tracts@data, 
                      listw = W, zero.policy = TRUE, etype = "emixed",tol.solve = 1e-16)
#obese_sdm_4 <- lagsarlm(update(parks_formula_3, OBESITY ~ .), data = tracts@data, 
                      #listw = W, zero.policy = TRUE, type = "mixed")
```

```{r spmodels obesity with age, echo=FALSE}
obese_sdm_age <- lagsarlm(update(base_formula_age, OBESITY ~ .), data = tracts@data, 
                      listw = W, zero.policy = TRUE, type = "mixed")
obese_sdm2_age <- lagsarlm(update(parks_formula_1a, OBESITY ~ .), data = tracts@data, 
                      listw = W, zero.policy = TRUE, type = "mixed")
obese_sdm3_age <- lagsarlm(update(parks_formula_2a, OBESITY ~ .), data = tracts@data, 
                      listw = W, zero.policy = TRUE, type = "mixed")
obese_sdem_age <- errorsarlm(update(base_formula_age, OBESITY ~ .), data = tracts@data, 
                      listw = W, zero.policy = TRUE, etype = "emixed",tol.solve = 1e-16)
obese_sdem2_age <- errorsarlm(update(parks_formula_1a, OBESITY ~ .), data = tracts@data, 
                      listw = W, zero.policy = TRUE, etype = "emixed",tol.solve = 1e-16)
obese_sdem3_age <- errorsarlm(update(parks_formula_2a, OBESITY ~ .), data = tracts@data, 
                      listw = W, zero.policy = TRUE, etype = "emixed",tol.solve = 1e-16)

```

Estimate SDM models for the `Phys_Act` variable. Model 1 does not include park access. Models 2-4 each use a different park access variable with a different combination of parameters:

```{r spmodels physact, echo=FALSE}
physact_sdm <- lagsarlm(update(base_formula, Phys_Act ~ .), data = tracts@data, 
                        listw = W, zero.policy = TRUE, type = "mixed")
physact_sdm2 <- lagsarlm(update(parks_formula_1, Phys_Act ~ .), data = tracts@data,
                         listw = W, zero.policy = TRUE, type = "mixed")
physact_sdm3 <- lagsarlm(update(parks_formula_2, Phys_Act ~ .), data = tracts@data,
                         listw = W, zero.policy = TRUE, type = "mixed")
physact_sdem <- errorsarlm(update(base_formula, OBESITY ~ .), data = tracts@data, 
                      listw = W, zero.policy = TRUE, etype = "emixed",tol.solve = 1e-16)
physact_sdem2 <- errorsarlm(update(parks_formula_1, OBESITY ~ .), data = tracts@data, 
                      listw = W, zero.policy = TRUE, etype = "emixed",tol.solve = 1e-16)
physact_sdem3 <- errorsarlm(update(parks_formula_2, OBESITY ~ .), data = tracts@data, 
                      listw = W, zero.policy = TRUE, etype = "emixed",tol.solve = 1e-16)
#physact_sdm4 <- lagsarlm(update(parks_formula_3, Phys_Act ~ .), data = tracts@data,
                         #listw = W, zero.policy = TRUE, type = "mixed")
```

```{r spmodels physact with age, echo=FALSE}
physact_sdm_age <- lagsarlm(update(base_formula_age, Phys_Act ~ .), data = tracts@data, 
                        listw = W, zero.policy = TRUE, type = "mixed")
physact_sdm2_age <- lagsarlm(update(parks_formula_1a, Phys_Act ~ .), data = tracts@data,
                         listw = W, zero.policy = TRUE, type = "mixed")
physact_sdm3_age <- lagsarlm(update(parks_formula_2a, Phys_Act ~ .), data = tracts@data,
                         listw = W, zero.policy = TRUE, type = "mixed")
physact_sdem_age <- errorsarlm(update(base_formula_age, OBESITY ~ .), data = tracts@data, 
                      listw = W, zero.policy = TRUE, etype = "emixed",tol.solve = 1e-16)
physact_sdem2_age <- errorsarlm(update(parks_formula_1a, OBESITY ~ .), data = tracts@data, 
                      listw = W, zero.policy = TRUE, etype = "emixed",tol.solve = 1e-16)
physact_sdem3_age <- errorsarlm(update(parks_formula_2a, OBESITY ~ .), data = tracts@data, 
                      listw = W, zero.policy = TRUE, etype = "emixed",tol.solve = 1e-16)

```

```{r spmodels physact new betas, echo=FALSE}
physact_sdm_new <- lagsarlm(update(base_formula, Phys_Act ~ .), data = tracts@data, 
                        listw = W, zero.policy = TRUE, type = "mixed")
physact_sdm2_new <- lagsarlm(update(parks_formula_3, Phys_Act ~ .), data = tracts@data,
                         listw = W, zero.policy = TRUE, type = "mixed")
physact_sdm3_new <- lagsarlm(update(parks_formula_4, Phys_Act ~ .), data = tracts@data,
                         listw = W, zero.policy = TRUE, type = "mixed")
physact_sdem_new <- errorsarlm(update(base_formula, OBESITY ~ .), data = tracts@data, 
                      listw = W, zero.policy = TRUE, etype = "emixed",tol.solve = 1e-16)
physact_sdem2_new <- errorsarlm(update(parks_formula_3, OBESITY ~ .), data = tracts@data, 
                      listw = W, zero.policy = TRUE, etype = "emixed",tol.solve = 1e-16)
physact_sdem3_new <- errorsarlm(update(parks_formula_4, OBESITY ~ .), data = tracts@data, 
                      listw = W, zero.policy = TRUE, etype = "emixed",tol.solve = 1e-16)
```

```{r spmodels physact age new betas, echo=FALSE}
physact_sdm_new_age <- lagsarlm(update(base_formula_age, Phys_Act ~ .), data = tracts@data, 
                        listw = W, zero.policy = TRUE, type = "mixed")
physact_sdm2_new_age <- lagsarlm(update(parks_formula_3a, Phys_Act ~ .), data = tracts@data,
                         listw = W, zero.policy = TRUE, type = "mixed")
physact_sdm3_new_age <- lagsarlm(update(parks_formula_4a, Phys_Act ~ .), data = tracts@data,
                         listw = W, zero.policy = TRUE, type = "mixed")
physact_sdem_new_age <- errorsarlm(update(base_formula_age, OBESITY ~ .), data = tracts@data, 
                      listw = W, zero.policy = TRUE, etype = "emixed",tol.solve = 1e-16)
physact_sdem2_new_age <- errorsarlm(update(parks_formula_3a, OBESITY ~ .), data = tracts@data, 
                      listw = W, zero.policy = TRUE, etype = "emixed",tol.solve = 1e-16)
physact_sdem3_new_age <- errorsarlm(update(parks_formula_4a, OBESITY ~ .), data = tracts@data, 
                      listw = W, zero.policy = TRUE, etype = "emixed",tol.solve = 1e-16)
```
Compare the models with `htmlreg()` in
the html output. First, the obesity models. The first table does not include the age variables.

```{r obesitysummary, results="asis", echo=FALSE}
obesity_models <- list(obese_sdm, obese_sdm2, obese_sdm3, obese_sdem, obese_sdem2, obese_sdem3)
names(obesity_models) <-  paste(c("SDM No Parks", "SDM Size & Distance", "SDM Size, Distance, Tweets", "SDEM 1", "SDEM 2", "SDEM 3"))
#screenreg(obesity_models, digits = 4)
htmlreg(obesity_models, caption = "NYC: SDM & SDEM Models of Obesity Prevalence")
```

```{r obesitysummaryage, results="asis", echo=FALSE}
obesity_models_age <- list(obese_sdm_age, obese_sdm2_age, obese_sdm3_age, obese_sdem_age, obese_sdem2_age, obese_sdem3_age)
names(obesity_models_age) <-  paste(c("SDM No Parks", "SDM Size & Distance", "SDM Size, Distance, Tweets", "SDEM 1", "SDEM 2", "SDEM 3"))
#screenreg(obesity_models, digits = 4)
htmlreg(obesity_models_age, caption = "NYC: SDM & SDEM Models of Obesity Prevalence")
```

Estimate the impacts for `obese_sdm`.
```{r obese_sdm_impacts}

listw <- as(W, "CsparseMatrix")
trMC <- trW(listw, type="MC")
set.seed(1)
summary(impacts(obese_sdm, tr=trMC, R=1000), zstats=TRUE)

```

Estimate the impacts for `obese_sdm_age`.
```{r obese_sdm_age_impacts}

listw <- as(W, "CsparseMatrix")
trMC <- trW(listw, type="MC")
set.seed(1)
summary(impacts(obese_sdm_age, tr=trMC, R=1000), zstats=TRUE)

```

Estimate the impacts for `obese_sdm2`.
```{r obese_sdm2_impacts}

summary(impacts(obese_sdm2, tr=trMC, R=1000), zstats=TRUE)

```

Estimate the impacts for `obese_sdm2_age`.
```{r obese_sdm2_age_impacts}

summary(impacts(obese_sdm2_age, tr=trMC, R=1000), zstats=TRUE)

```

Estimate the impacts for `obese_sdm3`.
```{r obese_sdm3_impacts}

summary(impacts(obese_sdm3, tr=trMC, R=1000), zstats=TRUE)

```

Estimate the impacts for `obese_sdm3_age`.
```{r obese_sdm3_age_impacts}

summary(impacts(obese_sdm3_age, tr=trMC, R=1000), zstats=TRUE)

```

Next, the physical activity models. Again, the first table does not contain the age variables. The third and fourth tables use new beta parameters.

```{r physactsummary, results="asis", echo=FALSE}
physact_models <- list(physact_sdm, physact_sdm2, physact_sdm3, physact_sdem, physact_sdem2, physact_sdem3)
names(physact_models) <-  paste(c("SDM No Parks", "SDM Size & Distance", "SDM Size, Distance, Tweets", "SDEM 1", "SDEM 2", "SDEM 3"))
#screenreg(physact_models)
htmlreg(physact_models, caption = "NYC: SDM Models of Physical Activity Participation")
```

```{r physactsummaryage, results="asis", echo=FALSE}
physact_models_age <- list(physact_sdm_age, physact_sdm2_age, physact_sdm3_age, physact_sdem_age, physact_sdem2_age, physact_sdem3_age)
names(physact_models_age) <-  paste(c("SDM No Parks", "SDM Size & Distance", "SDM Size, Distance, Tweets", "SDEM 1", "SDEM 2", "SDEM 3"))
#screenreg(physact_models)
htmlreg(physact_models_age, caption = "NYC: SDM Models of Physical Activity Participation")
```

The following two tables use a size beta of 0.00001 and a distance beta of -15.
```{r physact new betas, results="asis", echo=FALSE}
physact_models_new <- list(physact_sdm_new, physact_sdm2_new, physact_sdm3_new, physact_sdem_new, physact_sdem2_new, physact_sdem3_new)
names(physact_models_new) <-  paste(c("SDM No Parks", "SDM Size & Distance", "SDM Size, Distance, Tweets", "SDEM 1", "SDEM 2", "SDEM 3"))
#screenreg(physact_models)
htmlreg(physact_models_new, caption = "NYC: SDM Models of Physical Activity Participation")

```

```{r physact new betas with age, results="asis", echo=FALSE}
physact_models_new_age <- list(physact_sdm_new_age, physact_sdm2_new_age, physact_sdm3_new_age, physact_sdem_new_age, physact_sdem2_new_age, physact_sdem3_new_age)
names(physact_models_new_age) <-  paste(c("SDM No Parks", "SDM Size & Distance", "SDM Size, Distance, Tweets", "SDEM 1", "SDEM 2", "SDEM 3"))
#screenreg(physact_models)
htmlreg(physact_models_new_age, caption = "NYC: SDM Models of Physical Activity Participation")

```

Estimate the impacts for `physact_sdm`.
```{r physact_sdm_impacts}

summary(impacts(physact_sdm, tr=trMC, R=1000), zstats=TRUE)

```

Estimate the impacts for `physact_sdm_age`.
```{r physact_sdm_age_impacts}

summary(impacts(physact_sdm_age, tr=trMC, R=1000), zstats=TRUE)

```

Estimate the impacts for `physact_sdm2`.
```{r physact_sdm2_impacts}

summary(impacts(physact_sdm2, tr=trMC, R=1000), zstats=TRUE)

```

Estimate the impacts for `physact_sdm2_age`.
```{r physact_sdm2_age_impacts}

summary(impacts(physact_sdm2_age, tr=trMC, R=1000), zstats=TRUE)

```

Estimate the impacts for `physact_sdm3`.
```{r physact_sdm3_impacts}

summary(impacts(physact_sdm3, tr=trMC, R=1000), zstats=TRUE)

```

Estimate the impacts for `physact_sdm3_age`.
```{r physact_sdm3_age_impacts}

summary(impacts(physact_sdm3_age, tr=trMC, R=1000), zstats=TRUE)

```

Estimate the impacts for `physact_sdm3_new_age`.
```{r physact_sdm3_new_age_impacts}

summary(impacts(physact_sdm3_new_age, tr = trMC, R = 1000), zstats=TRUE)

```

### 2018-08-10 size00001dist15.R

@@ -1,0 +1,51 @@
+#' Calculate destination choice accessibility logsum
+#' 
+#' @param size A vector of park sizes in acres
+#' @param distance A vector of distances to the parks in `size` in miles
+#' @param betas A vector containing the size and distance coefficients
+#'
+#' @return The accessibility logsum
+#' 
+park_logsum <- function(size, distance, betas = c(.00001, -15)){
+  
+  # calculate exponentiated utility of all parks
+  U <- betas[1] * size + betas[2] * distance 
+  eU <- exp(U)
+  
+  # log-sum
+  log(sum(eU))
+}
+
+
+#' Calculate destination choice logsums from a distance matrix
+#' 
+#' @param tracts A spatial points dataframe containing the 
+#'   tract geographic information and their attributes.
+#' @param parks A spatial points dataframe containing the 
+#'   park geographic information and their attributes.
+#' @param betas A vector containing the size and distance coefficients
+#'   
+#' @return An accessibility logsum calculation for each tract to every park.
+#' 
+calculate_park_logsums <- function(tracts, parks, betas = c(.00001, -15)){
+  
+  # calculate centroids of tracts and parks
+  tract_centroids <- gCentroid(tracts, byid = TRUE) 
+  park_centroids <- gCentroid(parks, byid = TRUE)
+  
+  # distances from all tracts to all zones
+  distances <- gDistance(park_centroids, tract_centroids, byid = TRUE) / 5280
+  
+  # vector containing park sizes 
+  park_sizes <- parks$Park_Acres
+  
+  # calculate observed utility 
+  #   beta_size * size + beta_distane * distance
+  V <- sweep(betas[2] * distances, MARGIN = 2, 
+             betas[1] * park_sizes, `+`)
+  
+  # log-sum of exponentiated utility
+  log(rowSums(exp(V)))
+  
+  
+}

### 2018-08-10 size00001dist15tweet001.R

@@ -1,0 +1,55 @@
+#' Calculate destination choice accessibility logsum
+#' 
+#' @param size A vector of park sizes in acres
+#' @param distance A vector of distances to the parks in `size` in miles
+#' @param tweets A vector of the number of tweets recorded within each park in September 2014
+#' @param betas A vector containing the size and distance coefficients
+#'
+#' @return The accessibility logsum
+#' 
+park_logsum <- function(size, distance, betas = c(.00001, -15, 1)){
+  
+  # calculate exponentiated utility of all parks
+  U <- betas[1] * size + betas[2] * distance + betas[3] * park_tweets
+  eU <- exp(U)
+  
+  # log-sum
+  log(sum(eU))
+}
+
+
+#' Calculate destination choice logsums from a distance matrix
+#' 
+#' @param tracts A spatial points dataframe containing the 
+#'   tract geographic information and their attributes.
+#' @param parks A spatial points dataframe containing the 
+#'   park geographic information and their attributes.
+#' @param betas A vector containing the size and distance coefficients
+#'   
+#' @return An accessibility logsum calculation for each tract to every park.
+#' 
+calculate_park_logsums <- function(tracts, parks, betas = c(.00001, -15, .001)){
+  
+  # calculate centroids of tracts and parks
+  tract_centroids <- gCentroid(tracts, byid = TRUE) 
+  park_centroids <- gCentroid(parks, byid = TRUE)
+  
+  # distances from all tracts to all zones
+  distances <- gDistance(park_centroids, tract_centroids, byid = TRUE) / 5280
+  
+  # vector containing park sizes 
+  park_sizes <- parks$Park_Acres
+  
+  # vector containing tweet counts
+  park_tweets <- parks$TWEET_COUNT
+  
+  # calculate observed utility 
+  #   beta_size * size + beta_distance * distance + beta_tweet * tweets
+  V <- sweep(betas[2] * distances, MARGIN = 2, 
+             betas[1] * park_sizes + betas[3] * log(park_tweets + 0.01), `+`)
+  
+  # log-sum of exponentiated utility
+  log(rowSums(exp(V)))
+  
+  
+}

### 2018-08-10 size00001dist5tweet001.R

@@ -1,0 +1,55 @@
+#' Calculate destination choice accessibility logsum
+#' 
+#' @param size A vector of park sizes in acres
+#' @param distance A vector of distances to the parks in `size` in miles
+#' @param tweets A vector of the number of tweets recorded within each park in September 2014
+#' @param betas A vector containing the size and distance coefficients
+#'
+#' @return The accessibility logsum
+#' 
+park_logsum <- function(size, distance, betas = c(.00001, -5, 1)){
+  
+  # calculate exponentiated utility of all parks
+  U <- betas[1] * size + betas[2] * distance + betas[3] * park_tweets
+  eU <- exp(U)
+  
+  # log-sum
+  log(sum(eU))
+}
+
+
+#' Calculate destination choice logsums from a distance matrix
+#' 
+#' @param tracts A spatial points dataframe containing the 
+#'   tract geographic information and their attributes.
+#' @param parks A spatial points dataframe containing the 
+#'   park geographic information and their attributes.
+#' @param betas A vector containing the size and distance coefficients
+#'   
+#' @return An accessibility logsum calculation for each tract to every park.
+#' 
+calculate_park_logsums <- function(tracts, parks, betas = c(.00001, -5, .001)){
+  
+  # calculate centroids of tracts and parks
+  tract_centroids <- gCentroid(tracts, byid = TRUE) 
+  park_centroids <- gCentroid(parks, byid = TRUE)
+  
+  # distances from all tracts to all zones
+  distances <- gDistance(park_centroids, tract_centroids, byid = TRUE) / 5280
+  
+  # vector containing park sizes 
+  park_sizes <- parks$Park_Acres
+  
+  # vector containing tweet counts
+  park_tweets <- parks$TWEET_COUNT
+  
+  # calculate observed utility 
+  #   beta_size * size + beta_distance * distance + beta_tweet * tweets
+  V <- sweep(betas[2] * distances, MARGIN = 2, 
+             betas[1] * park_sizes + betas[3] * log(park_tweets + 0.01), `+`)
+  
+  # log-sum of exponentiated utility
+  log(rowSums(exp(V)))
+  
+  
+}

### 2018-08-10 size00001dist5tweetneg001.R

@@ -1,0 +1,55 @@
+#' Calculate destination choice accessibility logsum
+#' 
+#' @param size A vector of park sizes in acres
+#' @param distance A vector of distances to the parks in `size` in miles
+#' @param tweets A vector of the number of tweets recorded within each park in September 2014
+#' @param betas A vector containing the size and distance coefficients
+#'
+#' @return The accessibility logsum
+#' 
+park_logsum <- function(size, distance, betas = c(.00001, -5, 1)){
+  
+  # calculate exponentiated utility of all parks
+  U <- betas[1] * size + betas[2] * distance + betas[3] * park_tweets
+  eU <- exp(U)
+  
+  # log-sum
+  log(sum(eU))
+}
+
+
+#' Calculate destination choice logsums from a distance matrix
+#' 
+#' @param tracts A spatial points dataframe containing the 
+#'   tract geographic information and their attributes.
+#' @param parks A spatial points dataframe containing the 
+#'   park geographic information and their attributes.
+#' @param betas A vector containing the size and distance coefficients
+#'   
+#' @return An accessibility logsum calculation for each tract to every park.
+#' 
+calculate_park_logsums <- function(tracts, parks, betas = c(.00001, -5, -10)){
+  
+  # calculate centroids of tracts and parks
+  tract_centroids <- gCentroid(tracts, byid = TRUE) 
+  park_centroids <- gCentroid(parks, byid = TRUE)
+  
+  # distances from all tracts to all zones
+  distances <- gDistance(park_centroids, tract_centroids, byid = TRUE) / 5280
+  
+  # vector containing park sizes 
+  park_sizes <- parks$Park_Acres
+  
+  # vector containing tweet counts
+  park_tweets <- parks$TWEET_COUNT
+  
+  # calculate observed utility 
+  #   beta_size * size + beta_distance * distance + beta_tweet * tweets
+  V <- sweep(betas[2] * distances, MARGIN = 2, 
+             betas[1] * park_sizes + betas[3] * log(park_tweets), `+`)
+  
+  # log-sum of exponentiated utility
+  log(rowSums(exp(V)))
+  
+  
+}

### 2018-08-14 NYC_Twitter_SDM_models - Copy.Rmd
file swap
deleted wrong file and uploaded correct one

---
title: "NYC Spatial Models - Comparison of Spatial Models With and Without Twitter Data"
author: "Nico Boyd"
date: "August 10, 2017"
output: html_document
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
library(geojsonio)
library(tidyverse)
library(maptools)
library(rgdal)
library(rgeos)
library(spdep)
library(texreg)
library(leaflet)
```

```{r read_tracts, echo=FALSE}
tracts <- geojson_read("nyc_tracts.geojson", what = "sp")
#plot(tracts)
```

```{r read_parks, echo=FALSE}
parks <- geojson_read("nyc_parks.geojson", what = "sp")
#plot(parks)
```

```{r show_data, echo=FALSE}
tracts@data <- tbl_df(tracts@data) %>%
  select(GEOID, OBESITY, Park_Percent, Phys_Act, MENTAL, Income1, Income2, Income3, Income4, Income5, Income6, Income7, Income8, Income9, Income10, Pop_Density, FulltimeWork, CollegeDeg, Pct0to17, Pct18to29, Pct30to64, Pct65plus, Single_Percent, PctWhite, PctBlack, PctNative, PctAsian, PctPacific, PctOther, PctHispanic) %>%
    mutate_each(funs(ifelse(is.na(.), 0, .)), Park_Percent)

parks@data <- tbl_df(parks@data)

tracts@data
#parks@data
```

Calculate the logsum given a tracts and parks dataset. `tracts$park_ls1` uses only size and distance parameters. `tracts$park_ls2` uses size, distance and tweet parameters. `tracts$park_ls4` and `tracts$park_ls5` use a distance parameter of -15.
```{r accessibility}
source('size00001dist5.R', echo=TRUE)
tracts$park_ls1 <- calculate_park_logsums(tracts, parks, betas = c(.00001, -5))

source('size00001dist5tweet001.R', echo=TRUE)
tracts$park_ls2 <- calculate_park_logsums(tracts, parks, betas = c(.00001, -5, .001))

#source('size00001dist5tweetneg001.R', echo=TRUE)
#tracts$park_ls3 <- calculate_park_logsums(tracts, parks, betas = c(.00001, -5, -.001))

source('size00001dist15.R', echo=TRUE)
tracts$park_ls4 <- calculate_park_logsums(tracts, parks, betas = c(.00001, -15))

source('size00001dist15tweet001.R', echo=TRUE)
tracts$park_ls5 <- calculate_park_logsums(tracts, parks, betas = c(.00001, -15, .001))
```

Map the newly created Park Access variables. Here is a map of accessibility using only the size and distance parameters that are most likely when using obesity as the dependent variable.
```{r accessibility_leaflet_1, echo=FALSE}
# Create a continuous palette function
pal <- colorQuantile(
  palette = "Greens",
  domain = c(0, tracts$park_ls1)
)

pop_content <- function(shp){
  paste0(
    "Tract: ", shp$NAME, "<br/>",
    "% Park: ", round(shp$Park_Percent, 2), "<br/>",
    "Park Access: ", round(shp$park_ls1, 2), "<br/>"
  )
}

leaflet(tracts %>% spTransform(., CRS("+init=epsg:4326"))) %>%
  addProviderTiles("CartoDB.Positron") %>%
  addPolygons(weight = 1, popup = ~pop_content(tracts),
              color = ~pal(park_ls1))
```

And here is a map of accessibility using the size, distance and tweet parameters (using obesity as the DV). 
```{r accessibility_leaflet_2, echo=FALSE}
# Create a continuous palette function
pal <- colorQuantile(
  palette = "Greens",
  domain = c(0, tracts$park_ls2)
)

pop_content <- function(shp){
  paste0(
    "Tract: ", shp$NAME, "<br/>",
    "% Park: ", round(shp$Park_Percent, 2), "<br/>",
    "Park Access: ", round(shp$park_ls2, 2), "<br/>"
  )
}

leaflet(tracts %>% spTransform(., CRS("+init=epsg:4326"))) %>%
  addProviderTiles("CartoDB.Positron") %>%
  addPolygons(weight = 1, popup = ~pop_content(tracts),
              color = ~pal(park_ls2))
```

Here is a map of accessibility using the betas that produce the most likely model when using physical activity as the dependent variable. Only size and distance parameters are used.
```{r accessibility_leaflet_3, echo=FALSE}
# Create a continuous palette function
pal <- colorQuantile(
  palette = "Greens",
  domain = c(0, tracts$park_ls4)
)

pop_content <- function(shp){
  paste0(
    "Tract: ", shp$NAME, "<br/>",
    "% Park: ", round(shp$Park_Percent, 2), "<br/>",
    "Park Access: ", round(shp$park_ls4, 2), "<br/>"
  )
}

leaflet(tracts %>% spTransform(., CRS("+init=epsg:4326"))) %>%
  addProviderTiles("CartoDB.Positron") %>%
  addPolygons(weight = 1, popup = ~pop_content(tracts),
              color = ~pal(park_ls4))
```

And here is a map of accessibility using size, distance and tweet parameters (with the combination of betas that is most likely for physical activity as the DV).
```{r accessibility_leaflet_4, echo=FALSE}
# Create a continuous palette function
pal <- colorQuantile(
  palette = "Greens",
  domain = c(0, tracts$park_ls5)
)

pop_content <- function(shp){
  paste0(
    "Tract: ", shp$NAME, "<br/>",
    "% Park: ", round(shp$Park_Percent, 2), "<br/>",
    "Park Access: ", round(shp$park_ls5, 2), "<br/>"
  )
}

leaflet(tracts %>% spTransform(., CRS("+init=epsg:4326"))) %>%
  addProviderTiles("CartoDB.Positron") %>%
  addPolygons(weight = 1, popup = ~pop_content(tracts),
              color = ~pal(park_ls5))
```

Create regression formulas:
```{r obese_lm, echo=FALSE, message=FALSE, include=FALSE}
#base_formula <- formula(~log(Pop_Density) + FulltimeWork + 
    #CollegeDeg + Single_Percent + PctWhite + PctBlack + PctAsian + PctOther + PctHispanic)
base_formula_age <- formula(~log(Pop_Density) + FulltimeWork + 
    CollegeDeg + Single_Percent + Pct0to17 + Pct18to29 + Pct30to64 + Pct65plus + PctWhite + PctBlack + PctAsian + PctOther + PctHispanic)

#parks_formula_1 <- update(base_formula, . ~ . + log(park_ls1 + 0.1))
parks_formula_1a <- update(base_formula_age, . ~ . + log(park_ls1 + 0.1))
#parks_formula_2 <- update(base_formula, . ~ . + log(park_ls2 + 0.1))
parks_formula_2a <- update(base_formula_age, . ~ . + log(park_ls2 + 0.1))
#parks_formula_3 <- update(base_formula, . ~ . + log(park_ls4 + 0.01))
parks_formula_3a <- update(base_formula_age, . ~ . + log(park_ls4 + 5))
#parks_formula_4 <- update(base_formula, . ~ . + log(park_ls5 + 0.1))
parks_formula_4a <- update(base_formula_age, . ~ . + log(park_ls5 + 5))

```

Create a spatial weights matrix:

```{r W, echo=FALSE}
neighbors <- poly2nb(tracts, queen = FALSE)
W <- nb2listw(neighbours = neighbors, zero.policy = TRUE)
print(W, zero.policy = TRUE)
```

Estimate spatial models for the `OBESITY` variable. Model 1 of each type does not include park access. Models 2 & 3 each use a different park access variable with a different combination of parameters:

```{r spmodels obesity, echo=FALSE, message=FALSE, warning=FALSE}
obese_sar_age <- lagsarlm(update(base_formula_age, OBESITY ~ .), data = tracts@data, 
                      listw = W, zero.policy = TRUE)
obese_sar2_age <- lagsarlm(update(parks_formula_1a, OBESITY ~ .), data = tracts@data,
                      listw = W, zero.policy = TRUE)
obese_sar3_age <- lagsarlm(update(parks_formula_2a, OBESITY ~ .), data = tracts@data,
                     listw = W, zero.policy = TRUE)
obese_sem_age <- errorsarlm(update(base_formula_age, OBESITY ~ .), data = tracts@data, 
                      listw = W, zero.policy = TRUE)
obese_sem2_age <- errorsarlm(update(parks_formula_1a, OBESITY ~ .), data = tracts@data, 
                      listw = W, zero.policy = TRUE)
obese_sem3_age <- errorsarlm(update(parks_formula_2a, OBESITY ~ .), data = tracts@data, 
                      listw = W, zero.policy = TRUE)
obese_sdm_age <- lagsarlm(update(base_formula_age, OBESITY ~ .), data = tracts@data, 
                      listw = W, zero.policy = TRUE, type = "mixed")
obese_sdm2_age <- lagsarlm(update(parks_formula_1a, OBESITY ~ .), data = tracts@data, 
                      listw = W, zero.policy = TRUE, type = "mixed")
obese_sdm3_age <- lagsarlm(update(parks_formula_2a, OBESITY ~ .), data = tracts@data, 
                      listw = W, zero.policy = TRUE, type = "mixed")
obese_sdem_age <- errorsarlm(update(base_formula_age, OBESITY ~ .), data = tracts@data, 
                      listw = W, zero.policy = TRUE, etype = "emixed",tol.solve = 1e-16)
obese_sdem2_age <- errorsarlm(update(parks_formula_1a, OBESITY ~ .), data = tracts@data, 
                      listw = W, zero.policy = TRUE, etype = "emixed",tol.solve = 1e-16)
obese_sdem3_age <- errorsarlm(update(parks_formula_2a, OBESITY ~ .), data = tracts@data, 
                      listw = W, zero.policy = TRUE, etype = "emixed",tol.solve = 1e-16)

```

Estimate spatial models for the `Phys_Act` variable. Model 1 does not include park access. Models 2 & 3 each use a different park access variable with a different combination of parameters:

```{r spmodels physact with age, echo=FALSE, message=FALSE, warning=FALSE}

physact_sar_age <- lagsarlm(update(base_formula_age, Phys_Act ~ .), data = tracts@data, 
                      listw = W, zero.policy = TRUE)
physact_sar2_age <- lagsarlm(update(parks_formula_1a, Phys_Act ~ .), data = tracts@data,
                      listw = W, zero.policy = TRUE)
physact_sar3_age <- lagsarlm(update(parks_formula_2a, Phys_Act ~ .), data = tracts@data,
                     listw = W, zero.policy = TRUE)
physact_sem_age <- errorsarlm(update(base_formula_age, Phys_Act ~ .), data = tracts@data, 
                      listw = W, zero.policy = TRUE)
physact_sem2_age <- errorsarlm(update(parks_formula_1a, Phys_Act ~ .), data = tracts@data, 
                      listw = W, zero.policy = TRUE)
physact_sem3_age <- errorsarlm(update(parks_formula_2a, Phys_Act ~ .), data = tracts@data, 
                      listw = W, zero.policy = TRUE)
physact_sdm_age <- lagsarlm(update(base_formula_age, Phys_Act ~ .), data = tracts@data, 
                        listw = W, zero.policy = TRUE, type = "mixed")
physact_sdm2_age <- lagsarlm(update(parks_formula_1a, Phys_Act ~ .), data = tracts@data,
                         listw = W, zero.policy = TRUE, type = "mixed")
physact_sdm3_age <- lagsarlm(update(parks_formula_2a, Phys_Act ~ .), data = tracts@data,
                         listw = W, zero.policy = TRUE, type = "mixed")
physact_sdem_age <- errorsarlm(update(base_formula_age, Phys_Act ~ .), data = tracts@data, 
                      listw = W, zero.policy = TRUE, etype = "emixed",tol.solve = 1e-16)
physact_sdem2_age <- errorsarlm(update(parks_formula_1a, Phys_Act ~ .), data = tracts@data, 
                      listw = W, zero.policy = TRUE, etype = "emixed",tol.solve = 1e-16)
physact_sdem3_age <- errorsarlm(update(parks_formula_2a, Phys_Act ~ .), data = tracts@data, 
                      listw = W, zero.policy = TRUE, etype = "emixed",tol.solve = 1e-16)

```

Estimate additional spatial models for the `Phys_Act` variable using a new distance parameter of -15.
```{r spmodels physact age new betas, echo=FALSE}
physact_sar_new_age <- lagsarlm(update(base_formula_age, Phys_Act ~ .), data = tracts@data, 
                      listw = W, zero.policy = TRUE)
physact_sar2_new_age <- lagsarlm(update(parks_formula_3a, Phys_Act ~ .), data = tracts@data,
                           listw = W, zero.policy = TRUE)
physact_sar3_new_age <- lagsarlm(update(parks_formula_4a, Phys_Act ~ .), data = tracts@data,
                           listw = W, zero.policy = TRUE)
physact_sem_new_age <- errorsarlm(update(base_formula_age, Phys_Act ~ .), data = tracts@data, 
                      listw = W, zero.policy = TRUE)
physact_sem2_new_age <- errorsarlm(update(parks_formula_3a, Phys_Act ~ .), data = tracts@data, 
                      listw = W, zero.policy = TRUE)
physact_sem3_new_age <- errorsarlm(update(parks_formula_4a, Phys_Act ~ .), data = tracts@data, 
                      listw = W, zero.policy = TRUE)
physact_sdm_new_age <- lagsarlm(update(base_formula_age, Phys_Act ~ .), data = tracts@data, 
                        listw = W, zero.policy = TRUE, type = "mixed")
physact_sdm2_new_age <- lagsarlm(update(parks_formula_3a, Phys_Act ~ .), data = tracts@data,
                         listw = W, zero.policy = TRUE, type = "mixed")
physact_sdm3_new_age <- lagsarlm(update(parks_formula_4a, Phys_Act ~ .), data = tracts@data,
                         listw = W, zero.policy = TRUE, type = "mixed")
physact_sdem_new_age <- errorsarlm(update(base_formula_age, Phys_Act ~ .), data = tracts@data, 
                      listw = W, zero.policy = TRUE, etype = "emixed",tol.solve = 1e-16)
physact_sdem2_new_age <- errorsarlm(update(parks_formula_3a, Phys_Act ~ .), data = tracts@data, 
                      listw = W, zero.policy = TRUE, etype = "emixed",tol.solve = 1e-16)
physact_sdem3_new_age <- errorsarlm(update(parks_formula_4a, Phys_Act ~ .), data = tracts@data, 
                      listw = W, zero.policy = TRUE, etype = "emixed",tol.solve = 1e-16)
```

Compare the models with `htmlreg()` in
the html output. First, the obesity models.

```{r obesitysummaryage, results="asis", echo=FALSE}
obesity_models_age <- list(obese_sar_age, obese_sar2_age, obese_sar3_age, obese_sem_age, obese_sem2_age, obese_sem3_age, obese_sdm_age, obese_sdm2_age, obese_sdm3_age, obese_sdem_age, obese_sdem2_age, obese_sdem3_age)
names(obesity_models_age) <-  paste(rep(c("SAR", "SEM" , "SDM", "SDEM"), each = 3),
                                rep(c("1", "2", "3")))
htmlreg(obesity_models_age, caption = "NYC: Spatial Models of Obesity Prevalence")
```

Estimate the impacts for `obese_sdm_age`.
```{r obese_sdm_age_impacts}

listw <- as(W, "CsparseMatrix")
trMC <- trW(listw, type="MC")
set.seed(1)
#summary(impacts(obese_sdm_age, tr=trMC, R=1000), zstats=TRUE)

```

Estimate the impacts for `obese_sdm2_age`.
```{r obese_sdm2_age_impacts}

#summary(impacts(obese_sdm2_age, tr=trMC, R=1000), zstats=TRUE)

```

Estimate the impacts for `obese_sdm3_age`.
```{r obese_sdm3_age_impacts}

#summary(impacts(obese_sdm3_age, tr=trMC, R=1000), zstats=TRUE)

```

Next, the physical activity models.  The second table uses new beta parameters.

```{r physactsummaryage, results="asis", echo=FALSE}
physact_models_age <- list(physact_sar_age, physact_sar2_age, physact_sar3_age, physact_sem_age, physact_sem2_age, physact_sem3_age, physact_sdm_age, physact_sdm2_age, physact_sdm3_age, physact_sdem_age, physact_sdem2_age, physact_sdem3_age)
names(physact_models_age) <-  paste(rep(c("SAR", "SEM" , "SDM", "SDEM"), each = 3),
                                rep(c("1", "2", "3")))
htmlreg(physact_models_age, caption = "NYC: SDM Models of Physical Activity Participation")
```

The following table use a size beta of 0.00001 and a distance beta of -15.

```{r physact new betas with age, results="asis", echo=FALSE}
physact_models_new_age <- list(physact_sar_new_age, physact_sar2_new_age, physact_sar3_new_age, physact_sem_new_age, physact_sem2_new_age, physact_sem3_new_age, physact_sdm_new_age, physact_sdm2_new_age, physact_sdm3_new_age, physact_sdem_new_age, physact_sdem2_new_age, physact_sdem3_new_age)
names(physact_models_new_age) <-  paste(rep(c("SAR", "SEM" , "SDM", "SDEM"), each = 3),
                                rep(c("1", "2", "3")))
htmlreg(physact_models_new_age, caption = "NYC: Spatial Models of Physical Activity Participation")

```

Estimate the impacts for `physact_sdm_age`.
```{r physact_sdm_age_impacts}

#summary(impacts(physact_sdm_age, tr=trMC, R=1000), zstats=TRUE)

```

Estimate the impacts for `physact_sdm2_age`.
```{r physact_sdm2_age_impacts}

#summary(impacts(physact_sdm2_age, tr=trMC, R=1000), zstats=TRUE)

```

Estimate the impacts for `physact_sdm3_age`.
```{r physact_sdm3_age_impacts}

#summary(impacts(physact_sdm3_age, tr=trMC, R=1000), zstats=TRUE)

```

Estimate the impacts for `physact_sdm_new_age`.
```{r physact_sdm_new_age_impacts}

#summary(impacts(physact_sdm_new_age, tr = trMC, R = 1000), zstats=TRUE)

```

Estimate the impacts for `physact_sdm2_new_age`.
```{r physact_sdm2_new_age_impacts}

#summary(impacts(physact_sdm2_new_age, tr = trMC, R = 1000), zstats=TRUE)

```

Estimate the impacts for `physact_sdm3_new_age`.
```{r physact_sdm3_new_age_impacts}

#summary(impacts(physact_sdm3_new_age, tr = trMC, R = 1000), zstats=TRUE)

```

### 2018-11-02 NYC_Twitter_SDM_models - Copy.Rmd
Look for missingness in tracts and parks datasets

@@
 ---
+
+This document contains the model estimation and impacts results for the paper.
 
@@
 library(leaflet)
+library(sf)
 ```
 
-```{r read_tracts, echo=FALSE}
+Nico cleaned the tracts dataset into a geoJSON file stored in the repository. We
+should read it in twice because the sf library is better for visualization but
+spdep still requires sp objects for now.
+
+```{r read_tracts}
 tracts <- geojson_read("nyc_tracts.geojson", what = "sp")
-#plot(tracts)
+tracts_sf <- st_read("nyc_tracts.geojson", quiet = TRUE) %>%
+  st_transform(4326)
 ```
+
+## Complete Cases
+We noticed that our models had different numbers of observations / degrees of 
+freedom. Obviously we need to have comparable econometric models. The first thing
+we want to look for is missingness in the tracts variables, and which tracts
+are missing those variables. The map below shows the tracts that are missing
+this information.
+
+```{r tracts_with_missing}
+# identify complete cases
+complete_tracts_index <- tracts_sf %>%
+  st_set_geometry(NULL) %>%
+  tbl_df() %>%
+  complete.cases()
+
+# make a map of 
+leaflet( tracts_sf[!complete_tracts_index, ]) %>%
+  addProviderTiles(providers$OpenStreetMap.Mapnik) %>% 
+  addPolygons(label = ~ as.character(GEOID))
+```
+
+These tracts seem to fall into two basic categories:
+
+  - Tracts outside of New York City that were clipped into the shape layer 
+  and remain as edge remnants. It is generally better to select by FIPS code rather
+  than geometric operations, but we can just discard these at this point and
+  they will not affect anything.
+
+  - Tracts that are completely occupied by parks, airports, or other pieces
+  of non-residential infrastructure.
+
+The exceptions to this appear to be 3606100940 and 36061010200 in Midtown between
+5th Avenue and Park Avenue. Greg will assume for this analysis that these are 
+entirely commercial tracts and therefore have no residents, but Nico needs to 
+verify these.
+
+```{r complete_cases}
+tracts <- tracts[complete_tracts_index, ]
+```
+
+The parks data seems relatively complete, with the exception of a few variables
+that are missing everywhere. But we'll ignore this for now.
 
@@
 parks <- geojson_read("nyc_parks.geojson", what = "sp")
+parks_sf <- st_read("nyc_parks.geojson") %>%
+  st_transform(4326)
 #plot(parks)
@@
 tracts@data <- tbl_df(tracts@data) %>%
-  select(GEOID, OBESITY, Park_Percent, Phys_Act, MENTAL, Income1, Income2, Income3, Income4, Income5, Income6, Income7, Income8, Income9, Income10, Pop_Density, FulltimeWork, CollegeDeg, Pct0to17, Pct18to29, Pct30to64, Pct65plus, Single_Percent, PctWhite, PctBlack, PctNative, PctAsian, PctPacific, PctOther, PctHispanic) %>%
-    mutate_each(funs(ifelse(is.na(.), 0, .)), Park_Percent)
+  select(
+    GEOID, OBESITY, Park_Percent, Phys_Act, MENTAL, 
+    Income1, Income2, Income3, Income4, Income5, Income6, 
+    Income7, Income8, Income9, Income10, Pop_Density, 
+    FulltimeWork, CollegeDeg, Pct0to17, Pct18to29, Pct30to64, Pct65plus, 
+    Single_Percent, PctWhite, PctBlack, PctNative, PctAsian, PctPacific,
+    PctOther, PctHispanic
+  )
 
@@
 tracts@data
-#parks@data
+parks@data
 ```

### 2018-11-02 NYC_Twitter_SDM_models - Copy.Rmd
Replace multiple scripts with a single function

@@
 
-Calculate the logsum given a tracts and parks dataset.
-`tracts$park_ls1` uses only size and distance parameters.
-`tracts$park_ls2` uses size, distance and tweet parameters.
-`tracts$park_ls4` and `tracts$park_ls5` use a distance parameter of -15.
+Instead of keeping those functions in separate scripts, we are going to put
+everything right here in the document.
+
+```{r logsumfunction}
+
+#' Calculate destination choice logsums from a distance matrix
+#' 
+#' @param tracts A spatial points dataframe containing $n$
+#'   tract geographic information and their attributes.
+#' @param parks A spatial points dataframe containing $p$
+#'   park geographic information and their attributes.
+#' @param distances An $n\times p$ matrix with the distance from all tracts to
+#'   all parks in miles.
+#' @param sizes A p-length vector of park sizes
+#' @param tweets A p-length vector of tweets at parks
+#' @param betas A vector containing the size, distance, and tweet coefficients. 
+#'   If we submit two variables the tweets are ignored.
+#'   
+#' @return An n-length vector containing the weighted log-sum based
+#'   accessibility between a tract and all parks.
+#' @details If we have n tracts and p parks, distances needs to be a 
+#' 
+calculate_park_logsums <- function(tracts, parks, distances, sizes, tweets = NULL,
+                                   betas = c(.00001, -15, .001)){
+
+  # A is n x p
+  a <- betas[2] * distances
+
+  # B is p x 1
+  b <- betas[1] * sizes
+
+  if(!is.null(tweets)) 
+    b <- b + betas[3] * tweets
+
+  # calculate observed utility by adding the weighted park-level attributes 
+  # to the columns of the matrix
+  # V is n x p, with b added by-column to each element in a
+  V <- sweep(a, MARGIN = 2, b, `+`)
+
+  # log-sum of exponentiated utility, Output is n-length vector
+  log(rowSums(exp(V)))
+
+}
+```
+
+Calculate the logsum given a tracts and parks dataset. `tracts$park_ls1` uses
+only size and distance parameters. `tracts$park_ls2` uses size, distance and
+tweet parameters. `tracts$park_ls4` and `tracts$park_ls5` use a distance
+parameter of -15.
 ```{r accessibility}
-source('size00001dist5.R', echo=TRUE)
-tracts$park_ls1 <- calculate_park_logsums(tracts, parks, betas = c(.00001, -5))
+# calculate centroids of tracts and parks
+tract_centroids <- gCentroid(tracts, byid = TRUE) 
+park_centroids <- gCentroid(parks, byid = TRUE)
 
-source('size00001dist5tweet001.R', echo=TRUE)
-tracts$park_ls2 <- calculate_park_logsums(tracts, parks, betas = c(.00001, -5, .001))
+# distances from all tracts to all zones
+distances <- gDistance(park_centroids, tract_centroids, byid = TRUE) / 5280
 
-#source('size00001dist5tweetneg001.R', echo=TRUE)
-#tracts$park_ls3 <- calculate_park_logsums(tracts, parks, betas = c(.00001, -5, -.001))
+tracts$park_ls1 <- calculate_park_logsums(tracts, parks, distances,
+                                          parks$Park_Acres, betas = c(.00001, -5))
 
-source('size00001dist15.R', echo=TRUE)
-tracts$park_ls4 <- calculate_park_logsums(tracts, parks, betas = c(.00001, -15))
+tracts$park_ls2 <- calculate_park_logsums(tracts, parks, distances, 
+                                          parks$Park_Acres, parks$TWEET_COUNT, 
+                                          betas = c(.00001, -5, .001))
 
-source('size00001dist15tweet001.R', echo=TRUE)
-tracts$park_ls5 <- calculate_park_logsums(tracts, parks, betas = c(.00001, -15, .001))
+tracts$park_ls4 <- calculate_park_logsums(tracts, parks, distances, 
+                                          parks$Park_Acres, betas = c(.00001, -15))
+
+tracts$park_ls5 <- calculate_park_logsums(tracts, parks, 
+                                          parks$Park_Acres, parks$TWEET_COUNT, 
+                                          betas = c(.00001, -15, .001))
 ```

### 2018-11-02 size00001dist15.R

@@ -1,51 +1,0 @@
-#' Calculate destination choice accessibility logsum
-#' 
-#' @param size A vector of park sizes in acres
-#' @param distance A vector of distances to the parks in `size` in miles
-#' @param betas A vector containing the size and distance coefficients
-#'
-#' @return The accessibility logsum
-#' 
-park_logsum <- function(size, distance, betas = c(.00001, -15)){
-  
-  # calculate exponentiated utility of all parks
-  U <- betas[1] * size + betas[2] * distance 
-  eU <- exp(U)
-  
-  # log-sum
-  log(sum(eU))
-}
-
-
-#' Calculate destination choice logsums from a distance matrix
-#' 
-#' @param tracts A spatial points dataframe containing the 
-#'   tract geographic information and their attributes.
-#' @param parks A spatial points dataframe containing the 
-#'   park geographic information and their attributes.
-#' @param betas A vector containing the size and distance coefficients
-#'   
-#' @return An accessibility logsum calculation for each tract to every park.
-#' 
-calculate_park_logsums <- function(tracts, parks, betas = c(.00001, -15)){
-  
-  # calculate centroids of tracts and parks
-  tract_centroids <- gCentroid(tracts, byid = TRUE) 
-  park_centroids <- gCentroid(parks, byid = TRUE)
-  
-  # distances from all tracts to all zones
-  distances <- gDistance(park_centroids, tract_centroids, byid = TRUE) / 5280
-  
-  # vector containing park sizes 
-  park_sizes <- parks$Park_Acres
-  
-  # calculate observed utility 
-  #   beta_size * size + beta_distane * distance
-  V <- sweep(betas[2] * distances, MARGIN = 2, 
-             betas[1] * park_sizes, `+`)
-  
-  # log-sum of exponentiated utility
-  log(rowSums(exp(V)))
-  
-  
-}

### 2018-11-02 size00001dist15tweet001.R

@@ -1,55 +1,0 @@
-#' Calculate destination choice accessibility logsum
-#' 
-#' @param size A vector of park sizes in acres
-#' @param distance A vector of distances to the parks in `size` in miles
-#' @param tweets A vector of the number of tweets recorded within each park in September 2014
-#' @param betas A vector containing the size and distance coefficients
-#'
-#' @return The accessibility logsum
-#' 
-park_logsum <- function(size, distance, betas = c(.00001, -15, 1)){
-  
-  # calculate exponentiated utility of all parks
-  U <- betas[1] * size + betas[2] * distance + betas[3] * park_tweets
-  eU <- exp(U)
-  
-  # log-sum
-  log(sum(eU))
-}
-
-
-#' Calculate destination choice logsums from a distance matrix
-#' 
-#' @param tracts A spatial points dataframe containing the 
-#'   tract geographic information and their attributes.
-#' @param parks A spatial points dataframe containing the 
-#'   park geographic information and their attributes.
-#' @param betas A vector containing the size and distance coefficients
-#'   
-#' @return An accessibility logsum calculation for each tract to every park.
-#' 
-calculate_park_logsums <- function(tracts, parks, betas = c(.00001, -15, .001)){
-  
-  # calculate centroids of tracts and parks
-  tract_centroids <- gCentroid(tracts, byid = TRUE) 
-  park_centroids <- gCentroid(parks, byid = TRUE)
-  
-  # distances from all tracts to all zones
-  distances <- gDistance(park_centroids, tract_centroids, byid = TRUE) / 5280
-  
-  # vector containing park sizes 
-  park_sizes <- parks$Park_Acres
-  
-  # vector containing tweet counts
-  park_tweets <- parks$TWEET_COUNT
-  
-  # calculate observed utility 
-  #   beta_size * size + beta_distance * distance + beta_tweet * tweets
-  V <- sweep(betas[2] * distances, MARGIN = 2, 
-             betas[1] * park_sizes + betas[3] * log(park_tweets + 0.01), `+`)
-  
-  # log-sum of exponentiated utility
-  log(rowSums(exp(V)))
-  
-  
-}

### 2018-11-02 size00001dist5tweet001.R

@@ -1,55 +1,0 @@
-#' Calculate destination choice accessibility logsum
-#' 
-#' @param size A vector of park sizes in acres
-#' @param distance A vector of distances to the parks in `size` in miles
-#' @param tweets A vector of the number of tweets recorded within each park in September 2014
-#' @param betas A vector containing the size and distance coefficients
-#'
-#' @return The accessibility logsum
-#' 
-park_logsum <- function(size, distance, betas = c(.00001, -5, 1)){
-  
-  # calculate exponentiated utility of all parks
-  U <- betas[1] * size + betas[2] * distance + betas[3] * park_tweets
-  eU <- exp(U)
-  
-  # log-sum
-  log(sum(eU))
-}
-
-
-#' Calculate destination choice logsums from a distance matrix
-#' 
-#' @param tracts A spatial points dataframe containing the 
-#'   tract geographic information and their attributes.
-#' @param parks A spatial points dataframe containing the 
-#'   park geographic information and their attributes.
-#' @param betas A vector containing the size and distance coefficients
-#'   
-#' @return An accessibility logsum calculation for each tract to every park.
-#' 
-calculate_park_logsums <- function(tracts, parks, betas = c(.00001, -5, .001)){
-  
-  # calculate centroids of tracts and parks
-  tract_centroids <- gCentroid(tracts, byid = TRUE) 
-  park_centroids <- gCentroid(parks, byid = TRUE)
-  
-  # distances from all tracts to all zones
-  distances <- gDistance(park_centroids, tract_centroids, byid = TRUE) / 5280
-  
-  # vector containing park sizes 
-  park_sizes <- parks$Park_Acres
-  
-  # vector containing tweet counts
-  park_tweets <- parks$TWEET_COUNT
-  
-  # calculate observed utility 
-  #   beta_size * size + beta_distance * distance + beta_tweet * tweets
-  V <- sweep(betas[2] * distances, MARGIN = 2, 
-             betas[1] * park_sizes + betas[3] * log(park_tweets + 0.01), `+`)
-  
-  # log-sum of exponentiated utility
-  log(rowSums(exp(V)))
-  
-  
-}

### 2018-11-02 size00001dist5tweetneg001.R

@@ -1,55 +1,0 @@
-#' Calculate destination choice accessibility logsum
-#' 
-#' @param size A vector of park sizes in acres
-#' @param distance A vector of distances to the parks in `size` in miles
-#' @param tweets A vector of the number of tweets recorded within each park in September 2014
-#' @param betas A vector containing the size and distance coefficients
-#'
-#' @return The accessibility logsum
-#' 
-park_logsum <- function(size, distance, betas = c(.00001, -5, 1)){
-  
-  # calculate exponentiated utility of all parks
-  U <- betas[1] * size + betas[2] * distance + betas[3] * park_tweets
-  eU <- exp(U)
-  
-  # log-sum
-  log(sum(eU))
-}
-
-
-#' Calculate destination choice logsums from a distance matrix
-#' 
-#' @param tracts A spatial points dataframe containing the 
-#'   tract geographic information and their attributes.
-#' @param parks A spatial points dataframe containing the 
-#'   park geographic information and their attributes.
-#' @param betas A vector containing the size and distance coefficients
-#'   
-#' @return An accessibility logsum calculation for each tract to every park.
-#' 
-calculate_park_logsums <- function(tracts, parks, betas = c(.00001, -5, -10)){
-  
-  # calculate centroids of tracts and parks
-  tract_centroids <- gCentroid(tracts, byid = TRUE) 
-  park_centroids <- gCentroid(parks, byid = TRUE)
-  
-  # distances from all tracts to all zones
-  distances <- gDistance(park_centroids, tract_centroids, byid = TRUE) / 5280
-  
-  # vector containing park sizes 
-  park_sizes <- parks$Park_Acres
-  
-  # vector containing tweet counts
-  park_tweets <- parks$TWEET_COUNT
-  
-  # calculate observed utility 
-  #   beta_size * size + beta_distance * distance + beta_tweet * tweets
-  V <- sweep(betas[2] * distances, MARGIN = 2, 
-             betas[1] * park_sizes + betas[3] * log(park_tweets), `+`)
-  
-  # log-sum of exponentiated utility
-  log(rowSums(exp(V)))
-  
-  
-}

### 2018-11-02 NYC_Twitter_SDM_models - Copy.Rmd
Look for a better transformation for tweets variable

@@
 
+The Tweets count variable has a high degree of zeros (virtually all parks have
+no tweets originating there). We should identify if there is a transformation
+that will make the distribution more normal (log is typical), but that will allow
+for zero variables as $log(0)$ is undefined.
+
+```{r tweettransformation}
+parks_sf %>%
+  st_set_geometry(NULL) %>%
+  tbl_df() %>%
+  select(raw = TWEET_COUNT) %>%
+  mutate(
+    `0.5` = yeo.johnson(raw, 0.5),
+    `0` = yeo.johnson(raw, 0),
+    log = log(raw + 1)
+  ) %>%
+  sample_n(200) %>%
+  gather(lambda, x, -raw) %>%
+  ggplot(aes(x = raw, y = x, color = lambda)) + 
+  geom_point()
+```
+
 ```{r show_data, echo=FALSE}
@@
 
-Calculate the logsum given a tracts and parks dataset. `tracts$park_ls1` uses
-only size and distance parameters. `tracts$park_ls2` uses size, distance and
-tweet parameters. `tracts$park_ls4` and `tracts$park_ls5` use a distance
-parameter of -15.
-```{r accessibility}
+```{r distancematrix}
 # calculate centroids of tracts and parks
@@
 distances <- gDistance(park_centroids, tract_centroids, byid = TRUE) / 5280
+```
 
+Calculate the logsum given a tracts and parks dataset. `tracts$park_ls1` uses
+only size and distance parameters. `tracts$park_ls2` uses size, distance and
+tweet parameters. `tracts$park_ls4` and `tracts$park_ls5` use a distance
+parameter of -15.
+
+We need to firm up these calculations. Are we using log of acres as well?
+
+```{r logsums}
 tracts$park_ls1 <- calculate_park_logsums(tracts, parks, distances,
@@
 tracts$park_ls2 <- calculate_park_logsums(tracts, parks, distances,
-                                          parks$Park_Acres, parks$TWEET_COUNT, 
-                                          betas = c(.00001, -5, .001))
+                                          parks$Park_Acres,
+                                          yeo.johnson(parks$TWEET_COUNT, 0),
+                                          betas = c(.00001, -5, 2))
 
@@
 
-tracts$park_ls5 <- calculate_park_logsums(tracts, parks, 
-                                          parks$Park_Acres, parks$TWEET_COUNT, 
-                                          betas = c(.00001, -15, .001))
+tracts$park_ls5 <- calculate_park_logsums(tracts, parks, distances,
+                                          parks$Park_Acres,
+                                          yeo.johnson(parks$TWEET_COUNT, 0),
+                                          betas = c(.00001, -15, 2))
+```
+
+I'm interested in looking at the distribution of the log-sum statistics so we can
+know what if anything the different variables are doing, also to make sure that
+there is no missingness in the data.
+
+  - `ls1` and `ls2` have virtually the same distribution, along with `ls4` and
+  `ls5`. This suggests that tweets have virtually no effect on the rankings, at
+  least in this iteration.
+
+```{r distribution}
+logsums <- tracts@data %>%
+  select(park_ls1:park_ls6) %>%
+  gather(logsum_type, value)
+
+# check missingness
+table(is.na(logsums$value))
+
+# histogram
+ggplot(logsums, aes(x = value, fill = logsum_type)) + geom_histogram()
 ```
@@
 # Create a continuous palette function
-pal <- colorQuantile(
-  palette = "Greens",
-  domain = c(0, tracts$park_ls1)
-)
-
-pop_content <- function(shp){
-  paste0(
-    "Tract: ", shp$NAME, "<br/>",
-    "% Park: ", round(shp$Park_Percent, 2), "<br/>",
-    "Park Access: ", round(shp$park_ls1, 2), "<br/>"
-  )
-}
+pal1 <- colorQuantile(palette = "Greens", domain = tracts$park_ls1)
+pal2 <- colorQuantile(palette = "Greens", domain = tracts$park_ls2)
+pal4 <- colorQuantile(palette = "Greens", domain = tracts$park_ls4)
+pal5 <- colorQuantile(palette = "Greens", domain = tracts$park_ls5)
 
 leaflet(tracts %>% spTransform(., CRS("+init=epsg:4326"))) %>%
-  addProviderTiles("CartoDB.Positron") %>%
-  addPolygons(weight = 1, popup = ~pop_content(tracts),
-              color = ~pal(park_ls1))
+  addProviderTiles(providers$CartoDB.Positron) %>%
+  addPolygons(group = "Long distance, no tweets", weight = 1, color = ~pal1(park_ls1)) %>%
+  addPolygons(group = "Long distance, tweets", weight = 1, color = ~pal2(park_ls2)) %>%
+  addPolygons(group = "Short distance, no tweets", weight = 1, color = ~pal4(park_ls4)) %>%
+  addPolygons(group = "Short distance, tweets", weight = 1, color = ~pal5(park_ls5)) %>%
+  addLayersControl(baseGroups =  c("Short distance, no tweets", 
+                                   "Short distance, tweets", 
+                                   "Long distance, no tweets", 
+                                   "Long distance, tweets"))
 ```
 
-And here is a map of accessibility using the size, distance and tweet parameters (using obesity as the DV). 
-```{r accessibility_leaflet_2, echo=FALSE}
-# Create a continuous palette function
-pal <- colorQuantile(
-  palette = "Greens",
-  domain = c(0, tracts$park_ls2)
-)
-
-pop_content <- function(shp){
-  paste0(
-    "Tract: ", shp$NAME, "<br/>",
-    "% Park: ", round(shp$Park_Percent, 2), "<br/>",
-    "Park Access: ", round(shp$park_ls2, 2), "<br/>"
-  )
-}
-
-leaflet(tracts %>% spTransform(., CRS("+init=epsg:4326"))) %>%
-  addProviderTiles("CartoDB.Positron") %>%
-  addPolygons(weight = 1, popup = ~pop_content(tracts),
-              color = ~pal(park_ls2))
-```
-
-Here is a map of accessibility using the betas that produce the most likely model when using physical activity as the dependent variable. Only size and distance parameters are used.
-```{r accessibility_leaflet_3, echo=FALSE}
-# Create a continuous palette function
-pal <- colorQuantile(
-  palette = "Greens",
-  domain = c(0, tracts$park_ls4)
-)
-
-pop_content <- function(shp){
-  paste0(
-    "Tract: ", shp$NAME, "<br/>",
-    "% Park: ", round(shp$Park_Percent, 2), "<br/>",
-    "Park Access: ", round(shp$park_ls4, 2), "<br/>"
-  )
-}
-
-leaflet(tracts %>% spTransform(., CRS("+init=epsg:4326"))) %>%
-  addProviderTiles("CartoDB.Positron") %>%
-  addPolygons(weight = 1, popup = ~pop_content(tracts),
-              color = ~pal(park_ls4))
-```
-
-And here is a map of accessibility using size, distance and tweet parameters (with the combination of betas that is most likely for physical activity as the DV).
-```{r accessibility_leaflet_4, echo=FALSE}
-# Create a continuous palette function
-pal <- colorQuantile(
-  palette = "Greens",
-  domain = c(0, tracts$park_ls5)
-)
-
-pop_content <- function(shp){
-  paste0(
-    "Tract: ", shp$NAME, "<br/>",
-    "% Park: ", round(shp$Park_Percent, 2), "<br/>",
-    "Park Access: ", round(shp$park_ls5, 2), "<br/>"
-  )
-}
-
-leaflet(tracts %>% spTransform(., CRS("+init=epsg:4326"))) %>%
-  addProviderTiles("CartoDB.Positron") %>%
-  addPolygons(weight = 1, popup = ~pop_content(tracts),
-              color = ~pal(park_ls5))
-```
-
+A
 Create regression formulas:
@@
 
-#parks_formula_1 <- update(base_formula, . ~ . + log(park_ls1 + 0.1))
-parks_formula_1a <- update(base_formula_age, . ~ . + log(park_ls1 + 0.1))
-#parks_formula_2 <- update(base_formula, . ~ . + log(park_ls2 + 0.1))
-parks_formula_2a <- update(base_formula_age, . ~ . + log(park_ls2 + 0.1))
-#parks_formula_3 <- update(base_formula, . ~ . + log(park_ls4 + 0.01))
-parks_formula_3a <- update(base_formula_age, . ~ . + log(park_ls4 + 5))
-#parks_formula_4 <- update(base_formula, . ~ . + log(park_ls5 + 0.1))
-parks_formula_4a <- update(base_formula_age, . ~ . + log(park_ls5 + 5))
+parks_formula_1a <- update(base_formula_age, . ~ . + park_ls1 )
+parks_formula_2a <- update(base_formula_age, . ~ . + park_ls2 )
+parks_formula_3a <- update(base_formula_age, . ~ . + park_ls4 )
+parks_formula_4a <- update(base_formula_age, . ~ . + park_ls5 )
 

### 2018-11-05 NYC_Twitter_SDM_models - Copy.Rmd
Adjust transformation of logsums

@@
 library(sf)
+library(VGAM)
 ```
@@
 that will make the distribution more normal (log is typical), but that will allow
-for zero variables as $log(0)$ is undefined.
+for zero variables as $log(0)$ is undefined. Using a Yeo-Johnson transformation
+with a lambda value of zero transforms the tweet count into a distribution
+asymptotically equivalent to log, and with zero values equal to zero.
 
@@
     `0` = yeo.johnson(raw, 0),
-    log = log(raw + 1)
+    log = log(raw)
   ) %>%
@@
   ggplot(aes(x = raw, y = x, color = lambda)) + 
-  geom_point()
+  geom_line()
 ```
@@
                                           betas = c(.00001, -15, 2))
+ls_names <- c("Short, no tweets",  "Short, tweets",  
+              "Long, no tweets",  "Long, tweets")
 ```
@@
 logsums <- tracts@data %>%
-  select(park_ls1:park_ls6) %>%
+  select(park_ls1:park_ls5) %>%
   gather(logsum_type, value)

### 2018-11-05 NYC_Twitter_SDM_models - Copy.Rmd
Reorganize model estimation and impacts comparison

@@
   addProviderTiles(providers$CartoDB.Positron) %>%
-  addPolygons(group = "Long distance, no tweets", weight = 1, color = ~pal1(park_ls1)) %>%
-  addPolygons(group = "Long distance, tweets", weight = 1, color = ~pal2(park_ls2)) %>%
-  addPolygons(group = "Short distance, no tweets", weight = 1, color = ~pal4(park_ls4)) %>%
-  addPolygons(group = "Short distance, tweets", weight = 1, color = ~pal5(park_ls5)) %>%
-  addLayersControl(baseGroups =  c("Short distance, no tweets", 
-                                   "Short distance, tweets", 
-                                   "Long distance, no tweets", 
-                                   "Long distance, tweets"))
+  addPolygons(group = ls_names[1], weight = 1, color = ~pal1(park_ls1)) %>%
+  addPolygons(group = ls_names[2], weight = 1, color = ~pal2(park_ls2)) %>%
+  addPolygons(group = ls_names[3], weight = 1, color = ~pal4(park_ls4)) %>%
+  addPolygons(group = ls_names[4], weight = 1, color = ~pal5(park_ls5)) %>%
+  addLayersControl(baseGroups = ls_names)
 ```
 
-A
-Create regression formulas:
-```{r obese_lm, echo=FALSE, message=FALSE, include=FALSE}
-#base_formula <- formula(~log(Pop_Density) + FulltimeWork + 
-    #CollegeDeg + Single_Percent + PctWhite + PctBlack + PctAsian + PctOther + PctHispanic)
-base_formula_age <- formula(~log(Pop_Density) + FulltimeWork + 
-    CollegeDeg + Single_Percent + Pct0to17 + Pct18to29 + Pct30to64 + Pct65plus + PctWhite + PctBlack + PctAsian + PctOther + PctHispanic)
+# Obesity model estimation
 
-parks_formula_1a <- update(base_formula_age, . ~ . + park_ls1 )
-parks_formula_2a <- update(base_formula_age, . ~ . + park_ls2 )
-parks_formula_3a <- update(base_formula_age, . ~ . + park_ls4 )
-parks_formula_4a <- update(base_formula_age, . ~ . + park_ls5 )
+Create a spatial weights matrix. It's showing that we have three regions with
+no links, so we need to ensure that these are estimated in the model.
 
+```{r W, echo=FALSE}
+neighbors <- poly2nb(tracts, queen = TRUE)
+W <- nb2listw(neighbours = neighbors, zero.policy = TRUE)
+print(W, zero.policy = TRUE)
+trMC <- trW(as(W, "CsparseMatrix"), type="MC") # trace used in montecarlo impacts
 ```
 
-Create a spatial weights matrix:
+```{r obese_lm, echo=FALSE, message=FALSE, include=FALSE}
+base_formula <- formula( 
+  ~ log(Pop_Density) + 
+    FulltimeWork + CollegeDeg + Single_Percent + 
+    Pct0to17 + Pct18to29 + Pct65plus + # need to leave out a category for collinearity  
+    PctBlack + PctAsian + PctOther + PctHispanic)
 
-```{r W, echo=FALSE}
-neighbors <- poly2nb(tracts, queen = FALSE)
-W <- nb2listw(neighbours = neighbors, zero.policy = TRUE)
-print(W, zero.policy = TRUE)
+parks_formula_1 <- update(base_formula, . ~ . + park_ls1 )
+parks_formula_2 <- update(base_formula, . ~ . + park_ls2 )
+parks_formula_4 <- update(base_formula, . ~ . + park_ls4 )
+parks_formula_5 <- update(base_formula, . ~ . + park_ls5 )
 ```
 
-Estimate spatial models for the `OBESITY` variable.
-Model 1 of each type does not include park access.
-Models 2 & 3 each use a different park access variable with a different combination of parameters:
+First, let's estimate the base model as an LM, SDM, and SEM model
+
+```{r basemodels}
+obese_base_lm <- lm(update(base_formula, OBESITY ~ .), data = tracts@data)
+obese_base_sar <- lagsarlm(update(base_formula, OBESITY ~ .), 
+                           data = tracts@data, listw = W, zero.policy = TRUE)
+obese_base_sem <- errorsarlm(update(base_formula, OBESITY ~ .), 
+                             data = tracts@data, listw = W, zero.policy = TRUE)
+obese_base_sdm <- lagsarlm(update(base_formula, OBESITY ~ .), 
+                           data = tracts@data, listw = W, zero.policy = TRUE, 
+                           type = "mixed")
+```
+
+A likelihood ratio test reveals that the SEM is not preferred, so we use the SDM
+only going forward.
+
+```{r lrtest}
+lmtest::lrtest(obese_base_sem, obese_base_sdm)
+```
 
 ```{r spmodels obesity, echo=FALSE, message=FALSE, warning=FALSE}
-obese_sar_age <- lagsarlm(update(base_formula_age, OBESITY ~ .), data = tracts@data, 
-                      listw = W, zero.policy = TRUE)
-obese_sar2_age <- lagsarlm(update(parks_formula_1a, OBESITY ~ .), data = tracts@data,
-                      listw = W, zero.policy = TRUE)
-obese_sar3_age <- lagsarlm(update(parks_formula_2a, OBESITY ~ .), data = tracts@data,
-                     listw = W, zero.policy = TRUE)
-obese_sem_age <- errorsarlm(update(base_formula_age, OBESITY ~ .), data = tracts@data, 
-                      listw = W, zero.policy = TRUE)
-obese_sem2_age <- errorsarlm(update(parks_formula_1a, OBESITY ~ .), data = tracts@data, 
-                      listw = W, zero.policy = TRUE)
-obese_sem3_age <- errorsarlm(update(parks_formula_2a, OBESITY ~ .), data = tracts@data, 
-                      listw = W, zero.policy = TRUE)
-obese_sdm_age <- lagsarlm(update(base_formula_age, OBESITY ~ .), data = tracts@data, 
-                      listw = W, zero.policy = TRUE, type = "mixed")
-obese_sdm2_age <- lagsarlm(update(parks_formula_1a, OBESITY ~ .), data = tracts@data, 
-                      listw = W, zero.policy = TRUE, type = "mixed")
-obese_sdm3_age <- lagsarlm(update(parks_formula_2a, OBESITY ~ .), data = tracts@data, 
-                      listw = W, zero.policy = TRUE, type = "mixed")
-obese_sdem_age <- errorsarlm(update(base_formula_age, OBESITY ~ .), data = tracts@data, 
-                      listw = W, zero.policy = TRUE, etype = "emixed",tol.solve = 1e-16)
-obese_sdem2_age <- errorsarlm(update(parks_formula_1a, OBESITY ~ .), data = tracts@data, 
-                      listw = W, zero.policy = TRUE, etype = "emixed",tol.solve = 1e-16)
-obese_sdem3_age <- errorsarlm(update(parks_formula_2a, OBESITY ~ .), data = tracts@data, 
-                      listw = W, zero.policy = TRUE, etype = "emixed",tol.solve = 1e-16)
+obese_ls1_sdm <- lagsarlm(update(parks_formula_1, OBESITY ~ .), 
+                          data = tracts@data,  listw = W, zero.policy = TRUE, type = "mixed")
+obese_ls2_sdm <- lagsarlm(update(parks_formula_2, OBESITY ~ .), 
+                          data = tracts@data,  listw = W, zero.policy = TRUE, type = "mixed")
+obese_ls4_sdm <- lagsarlm(update(parks_formula_4, OBESITY ~ .), 
+                          data = tracts@data,  listw = W, zero.policy = TRUE, type = "mixed")
+obese_ls5_sdm <- lagsarlm(update(parks_formula_5, OBESITY ~ .), 
+                          data = tracts@data,  listw = W, zero.policy = TRUE, type = "mixed")
 
+obesity_models <- list(obese_base_sdm, obese_ls1_sdm, obese_ls2_sdm, obese_ls4_sdm, obese_ls5_sdm)
+names(obesity_models) <- c("Base", ls_names)
 ```
@@
 Estimate spatial models for the `Phys_Act` variable. Model 1 does not include park access. Models 2 & 3 each use a different park access variable with a different combination of parameters:
+
+```{r}
+screenreg(obesity_models)
+```
+
+Estimate the impacts for `obese_sdm_age`.
+
+```{r obese_sdm_age_impacts}
+impacts_extractor <- function(sdm) {
+
+  impacts_summary <- summary(impacts(sdm, tr=trMC, R=1000), zstats = TRUE)
+
+  coef <- list(
+    direct = impacts_summary$direct_sum,
+    indirect = impacts_summary$indirect_sum,
+    total = impacts_summary$total_sum
+  ) %>%
+  lapply(function(s) {
+    data_frame(
+      var = names(s$statistics[,1]),
+      impact = s$statistics[,1]
+    )
+  }) %>%
+    bind_rows(.id = "effect")
+
+  pval <- as_data_frame(impacts_summary$pzmat) %>%
+    mutate(
+     var = rownames(impacts_summary$pzmat)
+    ) %>%
+    gather("effect", "p-val", -var) %>%
+    mutate(effect = tolower(effect))
+
+  left_join(coef, pval, by = c("effect", "var"))
+}
+
+obesity_models %>%
+  lapply(impacts_extractor) %>%
+  bind_rows(.id = "model") %>%
+  transmute(model, var, effect, output = paste(round(impact, 5), gtools::stars.pval(`p-val`))) %>%
+  spread(model, output)
+
+```
 
@@
                       listw = W, zero.policy = TRUE, etype = "emixed",tol.solve = 1e-16)
-```
-
-Compare the models with `htmlreg()` in
-the html output. First, the obesity models.
-
-```{r obesitysummaryage, results="asis", echo=FALSE}
-obesity_models_age <- list(obese_sar_age, obese_sar2_age, obese_sar3_age, obese_sem_age, obese_sem2_age, obese_sem3_age, obese_sdm_age, obese_sdm2_age, obese_sdm3_age, obese_sdem_age, obese_sdem2_age, obese_sdem3_age)
-names(obesity_models_age) <-  paste(rep(c("SAR", "SEM" , "SDM", "SDEM"), each = 3),
-                                rep(c("1", "2", "3")))
-htmlreg(obesity_models_age, caption = "NYC: Spatial Models of Obesity Prevalence")
-```
-
-Estimate the impacts for `obese_sdm_age`.
-```{r obese_sdm_age_impacts}
-
-listw <- as(W, "CsparseMatrix")
-trMC <- trW(listw, type="MC")
-set.seed(1)
-#summary(impacts(obese_sdm_age, tr=trMC, R=1000), zstats=TRUE)
-
 ```

### 2018-11-06 NYC_Twitter_SDM_models - Copy.Rmd
Cleanup physical activity models

@@
 
-```{r spmodels obesity, echo=FALSE, message=FALSE, warning=FALSE}
+```{r obesity, echo=FALSE, message=FALSE, warning=FALSE}
 obese_ls1_sdm <- lagsarlm(update(parks_formula_1, OBESITY ~ .), 
@@
 
-obesity_models <- list(obese_base_sdm, obese_ls1_sdm, obese_ls2_sdm, obese_ls4_sdm, obese_ls5_sdm)
+obesity_models <- list(obese_base_sdm, obese_ls1_sdm, obese_ls2_sdm,
+                       obese_ls4_sdm, obese_ls5_sdm)
 names(obesity_models) <- c("Base", ls_names)
@@
   bind_rows(.id = "model") %>%
-  transmute(model, var, effect, output = paste(round(impact, 5), gtools::stars.pval(`p-val`))) %>%
-  spread(model, output)
+  transmute(
+    model, var, effect, 
+    output = paste(round(impact, 5), 
+                   gtools::stars.pval(`p-val`))
+  ) %>%
+  spread(model, output) %>% View
 
@@
 
-```{r spmodels physact with age, echo=FALSE, message=FALSE, warning=FALSE}
+# Estimate Physical Activity Models
 
-physact_sar_age <- lagsarlm(update(base_formula_age, Phys_Act ~ .), data = tracts@data, 
-                      listw = W, zero.policy = TRUE)
-physact_sar2_age <- lagsarlm(update(parks_formula_1a, Phys_Act ~ .), data = tracts@data,
-                      listw = W, zero.policy = TRUE)
-physact_sar3_age <- lagsarlm(update(parks_formula_2a, Phys_Act ~ .), data = tracts@data,
-                     listw = W, zero.policy = TRUE)
-physact_sem_age <- errorsarlm(update(base_formula_age, Phys_Act ~ .), data = tracts@data, 
-                      listw = W, zero.policy = TRUE)
-physact_sem2_age <- errorsarlm(update(parks_formula_1a, Phys_Act ~ .), data = tracts@data, 
-                      listw = W, zero.policy = TRUE)
-physact_sem3_age <- errorsarlm(update(parks_formula_2a, Phys_Act ~ .), data = tracts@data, 
-                      listw = W, zero.policy = TRUE)
-physact_sdm_age <- lagsarlm(update(base_formula_age, Phys_Act ~ .), data = tracts@data, 
-                        listw = W, zero.policy = TRUE, type = "mixed")
-physact_sdm2_age <- lagsarlm(update(parks_formula_1a, Phys_Act ~ .), data = tracts@data,
-                         listw = W, zero.policy = TRUE, type = "mixed")
-physact_sdm3_age <- lagsarlm(update(parks_formula_2a, Phys_Act ~ .), data = tracts@data,
-                         listw = W, zero.policy = TRUE, type = "mixed")
-physact_sdem_age <- errorsarlm(update(base_formula_age, Phys_Act ~ .), data = tracts@data, 
-                      listw = W, zero.policy = TRUE, etype = "emixed",tol.solve = 1e-16)
-physact_sdem2_age <- errorsarlm(update(parks_formula_1a, Phys_Act ~ .), data = tracts@data, 
-                      listw = W, zero.policy = TRUE, etype = "emixed",tol.solve = 1e-16)
-physact_sdem3_age <- errorsarlm(update(parks_formula_2a, Phys_Act ~ .), data = tracts@data, 
-                      listw = W, zero.policy = TRUE, etype = "emixed",tol.solve = 1e-16)
+```{r phy, echo=FALSE, message=FALSE, warning=FALSE}
+physact_base_sdm <- lagsarlm(update(base_formula, Phys_Act ~ .), 
+                             data = tracts@data,  listw = W, zero.policy = TRUE, 
+                             type = "mixed")
+physact_ls1_sdm <- lagsarlm(update(parks_formula_1, Phys_Act ~ .), 
+                            data = tracts@data,  listw = W, zero.policy = TRUE, 
+                            type = "mixed")
+physact_ls2_sdm <- lagsarlm(update(parks_formula_2, Phys_Act ~ .), 
+                            data = tracts@data,  listw = W, zero.policy = TRUE, 
+                            type = "mixed")
+physact_ls4_sdm <- lagsarlm(update(parks_formula_4, Phys_Act ~ .), 
+                            data = tracts@data,  listw = W, zero.policy = TRUE, 
+                            type = "mixed")
+physact_ls5_sdm <- lagsarlm(update(parks_formula_5, Phys_Act ~ .), 
+                            data = tracts@data,  listw = W, zero.policy = TRUE, 
+                            type = "mixed")
 
+physact_models <- list(physact_base_sdm, physact_ls1_sdm, physact_ls2_sdm,
+                       physact_ls4_sdm, physact_ls5_sdm)
+names(physact_models) <- c("Base", ls_names)
 ```
 
-Estimate additional spatial models for the `Phys_Act` variable using a new distance parameter of -15.
-```{r spmodels physact age new betas, echo=FALSE}
-physact_sar_new_age <- lagsarlm(update(base_formula_age, Phys_Act ~ .), data = tracts@data, 
-                      listw = W, zero.policy = TRUE)
-physact_sar2_new_age <- lagsarlm(update(parks_formula_3a, Phys_Act ~ .), data = tracts@data,
-                           listw = W, zero.policy = TRUE)
-physact_sar3_new_age <- lagsarlm(update(parks_formula_4a, Phys_Act ~ .), data = tracts@data,
-                           listw = W, zero.policy = TRUE)
-physact_sem_new_age <- errorsarlm(update(base_formula_age, Phys_Act ~ .), data = tracts@data, 
-                      listw = W, zero.policy = TRUE)
-physact_sem2_new_age <- errorsarlm(update(parks_formula_3a, Phys_Act ~ .), data = tracts@data, 
-                      listw = W, zero.policy = TRUE)
-physact_sem3_new_age <- errorsarlm(update(parks_formula_4a, Phys_Act ~ .), data = tracts@data, 
-                      listw = W, zero.policy = TRUE)
-physact_sdm_new_age <- lagsarlm(update(base_formula_age, Phys_Act ~ .), data = tracts@data, 
-                        listw = W, zero.policy = TRUE, type = "mixed")
-physact_sdm2_new_age <- lagsarlm(update(parks_formula_3a, Phys_Act ~ .), data = tracts@data,
-                         listw = W, zero.policy = TRUE, type = "mixed")
-physact_sdm3_new_age <- lagsarlm(update(parks_formula_4a, Phys_Act ~ .), data = tracts@data,
-                         listw = W, zero.policy = TRUE, type = "mixed")
-physact_sdem_new_age <- errorsarlm(update(base_formula_age, Phys_Act ~ .), data = tracts@data, 
-                      listw = W, zero.policy = TRUE, etype = "emixed",tol.solve = 1e-16)
-physact_sdem2_new_age <- errorsarlm(update(parks_formula_3a, Phys_Act ~ .), data = tracts@data, 
-                      listw = W, zero.policy = TRUE, etype = "emixed",tol.solve = 1e-16)
-physact_sdem3_new_age <- errorsarlm(update(parks_formula_4a, Phys_Act ~ .), data = tracts@data, 
-                      listw = W, zero.policy = TRUE, etype = "emixed",tol.solve = 1e-16)
+```{r physact_impacts}
+physact_models %>%
+  lapply(impacts_extractor) %>%
+  bind_rows(.id = "model") %>%
+  transmute(
+    model, var, effect, 
+    output = paste(round(impact, 5), 
+                   gtools::stars.pval(`p-val`))
+  ) %>%
+  spread(model, output) %>% View
 ```
-
-Estimate the impacts for `obese_sdm2_age`.
-```{r obese_sdm2_age_impacts}
-
-#summary(impacts(obese_sdm2_age, tr=trMC, R=1000), zstats=TRUE)
-
-```
-
-Estimate the impacts for `obese_sdm3_age`.
-```{r obese_sdm3_age_impacts}
-
-#summary(impacts(obese_sdm3_age, tr=trMC, R=1000), zstats=TRUE)
-
-```
-
-Next, the physical activity models.  The second table uses new beta parameters.
-
-```{r physactsummaryage, results="asis", echo=FALSE}
-physact_models_age <- list(physact_sar_age, physact_sar2_age, physact_sar3_age, physact_sem_age, physact_sem2_age, physact_sem3_age, physact_sdm_age, physact_sdm2_age, physact_sdm3_age, physact_sdem_age, physact_sdem2_age, physact_sdem3_age)
-names(physact_models_age) <-  paste(rep(c("SAR", "SEM" , "SDM", "SDEM"), each = 3),
-                                rep(c("1", "2", "3")))
-htmlreg(physact_models_age, caption = "NYC: SDM Models of Physical Activity Participation")
-```
-
-The following table use a size beta of 0.00001 and a distance beta of -15.
-
-```{r physact new betas with age, results="asis", echo=FALSE}
-physact_models_new_age <- list(physact_sar_new_age, physact_sar2_new_age, physact_sar3_new_age, physact_sem_new_age, physact_sem2_new_age, physact_sem3_new_age, physact_sdm_new_age, physact_sdm2_new_age, physact_sdm3_new_age, physact_sdem_new_age, physact_sdem2_new_age, physact_sdem3_new_age)
-names(physact_models_new_age) <-  paste(rep(c("SAR", "SEM" , "SDM", "SDEM"), each = 3),
-                                rep(c("1", "2", "3")))
-htmlreg(physact_models_new_age, caption = "NYC: Spatial Models of Physical Activity Participation")
-
-```
-
-Estimate the impacts for `physact_sdm_age`.
-```{r physact_sdm_age_impacts}
-
-#summary(impacts(physact_sdm_age, tr=trMC, R=1000), zstats=TRUE)
-
-```
-
-Estimate the impacts for `physact_sdm2_age`.
-```{r physact_sdm2_age_impacts}
-
-#summary(impacts(physact_sdm2_age, tr=trMC, R=1000), zstats=TRUE)
-
-```
-
-Estimate the impacts for `physact_sdm3_age`.
-```{r physact_sdm3_age_impacts}
-
-#summary(impacts(physact_sdm3_age, tr=trMC, R=1000), zstats=TRUE)
-
-```
-
-Estimate the impacts for `physact_sdm_new_age`.
-```{r physact_sdm_new_age_impacts}
-
-#summary(impacts(physact_sdm_new_age, tr = trMC, R = 1000), zstats=TRUE)
-
-```
-
-Estimate the impacts for `physact_sdm2_new_age`.
-```{r physact_sdm2_new_age_impacts}
-
-#summary(impacts(physact_sdm2_new_age, tr = trMC, R = 1000), zstats=TRUE)
-
-```
-
-Estimate the impacts for `physact_sdm3_new_age`.
-```{r physact_sdm3_new_age_impacts}
-
-#summary(impacts(physact_sdm3_new_age, tr = trMC, R = 1000), zstats=TRUE)
-
-```

### 2018-11-28 model_estimation.Rmd
Add notebook for model estimation tasks

---
title: "Model Estimation"
output:
  html_document:
    df_print: paged
editor_options: 
  chunk_output_type: console
---
```{r setup}
library(geojsonio)
library(tidyverse)
library(maptools)
library(rgdal)
library(rgeos)
library(spdep)
library(texreg)
library(leaflet)
library(sf)
library(VGAM)
```

# Data cleaning

Nico cleaned the tracts dataset into a geoJSON file stored in the repository. We
should read it in twice because the sf library is better for visualization but
spdep still requires sp objects for now.

```{r read_tracts}
tracts <- geojson_read("data/nyc_tracts.geojson", what = "sp")
tracts_sf <- st_read("data/nyc_tracts.geojson", quiet = TRUE) %>%
  st_transform(4326)
```

## Complete Cases
We noticed that our models had different numbers of observations / degrees of 
freedom. Obviously we need to have comparable econometric models. The first thing
we want to look for is missingness in the tracts variables, and which tracts
are missing those variables. The map below shows the tracts that are missing
this information.

```{r tracts_with_missing}
# identify complete cases
complete_tracts_index <- tracts_sf %>%
  st_set_geometry(NULL) %>%
  tbl_df() %>%
  complete.cases()

# make a map of 
leaflet( tracts_sf[!complete_tracts_index, ]) %>%
  addProviderTiles(providers$OpenStreetMap.Mapnik) %>% 
  addPolygons(label = ~ as.character(GEOID))
```

These tracts seem to fall into two basic categories:

  - Tracts outside of New York City that were clipped into the shape layer 
  and remain as edge remnants. It is generally better to select by FIPS code rather
  than geometric operations, but we can just discard these at this point and
  they will not affect anything.

  - Tracts that are completely occupied by parks, airports, or other pieces
  of non-residential infrastructure.

The exceptions to this appear to be 3606100940 and 36061010200 in Midtown between
5th Avenue and Park Avenue. Greg will assume for this analysis that these are 
entirely commercial tracts and therefore have no residents, but Nico needs to 
verify these.

```{r complete_cases}
tracts <- tracts[complete_tracts_index, ]
```

The parks data seems relatively complete, with the exception of a few variables
that are missing everywhere. But we'll ignore this for now.

```{r read_parks, echo=FALSE}
parks <- geojson_read("data/nyc_parks.geojson", what = "sp")
```

## Tweets Transformation

The Tweets count variable has a high degree of zeros (virtually all parks have
no tweets originating there). We should identify if there is a transformation
that will make the distribution more normal (log is typical), but that will allow
for zero variables as $log(0)$ is undefined. Using a Yeo-Johnson transformation
with a lambda value of zero transforms the tweet count into a distribution
asymptotically equivalent to log, and with zero values equal to zero.

```{r tweettransformation}
parks@data %>%
  tbl_df() %>%
  select(raw = TWEET_COUNT) %>%
  filter(raw < 100) %>%
  mutate(
    `0.5` = yeo.johnson(raw, 0.5),
    `0` = yeo.johnson(raw, 0),
    log = log(raw)
  ) %>%
  gather(lambda, x, -raw) %>%
  ggplot(aes(x = raw, y = x, color = lambda)) + 
  geom_line()
```

```{r show_data, echo=FALSE}
tracts@data <- tbl_df(tracts@data) %>%
  select(
    GEOID, OBESITY, Park_Percent, Phys_Act, MENTAL, 
    Income1, Income2, Income3, Income4, Income5, Income6, 
    Income7, Income8, Income9, Income10, Pop_Density, 
    FulltimeWork, CollegeDeg, Pct0to17, Pct18to29, Pct30to64, Pct65plus, 
    Single_Percent, PctWhite, PctBlack, PctNative, PctAsian, PctPacific,
    PctOther, PctHispanic
  )

parks@data <- tbl_df(parks@data) 
```

# Spatial Model Selection

Deciding which spatial model to select is an empirical question as outlined by
both LeSage and Pace (2010?) and Macfarlane et al. (2015) We estimate a model
without park accessibility to determine which spatial structure is correct
for our analysis.

We assert a first-order queen adjacency neighbors matrix, which is
row-standardized such that the total weight of each row in the matrix sums to 1.
There are three tracts in the data that have no adjacent neighbors, meaning that 
these tracts will have no spatial spillover effects.

```{r W, echo=FALSE}
neighbors <- poly2nb(tracts, queen = TRUE)
W <- nb2listw(neighbours = neighbors, zero.policy = TRUE, style = "W")
print(W, zero.policy = TRUE)
trMC <- trW(as(W, "CsparseMatrix"), type="MC") # trace used in montecarlo impacts
ev <- eigenw(W)
```

The base model regresses the obesity rate in the tract against covariates for
work status and educational attainment, minority and young/elderly population, 
and marital status.

```{r obese_lm, echo=FALSE, message=FALSE, include=FALSE}
base_formula <- formula( 
  ~ log(Pop_Density) + 
    FulltimeWork + CollegeDeg + Single_Percent + 
    Pct0to17 + Pct18to29 + Pct65plus + # need to leave out a category for collinearity  
    PctBlack + PctAsian + PctOther + PctHispanic)
```

We estimate four models with this regression specification:

  - Ordinary least squares
  - Spatial autoregressive
  - Spatial error
  - Spatial Durbin

```{r spatial selection}
obese_base_lm <- lm(update(base_formula, OBESITY ~ .), data = tracts@data)
obese_base_sar <- lagsarlm(update(base_formula, OBESITY ~ .), 
                           data = tracts@data, listw = W, zero.policy = TRUE)
obese_base_sem <- errorsarlm(update(base_formula, OBESITY ~ .), 
                             data = tracts@data, listw = W, zero.policy = TRUE)
obese_base_sdm <- lagsarlm(update(base_formula, OBESITY ~ .), 
                           data = tracts@data, listw = W, zero.policy = TRUE, 
                           type = "mixed")
```

A likelihood ratio test reveals that the SEM is not preferred, so we use the SDM
only going forward.

```{r lrtest}
lmtest::lrtest(obese_base_sem, obese_base_sdm)
```

# Logsum Calculation

Consider that an individual is choosing a park for a recreation activity.
According to basic choice theory (Mcfadden 1974), the probability of choosing
park $p$ from the set of all regional parks $J$ is:

$$ P(p | V_p) = \frac{\exp(V_p)}{\sum_{j \in J}\exp(V_j)}$$
where parks are differentiated from each other by their relative measurable
utilitie $V$. In principle, $V$ may include any measurable attributes of either
the choice maker or the park. In this study we use a linear formulation of

$$V_{ij} = size_j\beta_s + distance_{ij}\beta_d$$
incorporating the size of the park in acres and the distance of the park from
the census tract $i$ in miles. The coefficients $\beta$ are typically estimated from 
surveys, though in the absence of a survey we apply a maximum likelihood
technique described below.

A key theoretical understanding of random utility choice models is that the
consumer surplus of the choice set can be otained as the log-sum of the
denominator of the choice probability equation. In plainer terms, the *total value*
of an individual's park accessibility is defined to be:

$$CS_i = \ln\left({\sum_{j \in J}\exp(V_{ij})}\right)$$

There are several advantages to a log-sum defined metric relative to
buffer-based accessibility metrics more commonly found in the literature. First,
all individuals are defined as having some access to all parks, rather than an
arbitrary limit of 1/2 mile or so. This allows for the fact that some people are
more or less sensitive to distances, and that distance is a continuous, and not
a binary, phenomenon. Second, the random utility formulation allows the
researcher to include any attribute of the park; in this case, we consider the
size of the park as an element of accessibility. This suggests that not all
parks are equal, and that a large park such as New York City's Central Park may
provide health and activity benefits over a much larger area than a smaller
community square.

McFadden, D. (1974). The measurement of urban travel demand. Journal of public economics, 3(4), 303-328.

In the absence of a park choice survey, we estimated likely values of the $\beta$ 
coefficients by iteratively searching for the values which produced the highest
model likelihood in *which model did we use to estimate likelihood?*. The search
was constrained by two assertions. First, we required that the coefficient on
size be positive and that on distance negative; all else equal, people will
prefer to use larger parks that are nearer to their residence. Second, we visually
inspected the resulting accessibility scores to ensure that the scores produced a 
reasonably varied pattern of access throughout the metropolitan region.

$$ A_i = log\sum_{j \in J}\exp(distance_{ij} \beta_d + size_j \beta_j + tweets_j \beta_t)$$
We wrote a function to calculate this value for an arbitrary distance matrix, size
and tweets vector, and coefficients.

```{r logsums_function}

#' Calculate destination choice logsums from a distance matrix
#' 
#' @param distances An $n\times p$ matrix with the distance from all tracts to
#'   all parks in miles.
#' @param sizes A p-length vector of park sizes
#' @param tweets A p-length vector of tweets at parks
#' @param betas A vector containing the size, distance, and tweet coefficients. 
#'   If we submit two variables the tweets are ignored.
#'   
#' @return An n-length vector containing the weighted log-sum based
#'   accessibility between a tract and all parks.
#' @details If we have n tracts and p parks, distances needs to be a 
#' 
calculate_park_logsums <- function(distances, sizes, tweets = NULL,
                                   betas = c(-15,.00001, .001)){

  # A is n x p
  a <- betas[1] * distances

  # B is p x 1
  b <- betas[2] * sizes

  if(!is.null(tweets)) 
    b <- b + betas[3] * tweets

  # calculate observed utility by adding the weighted park-level attributes 
  # to the columns of the matrix
  # V is n x p, with b added by-column to each element in a
  V <- sweep(a, MARGIN = 2, b, `+`)

  # log-sum of exponentiated utility, Output is n-length vector
  log(rowSums(exp(V)))

}
```

First, we calculate the distance between the tracts and the parks.
```{r distancematrix}
# calculate centroids of tracts and parks
tract_centroids <- gCentroid(tracts, byid = TRUE) 
park_centroids <- gCentroid(parks, byid = TRUE)

# distances from all tracts to all zones
distances <- gDistance(park_centroids, tract_centroids, byid = TRUE) / 5280
```

Now we want to generate an experiment where we supply different $\beta$ values
to the logsum calculation in order to maximize the log-likelihood function of the
regression model.

## Model with tweets

```{r tweets_logsum}
tweets_logsum <- function(betas, distances, parks, tracts) {

  logsum <- calculate_park_logsums(
    log(distances), 
    log(parks$Park_Acres), 
    yeo.johnson(parks$TWEET_COUNT, lambda = 0),
    betas = betas
  )

  estimation_data <- tracts@data %>% mutate(logsum = logsum)

  # sdm <- lagsarlm(
  #   update(base_formula, OBESITY ~ . + logsum),   
  #   data = estimation_data,  listw = W, zero.policy = TRUE, type = "mixed")
  sdm <- lm(update(base_formula, OBESITY ~ . + logsum), data = estimation_data)

  return(-as.numeric(logLik(sdm)))
}

tweets_lm_val <- optim(fn = tweets_logsum, 
                       par = c(-1, 0, 0), # distance, size, tweets
                       lower = c(-Inf, 0, -Inf), # size coefficient must be positive
                       upper = c(0, Inf, Inf), # distance coefficient must be negative
                       hessian=TRUE, 
                       distances = distances, parks = parks, tracts = tracts,
                       method = "L-BFGS-B", control = list(trace = 6))
```

```{r tweets_sdm}
tweets_ls <- calculate_park_logsums(
    log(distances), 
    log(parks$Park_Acres), 
    yeo.johnson(parks$TWEET_COUNT, lambda = 0),
    betas = tweets_lm_val$par)
tracts$tweets_ls <- tweets_ls - mean(tweets_ls)
tweets_sdm <- lagsarlm(update(base_formula, OBESITY ~ . + tweets_ls),  
         data = tracts@data,  listw = W, zero.policy = TRUE, type = "mixed")
```

## Model without Tweets

```{r notweets_logsum}
notweets_logsum <- function(betas, distances, parks, tracts) {

  logsum <- calculate_park_logsums(
    log(distances), 
    log(parks$Park_Acres), 
    betas = betas
  )

  estimation_data <- tracts@data %>% mutate(logsum = logsum)

  # sdm <- lagsarlm(
  #   update(base_formula, OBESITY ~ . + logsum),   
  #   data = estimation_data,  listw = W, zero.policy = TRUE, type = "mixed")
  sdm <- lm(update(base_formula, OBESITY ~ . + logsum), data = estimation_data)

  return(-as.numeric(logLik(sdm)))
}

notweets_lm_val <- optim(fn = notweets_logsum, 
                       par = c(-1, 0), # distance, size
                       lower = c(-100, 0), # size coefficient must be positive
                       upper = c(0, 100), # distance coefficient must be negative
                       hessian=TRUE, 
                       distances = distances, parks = parks, tracts = tracts,
                       method = "L-BFGS-B", control = list(trace = 6))
```

```{r tweets_sdm}
notweets_ls <- calculate_park_logsums( 
  log(distances),  
  log(parks$Park_Acres),  
  betas = notweets_lm_val$par)

tracts@data$notweets_ls <- notweets_ls - mean(notweets_ls)

notweets_sdm <- lagsarlm(update(base_formula, OBESITY ~ . + notweets_ls),  
         data = tracts@data,  listw = W, zero.policy = TRUE, type = "mixed")
```

```{r modelchart}
obesity_models <- list(Base = obese_base_sdm, `Access` = notweets_sdm, 
                       `Access + Tweets` = tweets_sdm)
screenreg(obesity_models, digits = 4)
```

```{r impacts_extractor}
#' Get the MC simulated impact coefficients and significance from a model
#' @param sdm An autoregressive lag (SAR or SDM) model object `lagsarlm`
#' @return A data frame with the effect (direct, indirect, total), variable, 
#'   simulated mean impact, and significance p-value
impacts_extractor <- function(sdm) {

  impacts_summary <- summary(
    impacts(sdm, tr=trMC,  R = 1000, useHESS = !is.logical(sdm$fdHess)), 
    zstats = TRUE)

  coef <- list(
    direct = impacts_summary$direct_sum,
    indirect = impacts_summary$indirect_sum,
    total = impacts_summary$total_sum
  ) %>%
  lapply(function(s) {
    data_frame(
      var = names(s$statistics[,1]),
      impact = s$statistics[,1]
    )
  }) %>%
    bind_rows(.id = "effect")

  pval <- as_data_frame(impacts_summary$pzmat) %>%
    mutate(
     var = rownames(impacts_summary$pzmat)
    ) %>%
    gather("effect", "p-val", -var) %>%
    mutate(effect = tolower(effect))

  left_join(coef, pval, by = c("effect", "var"))
}
```

```{r obesity_impacts}
obesity_impacts <- obesity_models %>%
  lapply(impacts_extractor) %>%
  bind_rows(.id = "model") %>%
  transmute(
    model, var, effect, 
    output = paste(round(impact, 5), 
                   gtools::stars.pval(`p-val`))
  ) %>%
  spread(model, output) 
```

# Physical Activity

We cannot necessarily assume that the logsum parameter will be constructed the
same way for physical activity as for obesity. (maybe?) But that's the assumption
we'll make for now.

```{r physact_estimation}
# turn around the metric to make sense
tracts$PHYSACT = 100 - tracts$Phys_Act
pa_base_sdm <- lagsarlm(update(base_formula, PHYSACT ~ .),
         data = tracts@data,  listw = W, zero.policy = TRUE, type = "mixed")
pa_notweets_sdm <- lagsarlm(update(base_formula, PHYSACT ~ . + notweets_ls),  
          data = tracts@data,  listw = W, zero.policy = TRUE, type = "mixed")
pa_tweets_sdm <- lagsarlm(update(base_formula, PHYSACT ~ . + tweets_ls),  
          data = tracts@data,  listw = W, zero.policy = TRUE, type = "mixed")
```

```{r physact_models}
physact_models <- list(Base = pa_base_sdm, Access = pa_notweets_sdm, 
                       `Access + Tweets` = pa_tweets_sdm)
screenreg(physact_models, digits = 4)
```

```{r physact_impacts}
physact_impacts <- physact_models %>%
  lapply(impacts_extractor) %>%
  bind_rows(.id = "model") %>%
  transmute(
    model, var, effect, 
    output = paste(round(impact, 5), 
                   gtools::stars.pval(`p-val`))
  ) %>%
  spread(model, output) 
```

### 2018-12-05 model_estimation.Rmd
Iteratively optimize function with tweets and with no tweets

@@
   tbl_df() %>%
-  select(raw = TWEET_COUNT) %>%
+  dplyr::select(raw = TWEET_COUNT) %>%
   filter(raw < 100) %>%
@@
   geom_line()
+
+parks@data <- parks@data %>%
+  mutate(
+    log_size = log(Park_Acres),
+    log_tweets = yeo.johnson(TWEET_COUNT, 0)
+  )
+
 ```
+
+## Other variables
 
@@
 tracts@data <- tbl_df(tracts@data) %>%
-  select(
+  dplyr::select(
     GEOID, OBESITY, Park_Percent, Phys_Act, MENTAL, 
@@
     PctOther, PctHispanic
+  )  %>%
+  mutate(
+    log_obesity = log(OBESITY),
+    log_physact = log(100 - Phys_Act)
   )
@@
 ```{r spatial selection}
-obese_base_lm <- lm(update(base_formula, OBESITY ~ .), data = tracts@data)
-obese_base_sar <- lagsarlm(update(base_formula, OBESITY ~ .), 
+obese_base_lm <- lm(update(base_formula, log_obesity ~ .), data = tracts@data)
+obese_base_sar <- lagsarlm(update(base_formula, log_obesity ~ .), 
                            data = tracts@data, listw = W, zero.policy = TRUE)
-obese_base_sem <- errorsarlm(update(base_formula, OBESITY ~ .), 
+obese_base_sem <- errorsarlm(update(base_formula, log_obesity ~ .), 
                              data = tracts@data, listw = W, zero.policy = TRUE)
-obese_base_sdm <- lagsarlm(update(base_formula, OBESITY ~ .), 
+obese_base_sdm <- lagsarlm(update(base_formula, log_obesity ~ .), 
                            data = tracts@data, listw = W, zero.policy = TRUE, 
@@
 # distances from all tracts to all zones
-distances <- gDistance(park_centroids, tract_centroids, byid = TRUE) / 5280
+distances <- log(gDistance(park_centroids, tract_centroids, byid = TRUE) / 5280)
 ```
@@
 to the logsum calculation in order to maximize the log-likelihood function of the
-regression model.
+regression model. This will help us identify the proper coefficients to use for
+the logsum construction.
 
-## Model with tweets
+## Model without Tweets
+
+```{r notweets_logsum}
+notweets_logsum <- function(betas, distances, parks, tracts, model) {
+
+  estimation_data <- tracts@data %>% 
+    mutate(logsum = calculate_park_logsums(
+      distances,  parks$log_size,
+      betas = betas))
+
+  sdm <- lagsarlm(update(model, . ~ . + logsum), data = estimation_data, 
+                  listw = W, zero.policy = TRUE, type = "mixed")
+  return(-as.numeric(logLik(sdm)))
+}
+
+access_sdm_val <- optim(
+  fn = notweets_logsum,  
+  par = c(-1, 0), # distance, size 
+  lower = c(-100, 0), # size coefficient must be positive 
+  upper = c(0, 100), # distance coefficient must be negative 
+  hessian=TRUE, distances = distances, parks = parks, tracts = tracts, 
+  model = obese_base_lm, method = "L-BFGS-B", control = list(trace = 6))
+```
 
 ```{r tweets_logsum}
-tweets_logsum <- function(betas, distances, parks, tracts) {
+tweets_logsum <- function(betas, distances, parks, tracts, model) {
 
-  logsum <- calculate_park_logsums(
-    log(distances), 
-    log(parks$Park_Acres), 
-    yeo.johnson(parks$TWEET_COUNT, lambda = 0),
-    betas = betas
-  )
+  estimation_data <- tracts@data %>% 
+    mutate(logsum = calculate_park_logsums(
+      distances,  parks$log_size,  parks$log_tweets,
+      betas = betas))
 
-  estimation_data <- tracts@data %>% mutate(logsum = logsum)
-
-  # sdm <- lagsarlm(
-  #   update(base_formula, OBESITY ~ . + logsum),   
-  #   data = estimation_data,  listw = W, zero.policy = TRUE, type = "mixed")
-  sdm <- lm(update(base_formula, OBESITY ~ . + logsum), data = estimation_data)
+  sdm <- lagsarlm(update(model, . ~ . + logsum), data = estimation_data, 
+                  listw = W, zero.policy = TRUE, type = "mixed")
 
@@
 
-tweets_lm_val <- optim(fn = tweets_logsum, 
-                       par = c(-1, 0, 0), # distance, size, tweets
-                       lower = c(-Inf, 0, -Inf), # size coefficient must be positive
-                       upper = c(0, Inf, Inf), # distance coefficient must be negative
-                       hessian=TRUE, 
-                       distances = distances, parks = parks, tracts = tracts,
-                       method = "L-BFGS-B", control = list(trace = 6))
+tweets_sdm_val <- optim(
+  fn = notweets_logsum,  
+  par = c(-1, 1, 0),
+  lower = c(-100, 0, -100), # size coefficient must be positive
+  upper = c(-1e-9, 100, 100), # distance coefficient must be negative
+  hessian=TRUE, distances = distances, parks = parks, tracts = tracts, 
+  model = obese_base_lm, method = "L-BFGS-B", control = list(trace = 6))
+
 ```
 
-```{r tweets_sdm}
-tweets_ls <- calculate_park_logsums(
-    log(distances), 
-    log(parks$Park_Acres), 
-    yeo.johnson(parks$TWEET_COUNT, lambda = 0),
-    betas = tweets_lm_val$par)
-tracts$tweets_ls <- tweets_ls - mean(tweets_ls)
-tweets_sdm <- lagsarlm(update(base_formula, OBESITY ~ . + tweets_ls),  
-         data = tracts@data,  listw = W, zero.policy = TRUE, type = "mixed")
-```
-
-## Model without Tweets
-
-```{r notweets_logsum}
-notweets_logsum <- function(betas, distances, parks, tracts) {
-
-  logsum <- calculate_park_logsums(
-    log(distances), 
-    log(parks$Park_Acres), 
-    betas = betas
+```{r obesity_access_models}
+tracts@data <- tracts@data %>%
+  mutate(
+    access_ls = calculate_park_logsums(distances, parks$log_size, 
+                                       betas = notweets_lm_val$par),
+    tweets_ls = calculate_park_logsums(distances, parks$log_size, parks$log_tweets, 
+                                       betas = tweets_lm_val$par),
+    # demean
+    access_ls = access_ls - mean(access_ls),
+    tweets_ls = tweets_ls - mean(tweets_ls)
   )
 
-  estimation_data <- tracts@data %>% mutate(logsum = logsum)
-
-  # sdm <- lagsarlm(
-  #   update(base_formula, OBESITY ~ . + logsum),   
-  #   data = estimation_data,  listw = W, zero.policy = TRUE, type = "mixed")
-  sdm <- lm(update(base_formula, OBESITY ~ . + logsum), data = estimation_data)
-
-  return(-as.numeric(logLik(sdm)))
-}
-
-notweets_lm_val <- optim(fn = notweets_logsum, 
-                       par = c(-1, 0), # distance, size
-                       lower = c(-100, 0), # size coefficient must be positive
-                       upper = c(0, 100), # distance coefficient must be negative
-                       hessian=TRUE, 
-                       distances = distances, parks = parks, tracts = tracts,
-                       method = "L-BFGS-B", control = list(trace = 6))
-```
-
-```{r tweets_sdm}
-notweets_ls <- calculate_park_logsums( 
-  log(distances),  
-  log(parks$Park_Acres),  
-  betas = notweets_lm_val$par)
-
-tracts@data$notweets_ls <- notweets_ls - mean(notweets_ls)
-
-notweets_sdm <- lagsarlm(update(base_formula, OBESITY ~ . + notweets_ls),  
-         data = tracts@data,  listw = W, zero.policy = TRUE, type = "mixed")
+access_sdm <- lagsarlm(update(obese_base_sdm, .~ . + access_ls),   
+                       data = tracts@data,  listw = W, zero.policy = TRUE, 
+                       type = "mixed")
+tweets_sdm <- lagsarlm(update(obese_base_sdm, .~ . + tweets_ls),   
+                       data = tracts@data,  listw = W, zero.policy = TRUE, 
+                       type = "mixed")
 ```
@@
 ```{r modelchart}
-obesity_models <- list(Base = obese_base_sdm, `Access` = notweets_sdm, 
+obesity_models <- list(Base = obese_base_sdm, `Access` = access_sdm, 
                        `Access + Tweets` = tweets_sdm)
@@
   coef <- list(
-    direct = impacts_summary$direct_sum,
-    indirect = impacts_summary$indirect_sum,
-    total = impacts_summary$total_sum
+    Direct = impacts_summary$direct_sum,
+    Indirect = impacts_summary$indirect_sum,
+    Total = impacts_summary$total_sum
   ) %>%
@@
     ) %>%
-    gather("effect", "p-val", -var) %>%
-    mutate(effect = tolower(effect))
+    gather("effect", "p-val", -var)
 
@@
 ```{r physact_estimation}
-# turn around the metric to make sense
-tracts$PHYSACT = 100 - tracts$Phys_Act
 pa_base_sdm <- lagsarlm(update(base_formula, PHYSACT ~ .),
          data = tracts@data,  listw = W, zero.policy = TRUE, type = "mixed")
-pa_notweets_sdm <- lagsarlm(update(base_formula, PHYSACT ~ . + notweets_ls),  
+pa_access_sdm <- lagsarlm(update(base_formula, PHYSACT ~ . + notweets_ls),  
           data = tracts@data,  listw = W, zero.policy = TRUE, type = "mixed")
@@
 ```
+
+# Results
+```{r impacts_table}
+list("Obesity" = obesity_impacts, "Physical activity" = physact_impacts) %>%
+  bind_rows(.id = "Outcome") %>%
+  filter(var %in% c("access_ls", "tweets_ls")) %>%
+  dplyr::select(-var, -Base)  %>%
+  gather("Variable", "Impact", -Outcome, -effect) %>%
+  filter(!is.na(Impact))  %>%
+  spread(effect, Impact)
+
+```
+
+```{r logsum_scales}
+data_frame(
+  Parameter = c("Distance", "Size", "Tweets"),
+  Access = c(notweets_lm_val$par, NA),
+  `Access + Tweets` = tweets_lm_val$par
+)
+
+```

### 2018-12-05 R/calculate_park_logsums.R
Function to calculate logsum

@@ -1,0 +1,35 @@
+
+#' Calculate destination choice logsums from a distance matrix
+#' 
+#' @param distances An $n\times p$ matrix with the distance from all tracts to
+#'   all parks in miles.
+#' @param sizes A p-length vector of park sizes
+#' @param tweets A p-length vector of tweets at parks
+#' @param betas A vector containing the size, distance, and tweet coefficients. 
+#'   If we submit two variables the tweets are ignored.
+#'   
+#' @return An n-length vector containing the weighted log-sum based
+#'   accessibility between a tract and all parks.
+#' @details If we have n tracts and p parks, distances needs to be a 
+#' 
+calculate_park_logsums <- function(distances, sizes, tweets = NULL,
+                                   betas = c(.00001, -15, .001)){
+  
+  # A is n x p
+  a <- betas[2] * distances 
+  
+  # B is p x 1
+  b <- betas[1] * sizes
+  
+  if(!is.null(tweets)) 
+    b <- b + betas[3] * tweets
+  
+  # calculate observed utility by adding the weighted park-level attributes 
+  # to the columns of the matrix
+  # V is n x p, with b added by-column to each element in a
+  V <- sweep(a, MARGIN = 2, b, `+`)
+  
+  # log-sum of exponentiated utility, Output is n-length vector
+  log(rowSums(exp(V)))
+  
+}

### 2018-12-06 manuscript.Rnw
Add manuscript shell

\documentclass[Afour,sageh,times]{sagej}
\def\journalname{Submitted to Environment and Planning B: Urban Analytics and City Science}

\begin{document}
\SweaveOpts{concordance=TRUE}

%% Front matter =========================
\title{The impacts of park access on health outcomes: a spatial comprehensive approach}

\author{Nico Boyd\affilnum{1}, Gregory S. Macfarlane\affilnum{2}, and Kari Watkins\affilnum{1}}
\runninghead{Boyd, et al.}

\affiliation{\affilnum{1}School of Civil and Environmental Engineering, Georgia Institute of Technology\\
\affilnum{2}Civil and Environmental Engineering Department, Brigham Young University}

\corrauth{Gregory Macfarlane\\
430 Engineering Building\\
Brigham Young University\\
Provo, UT 84602\\
USA}
\begin{abstract}
This research identifies the correlation between size and access to urban parks
and physical activity and obesity outcomes at the neighborhood level. Proximity
to parks is associated with increased physical activity and reduced obesity, but
little research has been conducted on the relationship between accessibility to
parks and health outcomes. Using data for three urban areas, we created a new
measure for access to parks called ‘park choice accessibility.’ Park choice
accessibility uses a gravity model to interact distance to parks and the quality
of those parks as defined by their size. A small park very close to a
neighborhood can have a major impact, but a larger park at a similar distance
may have an even larger impact. Similarly, a large park can be further away and
still have an impact on health outcomes. Using spatial econometric analysis, we
assess whether park choice accessibility is associated with increased physical
activity or decreased prevalence of obesity at the neighborhood level. The
analysis controls for socioeconomic covariates such as age, marital status,
income, and educational attainment.
\end{abstract}

\maketitle

\section{Introduction}
The United States is currently facing an epidemic of obesity and chronic
diseases, which are non-communicable diseases of long duration and typically
slow development, including cardiovascular diseases, chronic respiratory
diseases, diabetes, stroke, joint and bone diseases, and cancer (World Health
Organization 2014). Recent statistics suggest that approximately 300,000
premature deaths each year can be attributed to chronic diseases (World Health
Organization 2014). According to the Trust for Public Land, “On average, an
obese American racks up nearly $1,500 more per year in health care costs than an
American of normal weight, for a national total of $147 billion in direct
medical expenses” (Harnik & Welle 2011). As obesity and chronic disease have
become rampant, it is no surprise that healthcare costs have risen to nearly
one-fifth of the United States’ gross domestic product (Harnik & Welle 2011).

While a moderate amount of regular physical activity has been established as an
effective strategy for reducing and managing obesity and many of the
aforementioned chronic diseases (Office of Disease Prevention and Health
Promotion 2008; Centers for Disease Control and Prevention 2009a; Durstine et
al. 2013), Wolf notes that “more than 50 percent of U.S. adults do not get
enough physical activity to provide health benefits; and 24 percent are not
active at all in their leisure time. Activity decreases with age and sufficient
activity is less common among women than men, and among those with lower incomes
and less education” (?? 2008). Similarly, the United States Department of Health
and Human Services notes that less than fifty-percent of Americans meet
established recommendations for moderate to vigorous physical activity, or MVPA
(U.S. DHHS 2010). As the epidemiological transition from infectious to chronic
diseases is now complete in the developed world, increasing physical activity
has become a vital public health task in the 21st century.

Research has shown that the design of the built environment influences a range
of behaviors, including those related to physical activity. The presence of
trees and other vegetation in outdoor environments have been shown to be
positively associated with physical activity (Pretty et al 2005); one finding
suggests that after sidewalks and trails have been constructed, the introduction
of natural elements positively impacts motivation to engage in physical activity
(Suminski et al. 2005). Additional evidence indicates that commonly vegetated
areas, such as parks and open space, support outdoor physical activity
(Giles-Corti et al. 2005; Wells et al. 2007). Perhaps most telling is the
finding that “people in large cities perceive themselves to be generally more
healthy if a greater percentage of the living environment is greenspace, are
inclined to be more active, and claim the ability to relax faster” (Wolf 2008).
Thus, by providing space for active recreation, public parks and greenspaces may
increase the likelihood of engaging in more physical activity. As such, public
investment in parks can be thought of as a public health intervention for
chronic diseases and conditions, as well as general population health.

To date, the literature exploring the relationship between parks and health
outcomes, specifically those related to physical activity and obesity, has
yielded mixed results. Importantly, there is considerable variation in the
design of past studies, including the spatial scale of analysis, the population
of interest, and the measure of proximity and/or accessibility. While studies of
neighborhood level health impacts do exist, they typically focus on discrete
populations of interest within individual parks instead of examining the impacts
of larger park networks. Other studies have used the city and metropolitan
statistical area as the spatial scale of analysis, which in some cases has
resulted in the discovery of positive associations between parks and health
impacts but may conceal more nuanced relationships that exist at smaller spatial
scales. Evaluating the health impacts of parks and greenspaces using a smaller
unit of analysis is an important research gap to fill, especially considering
that park use and physical activity within parks varies considerably according
to residential proximity to parks and park facilities as well as a number of
sociodemographic factors (Kaczynski et al. 2014). This study attempts to fill
that gap using New York City as a case study example. Using a new measure of
park access called ‘Park Choice Accessibility,’ we hypothesize that greater park
access is associated with an increase in physical activity and a reduction in
obesity at the census tract level.

\section{Existing Literature}

\section{Methodology}

\subsection{Spatial weights selection}

\section{Results}

\section{Conclusion}

\begin{funding}
This project was funded in part via \ldots.
\end{funding}

\bibliographystyle{SageH}

\end{document}

### 2018-12-06 model_estimation.Rmd
//WIP on estimation

@@
 ```{r read_parks, echo=FALSE}
-parks <- geojson_read("data/nyc_parks.geojson", what = "sp")
+parks <- sf::st_read("data/nyc_parks.geojson")
 ```
@@
 ```{r tweettransformation}
-parks@data %>%
+parks %>%
   tbl_df() %>%
@@
 
-parks@data <- parks@data %>%
+parks <- parks %>%
   mutate(
@@
 
-parks@data <- tbl_df(parks@data)
 ```
@@
 ```{r W, echo=FALSE}
-neighbors <- poly2nb(tracts, queen = TRUE)
-W <- nb2listw(neighbours = neighbors, zero.policy = TRUE, style = "W")
+tracts.dnn <- dnearneigh(gCentroid(tracts, byid = TRUE), 0, 1.8 * 5280)
+dists <- nbdists(tracts.dnn, gCentroid(tracts, byid = TRUE))
+dists.inv <- lapply(dists, function(x) 1 / x)
+W <- nb2listw(neighbours = tracts.dnn, glist = dists.inv, zero.policy = TRUE, style = "W")
 print(W, zero.policy = TRUE)
 trMC <- trW(as(W, "CsparseMatrix"), type="MC") # trace used in montecarlo impacts
-ev <- eigenw(W)
 ```
@@
 lmtest::lrtest(obese_base_sem, obese_base_sdm)
+```
+
+The residuals appear homoskedastic with few outliers.
+
+```{r residuals}
+tracts_sf <- tracts_sf[complete_tracts_index, ] %>%
+  mutate(
+    log_obesity = tracts$log_obesity,  
+    obese_base_lm_residuals = obese_base_lm$residuals,
+    obese_base_sdm_fitted = obese_base_sdm$fitted.values,  
+    obese_base_sdm_residuals = obese_base_sdm$residual
+  )
+
+ggplot(tracts_sf, aes(x = log_obesity, y = obese_base_sdm_fitted)) + 
+  geom_point()
+```
+
+The residuals also appear to be distributed wtihout serious spatial correlation
+```{r residuals_map}
+pal <- colorQuantile("BrBG", tracts_sf$obese_base_sdm_residuals, n = 9)
+leaflet(tracts_sf) %>%
+  addProviderTiles(providers$CartoDB.Positron) %>%
+  addPolygons(color = ~pal(obese_base_sdm_residuals), stroke = NA, opacity = 0.8)
 ```
@@
   # V is n x p, with b added by-column to each element in a
-  V <- sweep(a, MARGIN = 2, b, `+`)
+  V <- sweep(a, 2, b, `+`)
 
@@
 # calculate centroids of tracts and parks
-tract_centroids <- gCentroid(tracts, byid = TRUE) 
-park_centroids <- gCentroid(parks, byid = TRUE)
+tract_centroids <- tracts %>% st_as_sf() %>% st_centroid() 
+park_centroids <- parks %>% st_centroid() %>% st_transform(st_crs(tract_centroids))
 
 # distances from all tracts to all zones
-distances <- log(gDistance(park_centroids, tract_centroids, byid = TRUE) / 5280)
+distances <- st_distance(park_centroids, tract_centroids, byid = TRUE) %>%
+  units::set_units(miles) %>%
+  units::drop_units()
 ```
@@
   par = c(-1, 0), # distance, size 
-  lower = c(-100, 0), # size coefficient must be positive 
+  lower = c(-100, -100), # size coefficient must be positive 
   upper = c(0, 100), # distance coefficient must be negative 
-  hessian=TRUE, distances = distances, parks = parks, tracts = tracts, 
+  hessian=TRUE, distances = log(distances[1:10, 1:10]), parks = parks[1:10,], tracts = tracts[1:10, ], 
   model = obese_base_lm, method = "L-BFGS-B", control = list(trace = 6))

### 2018-12-07 model_estimation.Rmd
Fix bug where park/tract distance matrix was built inverted

@@
 #' 
-#' @param distances An $n\times p$ matrix with the distance from all tracts to
-#'   all parks in miles.
+#' @param d An $n\times p$ matrix with the distance from all tracts to
+#'   all parks
 #' @param sizes A p-length vector of park sizes
@@
 #' 
-calculate_park_logsums <- function(distances, sizes, tweets = NULL,
+calculate_park_logsums <- function(d, sizes, tweets = NULL,
                                    betas = c(-15,.00001, .001)){
@@
   # A is n x p
-  a <- betas[1] * distances
+  a <- betas[1] * d
 
@@
 # distances from all tracts to all zones
-distances <- st_distance(park_centroids, tract_centroids, byid = TRUE) %>%
+distances <- st_distance(tract_centroids, park_centroids, byid = TRUE) %>%
   units::set_units(miles) %>%
@@
     mutate(logsum = calculate_park_logsums(
-      distances,  parks$log_size,
+      d = distances,  parks$log_size,
       betas = betas))
@@
   upper = c(0, 100), # distance coefficient must be negative 
-  hessian=TRUE, distances = log(distances[1:10, 1:10]), parks = parks[1:10,], tracts = tracts[1:10, ], 
+  hessian=TRUE, distances = log(distances), parks = parks, tracts = tracts, 
   model = obese_base_lm, method = "L-BFGS-B", control = list(trace = 6))

### 2018-12-07 model_estimation.Rmd
Revised how logsum gets into model estimation

@@
 
-  estimation_data <- tracts@data %>% 
-    mutate(logsum = calculate_park_logsums(
-      d = distances,  parks$log_size,
-      betas = betas))
+  logsum <- calculate_park_logsums(
+    d = distances,  sizes = parks$log_size,
+    betas = betas)
 
-  sdm <- lagsarlm(update(model, . ~ . + logsum), data = estimation_data, 
+  sdm <- lagsarlm(update(model, . ~ . + logsum), data = tracts@data, 
                   listw = W, zero.policy = TRUE, type = "mixed")
@@
     mutate(logsum = calculate_park_logsums(
-      distances,  parks$log_size,  parks$log_tweets,
+      d = distances,  sizes = parks$log_size,  tweets = parks$log_tweets,
       betas = betas))
 
-  sdm <- lagsarlm(update(model, . ~ . + logsum), data = estimation_data, 
+  sdm <- lagsarlm(update(model, . ~ . + logsum, data = tracts@data), 
                   listw = W, zero.policy = TRUE, type = "mixed")

### 2018-12-07 model_estimation.Rmd
More edits to model optimization functions

@@
 
-## Model without Tweets
-
 ```{r notweets_logsum}
@@
 
-  sdm <- lagsarlm(update(model, . ~ . + logsum), data = tracts@data, 
-                  listw = W, zero.policy = TRUE, type = "mixed")
-  return(-as.numeric(logLik(sdm)))
+  tracts@data$logsum <- logsum
+
+  sdm <- update(model, . ~ . + logsum)
+
+  return(logLik(sdm))
 }
@@
   fn = notweets_logsum,  
-  par = c(-1, 0), # distance, size 
-  lower = c(-100, -100), # size coefficient must be positive 
+  par = c(-1, 2), # distance, size 
+  lower = c(-100, 0), # size coefficient must be positive 
   upper = c(0, 100), # distance coefficient must be negative 
-  hessian=TRUE, distances = log(distances), parks = parks, tracts = tracts, 
-  model = obese_base_lm, method = "L-BFGS-B", control = list(trace = 6))
+  distances = log(distances), parks = parks, tracts = tracts, 
+  model = obese_base_sdm, 
+  method = "L-BFGS-B", 
+  control = list(fnscale = -1, trace = 6))
 ```
@@
 
-  estimation_data <- tracts@data %>% 
-    mutate(logsum = calculate_park_logsums(
+  logsum <- calculate_park_logsums(
     d = distances,  sizes = parks$log_size, tweets = parks$log_tweets,
-      betas = betas))
+    betas = betas)
 
-  sdm <- lagsarlm(update(model, . ~ . + logsum, data = tracts@data), 
-                  listw = W, zero.policy = TRUE, type = "mixed")
+  tracts@data$logsum <- logsum
+  sdm <- update(model, . ~ . + logsum)
 
-  return(-as.numeric(logLik(sdm)))
+  return(logLik(sdm))
 }
@@
   lower = c(-100, 0, -100), # size coefficient must be positive
-  upper = c(-1e-9, 100, 100), # distance coefficient must be negative
+  upper = c(0, 100, 100), # distance coefficient must be negative
   hessian=TRUE, distances = distances, parks = parks, tracts = tracts, 
-  model = obese_base_lm, method = "L-BFGS-B", control = list(trace = 6))
+  model = obese_base_sdm, 
+  method = "L-BFGS-B", 
+  control = list(fnscale = -1, trace = 6))
 
@@
   mutate(
-    access_ls = calculate_park_logsums(distances, parks$log_size, 
-                                       betas = notweets_lm_val$par),
-    tweets_ls = calculate_park_logsums(distances, parks$log_size, parks$log_tweets, 
-                                       betas = tweets_lm_val$par),
+    access_ls = calculate_park_logsums(log(distances), parks$log_size,
+                                       betas = access_sdm_val$par),
+    tweets_ls = calculate_park_logsums(log(distances), parks$log_size, parks$log_tweets,
+                                       betas = tweets_sdm_val$par),
     # demean
@@
 
-access_sdm <- lagsarlm(update(obese_base_sdm, .~ . + access_ls),   
-                       data = tracts@data,  listw = W, zero.policy = TRUE, 
-                       type = "mixed")
-tweets_sdm <- lagsarlm(update(obese_base_sdm, .~ . + tweets_ls),   
-                       data = tracts@data,  listw = W, zero.policy = TRUE, 
-                       type = "mixed")
+access_sdm <- update(obese_base_sdm, .~ . + access_ls)
+tweets_sdm <- update(obese_base_sdm, .~ . + tweets_ls)
 ```
@@
 ```{r physact_estimation}
-pa_base_sdm <- lagsarlm(update(base_formula, PHYSACT ~ .),
-         data = tracts@data,  listw = W, zero.policy = TRUE, type = "mixed")
-pa_access_sdm <- lagsarlm(update(base_formula, PHYSACT ~ . + notweets_ls),  
-          data = tracts@data,  listw = W, zero.policy = TRUE, type = "mixed")
-pa_tweets_sdm <- lagsarlm(update(base_formula, PHYSACT ~ . + tweets_ls),  
-          data = tracts@data,  listw = W, zero.policy = TRUE, type = "mixed")
+pa_base_sdm <- update(obese_base_sdm, log_physact ~ .)
+pa_access_sdm <- update(access_sdm, log_physact ~ .)
+pa_tweets_sdm <- update(tweets_sdm, log_physact ~ .)
 ```
@@
 ```{r physact_models}
-physact_models <- list(Base = pa_base_sdm, Access = pa_notweets_sdm, 
+physact_models <- list(Base = pa_base_sdm, Access = pa_access_sdm, 
                        `Access + Tweets` = pa_tweets_sdm)
@@
   Parameter = c("Distance", "Size", "Tweets"),
-  Access = c(notweets_lm_val$par, NA),
-  `Access + Tweets` = tweets_lm_val$par
+  Access = c(access_sdm_val$par, NA),
+  `Access + Tweets` = tweets_sdm_val$par
 )

### 2018-12-10 model_estimation.Rmd
WIP

@@
   upper = c(0, 100), # distance coefficient must be negative 
-  distances = log(distances), parks = parks, tracts = tracts, 
-  model = obese_base_sdm, 
+  distances = distances, parks = parks, tracts = tracts, 
+  model = obese_base_lm, 
   method = "L-BFGS-B", 
@@
   logsum <- calculate_park_logsums(
-    d = distances,  sizes = parks$log_size, tweets = parks$log_tweets,
+    d = distances,  sizes = parks$Park_Acres, tweets = parks$TWEET_COUNT,
     betas = betas)
@@
   fn = notweets_logsum,  
-  par = c(-1, 1, 0),
+  par = c(access_sdm_val$par, 0),
   lower = c(-100, 0, -100), # size coefficient must be positive
@@
     access_ls = calculate_park_logsums(log(distances), parks$log_size,
-                                       betas = access_sdm_val$par),
+                                       betas = tweets_sdm_val$par),
     tweets_ls = calculate_park_logsums(log(distances), parks$log_size, parks$log_tweets,

### 2018-12-11 model_estimation.Rmd
WIP on estimation notebook

--- 2018-12-07 model_estimation.Rmd
@@
   fn = notweets_logsum,  
-  par = c(-1, 2), # distance, size 
+  par = c(-10, 0.001), # distance, size 
   lower = c(-100, 0), # size coefficient must be positive 
@@
   fn = notweets_logsum,  
-  par = c(-1, 1, 0),
+  par = c(access_sdm_val$par, 0),
   lower = c(-100, 0, -100), # size coefficient must be positive
   upper = c(0, 100, 100), # distance coefficient must be negative
-  hessian=TRUE, distances = distances, parks = parks, tracts = tracts, 
+  hessian=TRUE, distances = log(distances), parks = parks, tracts = tracts, 
   model = obese_base_sdm, 
@@
 
-```{r obesity_access_models}
+```{r show_accessibility}
 tracts@data <- tracts@data %>%
@@
     access_ls = calculate_park_logsums(log(distances), parks$log_size,
-                                       betas = access_sdm_val$par),
+                                       betas = c(-0.8, 0.8)),
     tweets_ls = calculate_park_logsums(log(distances), parks$log_size, parks$log_tweets,
-                                       betas = tweets_sdm_val$par),
-    # demean
-    access_ls = access_ls - mean(access_ls),
-    tweets_ls = tweets_ls - mean(tweets_ls)
+                                       betas = c(-0.8, 0.8, 0.2))
   )
 
+tracts_sf <- tracts_sf %>%
+  mutate(
+    access_ls = tracts$access_ls,
+    tweets_ls = tracts$tweets_ls
+  )
+
+pal <- colorQuantile("Spectral", c(tracts_sf$access_ls, tracts_sf$tweets_ls), 5)
+
+leaflet(tracts_sf) %>%
+  addProviderTiles(providers$CartoDB.Positron) %>%
+  addPolygons(group = "access", color = ~pal(access_ls)) %>%
+  addPolygons(group = "access + tweets", color = ~pal(tweets_ls)) %>%
+  addLayersControl(baseGroups = c("access", "access + tweets"))
+
+```
+
+```{r obesity_access_models}
 access_sdm <- update(obese_base_sdm, .~ . + access_ls)
@@
 ```{r impacts_table}
-list("Obesity" = obesity_impacts, "Physical activity" = physact_impacts) %>%
+list("Obesity" = obesity_impacts, "Physical Activity" = physact_impacts) %>%
   bind_rows(.id = "Outcome") %>%
@@
   dplyr::select(-var, -Base)  %>%
-  gather("Variable", "Impact", -Outcome, -effect) %>%
+  gather("Measure", "Impact", -Outcome, -effect) %>%
   filter(!is.na(Impact))  %>%

### 2018-12-11 manuscript.Rnw
Add first draft of literature review as Sweave doc

@@
 \maketitle
+<<setup, cache=FALSE, echo=FALSE, message=FALSE>>=
+library(tidyverse)
+knitr::opts_chunk$set(cache=TRUE)
+@
 
 \section{Introduction}
+
 The United States is currently facing an epidemic of obesity and chronic
@@
 slow development, including cardiovascular diseases, chronic respiratory
-diseases, diabetes, stroke, joint and bone diseases, and cancer (World Health
-Organization 2014). Recent statistics suggest that approximately 300,000
-premature deaths each year can be attributed to chronic diseases (World Health
-Organization 2014). According to the Trust for Public Land, “On average, an
-obese American racks up nearly $1,500 more per year in health care costs than an
-American of normal weight, for a national total of $147 billion in direct
-medical expenses” (Harnik & Welle 2011). As obesity and chronic disease have
-become rampant, it is no surprise that healthcare costs have risen to nearly
-one-fifth of the United States’ gross domestic product (Harnik & Welle 2011).
+diseases, diabetes, stroke, joint and bone diseases, and cancer. The \citet{WHO2014} estimates approximately 300,000
+premature deaths in the United States each year can be attributed to chronic diseases.
+\citet{Finkelstein2009} find an
+obese American spends nearly \$1,500 more per year in health care costs than an
+American of normal weight. As obesity and chronic disease
+have become rampant, it is no surprise that healthcare costs have risen to
+nearly one-fifth of the United States’ gross domestic product \citep{Harnik2011}.
 
@@
 effective strategy for reducing and managing obesity and many of the
-aforementioned chronic diseases (Office of Disease Prevention and Health
-Promotion 2008; Centers for Disease Control and Prevention 2009a; Durstine et
-al. 2013), Wolf notes that “more than 50 percent of U.S. adults do not get
-enough physical activity to provide health benefits; and 24 percent are not
-active at all in their leisure time. Activity decreases with age and sufficient
+aforementioned chronic diseases \citep{Gordon2013}.
+\citeauthor{Wolf2008} notes that “more than 50 percent of U.S. adults do not get
+enough physical activity to provide health benefits; 24 percent are not active
+at all in their leisure time. Activity decreases with age and sufficient
 activity is less common among women than men, and among those with lower incomes
-and less education” (?? 2008). Similarly, the United States Department of Health
-and Human Services notes that less than fifty-percent of Americans meet
+and less education” \cite[22]{Wolf2008}. Similarly, the United States Department of
+Health and Human Services notes that less than fifty-percent of Americans meet
 established recommendations for moderate to vigorous physical activity, or MVPA
@@
 trees and other vegetation in outdoor environments have been shown to be
-positively associated with physical activity (Pretty et al 2005); one finding
+positively associated with physical activity \citep{Pretty2005}; one finding
 suggests that after sidewalks and trails have been constructed, the introduction
 of natural elements positively impacts motivation to engage in physical activity
-(Suminski et al. 2005). Additional evidence indicates that commonly vegetated
+\citep{Suminski2005}. Additional evidence indicates that commonly vegetated
 areas, such as parks and open space, support outdoor physical activity
-(Giles-Corti et al. 2005; Wells et al. 2007). Perhaps most telling is the
-finding that “people in large cities perceive themselves to be generally more
-healthy if a greater percentage of the living environment is greenspace, are
-inclined to be more active, and claim the ability to relax faster” (Wolf 2008).
-Thus, by providing space for active recreation, public parks and greenspaces may
-increase the likelihood of engaging in more physical activity. As such, public
-investment in parks can be thought of as a public health intervention for
-chronic diseases and conditions, as well as general population health.
+\cite{Giles-Corti2005, Wells2007}. Perhaps most telling is the
+finding that “people in large cities perceive themselves to be generally
+healthier if a greater percentage of the living environment is greenspace, are
+inclined to be more active, and claim the ability to relax faster”
+\cite[24]{Wolf2008}. Thus, by providing space for active recreation, public
+parks and greenspaces may increase the likelihood of engaging in more physical
+activity. As such, public investment in parks can be thought of as a public
+health intervention for chronic diseases and conditions, as well as general
+population health.
 
@@
 populations of interest within individual parks instead of examining the impacts
-of larger park networks. Other studies have used the city and metropolitan
-statistical area as the spatial scale of analysis, which in some cases has
-resulted in the discovery of positive associations between parks and health
-impacts but may conceal more nuanced relationships that exist at smaller spatial
-scales. Evaluating the health impacts of parks and greenspaces using a smaller
+of larger park networks \citep{Bancroft2015, Coutts2008}.
+Other studies have used the city and metropolitan statistical area as the
+spatial scale of analysis, which
+in some cases has resulted in the discovery of positive associations between
+increased park space and positive health impacts but may conceal more nuanced
+relationships that exist at smaller spatial scales \citep{Larson2016}.
+Evaluating the health impacts of parks and greenspaces using a smaller
 unit of analysis is an important research gap to fill, especially considering
@@
 to residential proximity to parks and park facilities as well as a number of
-sociodemographic factors (Kaczynski et al. 2014). This study attempts to fill
-that gap using New York City as a case study example. Using a new measure of
-park access called ‘Park Choice Accessibility,’ we hypothesize that greater park
-access is associated with an increase in physical activity and a reduction in
-obesity at the census tract level.
+sociodemographic factors \citep{Kaczynski2014}. There are important urban
+design considerations here; is it better to build a single large park with many
+different amenities, or to build a series of smaller parks nearer people’s
+homes. Similarly, do parks placed near low-income households affect the behavior
+of those households? This study attempts to fill that gap using New York City as
+a case study example. Using a new measure of park access called ‘Park Choice
+Accessibility,’ we find
 
 \section{Existing Literature}
+
+\subsection{Urbanization and Public Health Outcomes}
+Across the world, the population of cities is growing at an unprecedented rate.
+With a current estimated global urban population of over 3.5 billion people, the
+United Nations Population Fund estimates that by 2030, 5 billion people will
+inhabit cities worldwide.  By 2050, an additional 3 billion people will live in
+cities – a sixty percent increase over twenty years – with much of this growth
+expected to take place in the developing world (United Nations Population Fund,
+2016). Despite the greater expected concentration of urban growth in other parts
+of the globe, the trend of urbanization is highly visible in the United States
+as well.  According to the U.S. Census Bureau, ''The nation's urban population
+increased by 12.1 percent from 2000 to 2010, outpacing the nation's overall
+growth rate of 9.7 percent for the same period.'' \citet{Nowak2005}
+ estimate that urban land area as a percentage of total land area in the
+U.S. will increase from 3.1\% in 2000 to approximately 8.1\% by the year 2050,
+with urbanized places collectively comprising a land area larger than the state
+of Montana. More than eighty percent of the U.S. population now lives in
+urban areas, compared to sixty-four percent in 1950 (United States Census Bureau
+2007).
+
+The global trend of urbanization has profound implications for population
+health, both positive and negative. Because urbanization corresponds with an
+increase in population density compared to rural and suburban settlement
+patterns, residents of dense urban areas throughout the world have better access
+on average to many health services and health-promoting amenities simply because
+of their closer proximity to such resources \citep{Larson2016}.
+However, despite this benefit of increased density and the ability of cities to
+provide many opportunities for innovation, economic growth and social progress,
+historical and current evidence suggests that rapid urban growth often leads to
+congestion and numerous negative environmental and human health outcomes.
+Interactions between growing urban populations and their environment, marked by
+intensive resource consumption and ecologically harmful patterns of development,
+lead to numerous undesired consequences including pollution and sanitation
+issues as well as racial and socioeconomic disparities \citep{Larson2016}.
+
+Because of the considerable growth of urbanization throughout the world, many
+city planning and public health professionals have begun to pay more attention
+to the role of the built environment in promoting or discouraging health
+behaviors. Until recently, most large-scale health promotion efforts focused on
+individual-level interventions intended to educate people about health
+lifestyles and behaviors, touching on topics including diet and exercise.
+According to \citet{Coutts2008} however, this trend is shifting as professionals
+''have begun adopting an ecological paradigm, accepting that both individual and
+environmental determinants play a role in health behavior. This new, arguably
+revisited, public-health paradigm accounts not only for the compositional (who
+you are) but also for the contextual (where you are) influences on physical
+activity.'' As professionals begin to operate from the assumption that the design
+and configuration of the built environment can facilitate or inhibit physical
+activity, they are increasingly looking to public spaces like parks and
+greenways as key elements of the built environment that can support exercise
+\citet{Coutts2008, Bedimo-Rung2005}.
+
+\subsection{Review of the Health Benefits of Parks}
+A large body of literature exists documenting the many benefits of urban parks
+and greenspaces to human health and wellbeing. Importantly, these benefits stem
+from the provision of ecosystem services, which occur when the natural
+environment supplies something that people demand, improving quality of life and
+well-being \citep{Larson2016}. These services can include the
+provisioning of goods such as fresh water and agricultural products; regulatory
+functions including protection of drinking water quality, heat mitigation, air
+purification, and stormwater management; and cultural functions, such as
+improving aesthetics, providing opportunities for recreation, tourism, and
+physical and mental health, and promoting biodiversity \citep{McDonald2015}.
+According to \citet{Wolch2014}, parks and greenspaces provide ''\ldots a wide
+range of ecosystem services that could help combat many urban ills and improve
+life for city dwellers—especially their health…Ecosystem services provided by
+urban greenspace not only support the ecological integrity of cities, but can
+also protect the public health of urban populations.'' Table
+\ref{tab:parkbenefits} summarizes commonly cited benefits of parks and
+greenspaces.
+
+\begin{table*}[ht]
+  \caption{Commonly cited benefits of parks and greenspaces}
+  \label{tab:parkbenefits}
+\begin{tabular}{lp{14cm}}
+  \hline
+Category & Summary of Benefits\\
+\hline
+Physical Health & Provide clean drinking water \citep{Benedict2006}\\
+ & Foster increased physical activity\\
+ & Promote faster healing in hospitals \citep{Akbari2001}\\
+ & Reduce of heat-related mortality \\
+ & Reduce incidence of cardiovascular-related mortality \\
+ & Improved air quality and related reductions in respiratory-related mortality \citep{Lovasi2008}\\
+ & Reduce incidence of childhood asthma \citep{Wolf2008}\\
+ & Improve birth outcomes\\
+\hline
+Mental Health & Reduce stress and mental fatigue\\
+  & Reduce aggression\\
+  & Enhance emotional and cognitive development\\
+  & Improve behavioral outcomes in youth \\
+\hline
+Social Health & Enhanced community aesthetics\\
+ & Reduce crime\\
+ & Increase social interaction \\
+\hline
+Economic Health & Provide of ecosystem services\\
+ & Increase residential property values and municipal property tax revenues\\
+ & Attract more shoppers and increase economic activity to commercial districts \\
+\hline
+\end{tabular}
+\end{table*}
+
+Given the general benefits of parks and greenspaces in Table \ref{tab:parkbenefits}, several studies
+have attempted to quantify the impacts of these spaces on different facets of
+health and wellbeing at the city level, yielding mixed results. \citet{Larson2016}
+used self-reported scores on the Gallup-Healthways Wellbeing
+Index to evaluate the relationship between different areas of wellbeing,
+including physical health, and park quantity, quality, and accessibility in 44
+U.S. cities. The authors found that “Park quantity (measured as the percentage
+of city area covered by public parks) was among the strongest predictors of
+overall wellbeing, and the strength of this relationship appeared to be driven
+by parks’ contributions to physical and community wellbeing” (Larson, Jennings \&
+Cloutier 2016). While the authors found positive associations between wellbeing
+and park quality and accessibility, these relationships were not statistically
+significant. A study by \citet{West2012} used park data from the
+Trust for Public Land’s 2010 City Park Facts and public health data from the
+Behavioral Risk Factor Surveillance System (BRFSS) to examine relationships
+between the density of parkland, parkland per capita, and levels of physical
+activity and obesity for 67 metropolitan statistical areas in the U.S. The study
+found a significant, positive association between park density and physical
+activity and a significant, negative association between park density and
+obesity. In a study of New York City, \citet{Stark2014} found that the
+“proportion of neighborhoods that was large or small park space and park
+cleanliness were associated with lower BMI among NYC adults after adjusting for
+other neighborhood features such as homicides and walkability, characteristics
+that could influence park usage.”
+
+Interestingly, in a study by \citet{Richardson2012} that examined the
+relationship between urban greenspace and selected mortality rates, the authors
+did not find a statistically significant relationship between the quantity of
+urban greenspace and mortality caused by lung cancer, diabetes, heart disease or
+car accidents, but found that all-cause mortality was significantly higher in
+greener cities. In a meta-analysis of 20 peer reviewed journal articles
+exploring the relationship between parks and objectively measured physical
+activity, \citet{Bancroft2015} found that five studies reported a significant
+positive association between the two, six studies produced mixed results, and
+nine studies found no association at all.
+
+While several of the aforementioned studies yielded results consistent with the
+hypothesis that park access and use confer health benefits stemming from
+increased physical activity and attendant decreases in obesity, all of the
+studies  -- except some that were included in the meta-analysis by
+\citet{Bancroft2015} -– were conducted at the level of the city or metropolitan statistical
+area, potentially concealing variation in health outcomes at more fine-grained
+levels of analysis \citep{Larson2016}.
+
+\subsection{Accessibility to Parks}
+To evaluate the impacts of parks and greenspaces on health outcomes at any
+scale, one of the fundamental tasks involved is the selection of a measure of
+accessibility. In a comprehensive meta-analysis of published studies that
+measure active accessibility (accessibility using non-motorized travel modes
+including walking and cycling), \citet{Vale2016} identified four
+broad categories of studies based on the active accessibility metric employed:
+“distance-based, gravity-based or potential, topological or
+infrastructure-based, and walkability and walk score-type measures.”
+Distance-based measures account only for the Euclidean distance between origins
+and destinations, while infrastructure-based measures explicitly incorporate
+relevant transportation networks like roads and sidewalks to more accurately
+measure travel time and distance. Gravity-based measures incorporate cost
+measures to model accessibility as a function of a destination’s attractiveness
+(i.e., size, commercial activity, etc.) and the cost of traveling to that
+destination from a given origin (i.e., travel time or distance).
+
+Importantly, \citeauthor{Vale2016} acknowledge that there is not yet a
+consensus on the most appropriate accessibility metric to use in a given
+setting, noting that ''ways to measure active accessibility are as varied as the
+number of scholars the measure them.'' However, their meta-analysis
+highlights some of the limitations associated with each of the types of metrics
+described. For example, in describing distance-based accessibility measures, the
+authors note that such measures are ''extremely sensitive to the way in which
+travel impedance (i.e., distance) is measured. Accordingly, four types of
+distance can be identified: Euclidean distance, Manhattan distance, shortest
+network distance, and shortest network time''. The
+appropriateness of one distance-based measure over another can vary
+significantly depending on the topography of the environment and the travel mode
+that is being employed. In describing gravity-based accessibility measures,
+which ''assume that travel is a derived demand and there is a tradeoff between
+the benefit of the opportunity and the cost to reach it from a given origin,''
+the authors note that such measures do not always explicitly account for land
+use characteristics near origins and destinations, which may impact that true
+accessibility of those places.
+
+In evaluating methodologies of measuring access to urban services, including
+parks and greenspaces, \citet{Logan2017} note that existing and commonly used
+approaches “often simplify their measure of proximity by using large areal
+units and by imposing arbitrary distance thresholds,” which often results in
+access-poor populations being overlooked. The authors concede that many
+existing approaches have long been necessary because the computational power
+required to use higher-resolution analytical techniques was unavailable.
+However, due to recent advances in computation power and the advent of
+municipal open-data policies, \citeauthor{Logan2017} recommend that future
+analyses of accessibility disaggregate population data to the building or
+parcel level and use network distance instead of Euclidean distance to measure
+proximity.
+
+\section{Choice-based Accessibility}
+
+Consider that an individual is choosing a park for a recreation activity.
+According to basic choice theory \citep{McFadden1974}, the probability of choosing
+park $p$ from the set of all regional parks $J$ is:
+
+$$ P(p | V_p) = \frac{\exp(V_p)}{\sum_{j \in J}\exp(V_j)}$$
+where parks are differentiated from each other by their relative measurable
+utilitie $V$. In principle, $V$ may include any measurable attributes of either
+the choice maker or the park. In this study we use a linear formulation of
+
+$$V_{ij} = size_j\beta_s + distance_{ij}\beta_d$$
+incorporating the size of the park in acres and the distance of the park from
+the census tract $i$ in miles. The coefficients $\beta$ are typically estimated from
+surveys, though in the absence of a survey we apply a maximum likelihood
+technique described below.
+
+A key theoretical understanding of random utility choice models is that the
+consumer surplus of the choice set can be otained as the log-sum of the
+denominator of the choice probability equation. In plainer terms, the *total value*
+of an individual's park accessibility is defined to be:
+
+$$CS_i = \ln\left({\sum_{j \in J}\exp(V_{ij})}\right)$$
+
+There are several advantages to a log-sum defined metric relative to
+buffer-based accessibility metrics more commonly found in the literature. First,
+all individuals are defined as having some access to all parks, rather than an
+arbitrary limit of 1/2 mile or so. This allows for the fact that some people are
+more or less sensitive to distances, and that distance is a continuous, and not
+a binary, phenomenon. Second, the random utility formulation allows the
+researcher to include any attribute of the park; in this case, we consider the
+size of the park as an element of accessibility. This suggests that not all
+parks are equal, and that a large park such as New York City's Central Park may
+provide health and activity benefits over a much larger area than a smaller
+community square.
+
+In the absence of a park choice survey, we estimated likely values of the $\beta$
+coefficients by iteratively searching for the values which produced the highest
+model likelihood in *which model did we use to estimate likelihood?*. The search
+was constrained by two assertions. First, we required that the coefficient on
+size be positive and that on distance negative; all else equal, people will
+prefer to use larger parks that are nearer to their residence. Second, we visually
+inspected the resulting accessibility scores to ensure that the scores produced a
+reasonably varied pattern of access throughout the metropolitan region.
+
+$$ A_i = log\sum_{j \in J}\exp(distance_{ij} \beta_d + size_j \beta_j + tweets_j \beta_t)$$
+We wrote a function to calculate this value for an arbitrary distance matrix, size
+and tweets vector, and coefficients.
 
@@
 \bibliographystyle{SageH}
+\bibliography{citations}
 

### 2018-12-11 model_estimation.Rmd
Remove likelihood experiment
When we maximize the likelihood, we end up with a distance value of 
-0.4, but a size value of 0. This doesn't accord with expectations, so 
we discard it.

@@
 
-Now we want to generate an experiment where we supply different $\beta$ values
-to the logsum calculation in order to maximize the log-likelihood function of the
-regression model. This will help us identify the proper coefficients to use for
-the logsum construction.
-
-```{r notweets_logsum}
-notweets_logsum <- function(betas, distances, parks, tracts, model) {
-
-  logsum <- calculate_park_logsums(
-    d = distances,  sizes = parks$log_size,
-    betas = betas)
-
-  tracts@data$logsum <- logsum
-
-  sdm <- update(model, . ~ . + logsum)
-
-  return(logLik(sdm))
-}
-
-access_sdm_val <- optim(
-  fn = notweets_logsum,  
-  par = c(-10, 0.001), # distance, size 
-  lower = c(-100, 0), # size coefficient must be positive 
-  upper = c(0, 100), # distance coefficient must be negative 
-  distances = log(distances), parks = parks, tracts = tracts, 
-  model = obese_base_sdm, 
-  method = "L-BFGS-B", 
-  control = list(fnscale = -1, trace = 6))
-```
-
-```{r tweets_logsum}
-tweets_logsum <- function(betas, distances, parks, tracts, model) {
-
-  logsum <- calculate_park_logsums(
-    d = distances,  sizes = parks$log_size, tweets = parks$log_tweets,
-    betas = betas)
-
-  tracts@data$logsum <- logsum
-  sdm <- update(model, . ~ . + logsum)
-
-  return(logLik(sdm))
-}
-
-tweets_sdm_val <- optim(
-  fn = notweets_logsum,  
-  par = c(access_sdm_val$par, 0),
-  lower = c(-100, 0, -100), # size coefficient must be positive
-  upper = c(0, 100, 100), # distance coefficient must be negative
-  hessian=TRUE, distances = log(distances), parks = parks, tracts = tracts, 
-  model = obese_base_sdm, 
-  method = "L-BFGS-B", 
-  control = list(fnscale = -1, trace = 6))
-
-```
-
 ```{r show_accessibility}
@@
 ```
-
-```{r logsum_scales}
-data_frame(
-  Parameter = c("Distance", "Size", "Tweets"),
-  Access = c(access_sdm_val$par, NA),
-  `Access + Tweets` = tweets_sdm_val$par
-)
-
-```

### 2018-12-11 manuscript.Rnw
Add some data to the manuscript

@@
 library(tidyverse)
+library(geojsonio)
+library(rgdal)
+library(rgeos)
+library(maptools)
+library(spdep)
+library(sf)
+library(VGAM)
 knitr::opts_chunk$set(cache=TRUE)
+
 @
@@
 
-\subsection{Spatial weights selection}
+<<tracts>>=
+tracts <- geojson_read("data/nyc_tracts.geojson", what = "sp")
+tracts_sf <- st_read("data/nyc_tracts.geojson", quiet = TRUE) %>%
+  st_transform(4326)
+
+# identify complete cases
+complete_tracts_index <- tracts_sf %>%
+  st_set_geometry(NULL) %>%
+  tbl_df() %>%
+  complete.cases()
+tracts <- tracts[complete_tracts_index, ]
+@
+
+<<parks>>=
+parks <- sf::st_read("data/nyc_parks.geojson", quiet = TRUE) %>%
+  mutate(
+    log_size = log(Park_Acres),
+    log_tweets = yeo.johnson(TWEET_COUNT, 0)
+  )
+@
+
+\subsection{Spatial weight and model selection}
+
+<<weights>>=
+tracts.dnn <- dnearneigh(gCentroid(tracts, byid = TRUE), 0, 1.8 * 5280)
+dists <- nbdists(tracts.dnn, gCentroid(tracts, byid = TRUE))
+dists.inv <- lapply(dists, function(x) 1 / x)
+W <- nb2listw(neighbours = tracts.dnn, glist = dists.inv, zero.policy = TRUE, style = "W")
+trMC <- trW(as(W, "CsparseMatrix"), type="MC") # trace used in montecarlo impacts
+@
 

### 2018-12-11 manuscript.Rnw
Change to single-column for draft
Add spatial econometrics discussion

@@
-\documentclass[Afour,sageh,times]{sagej}
+\documentclass[Crown,sageh,times]{sagej}
 \def\journalname{Submitted to Environment and Planning B: Urban Analytics and City Science}
 
+\usepackage{amsmath}
+     \def\Var{{\rm Var}\,}
+  	 \def\Cov{{\rm Cov}\,}
+     \def\E{{\rm E}\,}
+		 \def\Corr{{\rm Corr}\,}
+\newcommand{\vect}[1]{\boldsymbol{#1}}
+\newcommand{\?}{\overset{?}{=}}
+
 \begin{document}
-\SweaveOpts{concordance=TRUE}
 
@@
   \label{tab:parkbenefits}
-\begin{tabular}{lp{14cm}}
+\begin{tabular}{lp{11cm}}
   \hline
@@
 where parks are differentiated from each other by their relative measurable
-utilitie $V$. In principle, $V$ may include any measurable attributes of either
+utilities $V$. In principle, $V$ may include any measurable attributes of either
 the choice maker or the park. In this study we use a linear formulation of
@@
 
-In the absence of a park choice survey, we estimated likely values of the $\beta$
-coefficients by iteratively searching for the values which produced the highest
-model likelihood in *which model did we use to estimate likelihood?*. The search
-was constrained by two assertions. First, we required that the coefficient on
-size be positive and that on distance negative; all else equal, people will
-prefer to use larger parks that are nearer to their residence. Second, we visually
-inspected the resulting accessibility scores to ensure that the scores produced a
-reasonably varied pattern of access throughout the metropolitan region.
-
-$$ A_i = log\sum_{j \in J}\exp(distance_{ij} \beta_d + size_j \beta_j + tweets_j \beta_t)$$
-We wrote a function to calculate this value for an arbitrary distance matrix, size
-and tweets vector, and coefficients.
-
 \section{Methodology}
@@
 
+We obtained data on aggregate health outcomes from the Centers for Disease
+Control and Prevention (CDC) for census tracts in the New York City area. After
+removing tracts with missing population information, we have Sexpr{nrow(tracts)}
+complete cases.
+
 <<parks>>=
@@
 
-\subsection{Spatial weight and model selection}
+We also collected data on park geometry.
+
+\subsection{Park Choice Logsum}
+In the absence of a park choice survey, we estimated likely values of the $\beta$
+coefficients by iteratively searching for the values which produced the highest
+model likelihood in *which model did we use to estimate likelihood?*. The search
+was constrained by two assertions. First, we required that the coefficient on
+size be positive and that on distance negative; all else equal, people will
+prefer to use larger parks that are nearer to their residence. Second, we visually
+inspected the resulting accessibility scores to ensure that the scores produced a
+reasonably varied pattern of access throughout the metropolitan region.
+
+$$ A_i = log\sum_{j \in J}\exp(distance_{ij} \beta_d + size_j \beta_j + tweets_j \beta_t)$$
+We wrote a function to calculate this value for an arbitrary distance matrix, size
+and tweets vector, and coefficients.
+
+\subsection{Spatial dependence and correlation}
+
+Using tracts as the unit of observation in the model allows us to consider the
+availability of parks at a metropolitan level, but it also introduces the problem
+spatial correlation of unobserved errors and spatial dependence in the data
+generating process \citep{Cliff1970, Anselin1980}. Spatial correlation occurs
+when spatially-distributed unobservable or missing variables (school quality,
+neighborhood prestige, etc.) influence the modeled outcomes. Spatial dependence
+occurs when the outcome {\em depends} on the outcome in neighboring areas; for
+example, seeing neighbors exercising may encourage exercise. The relaxed spatial
+Durbin model (SDM) proposed by \citet{Burridge1981} controls for both processes:
+\begin{equation}\label{eq:sdm}
+  \vect{y}= \rho W\vect{y} + X\vect{\beta} + WX\vect{\gamma} + \vect{\epsilon}
+\end{equation}
+where $\rho$ is the estimated spatial correlation between the outcome variables
+$y_i, y_j, \ldots  \in \vect{y}$, $W$ is a matrix specifying the spatial
+relationship between observations $i$ and $j$ in the dataset,
+$\vect{\beta}$ is a vector of estimable coefficients relating the attributes $X$
+of an observation to its outcome, and $\vect{\gamma}$ is a vector of
+estimated coefficients relating the attributes of an observation's {\em neighbors}
+to its outcome.
+
+The researcher must assert $W$ {\em a priori}, though there are several
+common specifications \citep{Dubin1998}. In this study we specify that tracts
+with centroids located within 1.8 miles of each other are considered neighbors,
+and the strength of the relationship is the inverse of the distance between
+them:
+\begin{equation}\label{eq:weightsrule}
+W_{ij} = \begin{cases}
+  {\frac{1/d_{ij}} {\sum_{k=1}^n( 1/d_{ik})}} & \text{for $d_{ij} \leq 1.8$ miles}\\
+  0  & \text{for $d_{ij} > 1.8$ miles}
+  \end{cases}
+\end{equation}
 
@@
 dists.inv <- lapply(dists, function(x) 1 / x)
-W <- nb2listw(neighbours = tracts.dnn, glist = dists.inv, zero.policy = TRUE, style = "W")
+W <- nb2listw(neighbours = tracts.dnn, glist = dists.inv,
+              zero.policy = TRUE, style = "W")
 trMC <- trW(as(W, "CsparseMatrix"), type="MC") # trace used in montecarlo impacts
 @
+
+It may in some cases be more efficient to estimate a restricted spatial autoregressive
+model, that controls for only correlation or dependence. \citet{tra_autoregression_paper}
+illustrate a decision algorithm --- proposed by \citet{Florax2003} --- which uses
+a likelihood ratio test of the SDM against a restricted spatial error model (SEM):
+if the likelihood ratio test rejects that the models are equivalent, then the SDM
+should be used.
+
+A final point to note about spatial autoregressive models is that the partial
+derivative of the model with respect to a single regressor $\vect{x_k} \in X$
+does not equal the estimated coefficient $\beta_k$. Consequently, the estimated
+coefficients do not represent the {\em effect} of a variable on the outcome. The
+direct effects of an observation's attribute on its outcome, the indirect effects
+of neighbors' attributes on an outcome, and the total effect of own and neighbors'
+attributes can attained via Monte Carlo simulation. In this study we use the
+\texttt{spdep} package for R \citep{Bivand2013} for model estimation and impact
+simulation.
 

### 2018-12-12 manuscript.Rnw
Add data summary table

@@
 library(VGAM)
-knitr::opts_chunk$set(cache=TRUE)
+library(qwraps2)
+library(knitr)
+library(kableExtra)
+opts_chunk$set(cache=TRUE)
 
@@
 
+Note that this formulation is an extension of the gravity-based accessibility 
+statistics, in the same way that the gravity model is a specific case of 
+a destination-choice model \citep{Daly1982}. The extension
+
 \section{Methodology}
@@
 tracts <- tracts[complete_tracts_index, ]
+
+tracts@data <-  tbl_df(tracts@data) %>%
+  dplyr::transmute(
+    county_fips = substr(GEOID, 3, 5),
+    borough = case_when(
+      county_fips == "081" ~ "Queens",
+      county_fips == "047" ~ "Brooklyn",
+      county_fips == "061" ~ "Manhattan",
+      county_fips == "005" ~ "Bronx",
+      county_fips == "085" ~ "Staten Island",
+      TRUE ~ as.character(NA)
+    ),
+    obesity = OBESITY,
+    physact = 100 - Phys_Act,
+    log_obesity = log(OBESITY),
+    log_physact = log(100 - Phys_Act),
+    density = Pop_Density,
+    fulltime = FulltimeWork, college = CollegeDeg, 
+    single = Single_Percent,
+    black = PctBlack, asian = PctAsian, hispanic = PctHispanic,
+    other = PctNative + PctPacific,
+    Pct0to17, Pct18to29, Pct65plus, 
+    Income1, Income2, Income3, Income4, Income5, Income6, 
+    Income7, Income8, Income9, Income10
+  )
+
+tracts_sf <- st_as_sf(tracts)
 @
@@
 complete cases.
+
+\begin{table*}
+  \caption{Summary Statistics of Tract Data}
+  \label{tab:parkbenefits}
+<<data_summary, results="asis", message=FALSE, warning=FALSE, cache=FALSE>>=
+our_summary <- list(
+  "obesity"     = tab_summary(tracts$obesity),
+  "physact"     = tab_summary(tracts$physact),
+  "pop_density" = tab_summary(tracts$density),
+  "fulltime"    = tab_summary(tracts$fulltime),
+  "college"     = tab_summary(tracts$college)
+)
+
+summary_table(tracts@data, our_summary) %>%
+  kable("latex", booktabs = T, escape = FALSE) %>%
+  group_rows("Obese: percent of tract which is obese", 1, 4) %>%
+  group_rows("Physical Activity: Percent of tract reporting physical activity", 5, 8) %>%
+  group_rows("Population Density: individuals per square mile", 9, 12)  %>%
+  group_rows("Full Time: percent of adults who work full time", 13, 16) %>%
+  group_rows("College: percent of adults with a college degree", 17, 20)
+@
+\end{table*}
 
@@
 
-\subsection{Park Choice Logsum}
 In the absence of a park choice survey, we estimated likely values of the $\beta$
@@
 
-$$ A_i = log\sum_{j \in J}\exp(distance_{ij} \beta_d + size_j \beta_j + tweets_j \beta_t)$$
-We wrote a function to calculate this value for an arbitrary distance matrix, size
-and tweets vector, and coefficients.
+<<f_logsums>>=
+#' Calculate destination choice logsums from a distance matrix
+#' 
+#' @param d An $n\times p$ matrix with the distance from all tracts to
+#'   all parks
+#' @param sizes A p-length vector of park sizes
+#' @param tweets A p-length vector of tweets at parks
+#' @param betas A vector containing the size, distance, and tweet coefficients. 
+#'   If we submit two variables the tweets are ignored.
+#'   
+#' @return An n-length vector containing the weighted log-sum based
+#'   accessibility between a tract and all parks.
+#' @details If we have n tracts and p parks, distances needs to be a 
+#' 
+calculate_park_logsums <- function(d, sizes, tweets = NULL,
+                                   betas = c(-15,.00001, .001)){
+
+  # A is n x p
+  a <- betas[1] * d
+
+  # B is p x 1
+  b <- betas[2] * sizes
+
+  if(!is.null(tweets)) 
+    b <- b + betas[3] * tweets
+
+  # calculate observed utility by adding the weighted park-level attributes 
+  # to the columns of the matrix
+  # V is n x p, with b added by-column to each element in a
+  V <- sweep(a, 2, b, `+`)
+
+  # log-sum of exponentiated utility, Output is n-length vector
+  log(rowSums(exp(V)))
+
+}
+@
 
@@
 It may in some cases be more efficient to estimate a restricted spatial autoregressive
-model, that controls for only correlation or dependence. \citet{tra_autoregression_paper}
+model, that controls for only correlation or only dependence. \citet{tra_autoregression_paper}
 illustrate a decision algorithm --- proposed by \citet{Florax2003} --- which uses

### 2018-12-13 manuscript.Rnw
Various data cleaning and visualization
Add a map illustrating the logsum value. Inspecting this figure encouraged me to change the distance calculation. We previously had tract centroid to park centroid; we now use population-weighted tract centroid to edge of park. For cases where the population-weighted centroid is *in* a park, we set the minimum distance to 0.1 miles to avoid infinite logsums.

@@
 opts_chunk$set(cache=TRUE)
+library(leaflet)
 
@@
 
-\section{Methodology}
+\section{Empirical Application}
+
+\subsection{Data}
 
@@
 tracts <- geojson_read("data/nyc_tracts.geojson", what = "sp")
-tracts_sf <- st_read("data/nyc_tracts.geojson", quiet = TRUE) %>%
-  st_transform(4326)
 
 # identify complete cases
-complete_tracts_index <- tracts_sf %>%
-  st_set_geometry(NULL) %>%
+complete_tracts_index <- tracts@data %>%
   tbl_df() %>%
@@
   dplyr::transmute(
+    geoid = GEOID,
     county_fips = substr(GEOID, 3, 5),
@@
   ) 
-
-tracts_sf <- st_as_sf(tracts)
 @
 
+This study used data available to the public from a variety of agencies and
+data providers.
 We obtained data on aggregate health outcomes from the Centers for Disease
-Control and Prevention (CDC) for census tracts in the New York City area. After
-removing tracts with missing population information, we have Sexpr{nrow(tracts)}
-complete cases.
+Control and Prevention (CDC) for census tracts in five boroughs of New York City.
+After removing tracts with missing population information, we have
+Sexpr{nrow(tracts)} complete cases. Table \ref{tab:tractsdata} presents key
+descriptive statistics for these data.
 
@@
   \caption{Summary Statistics of Tract Data}
-  \label{tab:parkbenefits}
+  \label{tab:tractsdata}
 <<data_summary, results="asis", message=FALSE, warning=FALSE, cache=FALSE>>=
@@
 parks <- sf::st_read("data/nyc_parks.geojson", quiet = TRUE) %>%
-  mutate(
+  transmute(
+    size = Park_Acres,
     log_size = log(Park_Acres),
+    tweets = TWEET_COUNT, 
     log_tweets = yeo.johnson(TWEET_COUNT, 0)
@@
 
-In the absence of a park choice survey, we estimated likely values of the $\beta$
-coefficients by iteratively searching for the values which produced the highest
-model likelihood in *which model did we use to estimate likelihood?*. The search
-was constrained by two assertions. First, we required that the coefficient on
-size be positive and that on distance negative; all else equal, people will
-prefer to use larger parks that are nearer to their residence. Second, we visually
-inspected the resulting accessibility scores to ensure that the scores produced a
-reasonably varied pattern of access throughout the metropolitan region.
+\subsection{Model}
 
-<<f_logsums>>=
-#' Calculate destination choice logsums from a distance matrix
-#' 
-#' @param d An $n\times p$ matrix with the distance from all tracts to
-#'   all parks
-#' @param sizes A p-length vector of park sizes
-#' @param tweets A p-length vector of tweets at parks
-#' @param betas A vector containing the size, distance, and tweet coefficients. 
-#'   If we submit two variables the tweets are ignored.
-#'   
-#' @return An n-length vector containing the weighted log-sum based
-#'   accessibility between a tract and all parks.
-#' @details If we have n tracts and p parks, distances needs to be a 
-#' 
-calculate_park_logsums <- function(d, sizes, tweets = NULL,
-                                   betas = c(-15,.00001, .001)){
-
-  # A is n x p
-  a <- betas[1] * d
-
-  # B is p x 1
-  b <- betas[2] * sizes
-
-  if(!is.null(tweets)) 
-    b <- b + betas[3] * tweets
-
-  # calculate observed utility by adding the weighted park-level attributes 
-  # to the columns of the matrix
-  # V is n x p, with b added by-column to each element in a
-  V <- sweep(a, 2, b, `+`)
-
-  # log-sum of exponentiated utility, Output is n-length vector
-  log(rowSums(exp(V)))
-
-}
+We predict the obesity rate in a census tract as a function of the tract's
+sociodemographic characteristics and choice-based accessibility to parks with a
+semi-log specification:
+\begin{equation}\label{eq:themodel}
+\log(\mathit{Obesity})_i = f(\vect{X}_i, A_i, g(W, \mathit(Obesity), \vect{X}))
+\end{equation}
+<<base_formula>>=
+base_formula <- formula( 
+  ~ log(density) + 
+    fulltime + college + single + 
+    Pct0to17 + Pct18to29 + Pct65plus + # need to leave out a category for collinearity  
+    black + asian + hispanic + other)
 @
-
-\subsection{Spatial dependence and correlation}
-
+where $\vect{T}_i$ is a vector of tract-level attributes, $A_i$ is a choice-based
+accessibility measure for the tract, and $g(W, \mathit(Obesity), \vect{T})$
+represents the impact of spatial neighborhood effects on the outcome.
 Using tracts as the unit of observation in the model allows us to consider the
 availability of parks at a metropolitan level, but it also introduces the problem
-spatial correlation of unobserved errors and spatial dependence in the data
+of spatial correlation of unobserved errors and spatial dependence in the data
 generating process \citep{Cliff1970, Anselin1980}. Spatial correlation occurs
@@
 
-It may in some cases be more efficient to estimate a restricted spatial autoregressive
-model, that controls for only correlation or only dependence. \citet{tra_autoregression_paper}
-illustrate a decision algorithm --- proposed by \citet{Florax2003} --- which uses
+It may  be more efficient to estimate a restricted spatial autoregressive
+model -- that controls for only correlation or only dependence -- and therefore
+save degrees of freedom in the estimation.  \citet{tra_autoregression_paper}
+illustrate a decision algorithm originally proposed by \citet{Florax2003} which uses
 a likelihood ratio test of the SDM against a restricted spatial error model (SEM):
@@
 
+\subsection{Accessibility}
+
+Extant destination choice models used in practice typically use travel time in
+minutes as a travel impedance term, and employment by sector as a size term or
+attraction component. Typically such models handle recreational trips together with
+other non-work and non-school trips, and we therefore can find no previously
+estimated coefficients using size terms relevant to park acreage or amenities.
+Given this gap in the literature, we instead assert values for $\beta$ to generate
+the choice-based accessibility statistic used in this study.
+
+<<f_logsums>>=
+#' Calculate destination choice logsums from a distance matrix
+#'
+#' @param d An $n\times p$ matrix with the distance from all tracts to
+#'   all parks
+#' @param sizes A p-length vector of park sizes
+#' @param tweets A p-length vector of tweets at parks
+#' @param betas A vector containing the size, distance, and tweet coefficients.
+#'   If we submit two variables the tweets are ignored.
+#'
+#' @return An n-length vector containing the weighted log-sum based
+#'   accessibility between a tract and all parks.
+#' @details If we have n tracts and p parks, distances needs to be a
+#'
+calculate_park_logsums <- function(d, sizes, tweets = NULL,
+                                   betas = c(-0.4,0.2,0.001)){
+
+  # A is n x p
+  a <- betas[1] * d
+
+  # B is p x 1
+  b <- betas[2] * sizes
+
+  if(!is.null(tweets))
+    b <- b + betas[3] * tweets
+
+  # calculate observed utility by adding the weighted park-level attributes
+  # to the columns of the matrix
+  # V is n x p, with b added by-column to each element in a
+  V <- sweep(a, 2, b, `+`)
+
+  # log-sum of exponentiated utility, Output is n-length vector
+  log(rowSums(exp(V)))
+
+}
+@
+
+We initially estimated potential values of the $\beta$ coefficients by
+applying a limited-memory, box-constrained optimization algorithm maximizing the
+SDM model log-likehood. The constraint required that the coefficient on
+size be positive and that on distance negative; all else equal, we assert people
+will prefer to use larger parks and parks that are nearer to their residence. This
+resulted in $\beta_d = -0.42$  and $\beta_s = 0.000$. Given our expectation that
+park size is a valuable attribute, we subsequently adjusted this value to
+
+<<distances, warning=FALSE>>=
+# calculate centroids of tracts and parks
+pop_weighted_centroids <- read_csv("data/CenPop2010_Mean_TR36.csv") %>%
+  mutate(geoid = str_c(STATEFP, COUNTYFP, TRACTCE))
+pop_weighted_centroids <- left_join(tracts@data, pop_weighted_centroids, by = "geoid") %>%
+  dplyr::select(geoid, LONGITUDE, LATITUDE) %>%
+  st_as_sf(crs = 4326, coords = c("LONGITUDE", "LATITUDE")) %>%
+  st_transform(2263) 
+parks <- parks %>% st_transform(st_crs(pop_weighted_centroids))
+
+# distances from all tracts to all zones
+distances <- st_distance(pop_weighted_centroids, parks, byid = TRUE) %>%
+  units::set_units(miles) %>%
+  units::drop_units()
+
+# assert that a tract must be at least 1/10 mile from a park
+distances <- pmax(distances, 0.1)
+@
+
+<<logsum_betas>>=
+tracts@data <- tracts@data %>%
+  mutate(
+    access_ls = calculate_park_logsums(log(distances), parks$log_size,
+                                       betas = c(-0.42, 1.0)),
+    tweets_ls = calculate_park_logsums(log(distances), parks$log_size, parks$log_tweets,
+                                       betas = c(-0.42, 1.0)),
+    distance_ls = (distance_ls - mean(distance_ls) )/ sd(distance_ls),
+    access_ls = (access_ls - mean(access_ls) )/ sd(access_ls),
+    tweets_ls = (tweets_ls - mean(tweets_ls) )/ sd(tweets_ls)
+  )
+
+tracts_sf <- tracts %>% st_as_sf() %>% st_transform(4326)
+@
+
+<<leaflet, eval=FALSE>>=
+# this interactive map is much easier to create and explore
+pal <- colorQuantile("Spectral", tracts_sf$access_ls, n = 5)
+leaflet(tracts_sf) %>%
+  addProviderTiles(providers$CartoDB.Positron) %>%
+  addPolygons(group = "access", color = ~ pal(access_ls), label = ~as.character(round(access_ls,2))) 
+@
+
+\begin{figure*}
+
+<<map>>=
+map <- ggmap(get_map(unname(st_bbox(tracts_sf)), source = "stamen", maptype = "toner"),
+             extent = "device")
+map +
+  geom_sf(data = tracts_sf, inherit.aes=FALSE, 
+          aes(fill = cut(access_ls, quantile(access_ls))), alpha = 0.8) +
+  scale_fill_brewer("Relative Park Logsum", palette = "GnBu")
+@
+  \caption{Relative surplus-based accessibility in New York City}
+  \label{fig:map}
+\end{figure*}
+
 \section{Results}
+
+We estimate
+
+```{r spatial selection}
+obese_base_lm <- lm(update(base_formula, log_obesity ~ .), data = tracts@data)
+obese_base_sar <- lagsarlm(update(base_formula, log_obesity ~ .), 
+                           data = tracts@data, listw = W, zero.policy = TRUE)
+obese_base_sem <- errorsarlm(update(base_formula, log_obesity ~ .), 
+                             data = tracts@data, listw = W, zero.policy = TRUE)
+obese_base_sdm <- lagsarlm(update(base_formula, log_obesity ~ .), 
+                           data = tracts@data, listw = W, zero.policy = TRUE, 
+                           type = "mixed")
+```
+
+A likelihood ratio test reveals that the SEM is not preferred, so we use the SDM
+only going forward.
+
+```{r lrtest}
+lmtest::lrtest(obese_base_sem, obese_base_sdm)
+```
+
+```{r obesity_access_models}
+access_sdm <- update(obese_base_sdm, .~ . + access_ls)
+tweets_sdm <- update(obese_base_sdm, .~ . + tweets_ls)
+```
 

### 2018-12-13 manuscript.Rnw
Add estimated models to document

@@
 \begin{tabular}{lp{11cm}}
-  \hline
+\toprule
 Category & Summary of Benefits\\
-\hline
+\midrule
 Physical Health & Provide clean drinking water \citep{Benedict2006}\\
@@
  & Improve birth outcomes\\
-\hline
+\midrule
 Mental Health & Reduce stress and mental fatigue\\
@@
   & Improve behavioral outcomes in youth \\
-\hline
+\midrule
 Social Health & Enhanced community aesthetics\\
@@
  & Increase social interaction \\
-\hline
+\midrule
 Economic Health & Provide of ecosystem services\\
@@
  & Attract more shoppers and increase economic activity to commercial districts \\
-\hline
+\bottomrule
 \end{tabular}
@@
 <<weights>>=
-tracts.dnn <- dnearneigh(gCentroid(tracts, byid = TRUE), 0, 1.8 * 5280)
+pop_weighted_centroids <- read_csv("data/CenPop2010_Mean_TR36.csv") %>%
+  mutate(geoid = str_c(STATEFP, COUNTYFP, TRACTCE))
+pop_weighted_centroids <- left_join(tracts@data, pop_weighted_centroids, by = "geoid") %>%
+  dplyr::select(geoid, LONGITUDE, LATITUDE) %>%
+  st_as_sf(crs = 4326, coords = c("LONGITUDE", "LATITUDE")) %>%
+  st_transform(2263)
+
+tracts.dnn <- dnearneigh(gCentroid(as(pop_weighted_centroids, "Spatial"), byid = TRUE), 
+                         0, 1.8 * 5280)
 dists <- nbdists(tracts.dnn, gCentroid(tracts, byid = TRUE))
@@
 # calculate centroids of tracts and parks
-pop_weighted_centroids <- read_csv("data/CenPop2010_Mean_TR36.csv") %>%
-  mutate(geoid = str_c(STATEFP, COUNTYFP, TRACTCE))
-pop_weighted_centroids <- left_join(tracts@data, pop_weighted_centroids, by = "geoid") %>%
-  dplyr::select(geoid, LONGITUDE, LATITUDE) %>%
-  st_as_sf(crs = 4326, coords = c("LONGITUDE", "LATITUDE")) %>%
-  st_transform(2263) 
+
 parks <- parks %>% st_transform(st_crs(pop_weighted_centroids))
@@
     tweets_ls = calculate_park_logsums(log(distances), parks$log_size, parks$log_tweets,
-                                       betas = c(-0.42, 1.0)),
-    distance_ls = (distance_ls - mean(distance_ls) )/ sd(distance_ls),
+                                       betas = c(-0.42, 1.0, 0.1)),
     access_ls = (access_ls - mean(access_ls) )/ sd(access_ls),
@@
   addProviderTiles(providers$CartoDB.Positron) %>%
-  addPolygons(group = "access", color = ~ pal(access_ls), label = ~as.character(round(access_ls,2))) 
+  addPolygons(group = "access", color = ~ pal(access_ls), 
+              label = ~as.character(round(access_ls,2))) %>% 
+  addPolygons(group = "tweets", color = ~ pal(tweets_ls), 
+              label = ~as.character(round(tweets_ls,2))) %>%
+  addLayersControl(baseGroups = c("access", "tweets"))
 @
@@
 
-<<map>>=
+<<map, echo=FALSE, message=FALSE, warning=FALSE>>=
+library(ggmap)
 map <- ggmap(get_map(unname(st_bbox(tracts_sf)), source = "stamen", maptype = "toner"),
@@
 
-```{r spatial selection}
-obese_base_lm <- lm(update(base_formula, log_obesity ~ .), data = tracts@data)
-obese_base_sar <- lagsarlm(update(base_formula, log_obesity ~ .), 
-                           data = tracts@data, listw = W, zero.policy = TRUE)
+<<spatial selection>>=
 obese_base_sem <- errorsarlm(update(base_formula, log_obesity ~ .), 
@@
                            type = "mixed")
-```
+test_sdmsem <- lmtest::lrtest(obese_base_sem, obese_base_sdm)
+@
 
@@
 
-```{r lrtest}
-lmtest::lrtest(obese_base_sem, obese_base_sdm)
-```
-
-```{r obesity_access_models}
+<<access_sdm>>=
+# estimate models
 access_sdm <- update(obese_base_sdm, .~ . + access_ls)
 tweets_sdm <- update(obese_base_sdm, .~ . + tweets_ls)
-```
+@
+
+\begin{table*}
+  \caption{Estimated Coefficients on Obesity}
+  \label{fig:obeseity_models}
+<<modelchart, results="asis">>=
+library(texreg)
+obesity_models <- list(Base = obese_base_sdm, `Access` = access_sdm, 
+                       `Access + Tweets` = tweets_sdm)
+texreg(obesity_models, digits = 4, table = FALSE, single.row = TRUE, 
+       booktabs = TRUE, use.packages = FALSE)
+@
+\end{table*}
 

### 2019-01-02 manuscript.Rnw
Edit notation in methodology section

@@
 library(kableExtra)
-opts_chunk$set(cache=TRUE)
+opts_chunk$set(cache=TRUE, message = FALSE, echo = FALSE)
 library(leaflet)
@@
 
-\section{Choice-based Accessibility}
+\subsection{Choice-based Accessibility}
 
 Consider that an individual is choosing a park for a recreation activity.
-According to basic choice theory \citep{McFadden1974}, the probability of choosing
-park $p$ from the set of all regional parks $J$ is:
-
-$$ P(p | V_p) = \frac{\exp(V_p)}{\sum_{j \in J}\exp(V_j)}$$
+If we apply random utility choice theory, specifically the multinomial
+logit model \citep{McFadden1974}, the expected probability of individual 
+$i$ choosing park $j$ from the set of all regional parks $P$ is
+\begin{equation}\label{eq:mnl}
+P(j | V_{ij}) = \frac{\exp(V_{ij})}{\sum_{p \in P}\exp(V_{ip})}
+\end{equation}
 where parks are differentiated from each other by their relative measurable
-utilities $V$. In principle, $V$ may include any measurable attributes of either
+utilities $V_{ij}$. In principle, $V$ may include any measurable attributes of either
 the choice maker or the park. In this study we use a linear formulation of
-
-$$V_{ij} = size_j\beta_s + distance_{ij}\beta_d$$
+\begin{equation}\label{eq:v_sd}
+V_{ij} = size_j\beta_s + distance_{ij}\beta_d
+\end{equation}
 incorporating the size of the park in acres and the distance of the park from
-the census tract $i$ in miles. The coefficients $\beta$ are typically estimated from
-surveys, though in the absence of a survey we apply a maximum likelihood
-technique described below.
+the census tract $i$ in miles. The coefficients $\vec{\beta}$ can be 
+estimated from household surveys, though in the absence of a survey we 
+may assert reasonable values.
 
 A key theoretical understanding of random utility choice models is that the
-consumer surplus of the choice set can be otained as the log-sum of the
-denominator of the choice probability equation. In plainer terms, the *total value*
-of an individual's park accessibility is defined to be:
-
-$$CS_i = \ln\left({\sum_{j \in J}\exp(V_{ij})}\right)$$
-
+consumer surplus, or the total value of all alternatives of the choice set, can 
+be obtained as the log-sum of the denominator of the choice probability equation:
+\begin{equation}\label{eq:cs}
+CS_i = \ln\left({\sum_{p \in J}\exp(V_{ip})}\right)
+\end{equation}
 There are several advantages to a log-sum defined metric relative to
@@
 all individuals are defined as having some access to all parks, rather than an
-arbitrary limit of 1/2 mile or so. This allows for the fact that some people are
-more or less sensitive to distances, and that distance is a continuous, and not
-a binary, phenomenon. Second, the random utility formulation allows the
-researcher to include any attribute of the park; in this case, we consider the
+arbitrary cutoff asserted by the researcher. This allows for the fact that some 
+people are more or less sensitive to distances, and that distance is a 
+continuous, and not a binary, phenomenon. Second, the random utility 
+formulation allows the researcher to include any attribute of the park; 
+in this case, we consider the
 size of the park as an element of accessibility. This suggests that not all
-parks are equal, and that a large park such as New York City's Central Park may
+parks are equal, and that a large park such as Central Park on Manhattan may
 provide health and activity benefits over a much larger area than a smaller
@@
 statistics, in the same way that the gravity model is a specific case of
-a destination-choice model \citep{Daly1982}. The extension
+a destination-choice model \citep{Daly1982}. The extension allows us to consider
+more attributes of the park or the indivdual beyond size and impedance. For example,
+we can aggregate the number of tweets emanating from a park as a measure of the
+park's attractiveness beyond its size, and the utility for the choice maker from 
+Equation \ref{eq:v_sd} then becomes
+\begin{equation}\label{eq:v_sdt}
+V_{ij} = size_j\beta_s + distance_{ij}\beta_d + tweets_j\beta_t
+\end{equation}
 
@@
 \begin{equation}\label{eq:themodel}
-\log(\mathit{Obesity})_i = f(\vect{X}_i, A_i, g(W, \mathit(Obesity), \vect{X}))
+\log(\mathit{Obesity})_i = f(\vect{x}_i, CS_i, g(W, \mathit{Obesity}, \vect{X}))
 \end{equation}
@@
 @
-where $\vect{T}_i$ is a vector of tract-level attributes, $A_i$ is a choice-based
-accessibility measure for the tract, and $g(W, \mathit(Obesity), \vect{T})$
+where $\vect{x}_i$ is a vector of tract-level attributes, $CS_i$ is a choice-based
+accessibility measure for the tract computed from Equations \ref{eq:v_sd} or 
+\ref{eq:v_sdt}, and $g(W, \mathit{Obesity}, T)$
 represents the impact of spatial neighborhood effects on the outcome.
@@
 
-We estimate
-
 <<spatial selection>>=
@@
 
-A likelihood ratio test reveals that the SEM is not preferred, so we use the SDM
+We estimated the A likelihood ratio test rejects that the SEM is equivalent, so we use the SDM
 only going forward.

### 2019-01-04 manuscript.Rnw
Rewrite results to focus on physical activity
Physical activity is what we actually hope will happen with parks. Let's look at obesity as a secondary outcome.

@@
 library(kableExtra)
+library(texreg)
 opts_chunk$set(cache=TRUE, message = FALSE, echo = FALSE)
@@
 logit model \citep{McFadden1974}, the expected probability of individual 
-$i$ choosing park $j$ from the set of all regional parks $P$ is
+$i$ choosing park $j$ from the set of all regional parks $J$ is
 \begin{equation}\label{eq:mnl}
-P(j | V_{ij}) = \frac{\exp(V_{ij})}{\sum_{p \in P}\exp(V_{ip})}
+P(j | V_{ij}) = \frac{\exp(V_{ij})}{\sum_{p \in J}\exp(V_{ip})}
 \end{equation}
@@
 \begin{equation}\label{eq:v_sd}
-V_{ij} = size_j\beta_s + distance_{ij}\beta_d
+V_{ij} = size_j\lambda_s + distance_{ij}\lambda_d
 \end{equation}
 incorporating the size of the park in acres and the distance of the park from
-the census tract $i$ in miles. The coefficients $\vec{\beta}$ can be 
+the census tract $i$ in miles. The coefficients $\vect{\lambda}$ can be 
 estimated from household surveys, though in the absence of a survey we 
@@
 
-A key theoretical understanding of random utility choice models is that the
+A key theoretical outcome of random utility choice models is that the
 consumer surplus, or the total value of all alternatives of the choice set, can 
@@
 \begin{equation}\label{eq:v_sdt}
-V_{ij} = size_j\beta_s + distance_{ij}\beta_d + tweets_j\beta_t
+V_{ij} = size_j\lambda_s + distance_{ij}\lambda_d + tweets_j\lambda_t
 \end{equation}
 
-\section{Empirical Application}
+\section{Empirical Methodology}
 
@@
     log_tweets = yeo.johnson(TWEET_COUNT, 0)
-  )
+  ) %>% 
+  filter(size > 0.5)
 @
 
-We also collected data on park geometry.
+We also collected data on park geometry, limited to parks larger than 0.5 acres.
 
-\subsection{Model}
+\begin{table*}
+  \caption{Summary Statistics of Parks Data}
+  \label{tab:parksdata}
+<<parks_summary, results="asis", message=FALSE, warning=FALSE, eval = FALSE>>=
+park_summary <- list(
+  "size"     = tab_summary(parks$size),
+  "tweets"   = tab_summary(parks$tweets)
+)
 
-We predict the obesity rate in a census tract as a function of the tract's
-sociodemographic characteristics and choice-based accessibility to parks with a
-semi-log specification:
+summary_table(as_data_frame(parks), park_summary) %>%
+  kable("latex", booktabs = T, escape = FALSE, col.names = c("")) %>%
+  group_rows("Size: size of park in acres", 1, 4) %>%
+  group_rows("Tweets: number of tweets emanating from park", 5, 8)
+@
+\end{table*}
+
+\subsection{Model}\label{subsec:model}
+
+We predict the rate of physical activity ($y$) in a census tract as a 
+function of the tract's sociodemographic characteristics and choice-based 
+accessibility to parks with a semi-log specification:
 \begin{equation}\label{eq:themodel}
-\log(\mathit{Obesity})_i = f(\vect{x}_i, CS_i, g(W, \mathit{Obesity}, \vect{X}))
+\log(y_i) = f(\vect{x}_i, CS_i, g(W, \mathit{Obesity}, \vect{X}))
 \end{equation}
@@
 @
-where $\vect{x}_i$ is a vector of tract-level attributes, $CS_i$ is a choice-based
-accessibility measure for the tract computed from Equations \ref{eq:v_sd} or 
-\ref{eq:v_sdt}, and $g(W, \mathit{Obesity}, T)$
+where $\vect{x}_i$ is a vector of tract-level sociodemographic attributes, 
+$CS_i$ is a choice-based
+accessibility measure for the tract computed from Equation \ref{eq:cs} and 
+$g(W, \vect{y}, X)$
 represents the impact of spatial neighborhood effects on the outcome.
+
 Using tracts as the unit of observation in the model allows us to consider the
-availability of parks at a metropolitan level, but it also introduces the problem
+availability of parks at a sub-metropolitan level, but it also introduces the problem
 of spatial correlation of unobserved errors and spatial dependence in the data
@@
 
-\subsection{Accessibility}
+\subsection{Accessibility}\label{subsec:accessibility}
 
@@
 park size is a valuable attribute, we subsequently adjusted this value to
+a value of $1.00$. This implies that each mile
 
@@
 # this interactive map is much easier to create and explore
-pal <- colorQuantile("Spectral", tracts_sf$access_ls, n = 5)
+pal <- colorQuantile("GnBu", tracts_sf$access_ls, n = 5)
 leaflet(tracts_sf) %>%
@@
 library(ggmap)
-map <- ggmap(get_map(unname(st_bbox(tracts_sf)), source = "stamen", maptype = "toner"),
+map <- ggmap(get_map(unname(st_bbox(tracts_sf)), source = "stamen", maptype = "toner-lite"),
              extent = "device")
@@
   geom_sf(data = tracts_sf, inherit.aes=FALSE, 
-          aes(fill = cut(access_ls, quantile(access_ls))), alpha = 0.8) +
+          aes(fill = cut(access_ls, breaks = c(-Inf, -3, -2, -1, 0, 2, 3, Inf))),
+          alpha = 0.8) +
   scale_fill_brewer("Relative Park Logsum", palette = "GnBu")
@@
 <<spatial selection>>=
-obese_base_sem <- errorsarlm(update(base_formula, log_obesity ~ .), 
+physact_base_sem <- errorsarlm(update(base_formula, log_physact ~ .), 
                              data = tracts@data, listw = W, zero.policy = TRUE)
-obese_base_sdm <- lagsarlm(update(base_formula, log_obesity ~ .), 
+physact_base_sdm <- lagsarlm(update(base_formula, log_physact ~ .), 
                            data = tracts@data, listw = W, zero.policy = TRUE, 
                            type = "mixed")
-test_sdmsem <- lmtest::lrtest(obese_base_sem, obese_base_sdm)
+test_sdmsem <- lmtest::lrtest(physact_base_sem, physact_base_sdm)
 @
-
-We estimated the A likelihood ratio test rejects that the SEM is equivalent, so we use the SDM
-only going forward.
 
@@
 # estimate models
-access_sdm <- update(obese_base_sdm, .~ . + access_ls)
-tweets_sdm <- update(obese_base_sdm, .~ . + tweets_ls)
+pa_access_sdm <- update(physact_base_sdm, .~ . + access_ls)
+pa_tweets_sdm <- update(physact_base_sdm, .~ . + tweets_ls)
+@
+
+We estimated SDM and SEM models regressing the physical activity rate against
+against the base covariates (without the accessibility logsum). A likelihood 
+ratio test rejects that the SEM is equivalent 
+($\chi^2 = \Sexpr{test_sdmsem$Chisq[2]}$ on \Sexpr{test_sdmsem$Df[2]} degrees of 
+freedom), so we use the SDM specification only going forward. Table 
+\ref{tab:pa_models} presents the estimated coefficients for the base model as 
+well as models including logsum-based accessibility with and without Twitter
+activity. A few initial observations can be made from this table. First, adding
+the logsum measurement significantly improves the overall model fit, using a
+likelihood ratio test on the model log-likelihood statistics. Second, the
+coefficient estimates on the controlling variables remain virtually unchanged 
+from the base model.
+
+\begin{table*}
+  \caption{Estimated Coefficients Relating Park Access to Physical Activity}
+  \label{tab:pa_models}
+<<pamodelchart, results="asis">>=
+pa_models <- list(Base = physact_base_sdm, `Access` = pa_access_sdm, 
+                       `Access + Tweets` = pa_tweets_sdm)
+texreg(pa_models, digits = 4, table = FALSE, single.row = TRUE, 
+       booktabs = TRUE, use.packages = FALSE)
+#' screenreg(pa_models, digits = 4, table = FALSE) 
+@
+\end{table*}
+
+Recall from the presentation of spatial econometrics above that the total effect of a
+covariate on the dependent variable is different from the estimated coefficient
+and involves interaction between direct and indirect effects, or between 
+the $\beta$ and $\gamma$ parameters. This is particularly important when the
+$\beta$ and $\gamma$ estimates for a particular covariate show a different
+sign, such as in the proportion of black residents in a census tract, as 
+interpreting the direction and magnitude of the relationship is not straightforward.
+
+<<extract_impacts>>=
+#' Get the MC simulated impact coefficients and significance from a model
+#' @param sdm An autoregressive lag (SAR or SDM) model object `lagsarlm`
+#' @return A data frame with the effect (direct, indirect, total), variable, 
+#'   simulated mean impact, and significance p-value
+extract_impacts <- function(sdm) {
+
+  impacts_summary <- summary(
+    impacts(sdm, tr=trMC,  R = 1000, useHESS = !is.logical(sdm$fdHess)), 
+    zstats = TRUE)
+
+  estimate <- list(
+    Direct = impacts_summary$direct_sum,
+    Indirect = impacts_summary$indirect_sum,
+    Total = impacts_summary$total_sum
+  ) %>%
+  lapply(function(s) {
+    data_frame(
+      term = names(s$statistics[,1]),
+      estimate = s$statistics[,1],
+      std.error = s$statistics[,2]
+    )
+  }) %>%
+    bind_rows(.id = "effect")
+
+  pval <- as_data_frame(impacts_summary$pzmat) %>%
+    mutate(
+     term = rownames(impacts_summary$pzmat)
+    ) %>%
+    gather("effect", "p-val", -term)
+
+  left_join(estimate, pval, by = c("effect", "term"))
+}
 @
@@
 \begin{table*}
-  \caption{Estimated Coefficients on Obesity}
-  \label{fig:obeseity_models}
-<<modelchart, results="asis">>=
-library(texreg)
-obesity_models <- list(Base = obese_base_sdm, `Access` = access_sdm, 
-                       `Access + Tweets` = tweets_sdm)
-texreg(obesity_models, digits = 4, table = FALSE, single.row = TRUE, 
-       booktabs = TRUE, use.packages = FALSE)
+  \caption{Simulated Impact of Accessibility on Physical Activity}
+  \label{tab:pa_impacts}
+<<impacts>>=
+physact_impacts <- pa_models %>%
+  lapply(extract_impacts) %>%
+  bind_rows(.id = "model") %>% ungroup() %>%
+  rename(submodel = effect)
+
+effect_estimate <- physact_impacts %>% 
+  filter(model == "Access + Tweets", term == "tweets_ls", submodel == "Total") %>% 
+  pull(estimate)
+
+density_estimate <- physact_impacts %>% 
+  filter(model == "Access + Tweets", term == "log(density)", submodel == "Total") %>% 
+  pull(estimate)
+
+physact_impacts %>%
+  transmute(
+    model, term, submodel, estimate,
+    output = paste(round(`p-val`, 5), 
+                   gtools::stars.pval(`p-val`))
+  ) %>%
+  filter(term %in% c("access_ls", "tweets_ls")) %>%
+  select(-term) %>%
+  kable("latex", booktabs = T, escape = FALSE, 
+        col.names = c("Logsum Measure", "Effect", "Estimate", "$p$-value"))
 @
@@
 
-\section{Conclusion}
+Table \ref{tab:pa_impacts} shows the simulated marginal impact of the
+accessibility logsum measurements on the physical activity rate in each tract.
+Given that the logsum is a relatively continuous measure over space, and that it
+would be difficult to improve the park accessibility of a tract without also 
+improving the accessibility of the tract's neighbors, the total effect combining
+the direct effect and the indirect effects of all neighbors is 
+most appropriate. The total effect of tract-level accessiblity on physical 
+activity rates is strongly significant, regardless of whether we consider Twitter
+activity.
+
+It is of course necessary to also consider the magnitude of the effect. 
+As this is a semi-log model, the coefficient implies a unit increase
+in the access and tweets logsum leads to a 
+\Sexpr{round(effect_estimate * 100, 3)}\% increase in the physical activity rate. 
+Referring to the map of logsum measures in Figure \ref{fig:map}, the measures
+span from roughly $-3$ in heavily industrialized segments of southern Brooklyn
+to roughly $3$ in tracts immediately adjacent to Central Park on the Upper East
+Side. Thus moving from the least accessible parts of the city to the most
+accessible will reduce the expected physical activity rate by 
+\Sexpr{round(effect_estimate * 100 * 6, 3)}\%. For an aggregate statistic, this 
+is not an insubstantial marginal effect.
+
+\section{Discussion}
 

### 2019-01-05 manuscript.Rnw
Add literature review from Nico, revised and expand

@@
 slow development, including cardiovascular diseases, chronic respiratory
-diseases, diabetes, stroke, joint and bone diseases, and cancer. The \citet{WHO2014} estimates approximately 300,000
-premature deaths in the United States each year can be attributed to chronic diseases.
-\citet{Finkelstein2009} find an
-obese American spends nearly \$1,500 more per year in health care costs than an
-American of normal weight. As obesity and chronic disease
-have become rampant, it is no surprise that healthcare costs have risen to
-nearly one-fifth of the United States’ gross domestic product \citep{Harnik2011}.
+diseases, diabetes, stroke, joint and bone diseases, and cancer \citep{WHO2014}
+Recent statistics suggest that approximately 300,000 premature deaths each year
+can be attributed to chronic diseases \citep{WHO2014}.
+According to \citet{Finkelstein2009}, obese Americans spend approximately
+\$1,500 more in healthcare costs annually than Americans of
+average weight, totaling \$147 billion in direct medical expenses nationally.
 
@@
 effective strategy for reducing and managing obesity and many of the
-aforementioned chronic diseases \citep{Gordon2013}.
-\citeauthor{Wolf2008} notes that “more than 50 percent of U.S. adults do not get
-enough physical activity to provide health benefits; 24 percent are not active
-at all in their leisure time. Activity decreases with age and sufficient
-activity is less common among women than men, and among those with lower incomes
-and less education” \cite[22]{Wolf2008}. Similarly, the United States Department of
-Health and Human Services notes that less than fifty-percent of Americans meet
-established recommendations for moderate to vigorous physical activity, or MVPA
-(U.S. DHHS 2010). As the epidemiological transition from infectious to chronic
-diseases is now complete in the developed world, increasing physical activity
-has become a vital public health task in the 21st century.
+aforementioned chronic diseases \citet{CDC2009, Durstine2013},
+\citet{Wolf2008} notes that “more than 50 percent of U.S. adults do not get enough
+physical activity to provide health benefits; 24 percent are not active at all
+in their leisure time. Activity decreases with age and sufficient activity is
+less common among women than men, and among those with lower incomes
+and less education” (p. 22). As the epidemiological transition from
+infectious to chronic diseases is now complete in the developed world,
+increasing physical activity has become a vital public health task in the 21st
+century.
 
-Research has shown that the design of the built environment influences a range
-of behaviors, including those related to physical activity. The presence of
-trees and other vegetation in outdoor environments have been shown to be
-positively associated with physical activity \citep{Pretty2005}; one finding
-suggests that after sidewalks and trails have been constructed, the introduction
-of natural elements positively impacts motivation to engage in physical activity
-\citep{Suminski2005}. Additional evidence indicates that commonly vegetated
-areas, such as parks and open space, support outdoor physical activity
-\cite{Giles-Corti2005, Wells2007}. Perhaps most telling is the
-finding that “people in large cities perceive themselves to be generally
-healthier if a greater percentage of the living environment is greenspace, are
-inclined to be more active, and claim the ability to relax faster”
-\cite[24]{Wolf2008}. Thus, by providing space for active recreation, public
-parks and greenspaces may increase the likelihood of engaging in more physical
-activity. As such, public investment in parks can be thought of as a public
-health intervention for chronic diseases and conditions, as well as general
-population health.
-
-To date, the literature exploring the relationship between parks and health
-outcomes, specifically those related to physical activity and obesity, has
-yielded mixed results. Importantly, there is considerable variation in the
-design of past studies, including the spatial scale of analysis, the population
-of interest, and the measure of proximity and/or accessibility. While studies of
-neighborhood level health impacts do exist, they typically focus on discrete
-populations of interest within individual parks instead of examining the impacts
-of larger park networks \citep{Bancroft2015, Coutts2008}.
-Other studies have used the city and metropolitan statistical area as the
-spatial scale of analysis, which
-in some cases has resulted in the discovery of positive associations between
-increased park space and positive health impacts but may conceal more nuanced
-relationships that exist at smaller spatial scales \citep{Larson2016}.
-Evaluating the health impacts of parks and greenspaces using a smaller
-unit of analysis is an important research gap to fill, especially considering
-that park use and physical activity within parks varies considerably according
-to residential proximity to parks and park facilities as well as a number of
-sociodemographic factors \citep{Kaczynski2014}. There are important urban
-design considerations here; is it better to build a single large park with many
-different amenities, or to build a series of smaller parks nearer people’s
-homes. Similarly, do parks placed near low-income households affect the behavior
-of those households? This study attempts to fill that gap using New York City as
-a case study example. Using a new measure of park access called ‘Park Choice
-Accessibility,’ we find
-
-\section{Existing Literature}
-
-\subsection{Urbanization and Public Health Outcomes}
-Across the world, the population of cities is growing at an unprecedented rate.
-With a current estimated global urban population of over 3.5 billion people, the
-United Nations Population Fund estimates that by 2030, 5 billion people will
-inhabit cities worldwide.  By 2050, an additional 3 billion people will live in
-cities – a sixty percent increase over twenty years – with much of this growth
-expected to take place in the developing world (United Nations Population Fund,
-2016). Despite the greater expected concentration of urban growth in other parts
-of the globe, the trend of urbanization is highly visible in the United States
-as well.  According to the U.S. Census Bureau, ''The nation's urban population
-increased by 12.1 percent from 2000 to 2010, outpacing the nation's overall
-growth rate of 9.7 percent for the same period.'' \citet{Nowak2005}
- estimate that urban land area as a percentage of total land area in the
-U.S. will increase from 3.1\% in 2000 to approximately 8.1\% by the year 2050,
-with urbanized places collectively comprising a land area larger than the state
-of Montana. More than eighty percent of the U.S. population now lives in
-urban areas, compared to sixty-four percent in 1950 (United States Census Bureau
-2007).
-
-The global trend of urbanization has profound implications for population
-health, both positive and negative. Because urbanization corresponds with an
-increase in population density compared to rural and suburban settlement
-patterns, residents of dense urban areas throughout the world have better access
-on average to many health services and health-promoting amenities simply because
-of their closer proximity to such resources \citep{Larson2016}.
-However, despite this benefit of increased density and the ability of cities to
-provide many opportunities for innovation, economic growth and social progress,
-historical and current evidence suggests that rapid urban growth often leads to
-congestion and numerous negative environmental and human health outcomes.
-Interactions between growing urban populations and their environment, marked by
-intensive resource consumption and ecologically harmful patterns of development,
-lead to numerous undesired consequences including pollution and sanitation
-issues as well as racial and socioeconomic disparities \citep{Larson2016}.
-
-Because of the considerable growth of urbanization throughout the world, many
-city planning and public health professionals have begun to pay more attention
-to the role of the built environment in promoting or discouraging health
-behaviors. Until recently, most large-scale health promotion efforts focused on
-individual-level interventions intended to educate people about health
+Until recently, most large-scale health promotion efforts focused on
+individual-level interventions intended to educate people about healthy
 lifestyles and behaviors, touching on topics including diet and exercise.
 According to \citet{Coutts2008} however, this trend is shifting as professionals
-''have begun adopting an ecological paradigm, accepting that both individual and
+“have begun adopting an ecological paradigm, accepting that both individual and
 environmental determinants play a role in health behavior. This new, arguably
@@
 you are) but also for the contextual (where you are) influences on physical
-activity.'' As professionals begin to operate from the assumption that the design
+activity.” As professionals begin to operate from the assumption that the design
 and configuration of the built environment can facilitate or inhibit physical
@@
 greenways as key elements of the built environment that can support exercise
-\citet{Coutts2008, Bedimo-Rung2005}.
+\citep{Bedimo-Rung2005a}.
 
-\subsection{Review of the Health Benefits of Parks}
-A large body of literature exists documenting the many benefits of urban parks
-and greenspaces to human health and wellbeing. Importantly, these benefits stem
-from the provision of ecosystem services, which occur when the natural
-environment supplies something that people demand, improving quality of life and
-well-being \citep{Larson2016}. These services can include the
-provisioning of goods such as fresh water and agricultural products; regulatory
-functions including protection of drinking water quality, heat mitigation, air
-purification, and stormwater management; and cultural functions, such as
-improving aesthetics, providing opportunities for recreation, tourism, and
-physical and mental health, and promoting biodiversity \citep{McDonald2015}.
-According to \citet{Wolch2014}, parks and greenspaces provide ''\ldots a wide
-range of ecosystem services that could help combat many urban ills and improve
-life for city dwellers—especially their health…Ecosystem services provided by
-urban greenspace not only support the ecological integrity of cities, but can
-also protect the public health of urban populations.'' Table
-\ref{tab:parkbenefits} summarizes commonly cited benefits of parks and
-greenspaces.
+In this paper, we present a holistic and flexible measurement for park
+accessibility based in random utility theory. This measure can consider the continuous
+distance to all parks in the region, weighted by the size of the park and other
+amenities. We apply this measurement to study the link between park
+accessibility and attractiveness and Census tract-level aggregate physical
+activity participation and obesity rates in New York City, controlling for
+spatial correlation and dependence. We find a positive relationship where the
+least park-accessible tracts have an expected physical activity participation
+rate 3.8 percent lower than the most accessible tracts, controlling for
+socioeconomic characteristics. We also find that increased physical activity
+rates lead to reduced aggregate obesity rates, but that when controlling for
+physical activity rates, accessibility to parks does not reduce overall accessibility.
+This suggests that to be an effective tool for addressing chronic diseases,
+communities must do more than simply provide parks, but also incentivize more
+citizens to use them.
+
+\section{Literature Review}\label{sec:litreview}
+
+Numerous previous studies have found relationships between aspects of the built
+environment and aspects of physical, mental, social, and economic health. Many
+of these findings are summarized in Table \ref{tab:parkbenefits}.
+For the purposes of this study, it is clear that parks and other
+vegetated areas support outdoor physical activity in ways distinct from other
+urban environments \citep{Giles-Corti2005, Wells2007}.
 
@@
 Physical Health & Provide clean drinking water \citep{Benedict2006}\\
- & Foster increased physical activity\\
  & Promote faster healing in hospitals \citep{Akbari2001}\\
- & Reduce of heat-related mortality \\
- & Reduce incidence of cardiovascular-related mortality \\
+ & Reduce heat-related mortality \citep{Stone2006}\\
+ & Reduce incidence of cardiovascular-related mortality \citep{Mitchell2008}\\
  & Improved air quality and related reductions in respiratory-related mortality \citep{Lovasi2008}\\
- & Reduce incidence of childhood asthma \citep{Wolf2008}\\
- & Improve birth outcomes\\
 \midrule
-Mental Health & Reduce stress and mental fatigue\\
-  & Reduce aggression\\
+Mental Health & Reduce stress and mental fatigue \citep{Wolch2014}\\
+  & Reduce aggression \citep{Kuo2001a}\\
   & Enhance emotional and cognitive development\\
@@
 Social Health & Enhanced community aesthetics\\
- & Reduce crime\\
- & Increase social interaction \\
+ & Reduce crime \citep{Kuo2013,Wolfe2012}\\
+ & Increase social interaction \citep{Sullivan2004}\\
 \midrule
-Economic Health & Provide of ecosystem services\\
+Economic Health & Provide ecosystem services\\
  & Increase residential property values and municipal property tax revenues\\
@@
 
-Given the general benefits of parks and greenspaces in Table \ref{tab:parkbenefits}, several studies
-have attempted to quantify the impacts of these spaces on different facets of
-health and wellbeing at the city level, yielding mixed results. \citet{Larson2016}
+In spite of these generally positive findings, the evidence of a specific link
+between green space and physical activity is somewhat mixed.
+\citet{Wolf2008} report
+residents self-identify as more active and attaining a higher quality of life when
+a greater portion of the environment is greenspace.
+\citet{Larson2016}
 used self-reported scores on the Gallup-Healthways Wellbeing
-Index to evaluate the relationship between different areas of wellbeing,
+Index (citation) to evaluate the relationship between different areas of wellbeing,
 including physical health, and park quantity, quality, and accessibility in 44
-U.S. cities. The authors found that “Park quantity (measured as the percentage
-of city area covered by public parks) was among the strongest predictors of
-overall wellbeing, and the strength of this relationship appeared to be driven
-by parks’ contributions to physical and community wellbeing” (Larson, Jennings \&
-Cloutier 2016). While the authors found positive associations between wellbeing
+U.S. cities. While the authors found positive associations between wellbeing
 and park quality and accessibility, these relationships were not statistically
@@
 activity and a significant, negative association between park density and
-obesity. In a study of New York City, \citet{Stark2014} found that the
-“proportion of neighborhoods that was large or small park space and park
-cleanliness were associated with lower BMI among NYC adults after adjusting for
-other neighborhood features such as homicides and walkability, characteristics
-that could influence park usage.”
-
-Interestingly, in a study by \citet{Richardson2012} that examined the
-relationship between urban greenspace and selected mortality rates, the authors
+obesity.
+Conversely, \citet{Richardson2012} examined the
+relationship between urban greenspace and selected mortality rates, and though the authors
 did not find a statistically significant relationship between the quantity of
 urban greenspace and mortality caused by lung cancer, diabetes, heart disease or
-car accidents, but found that all-cause mortality was significantly higher in
-greener cities. In a meta-analysis of 20 peer reviewed journal articles
+car accidents, they did find that all-cause mortality was significantly higher in
+greener cities.
+In a meta-analysis of 20 peer reviewed journal articles
 exploring the relationship between parks and objectively measured physical
@@
 increased physical activity and attendant decreases in obesity, all of the
-studies  -- except some that were included in the meta-analysis by
-\citet{Bancroft2015} -– were conducted at the level of the city or metropolitan statistical
-area, potentially concealing variation in health outcomes at more fine-grained
-levels of analysis \citep{Larson2016}.
+studies -- except some included in the meta-analysis by \citet{Bancroft2015} -–
+were conducted at the level of the city or metropolitan statistical area.
+Metropolitan-level analyses do not capture the within-city variation in
+accessibility to parks that exists. Some cities with large amounts of total
+greenspace may nevertheless inequally distribute this space throughout the city
+leading to areas with poor access. Conversely, a city with a smaller overall
+proportion of greenspace may give all of its citizens better access to significant
+greenspace. Considering access at the neighborhood level within a single city
+may eliminate some regional or cultural fixed effects affecting
+metropolitan-level analyses.
 
-\subsection{Accessibility to Parks}
-To evaluate the impacts of parks and greenspaces on health outcomes at any
-scale, one of the fundamental tasks involved is the selection of a measure of
-accessibility. In a comprehensive meta-analysis of published studies that
-measure active accessibility (accessibility using non-motorized travel modes
-including walking and cycling), \citet{Vale2016} identified four
-broad categories of studies based on the active accessibility metric employed:
-“distance-based, gravity-based or potential, topological or
-infrastructure-based, and walkability and walk score-type measures.”
-Distance-based measures account only for the Euclidean distance between origins
-and destinations, while infrastructure-based measures explicitly incorporate
+In a study of neighborhoods in New York City, \citet{Stark2014} found that a
+higher proportion of park space in a neighborhood was associated with a lower
+BMI for its residents. \citet{Pretty2005} showed the presence of trees and other
+vegetation in outdoor environments is positively associated with physical
+activity; further, \citet{Suminski2005} showed that adding vegetation to
+existing sidewalks and trails increased local motivation to engage in physical
+activity.
+
+To evaluate the impacts of parks and greenspaces on health outcomes at a
+sub-metropolitan level, one of the fundamental tasks involved is the selection
+of a measure of accessibility. In a comprehensive meta-analysis of published
+studies that measure active accessibility (accessibility using non-motorized
+travel modes including walking and cycling), \citet{Vale2016} identified four
+broad categories of studies based on the metric employed:
+\begin{description}
+  \item[Distance-based]{Account only for the Euclidean distance between origins
+  and destinations}
+  \item[Infrastructure-based]{Explicitly incorporate
   relevant transportation networks like roads and sidewalks to more accurately
-measure travel time and distance. Gravity-based measures incorporate cost
-measures to model accessibility as a function of a destination’s attractiveness
-(i.e., size, commercial activity, etc.) and the cost of traveling to that
-destination from a given origin (i.e., travel time or distance).
+  measure travel time and distance}
+  \item[Gravity-based]{Incorporate travel impedance
+  measures to model accessibility as a function of a destination’s attractiveness}
+  \item[Walk-score]{Similar to Gravity-based, with the impedance measure
+  explicitly designed around walking}
+\end{description}
 
@@
 setting, noting that ''ways to measure active accessibility are as varied as the
-number of scholars the measure them.'' However, their meta-analysis
-highlights some of the limitations associated with each of the types of metrics
-described. For example, in describing distance-based accessibility measures, the
-authors note that such measures are ''extremely sensitive to the way in which
-travel impedance (i.e., distance) is measured. Accordingly, four types of
-distance can be identified: Euclidean distance, Manhattan distance, shortest
-network distance, and shortest network time''. The
-appropriateness of one distance-based measure over another can vary
-significantly depending on the topography of the environment and the travel mode
-that is being employed. In describing gravity-based accessibility measures,
-which ''assume that travel is a derived demand and there is a tradeoff between
-the benefit of the opportunity and the cost to reach it from a given origin,''
-the authors note that such measures do not always explicitly account for land
-use characteristics near origins and destinations, which may impact that true
-accessibility of those places.
-
+number of scholars the measure them.'' For example, infrastructure-based distances
+can vary widely depending on which modes of travel infrastructure are considered.
+Similarly, gravity-based measures may not always account for surrounding land
+uses when considering the attractiveness or impedance of a destination.
 In evaluating methodologies of measuring access to urban services, including
 parks and greenspaces, \citet{Logan2017} note that existing and commonly used
-approaches “often simplify their measure of proximity by using large areal
-units and by imposing arbitrary distance thresholds,” which often results in
-access-poor populations being overlooked. The authors concede that many
-existing approaches have long been necessary because the computational power
-required to use higher-resolution analytical techniques was unavailable.
-However, due to recent advances in computation power and the advent of
-municipal open-data policies, \citeauthor{Logan2017} recommend that future
-analyses of accessibility disaggregate population data to the building or
-parcel level and use network distance instead of Euclidean distance to measure
-proximity.
+approaches "often simplify their measure of proximity by using large areal
+units and by imposing arbitrary distance thresholds," which often results in
+access-poor populations being overlooked.
 
@@
 consumer surplus, or the total value of all alternatives of the choice set, can
-be obtained as the log-sum of the denominator of the choice probability equation:
+be obtained as the log-sum of the denominator of the choice probability equation
+\citep{Small1981}:
 \begin{equation}\label{eq:cs}
@@
 people are more or less sensitive to distances, and that distance is a
-continuous, and not a binary, phenomenon. Second, the random utility 
-formulation allows the researcher to include any attribute of the park; 
-in this case, we consider the
-size of the park as an element of accessibility. This suggests that not all
-parks are equal, and that a large park such as Central Park on Manhattan may
-provide health and activity benefits over a much larger area than a smaller
+continuous, and not a binary, phenomenon. Second, the random utility formulation
+allows the researcher to include any attribute of the park; in this case, we
+consider the size of the park as an element of accessibility. This suggests that
+not all parks are equal, and that a large park such as Central Park on Manhattan
+may provide health and activity benefits over a much larger area than a smaller
 community square.
@@
 statistics, in the same way that the gravity model is a specific case of
-a destination-choice model \citep{Daly1982}. The extension allows us to consider
-more attributes of the park or the indivdual beyond size and impedance. For example,
-we can aggregate the number of tweets emanating from a park as a measure of the
-park's attractiveness beyond its size, and the utility for the choice maker from 
-Equation \ref{eq:v_sd} then becomes
+a destination-choice model \citep{Daly1982}. But the extension is meaningful, and
+allows us to consider more attributes of the park or the indivdual beyond size
+and impedance. For example, we can aggregate the number of tweets emanating from
+a park as a measure of the park's attractiveness beyond its size, and the
+utility for the choice maker from Equation \ref{eq:v_sd} then becomes
 \begin{equation}\label{eq:v_sdt}
@@
 
+In spite of its advantages, logsum-based accessibilities have not received
+as much application in the accessibility literature as distance-based or even
+gravity-based measures. They are commonly used, however, in alternatives analyses of
+transit infrastructure improvements \citep{DeJong2007}. A reason for this is
+likely that such analyses are usually conducted in the context of a travel
+demand model, where calibrated and multimodal logsums are readily available
+\citep{Geurs2010}.
+
 \section{Empirical Methodology}
+In this section we describe an experiment with data for New York City, where
+we compute a holistic logsum-based accessibility component and model the
+relationship between this measure and physical activity rates, controlling for
+spatial effects and socioeconomic factors.
 
 \subsection{Data}
+This study uses data available to the public from a variety of agencies and
+data providers.
 
@@
 
-This study used data available to the public from a variety of agencies and
-data providers.
-We obtained data on aggregate health outcomes from the Centers for Disease
-Control and Prevention (CDC) for census tracts in five boroughs of New York City.
-After removing tracts with missing population information, we have
-Sexpr{nrow(tracts)} complete cases. Table \ref{tab:tractsdata} presents key
-descriptive statistics for these data.
+The primary dataset is a geographic polygons shapefile of Census tracts in
+New York City. We appended relevant sociodemographic data for each tract level
+from the American Community Survey 2011-2015 5-year estimates. Data on physical
+activity participation rate and obesity rates for each tract are available from
+the Centers for Disease Control and Prevention’s 500 Cities Project.
+After removing tracts with missing population information or outlying and unusual
+characteristics, we have Sexpr{nrow(tracts)} complete cases. Table
+\ref{tab:tractsdata} presents key descriptive statistics for these data.
 
@@
   group_rows("Obese: percent of tract which is obese", 1, 4) %>%
-  group_rows("Physical Activity: Percent of tract reporting physical activity", 5, 8) %>%
+  group_rows("Physical Activity: percent of tract reporting physical activity", 5, 8) %>%
   group_rows("Population Density: individuals per square mile", 9, 12)  %>%
@@
 \end{table*}
+
+In a destination choice framework, the tracts represent the "origins" and
+"destinations" are parks in New York City.
+We retrieved a shapefile of public parks and greenspaces within New York City’s
+municipal boundaries and checked it for accuracy and relevance. Upon inspection,
+we removed several facilities that do not qualify as publicly accessible green
+space, such as Yankee Stadium and its surrounding parking lots. We also removed
+parks of less than half an acre in size, as these appear to be predominantly
+vacant lots or parking strips rather than legitimate public green space.
 
@@
 
-We also collected data on park geometry, limited to parks larger than 0.5 acres.
-
 \begin{table*}
@@
   \label{tab:parksdata}
-<<parks_summary, results="asis", message=FALSE, warning=FALSE, eval = FALSE>>=
+<<parks_summary, results="asis", message=FALSE, warning=FALSE, cache = FALSE>>=
 park_summary <- list(
@@
 summary_table(as_data_frame(parks), park_summary) %>%
-  kable("latex", booktabs = T, escape = FALSE, col.names = c("")) %>%
-  group_rows("Size: size of park in acres", 1, 4) %>%
-  group_rows("Tweets: number of tweets emanating from park", 5, 8)
+  kable("latex", booktabs = T, escape = FALSE, col.names = c("Parks"))
 @
@@
 
+<<obesity_models>>=
+obese_base_sdm <- lagsarlm(update(base_formula, log_obesity ~ . + log_physact),
+                           data = tracts@data, listw = W, zero.policy = TRUE,
+                           type = "mixed")
+obese_access_sdm <- update(obese_base_sdm, . ~ . + access_ls + access_ls * log_physact)
+obese_tweets_sdm <- update(obese_base_sdm, . ~ . + tweets_ls + access_ls * log_physact)
+obesity_models <- list("Obesity" = obese_base_sdm, "Obesity Access" = obese_access_sdm, 
+                       "Obesity Tweets + Access" = obese_tweets_sdm)
+@
+
+<<obese_impacts>>=
+obese_impacts <- obesity_models %>%
+  lapply(extract_impacts) %>%
+  bind_rows(.id = "model") %>% ungroup() %>%
+  rename(submodel = effect)
+
+dwplot(bind_rows(physact_impacts, obese_impacts) %>% 
+                 filter(term %in% c("log_physact", "access_ls", "tweets_ls")) %>%
+                 filter(submodel == "Total")) +
+  theme_bw()
+
+obese_impacts %>%
+  transmute(
+    model, term, submodel, estimate,
+    output = paste(round(`p-val`, 5),
+                   gtools::stars.pval(`p-val`))
+  ) %>%
+  filter(term %in% c("log_physact", "access_ls", "tweets_ls"), submodel == "Total")
+@
+
 \section{Discussion}

### 2019-01-07 manuscript.Rnw
Restructure data tables in paper

@@
 
+knit_hooks$set(inline = function(x) {
+  prettyNum(x, big.mark=",", drop0trailing = TRUE)
+})
 @
@@
 
-\begin{table*}
-  \caption{Summary Statistics of Tract Data}
-  \label{tab:tractsdata}
-<<data_summary, results="asis", message=FALSE, warning=FALSE, cache=FALSE>>=
-our_summary <- list(
-  "obesity"     = tab_summary(tracts$obesity),
-  "physact"     = tab_summary(tracts$physact),
-  "pop_density" = tab_summary(tracts$density),
-  "fulltime"    = tab_summary(tracts$fulltime),
-  "college"     = tab_summary(tracts$college)
-)
-
-summary_table(tracts@data, our_summary) %>%
-  kable("latex", booktabs = T, escape = FALSE) %>%
-  group_rows("Obese: percent of tract which is obese", 1, 4) %>%
-  group_rows("Physical Activity: percent of tract reporting physical activity", 5, 8) %>%
-  group_rows("Population Density: individuals per square mile", 9, 12)  %>%
-  group_rows("Full Time: percent of adults who work full time", 13, 16) %>%
-  group_rows("College: percent of adults with a college degree", 17, 20)
-@
-\end{table*}
-
-In a destination choice framework, the tracts represent the "origins" and
-"destinations" are parks in New York City.
-We retrieved a shapefile of public parks and greenspaces within New York City’s
-municipal boundaries and checked it for accuracy and relevance. Upon inspection,
-we removed several facilities that do not qualify as publicly accessible green
-space, such as Yankee Stadium and its surrounding parking lots. We also removed
-parks of less than half an acre in size, as these appear to be predominantly
-vacant lots or parking strips rather than legitimate public green space.
-
 <<parks>>=
@@
 
-\begin{table*}
-  \caption{Summary Statistics of Parks Data}
-  \label{tab:parksdata}
-<<parks_summary, results="asis", message=FALSE, warning=FALSE, cache = FALSE>>=
-park_summary <- list(
-  "size"     = tab_summary(parks$size),
-  "tweets"   = tab_summary(parks$tweets)
-)
+In a destination choice framework, the tracts represent the "origins" and
+"destinations" are parks in New York City.
+We retrieved a shapefile of public parks and greenspaces within New York City’s
+municipal boundaries and checked it for accuracy and relevance. Upon inspection,
+we removed several facilities that do not qualify as publicly accessible green
+space, such as Yankee Stadium and its surrounding parking lots. We also removed
+parks of less than half an acre in size, as these appear to be predominantly
+vacant lots or parking strips rather than legitimate public green space. This 
+leaves us with \Sexpr{nrow(parks)} separate parks.
 
-summary_table(as_data_frame(parks), park_summary) %>%
-  kable("latex", booktabs = T, escape = FALSE, col.names = c("Parks"))
+<<tractsdata, results="asis", message=FALSE, warning=FALSE, cache=FALSE>>=
+tracts_data_index <- data_frame(
+  variable =  c(
+    "obesity", "physact", "density", "fulltime", "college", "single", 
+    "Pct0to17", "Pct18to29",  "Pct65plus", "black", "asian", "hispanic", 
+    "other")
+  ) %>% mutate(order = row_number())
+
+tract_summary <- tracts@data %>%
+  dplyr::select(obesity:Pct65plus) %>%
+  gather(variable, value) %>%
+  group_by(variable) %>%
+  summarise(
+    min = min(value),
+    `msd` = paste(median_iqr(value)),
+    max = max(value)
+  ) %>%
+  inner_join(tracts_data_index) %>%
+  arrange(order) %>%
+  mutate(
+    variable = c(
+      "Obesity Rate", "Physical Activity Rate",
+      "Density: population per square mile",
+      "Full-time workers", "College-educated", 
+      "Single Adults", "Share of children (0-17)", "Share of young adults (18-29)", 
+      "Share of seniors (65+)", "Black population share", "Asian population share", 
+      "Hispanic population share", "Other Minorities")
+  ) %>%
+  dplyr::select(-order)
+
+park_summary <- parks %>%
+  st_set_geometry(NULL) %>%
+  dplyr::select(size, tweets) %>%
+  gather(variable, value) %>%
+  group_by(variable) %>%
+  summarise(
+    min = min(value),
+    `msd` = paste(median_iqr(value)),
+    max = max(value)
+  ) %>%
+  mutate( variable = c("Park Size in Acres", "Tweets Emanating from Park"))
+
+bind_rows(tract_summary, park_summary) %>%
+  kable(escape = FALSE, digits = 2, 
+        table.envir = "table*", booktabs = TRUE,
+        caption = "Descriptive Statistics of Tract and Park Variables",
+        col.names = c(" ", "Minimum", "Median (IQR)", "Maximum")) %>%
+  kable_styling() %>%
+  group_rows(paste("Tract Variables, N =", nrow(tracts@data)), 1, 13) %>%
+  group_rows(paste("Park Variables, N =", nrow(parks)), 14, 15)
+
 @
-\end{table*}
 
@@
     black + asian + hispanic + other)
+
+termnames <- c("log(density)" , "Full-time workers", "College-educated", 
+               "Single Adults", "Share of children (0-17)", "Share of young adults (18-29)", 
+               "Share of seniors (65+)", "Black population share", "Asian population share", 
+               "Hispanic population share", "Other Minorities")
 @
@@
 
-<<weights>>=
+<<weights, warning = FALSE>>=
 pop_weighted_centroids <- read_csv("data/CenPop2010_Mean_TR36.csv") %>%
@@
 texreg(pa_models, digits = 4, table = FALSE, single.row = TRUE,
-       booktabs = TRUE, use.packages = FALSE)
+       booktabs = TRUE, use.packages = FALSE, 
+       custom.coef.names = c("Intercept", termnames, paste("lag", termnames), "$\\rho$",
+                             "Access logsum", "lag Access logsum", 
+                             "Access + Tweets logsum", "lag Access + Tweets logsum"))
 #' screenreg(pa_models, digits = 4, table = FALSE)
@@
   \label{tab:pa_impacts}
-<<impacts>>=
+<<impacts, results="asis">>=
 physact_impacts <- pa_models %>%
@@
   ) %>%
-  filter(term %in% c("access_ls", "tweets_ls")) %>%
-  select(-term) %>%
+  filter(term %in% c("access_ls", "tweets_ls"), submodel == "Total") %>%
+  dplyr::select(-term, -submodel) %>%
   kable("latex", booktabs = T, escape = FALSE,
-        col.names = c("Logsum Measure", "Effect", "Estimate", "$p$-value"))
+        col.names = c("Logsum Measure", "Estimate", "$p$-value")) %>%
+  footnote("$^{***}p<0.001$, $^{**}p<0.01$, $^*p<0.05$", escape = F)
 @
@@
 accessible will reduce the expected physical activity rate by
-\Sexpr{round(effect_estimate * 100 * 6, 3)}\%. For an aggregate statistic, this
-is not an insubstantial marginal effect.
+\Sexpr{round(effect_estimate * 100 * 6, 3)}\%. In other words, if a tract with 
+poor accessibility is expected to achieve the sample mean of 
+\Sexpr{round(mean(tracts$physact), 1)}\%, the same tract with excellent accessibility
+is expected to achieve a rate of 
+\Sexpr{round(mean(tracts$physact) * (1 + effect_estimate  * 6), 2)}\%. 
+For an aggregate statistic strongly influenced by other socioeconomic factors, 
+this is not an insubstantial marginal effect.
 
 <<obesity_models>>=
-obese_base_sdm <- lagsarlm(update(base_formula, log_obesity ~ . + log_physact),
-                           data = tracts@data, listw = W, zero.policy = TRUE,
-                           type = "mixed")
-obese_access_sdm <- update(obese_base_sdm, . ~ . + access_ls + access_ls * log_physact)
-obese_tweets_sdm <- update(obese_base_sdm, . ~ . + tweets_ls + access_ls * log_physact)
-obesity_models <- list("Obesity" = obese_base_sdm, "Obesity Access" = obese_access_sdm, 
-                       "Obesity Tweets + Access" = obese_tweets_sdm)
+obese_tweets_sdm <- update(pa_tweets_sdm, log_obesity ~ . + log_physact)
 @
 
-<<obese_impacts>>=
-obese_impacts <- obesity_models %>%
-  lapply(extract_impacts) %>%
-  bind_rows(.id = "model") %>% ungroup() %>%
-  rename(submodel = effect)
+We now consider the impacts of a model where the dependent variable is 
+the {\em obesity} rate, and physical activity becomes an independent covariate
+alongside the controlling variables and the tweets-augmented logsum variable. 
+Table \ref{tab:obese_impacts} presents the simulated impacts resulting from
+this model.
 
-dwplot(bind_rows(physact_impacts, obese_impacts) %>% 
-                 filter(term %in% c("log_physact", "access_ls", "tweets_ls")) %>%
-                 filter(submodel == "Total")) +
-  theme_bw()
-
-obese_impacts %>%
+\begin{table*}
+  \caption{Simulated Impacts of Variables on Obesity}
+  \label{tab:obese_impacts}
+<<tab:obese_impacts, results="asis">>=
+extract_impacts(obese_tweets_sdm) %>%
+  group_by(effect) %>%
   transmute(
-    model, term, submodel, estimate,
-    output = paste(round(`p-val`, 5),
-                   gtools::stars.pval(`p-val`))
+    term,
+    output = paste(round(`estimate`, 5),
+                   gtools::stars.pval(`p-val`)),
+    order = row_number()
   ) %>%
-  filter(term %in% c("log_physact", "access_ls", "tweets_ls"), submodel == "Total")
+  spread(effect, output) %>%
+  arrange(order) %>%
+  dplyr::select(-order) %>%
+  mutate(
+    term = c(termnames, "Access + Tweets logsum", "log(Physical Activity Rate)")
+  ) %>%
+  kable("latex", booktabs = T, escape = FALSE, digits = 3, 
+        col.names = c("", "Direct", "Indirect", "Total")) %>%
+  footnote("$^{***}p<0.001$, $^{**}p<0.01$, $^*p<0.05$", escape = F)
 @
+\end{table*}
 
 \section{Discussion}
+
+We readily acknowledge limitations in this study. To begin, it would be beneficial 
+to have estimated the logsum coefficients from a high-quality survey rather than
+asserting them to match {\em a priori expectations of}
 

### 2019-01-09 manuscript.Rnw
WIP

@@
     Pct0to17, Pct18to29, Pct65plus,
-    Income1, Income2, Income3, Income4, Income5, Income6,
-    Income7, Income8, Income9, Income10
+    income = NYC_Income,
+    lowincome = Income1 + Income2 + Income3,
+    highincome = Income8 + Income9 + Income10
   )
@@
   variable =  c(
-    "obesity", "physact", "density", "fulltime", "college", "single", 
+    "obesity", "physact", "income", "lowincome", "highincome",  
+    "density", "fulltime", "college", "single", 
     "Pct0to17", "Pct18to29",  "Pct65plus", "black", "asian", "hispanic", 
@@
 tract_summary <- tracts@data %>%
-  dplyr::select(obesity:Pct65plus) %>%
+  dplyr::select(obesity:highincome) %>%
   gather(variable, value) %>%
@@
       "Obesity Rate", "Physical Activity Rate",
+      "Median tract income",
+      "Share of low income",
+      "Share of high income",
       "Density: population per square mile",
@@
   ~ log(density) +
+    lowincome + 
     fulltime + college + single +
@@
 
-termnames <- c("log(density)" , "Full-time workers", "College-educated", 
-               "Single Adults", "Share of children (0-17)", "Share of young adults (18-29)", 
+termnames <- c("log(density)" , "log(Income)", 
+               "Full-time workers", "College-educated",  "Single Adults", 
+               "Share of children (0-17)", "Share of young adults (18-29)", 
                "Share of seniors (65+)", "Black population share", "Asian population share", 
@@
 # calculate centroids of tracts and parks
-
 parks <- parks %>% st_transform(st_crs(pop_weighted_centroids))
@@
     access_ls = calculate_park_logsums(log(distances), parks$log_size,
-                                       betas = c(-0.42, 1.0)),
+                                       betas = c(-1.79, 0)),
     tweets_ls = calculate_park_logsums(log(distances), parks$log_size, parks$log_tweets,
-                                       betas = c(-0.42, 1.0, 0.1)),
+                                       betas = c(-0.79, 0, 0.1)),
     access_ls = (access_ls - mean(access_ls) )/ sd(access_ls),
@@
 <<spatial selection>>=
-physact_base_sem <- errorsarlm(update(base_formula, log_physact ~ .),
+physact_base_sem <- errorsarlm(update(base_formula, physact ~ .),
                              data = tracts@data, listw = W, zero.policy = TRUE)
-physact_base_sdm <- lagsarlm(update(base_formula, log_physact ~ .),
+physact_base_sdm <- lagsarlm(update(base_formula, physact ~ .),
                            data = tracts@data, listw = W, zero.policy = TRUE,
@@
        booktabs = TRUE, use.packages = FALSE, 
-       custom.coef.names = c("Intercept", termnames, paste("lag", termnames), "$\\rho$",
-                             "Access logsum", "lag Access logsum", 
-                             "Access + Tweets logsum", "lag Access + Tweets logsum"))
+       custom.coef.names = c("Intercept", termnames, "$\\lambda$",
+                             "Access logsum", 
+                             "Access + Tweets logsum"))
 #' screenreg(pa_models, digits = 4, table = FALSE)
@@
 <<obesity_models>>=
-obese_tweets_sdm <- update(pa_tweets_sdm, log_obesity ~ . + log_physact)
+obese_access_sdm <- update(pa_access_sdm, obesity ~ .)
+obese_physact_sdm <- update(obese_access_sdm, . ~ . + physact)
 @
@@
 <<tab:obese_impacts, results="asis">>=
-extract_impacts(obese_tweets_sdm) %>%
+extract_impacts(obese_physact_sdm) %>%
   group_by(effect) %>%
@@
     output = paste(round(`estimate`, 5),
+                   "(",round(`std.error`, 5),")",
                    gtools::stars.pval(`p-val`)),

### 2019-01-09 model_estimation.Rmd
WIP

@@
 
-```{r read_tracts}
-tracts <- geojson_read("data/nyc_tracts.geojson", what = "sp")
-tracts_sf <- st_read("data/nyc_tracts.geojson", quiet = TRUE) %>%
-  st_transform(4326)
-```
-
 ## Complete Cases
@@
 this information.
+```{r read_tracts}
+tracts <- geojson_read("data/nyc_tracts.geojson", what = "sp")
+tracts_sf <- st_read("data/nyc_tracts.geojson")
+
+# identify complete cases
+complete_tracts_index <- tracts@data %>%
+  tbl_df() %>%
+  complete.cases()
+tracts <- tracts[complete_tracts_index, ]
+
+tracts@data <-  tbl_df(tracts@data) %>%
+  dplyr::transmute(
+    geoid = GEOID,
+    county_fips = substr(GEOID, 3, 5),
+    borough = case_when(
+      county_fips == "081" ~ "Queens",
+      county_fips == "047" ~ "Brooklyn",
+      county_fips == "061" ~ "Manhattan",
+      county_fips == "005" ~ "Bronx",
+      county_fips == "085" ~ "Staten Island",
+      TRUE ~ as.character(NA)
+    ),
+    obesity = OBESITY,
+    physact = 100 - Phys_Act,
+    log_obesity = log(OBESITY),
+    log_physact = log(100 - Phys_Act),
+    density = Pop_Density,
+    fulltime = FulltimeWork, college = CollegeDeg,
+    single = Single_Percent,
+    black = PctBlack, asian = PctAsian, hispanic = PctHispanic,
+    other = PctNative + PctPacific,
+    Pct0to17, Pct18to29, Pct65plus,
+    income = NYC_Income
+  )
+```
 
 ```{r tracts_with_missing}
-# identify complete cases
-complete_tracts_index <- tracts_sf %>%
-  st_set_geometry(NULL) %>%
-  tbl_df() %>%
-  complete.cases()
-
 # make a map of 
-leaflet( tracts_sf[!complete_tracts_index, ]) %>%
+leaflet( tracts_sf[!complete_tracts_index, ] %>% st_transform(4326)) %>%
   addProviderTiles(providers$OpenStreetMap.Mapnik) %>% 
@@
 
-```{r complete_cases}
-tracts <- tracts[complete_tracts_index, ]
+The parks data seems relatively complete, with the exception of a few variables
+that are missing everywhere. Also it seems that we are including parking strips 
+very small pieces of property, even though these facilities are unlikely to 
+generate physical activity - related trips. Finally, sports fields within larger
+parks probably do not need to be double-counted.
+
+```{r read_parks, echo=FALSE}
+allparks <- sf::st_read("data/nyc_parks.geojson", quiet = TRUE) %>%
+  transmute(
+    size = Park_Acres,
+    log_size = log(Park_Acres),
+    tweets = TWEET_COUNT,
+    log_tweets = yeo.johnson(TWEET_COUNT, 0)
+  )
+
+parks <- allparks %>%
+  filter(size > 0.5)
 ```
 
-The parks data seems relatively complete, with the exception of a few variables
-that are missing everywhere. But we'll ignore this for now.
-
-```{r read_parks, echo=FALSE}
-parks <- sf::st_read("data/nyc_parks.geojson")
+The map below shows the parks we are including.
+```{r parksmap}
+leaflet( allparks %>% filter(size > 0.5) %>% st_transform(4326)) %>%
+  addProviderTiles(providers$OpenStreetMap.Mapnik) %>% 
+  addPolygons()
 ```
@@
 ```{r tweettransformation}
-parks %>%
-  tbl_df() %>%
-  dplyr::select(raw = TWEET_COUNT) %>%
-  filter(raw < 100) %>%
-  mutate(
-    `0.5` = yeo.johnson(raw, 0.5),
-    `0` = yeo.johnson(raw, 0),
-    log = log(raw)
-  ) %>%
-  gather(lambda, x, -raw) %>%
-  ggplot(aes(x = raw, y = x, color = lambda)) + 
-  geom_line()
-
-parks <- parks %>%
-  mutate(
-    log_size = log(Park_Acres),
-    log_tweets = yeo.johnson(TWEET_COUNT, 0)
-  )
+data_frame(x = seq(0, 50, by = 0.5)) %>% 
+  mutate(log = log(x), `yeo-johnson` = yeo.johnson(x, 0)) %>%
+  gather(transform, y, -x) %>%
+  ggplot(aes(x = x, y = y, color= transform)) + geom_line()
 
@@
 ```{r show_data, echo=FALSE}
-tracts@data <- tbl_df(tracts@data) %>%
-  dplyr::select(
-    GEOID, OBESITY, Park_Percent, Phys_Act, MENTAL, 
-    Income1, Income2, Income3, Income4, Income5, Income6, 
-    Income7, Income8, Income9, Income10, Pop_Density, 
-    FulltimeWork, CollegeDeg, Pct0to17, Pct18to29, Pct30to64, Pct65plus, 
-    Single_Percent, PctWhite, PctBlack, PctNative, PctAsian, PctPacific,
-    PctOther, PctHispanic
-  )  %>%
-  mutate(
-    log_obesity = log(OBESITY),
-    log_physact = log(100 - Phys_Act)
-  )
-
+tracts@data
 ```
@@
 base_formula <- formula(
-  ~ log(Pop_Density) + 
-    FulltimeWork + CollegeDeg + Single_Percent + 
+  ~ log(density) + log(income) +
+    fulltime + college + single +
     Pct0to17 + Pct18to29 + Pct65plus + # need to leave out a category for collinearity
-    PctBlack + PctAsian + PctOther + PctHispanic)
+    black + asian + hispanic + other)
 ```
@@
 ```{r spatial selection}
-obese_base_lm <- lm(update(base_formula, log_obesity ~ .), data = tracts@data)
-obese_base_sar <- lagsarlm(update(base_formula, log_obesity ~ .), 
+physact_base_lm <- lm(update(base_formula, physact ~ .), data = tracts)
+physact_base_sar <- lagsarlm(update(base_formula, physact ~ .),
                            data = tracts@data, listw = W, zero.policy = TRUE)
-obese_base_sem <- errorsarlm(update(base_formula, log_obesity ~ .), 
+physact_base_sem <- errorsarlm(update(base_formula, physact ~ .),
                              data = tracts@data, listw = W, zero.policy = TRUE)
-obese_base_sdm <- lagsarlm(update(base_formula, log_obesity ~ .), 
+physact_base_sdm <- lagsarlm(update(base_formula, physact ~ .),
                            data = tracts@data, listw = W, zero.policy = TRUE,
                            type = "mixed")
+test_sdmsem <- lmtest::lrtest(physact_base_sem, physact_base_sdm)
 ```
 
-A likelihood ratio test reveals that the SEM is not preferred, so we use the SDM
+A likelihood ratio test reveals that the SEM actually preferred, so we use this
 only going forward.
 
-```{r lrtest}
-lmtest::lrtest(obese_base_sem, obese_base_sdm)
+```{r }
+physact_models <- list("OLS" = physact_base_lm, "SAR" = physact_base_sar, 
+                       "SEM" = physact_base_sem, "SDM" = physact_base_sdm)
+screenreg(physact_models, digits = 5)
+
 ```
@@
 ```{r residuals}
-tracts_sf <- tracts_sf[complete_tracts_index, ] %>%
+tracts_sf <- tracts_sf %>%
   mutate(
-    log_obesity = tracts$log_obesity,  
-    obese_base_lm_residuals = obese_base_lm$residuals,
-    obese_base_sdm_fitted = obese_base_sdm$fitted.values,  
-    obese_base_sdm_residuals = obese_base_sdm$residual
+    log_obesity = tracts$log_physact,  
+    physact_base_lm_residuals = physact_base_lm$residuals,
+    physact_base_sem_fitted = physact_base_sem$fitted.values,  
+    physact_base_sem_residuals = physact_base_sem$residual
   )
 
-ggplot(tracts_sf, aes(x = log_obesity, y = obese_base_sdm_fitted)) + 
+ggplot(tracts_sf, aes(x = log_obesity, y = physact_base_sem_fitted)) + 
   geom_point()
@@
 ```{r residuals_map}
-pal <- colorQuantile("BrBG", tracts_sf$obese_base_sdm_residuals, n = 9)
-leaflet(tracts_sf) %>%
+pal <- colorQuantile("BrBG", tracts_sf$physact_base_sem_residuals, n = 9)
+leaflet(tracts_sf %>% st_transform(4326)) %>%
   addProviderTiles(providers$CartoDB.Positron) %>%
-  addPolygons(color = ~pal(obese_base_sdm_residuals), stroke = NA, opacity = 0.8)
+  addPolygons(color = ~pal(physact_base_sem_residuals), stroke = NA, opacity = 0.8)
 ```
@@
 # Logsum Calculation
-
-Consider that an individual is choosing a park for a recreation activity.
-According to basic choice theory (Mcfadden 1974), the probability of choosing
-park $p$ from the set of all regional parks $J$ is:
-
-$$ P(p | V_p) = \frac{\exp(V_p)}{\sum_{j \in J}\exp(V_j)}$$
-where parks are differentiated from each other by their relative measurable
-utilitie $V$. In principle, $V$ may include any measurable attributes of either
-the choice maker or the park. In this study we use a linear formulation of
-
-$$V_{ij} = size_j\beta_s + distance_{ij}\beta_d$$
-incorporating the size of the park in acres and the distance of the park from
-the census tract $i$ in miles. The coefficients $\beta$ are typically estimated from 
-surveys, though in the absence of a survey we apply a maximum likelihood
-technique described below.
-
-A key theoretical understanding of random utility choice models is that the
-consumer surplus of the choice set can be otained as the log-sum of the
-denominator of the choice probability equation. In plainer terms, the *total value*
-of an individual's park accessibility is defined to be:
-
-$$CS_i = \ln\left({\sum_{j \in J}\exp(V_{ij})}\right)$$
-
-There are several advantages to a log-sum defined metric relative to
-buffer-based accessibility metrics more commonly found in the literature. First,
-all individuals are defined as having some access to all parks, rather than an
-arbitrary limit of 1/2 mile or so. This allows for the fact that some people are
-more or less sensitive to distances, and that distance is a continuous, and not
-a binary, phenomenon. Second, the random utility formulation allows the
-researcher to include any attribute of the park; in this case, we consider the
-size of the park as an element of accessibility. This suggests that not all
-parks are equal, and that a large park such as New York City's Central Park may
-provide health and activity benefits over a much larger area than a smaller
-community square.
-
-McFadden, D. (1974). The measurement of urban travel demand. Journal of public economics, 3(4), 303-328.
-
-In the absence of a park choice survey, we estimated likely values of the $\beta$ 
-coefficients by iteratively searching for the values which produced the highest
-model likelihood in *which model did we use to estimate likelihood?*. The search
-was constrained by two assertions. First, we required that the coefficient on
-size be positive and that on distance negative; all else equal, people will
-prefer to use larger parks that are nearer to their residence. Second, we visually
-inspected the resulting accessibility scores to ensure that the scores produced a 
-reasonably varied pattern of access throughout the metropolitan region.
 
@@
 # calculate centroids of tracts and parks
-tract_centroids <- tracts %>% st_as_sf() %>% st_centroid() 
-park_centroids <- parks %>% st_centroid() %>% st_transform(st_crs(tract_centroids))
+parks <- parks %>% st_transform(st_crs(pop_weighted_centroids))
 
 # distances from all tracts to all zones
-distances <- st_distance(tract_centroids, park_centroids, byid = TRUE) %>%
+distances <- st_distance(pop_weighted_centroids, parks, byid = TRUE) %>%
   units::set_units(miles) %>%
   units::drop_units()
+
+# assert that a tract must be at least 1/10 mile from a park
+distances <- pmax(distances, 0.1)
+```
+
+We now need to run a subroutine to determine the values of the beta coefficients
+which produce the highest model likelihood, and .
+
+```{r notweets_logsum, eval = FALSE}
+notweets_logsum <- function(betas, distances, parks, tracts, model) {
+
+  logsum <- calculate_park_logsums(
+    d = distances,  sizes = parks$log_size,
+    betas = betas)
+
+  tracts@data$access_ls <- (logsum - mean(logsum)) / sd(logsum)
+
+  sdm <- update(model, . ~ . + access_ls)
+  return(-logLik(sdm))
+}
+
+access_sdm_val <- optim(
+  fn = notweets_logsum,  
+  par = c(-0.42, 0.5), # distance, size 
+  lower = c(-10, 0), # size coefficient must be positive 
+  upper = c(0, 10), # distance coefficient must be negative 
+  distances = log(distances), parks = parks, tracts = tracts, 
+  model = physact_base_sem,
+  method = "L-BFGS-B", 
+  control = list(trace = 6))
+```
+
+```{r tweets_logsum, eval = FALSE}
+tweets_logsum <- function(betas, distances, parks, tracts, model) {
+
+  logsum <- calculate_park_logsums(
+    d = distances,  sizes = parks$log_size, tweets = parks$log_tweets,
+    betas = betas)
+
+  tracts@data$tweets_ls <- (logsum - mean(logsum)) / sd(logsum)
+
+  sdm <- update(model, . ~ . + tweets_ls)
+  return(-logLik(sdm))
+}
+
+tweets_sdm_val <- optim(
+  fn = tweets_logsum,  
+  par = c(-0.42, 0.5, 1), # distance, size 
+  lower = c(-10, 0, -10), # size coefficient must be positive 
+  upper = c(0, 10, 10), # distance coefficient must be negative 
+  distances = log(distances), parks = parks, tracts = tracts, 
+  model = physact_base_sem,
+  method = "L-BFGS-B", 
+  control = list(trace = 6))
+
 ```
@@
     access_ls = calculate_park_logsums(log(distances), parks$log_size,
-                                       betas = c(-0.8, 0.8)),
+                                       betas = c(-1.796, 1.0)),
     tweets_ls = calculate_park_logsums(log(distances), parks$log_size, parks$log_tweets,
-                                       betas = c(-0.8, 0.8, 0.2))
+                                       betas = c(-1.796, 1.0, 0.1)),
+    access_ls = (access_ls - mean(access_ls) )/ sd(access_ls),
+    tweets_ls = (tweets_ls - mean(tweets_ls) )/ sd(tweets_ls)
   )
@@
 
-leaflet(tracts_sf) %>%
+leaflet(tracts_sf %>% st_transform(4326)) %>%
   addProviderTiles(providers$CartoDB.Positron) %>%
@@
 ```{r obesity_access_models}
-access_sdm <- update(obese_base_sdm, .~ . + access_ls)
-tweets_sdm <- update(obese_base_sdm, .~ . + tweets_ls)
+access_sem <- update(physact_base_sem, .~ . + access_ls)
+tweets_sem <- update(physact_base_sem, .~ . + tweets_ls)
 ```
@@
 ```{r modelchart}
-obesity_models <- list(Base = obese_base_sdm, `Access` = access_sdm, 
-                       `Access + Tweets` = tweets_sdm)
-screenreg(obesity_models, digits = 4)
+physact_models <-  
+screenreg()
+
 ```

### 2019-02-11 manuscript.Rnw
Lock down coefficients on size and distance.

@@
 
-Numerous previous studies have found relationships between aspects of the built
+Numerous previous studies have found relationships between characteristics of the built
 environment and aspects of physical, mental, social, and economic health. Many
@@
 
-While several of the aforementioned studies yielded results consistent with the
+While several of the aforementioned studies reported results consistent with the
 hypothesis that park access and use confer health benefits stemming from
@@
 Metropolitan-level analyses do not capture the within-city variation in
-accessibility to parks that exists. Some cities with large amounts of total
-greenspace may nevertheless inequally distribute this space throughout the city
+accessibility to parks that exist in many cities. Some cities with large amounts of total
+greenspace may nevertheless inequally distribute this space throughout the city,
 leading to areas with poor access. Conversely, a city with a smaller overall
-proportion of greenspace may give all of its citizens better access to significant
-greenspace. Considering access at the neighborhood level within a single city
-may eliminate some regional or cultural fixed effects affecting
+proportion of greenspace may give all of its citizens better access to sufficient
+greenspace to meet their needs or wants for physical activity. Considering access at the neighborhood level within a single city
+may also eliminate some regional or cultural fixed effects affecting
 metropolitan-level analyses.
@@
 A key theoretical outcome of random utility choice models is that the
-consumer surplus, or the total value of all alternatives of the choice set, can
+consumer surplus, or the total weighted value of all alternatives of the choice set, can
 be obtained as the log-sum of the denominator of the choice probability equation
@@
 people are more or less sensitive to distances, and that distance is a
-continuous, and not a binary, phenomenon. Second, the random utility formulation
+continuous, non-binary phenomenon. Second, the random utility formulation
 allows the researcher to include any attribute of the park; in this case, we
@@
 transit infrastructure improvements \citep{DeJong2007}. A reason for this is
-likely that such analyses are usually conducted in the context of a travel
+likely that such alternatives analyses are usually conducted in the context of a travel
 demand model, where calibrated and multimodal logsums are readily available
@@
 
-\section{Empirical Methodology}
+\section{Empirical Application}
 In this section we describe an experiment with data for New York City, where
-we compute a holistic logsum-based accessibility component and model the
+we compute a holistic logsum-based accessibility to parks metric and model the
 relationship between this measure and physical activity rates, controlling for
-spatial effects and socioeconomic factors.
+spatial effects and socioeconomic factors. We subsequently model the effect
+of this accessibility on obesity.
 
@@
 activity participation rate and obesity rates for each tract are available from
-the Centers for Disease Control and Prevention’s 500 Cities Project.
+the Centers for Disease Control and Prevention’s 500 Cities Project (WE NEED A CITATION).
 After removing tracts with missing population information or outlying and unusual
@@
 We retrieved a shapefile of public parks and greenspaces within New York City’s
-municipal boundaries and checked it for accuracy and relevance. Upon inspection,
+municipal boundaries and checked it for accuracy and relevance (WE NEED A CITATION).
+Upon inspection,
 we removed several facilities that do not qualify as publicly accessible green
@@
 vacant lots or parking strips rather than legitimate public green space. This
-leaves us with \Sexpr{nrow(parks)} separate parks.
+leaves us with \Sexpr{nrow(parks)} parks. It is worth noting that the particular
+dataset classifies specific facilities within larger parks as independent
+features: baseball fields within Central Park, for example, are themselves
+unique greenspace. We elect to retain these independent features as they may
+generate additional activity.
 
@@
   kable_styling() %>%
-  group_rows(paste("Tract Variables, N =", nrow(tracts@data)), 1, 13) %>%
-  group_rows(paste("Park Variables, N =", nrow(parks)), 14, 15)
+  group_rows(paste("Tract Variables, N =", nrow(tracts@data)), 1, 16) %>%
+  group_rows(paste("Park Variables, N =", nrow(parks)), 17, 18)
 
@@
 function of the tract's sociodemographic characteristics and choice-based
-accessibility to parks with a semi-log specification:
+accessibility to parks with the following  specification:
 \begin{equation}\label{eq:themodel}
-\log(y_i) = f(\vect{x}_i, CS_i, g(W, \mathit{Obesity}, \vect{X}))
+ y_i = f(\vect{x}_i, CS_i, g(W, \vect{y} \vect{X}))
 \end{equation}
-<<base_formula>>=
-base_formula <- formula(
-  ~ log(density) +
-    lowincome + 
-    fulltime + college + single +
-    Pct0to17 + Pct18to29 + Pct65plus + # need to leave out a category for collinearity
-    black + asian + hispanic + other)
-
-termnames <- c("log(density)" , "log(Income)", 
-               "Full-time workers", "College-educated",  "Single Adults", 
-               "Share of children (0-17)", "Share of young adults (18-29)", 
-               "Share of seniors (65+)", "Black population share", "Asian population share", 
-               "Hispanic population share", "Other Minorities")
-@
 where $\vect{x}_i$ is a vector of tract-level sociodemographic attributes,
@@
 estimated coefficients relating the attributes of an observation's {\em neighbors}
-to its outcome.
+to its outcome. If only correlation and not dependence exists in the data
+generating process, it may  be more efficient to estimate a restricted
+spatial autoregressive model and therefore
+save degrees of freedom in the estimation. \citet{tra_autoregression_paper}
+illustrate a decision algorithm originally proposed by \citet{Florax2003} which uses
+a likelihood ratio test of the SDM against the restricted spatial error model (SEM):
+\begin{align}
+{\boldsymbol y} &= X{\boldsymbol \beta} + {\boldsymbol u},  {\boldsymbol u} = \lambda W {\boldsymbol u} +
+{\boldsymbol \epsilon} \label{eq:SEM}\\
+{\boldsymbol y} &= X{\boldsymbol \beta} + (I-\lambda W)^{-1}{\boldsymbol \epsilon}\notag
+\end{align}
+If the likelihood ratio test rejects that the models are equivalent, then the SDM
+should be used because its estimates are unbiased. If the test cannot reject
+equivalents, the SEM is more efficient.
 
@@
 @
-
-It may  be more efficient to estimate a restricted spatial autoregressive
-model -- that controls for only correlation or only dependence -- and therefore
-save degrees of freedom in the estimation.  \citet{tra_autoregression_paper}
-illustrate a decision algorithm originally proposed by \citet{Florax2003} which uses
-a likelihood ratio test of the SDM against a restricted spatial error model (SEM):
-if the likelihood ratio test rejects that the models are equivalent, then the SDM
-should be used.
-
-A final point to note about spatial autoregressive models is that the partial
-derivative of the model with respect to a single regressor $\vect{x_k} \in X$
-does not equal the estimated coefficient $\beta_k$. Consequently, the estimated
-coefficients do not represent the {\em effect} of a variable on the outcome. The
-direct effects of an observation's attribute on its outcome, the indirect effects
-of neighbors' attributes on an outcome, and the total effect of own and neighbors'
-attributes can attained via Monte Carlo simulation. In this study we use the
-\texttt{spdep} package for R \citep{Bivand2013} for model estimation and impact
-simulation.
 
@@
 applying a limited-memory, box-constrained optimization algorithm maximizing the
-SDM model log-likehood. The constraint required that the coefficient on
+SDM model log-likehood \cite. The constraint required that the coefficient on
 size be positive and that on distance negative; all else equal, we assert people
 will prefer to use larger parks and parks that are nearer to their residence. This
-resulted in $\beta_d = -0.42$  and $\beta_s = 0.000$. Given our expectation that
-park size is a valuable attribute, we subsequently adjusted this value to
-a value of $1.00$. This implies that each mile
+resulted in $\beta_d = -1.98$  and $\beta_s = 0.000$. Though the "most likely"
+parameter value, in this case it was unsatisfactory for two reasons: first,
+our prior assertion that park size is a valuable attribute of a park, and second
+the observation that in this specification tracts bordering on the corners of
+Central Park were measured with poor overall accessibility. we subsequently adjusted the $\beta_s$ value
+so that the
+
+to a value of $1.00$
 
@@
     access_ls = calculate_park_logsums(log(distances), parks$log_size,
-                                       betas = c(-1.79, 0)),
+                                       betas = c(-1.98, 0.3)),
     tweets_ls = calculate_park_logsums(log(distances), parks$log_size, parks$log_tweets,
-                                       betas = c(-0.79, 0, 0.1)),
+                                       betas = c(-1.98, 0.3, 0.1)),
     access_ls = (access_ls - mean(access_ls) )/ sd(access_ls),
@@
           aes(fill = cut(access_ls, breaks = c(-Inf, -3, -2, -1, 0, 2, 3, Inf))),
-          alpha = 0.8) +
+          alpha = 0.8, color = NA) +
   scale_fill_brewer("Relative Park Logsum", palette = "GnBu")
@@
 
-\section{Results}
+\subsection{Results}
+
+<<base_formula>>=
+base_formula <- formula(
+  ~ log(density) +
+    fulltime + college + single +
+    Pct0to17 + Pct18to29 + Pct65plus + # need to leave out a category for collinearity
+    black + asian + hispanic + other)
+
+@
 
 <<spatial selection>>=
-physact_base_sem <- errorsarlm(update(base_formula, physact ~ .),
+physact_base_sem <- errorsarlm(update(base_formula,
+                                      physact ~ . + highincome + lowincome),
                              data = tracts@data, listw = W, zero.policy = TRUE)
-physact_base_sdm <- lagsarlm(update(base_formula, physact ~ .),
+physact_base_sdm <- lagsarlm(update(base_formula,
+                                    physact ~ . + highincome + lowincome),
                            data = tracts@data, listw = W, zero.policy = TRUE,
                            type = "mixed")
+pa_base_logincome <- update(physact_base_sem, . ~ . +log(income) - highincome - lowincome)
 test_sdmsem <- lmtest::lrtest(physact_base_sem, physact_base_sdm)
-@
-
-<<access_sdm>>=
-# estimate models
-pa_access_sdm <- update(physact_base_sdm, .~ . + access_ls)
-pa_tweets_sdm <- update(physact_base_sdm, .~ . + tweets_ls)
 @
@@
 against the base covariates (without the accessibility logsum). A likelihood
-ratio test rejects that the SEM is equivalent
+ratio test failed to reject that the SEM is equivalent
 ($\chi^2 = \Sexpr{test_sdmsem$Chisq[2]}$ on \Sexpr{test_sdmsem$Df[2]} degrees of
-freedom), so we use the SDM specification only going forward. Table
-\ref{tab:pa_models} presents the estimated coefficients for the base model as
+freedom), so we use the SEM specification only going forward.
+Table \ref{tab:pa_models} presents the estimated coefficients for the base
+model with income represented in two ways. In the first, the share of the tract
+population in
+the bottom three income deciles and the share of the population in the top three
+income deciles are each included as independent predictors. In the second, the
+logarithm of the mean tract income is included as a predictor.
+Notwithstanding the higher model likelihood of the deciles model, comparing
+coefficients across the models shows that the predictors change magnitude
+or significance only marginally, if at all. Overall, the base model
+predictors are significant and of intuitive direction. Tracts with higher shares
+of full-time workers, college-educated adults, young adults, and high-income
+households all show a greater share of individuals engaging in regular physical
+activity. Conversely, tracts with greater population density and a greater
+share of single adults, children, seniors, minorities of all types, and low
+income households all report lower rates of physical activity.
+
+The other two models presented in Table \ref{tab:pa_models} include
+the log-sum accessibility statistic described in Section
+\ref{subsec:accessibility}  added as a predictor. In both income specifications,
+the
+
+base model as
 well as models including logsum-based accessibility with and without Twitter
@@
 
+<<access_sdm>>=
+# estimate models
+pa_access_sem <- update(physact_base_sem, .~ . + access_ls)
+pa_access_logincome <- update(pa_access_sem, . ~ . + log(income) - highincome - lowincome)
+pa_models <- list("Base" = physact_base_sem,
+                  "Base: log(Income)" = pa_base_logincome,
+                  "Access" = pa_access_sem,
+                  "Access: log(Income)" = pa_access_logincome)
+
+@
+
+<<screen_pa_models, eval = FALSE>>=
+screenreg(pa_models, digits = 4, table = FALSE,
+          stars = c(0.001, 0.01, 0.05, 0.1) )
+@
+
 \begin{table*}
-  \caption{Estimated Coefficients Relating Park Access to Physical Activity}
+  \caption{SEM Coefficients Predicting Physical Activity Rates}
   \label{tab:pa_models}
 <<pamodelchart, results="asis">>=
-pa_models <- list(Base = physact_base_sdm, `Access` = pa_access_sdm,
-                       `Access + Tweets` = pa_tweets_sdm)
+termnames <- c("(Intercept)", "log(density)" ,
+               "Full-time workers", "College-educated",  "Single Adults",
+               "Share of children (0-17)", "Share of young adults (18-29)",
+               "Share of seniors (65+)", "Black population share",
+               "Asian population share",
+               "Hispanic population share", "Other Minorities",
+               "Share high income", "Share low income",
+               "$\\lambda$", "$\\log($Income$)$",
+               "Access logsum")
 texreg(pa_models, digits = 4, table = FALSE, single.row = TRUE,
        booktabs = TRUE, use.packages = FALSE,
-       custom.coef.names = c("Intercept", termnames, "$\\lambda$",
-                             "Access logsum", 
-                             "Access + Tweets logsum"))
-#' screenreg(pa_models, digits = 4, table = FALSE)
+       custom.coef.names = termnames)
 @
@@
 
-Recall from the presentation of spatial econometrics above that the total effect of a
-covariate on the dependent variable is different from the estimated coefficient
-and involves interaction between direct and indirect effects, or between
-the $\beta$ and $\gamma$ parameters. This is particularly important when the
-$\beta$ and $\gamma$ estimates for a particular covariate show a different
-sign, such as in the proportion of black residents in a census tract, as
-interpreting the direction and magnitude of the relationship is not straightforward.
+<<obesity_models>>=
+obesity_models <- lapply(pa_models, function(model){
+  update(update(model, obesity ~ . ))
+})
 
-<<extract_impacts>>=
-#' Get the MC simulated impact coefficients and significance from a model
-#' @param sdm An autoregressive lag (SAR or SDM) model object `lagsarlm`
-#' @return A data frame with the effect (direct, indirect, total), variable,
-#'   simulated mean impact, and significance p-value
-extract_impacts <- function(sdm) {
-
-  impacts_summary <- summary(
-    impacts(sdm, tr=trMC,  R = 1000, useHESS = !is.logical(sdm$fdHess)),
-    zstats = TRUE)
-
-  estimate <- list(
-    Direct = impacts_summary$direct_sum,
-    Indirect = impacts_summary$indirect_sum,
-    Total = impacts_summary$total_sum
-  ) %>%
-  lapply(function(s) {
-    data_frame(
-      term = names(s$statistics[,1]),
-      estimate = s$statistics[,1],
-      std.error = s$statistics[,2]
-    )
-  }) %>%
-    bind_rows(.id = "effect")
-
-  pval <- as_data_frame(impacts_summary$pzmat) %>%
-    mutate(
-     term = rownames(impacts_summary$pzmat)
-    ) %>%
-    gather("effect", "p-val", -term)
-
-  left_join(estimate, pval, by = c("effect", "term"))
-}
+names(obesity_models) <- names(pa_models)
 @
 
-\begin{table*}
-  \caption{Simulated Impact of Accessibility on Physical Activity}
-  \label{tab:pa_impacts}
-<<impacts, results="asis">>=
-physact_impacts <- pa_models %>%
-  lapply(extract_impacts) %>%
-  bind_rows(.id = "model") %>% ungroup() %>%
-  rename(submodel = effect)
-
-effect_estimate <- physact_impacts %>%
-  filter(model == "Access + Tweets", term == "tweets_ls", submodel == "Total") %>%
-  pull(estimate)
-
-density_estimate <- physact_impacts %>%
-  filter(model == "Access + Tweets", term == "log(density)", submodel == "Total") %>%
-  pull(estimate)
-
-physact_impacts %>%
-  transmute(
-    model, term, submodel, estimate,
-    output = paste(round(`p-val`, 5),
-                   gtools::stars.pval(`p-val`))
-  ) %>%
-  filter(term %in% c("access_ls", "tweets_ls"), submodel == "Total") %>%
-  dplyr::select(-term, -submodel) %>%
-  kable("latex", booktabs = T, escape = FALSE,
-        col.names = c("Logsum Measure", "Estimate", "$p$-value")) %>%
-  footnote("$^{***}p<0.001$, $^{**}p<0.01$, $^*p<0.05$", escape = F)
+<<screen_obese_models, eval = FALSE>>=
+screenreg(obesity_models, digits = 4, table = FALSE,
+          stars = c(0.001, 0.01, 0.05, 0.1) )
 @
-\end{table*}
-
-Table \ref{tab:pa_impacts} shows the simulated marginal impact of the
-accessibility logsum measurements on the physical activity rate in each tract.
-Given that the logsum is a relatively continuous measure over space, and that it
-would be difficult to improve the park accessibility of a tract without also
-improving the accessibility of the tract's neighbors, the total effect combining
-the direct effect and the indirect effects of all neighbors is
-most appropriate. The total effect of tract-level accessiblity on physical
-activity rates is strongly significant, regardless of whether we consider Twitter
-activity.
-
-It is of course necessary to also consider the magnitude of the effect.
-As this is a semi-log model, the coefficient implies a unit increase
-in the access and tweets logsum leads to a
-\Sexpr{round(effect_estimate * 100, 3)}\% increase in the physical activity rate.
-Referring to the map of logsum measures in Figure \ref{fig:map}, the measures
-span from roughly $-3$ in heavily industrialized segments of southern Brooklyn
-to roughly $3$ in tracts immediately adjacent to Central Park on the Upper East
-Side. Thus moving from the least accessible parts of the city to the most
-accessible will reduce the expected physical activity rate by
-\Sexpr{round(effect_estimate * 100 * 6, 3)}\%. In other words, if a tract with 
-poor accessibility is expected to achieve the sample mean of 
-\Sexpr{round(mean(tracts$physact), 1)}\%, the same tract with excellent accessibility
-is expected to achieve a rate of 
-\Sexpr{round(mean(tracts$physact) * (1 + effect_estimate  * 6), 2)}\%. 
-For an aggregate statistic strongly influenced by other socioeconomic factors, 
-this is not an insubstantial marginal effect.
-
-<<obesity_models>>=
-obese_access_sdm <- update(pa_access_sdm, obesity ~ .)
-obese_physact_sdm <- update(obese_access_sdm, . ~ . + physact)
-@
-
 We now consider the impacts of a model where the dependent variable is
 the {\em obesity} rate, and physical activity becomes an independent covariate
-alongside the controlling variables and the tweets-augmented logsum variable. 
-Table \ref{tab:obese_impacts} presents the simulated impacts resulting from
-this model.
+alongside the controlling variables and the accessibility logsum variable.
 
 \begin{table*}
-  \caption{Simulated Impacts of Variables on Obesity}
-  \label{tab:obese_impacts}
-<<tab:obese_impacts, results="asis">>=
-extract_impacts(obese_physact_sdm) %>%
-  group_by(effect) %>%
-  transmute(
-    term,
-    output = paste(round(`estimate`, 5),
-                   "(",round(`std.error`, 5),")",
-                   gtools::stars.pval(`p-val`)),
-    order = row_number()
-  ) %>%
-  spread(effect, output) %>%
-  arrange(order) %>%
-  dplyr::select(-order) %>%
-  mutate(
-    term = c(termnames, "Access + Tweets logsum", "log(Physical Activity Rate)")
-  ) %>%
-  kable("latex", booktabs = T, escape = FALSE, digits = 3, 
-        col.names = c("", "Direct", "Indirect", "Total")) %>%
-  footnote("$^{***}p<0.001$, $^{**}p<0.01$, $^*p<0.05$", escape = F)
+  \caption{SEM Coefficients Predicting Obesity Rates}
+  \label{tab:obese_models}
+<<obesemodelchart, results="asis">>=
+termnames <- c("(Intercept)", "log(density)" ,
+               "Full-time workers", "College-educated",  "Single Adults",
+               "Share of children (0-17)", "Share of young adults (18-29)",
+               "Share of seniors (65+)", "Black population share",
+               "Asian population share",
+               "Hispanic population share", "Other Minorities",
+               "Share high income", "Share low income",
+               "$\\lambda$", "$\\log($Income$)$",
+               "Access logsum")
+texreg(obesity_models, digits = 4, table = FALSE, single.row = TRUE,
+       booktabs = TRUE, use.packages = FALSE,
+       custom.coef.names = termnames)
 @

### 2019-02-11 model_estimation.Rmd
Lock down coefficients on size and distance.

@@
     Pct0to17, Pct18to29, Pct65plus,
-    income = NYC_Income
+    income = NYC_Income,
+    lowincome = Income1 + Income2 + Income3,
+    highincome = Income8 + Income9 + Income10
   )
+
+tracts_sf <- tracts_sf %>% filter(GEOID %in% tracts@data$geoid)
 ```
@@
 ```{r W, echo=FALSE}
-tracts.dnn <- dnearneigh(gCentroid(tracts, byid = TRUE), 0, 1.8 * 5280)
+pop_weighted_centroids <- read_csv("data/CenPop2010_Mean_TR36.csv") %>%
+  mutate(geoid = str_c(STATEFP, COUNTYFP, TRACTCE))
+pop_weighted_centroids <- left_join(tracts@data, pop_weighted_centroids, by = "geoid") %>%
+  dplyr::select(geoid, LONGITUDE, LATITUDE) %>%
+  st_as_sf(crs = 4326, coords = c("LONGITUDE", "LATITUDE")) %>%
+  st_transform(2263)
+
+tracts.dnn <- dnearneigh(gCentroid(as(pop_weighted_centroids, "Spatial"), byid = TRUE),
+                         0, 1.8 * 5280)
 dists <- nbdists(tracts.dnn, gCentroid(tracts, byid = TRUE))
 dists.inv <- lapply(dists, function(x) 1 / x)
-W <- nb2listw(neighbours = tracts.dnn, glist = dists.inv, zero.policy = TRUE, style = "W")
-print(W, zero.policy = TRUE)
+W <- nb2listw(neighbours = tracts.dnn, glist = dists.inv,
+              zero.policy = TRUE, style = "W")
 trMC <- trW(as(W, "CsparseMatrix"), type="MC") # trace used in montecarlo impacts
@@
 
-```{r obese_lm, echo=FALSE, message=FALSE, include=FALSE}
+```{r base_formula, message=FALSE, include=FALSE}
 base_formula <- formula(
-  ~ log(density) + log(income) +
+  ~ log(density) + highincome + lowincome +
     fulltime + college + single +
@@
 
-```{r spatial selection}
+```{r spatial_selection}
 physact_base_lm <- lm(update(base_formula, physact ~ .), data = tracts)
@@
 
-```{r }
+```{r physact_models, results='asis'}
 physact_models <- list("OLS" = physact_base_lm, "SAR" = physact_base_sar, 
                        "SEM" = physact_base_sem, "SDM" = physact_base_sdm)
-screenreg(physact_models, digits = 5)
-
+htmlreg(physact_models, digits = 5)
 ```
@@
   mutate(
-    log_obesity = tracts$log_physact,  
+    physact = tracts$physact,  
     physact_base_lm_residuals = physact_base_lm$residuals,
@@
 
-ggplot(tracts_sf, aes(x = log_obesity, y = physact_base_sem_fitted)) + 
+ggplot(tracts_sf, aes(x = physact, y = physact_base_sem_fitted)) + 
   geom_point()
@@
 ```{r logsums_function}
-
 #' Calculate destination choice logsums from a distance matrix
@@
 calculate_park_logsums <- function(d, sizes, tweets = NULL,
-                                   betas = c(-15,.00001, .001)){
+                                   betas = c(-0.4,0.2,0.001)){
 
@@
   fn = notweets_logsum,  
-  par = c(-0.42, 0.5), # distance, size 
+  par = c(-1.79, 0.5), # distance, size 
   lower = c(-10, 0), # size coefficient must be positive 
@@
 
-```{r tweets_logsum, eval = FALSE}
-tweets_logsum <- function(betas, distances, parks, tracts, model) {
-
-  logsum <- calculate_park_logsums(
-    d = distances,  sizes = parks$log_size, tweets = parks$log_tweets,
-    betas = betas)
-
-  tracts@data$tweets_ls <- (logsum - mean(logsum)) / sd(logsum)
-
-  sdm <- update(model, . ~ . + tweets_ls)
-  return(-logLik(sdm))
-}
-
-tweets_sdm_val <- optim(
-  fn = tweets_logsum,  
-  par = c(-0.42, 0.5, 1), # distance, size 
-  lower = c(-10, 0, -10), # size coefficient must be positive 
-  upper = c(0, 10, 10), # distance coefficient must be negative 
-  distances = log(distances), parks = parks, tracts = tracts, 
-  model = physact_base_sem,
-  method = "L-BFGS-B", 
-  control = list(trace = 6))
-
-```
+The final estimated values are $\beta_d$= `r access_sdm_val$par[1]`,
+$\beta_s$= `r access_sdm_val$par[2]`, though this presents problems from 
+two perspectives. First, we assert that the size of a park is not 
+immaterial in its attractiveness. Second, ignoring the size of a park results in
+tracts immediately adjacent to Central Park with poor accessibility, which is
+unreasonable. So we assert a value of 0.8.
 
@@
     access_ls = calculate_park_logsums(log(distances), parks$log_size,
-                                       betas = c(-1.796, 1.0)),
-    tweets_ls = calculate_park_logsums(log(distances), parks$log_size, parks$log_tweets,
-                                       betas = c(-1.796, 1.0, 0.1)),
+                                       betas = c(-1.96805, 0.8)),
+    estimd_ls = calculate_park_logsums(log(distances), parks$log_size,
+                                       betas = c(-1.96805, 0.0)),
+    tweets_ls = calculate_park_logsums(log(distances), parks$log_size,
+                                       parks$log_tweets,
+                                       betas = c(-1.96805, 0.8, 0.1)),
     access_ls = (access_ls - mean(access_ls) )/ sd(access_ls),
+    estimd_ls = (estimd_ls - mean(estimd_ls) )/ sd(estimd_ls),
     tweets_ls = (tweets_ls - mean(tweets_ls) )/ sd(tweets_ls)
@@
 tracts_sf <- tracts_sf %>%
-  mutate(
-    access_ls = tracts$access_ls,
-    tweets_ls = tracts$tweets_ls
-  )
+  select(-access_ls, -estimd_ls, -tweets_ls) %>%
+  left_join(tracts@data %>% select(geoid, access_ls, estimd_ls, tweets_ls), 
+            by = c("GEOID" = "geoid"))
 
-pal <- colorQuantile("Spectral", c(tracts_sf$access_ls, tracts_sf$tweets_ls), 5)
+pal <- colorQuantile("Spectral", c(tracts_sf$access_ls, tracts_sf$estimd_ls, 
+                                   tracts_sf$tweets_ls), 5)
 
@@
   addPolygons(group = "access", color = ~pal(access_ls)) %>%
+  addPolygons(group = "maximum likelihood", color = ~pal(estimd_ls)) %>%
   addPolygons(group = "access + tweets", color = ~pal(tweets_ls)) %>%
-  addLayersControl(baseGroups = c("access", "access + tweets"))
+  addLayersControl(baseGroups = c("access", "maximum likelihood", 
+                                  "access + tweets"))
 
 ```
+
+Now that we have the dependent variable, we can look at the relationship between
+access and physical activity.
 
@@
 ```
-
-```{r modelchart}
-physact_models <-  
-screenreg()
-
-```
-
-```{r impacts_extractor}
-#' Get the MC simulated impact coefficients and significance from a model
-#' @param sdm An autoregressive lag (SAR or SDM) model object `lagsarlm`
-#' @return A data frame with the effect (direct, indirect, total), variable, 
-#'   simulated mean impact, and significance p-value
-impacts_extractor <- function(sdm) {
-
-  impacts_summary <- summary(
-    impacts(sdm, tr=trMC,  R = 1000, useHESS = !is.logical(sdm$fdHess)), 
-    zstats = TRUE)
-
-  coef <- list(
-    Direct = impacts_summary$direct_sum,
-    Indirect = impacts_summary$indirect_sum,
-    Total = impacts_summary$total_sum
-  ) %>%
-  lapply(function(s) {
-    data_frame(
-      var = names(s$statistics[,1]),
-      impact = s$statistics[,1]
-    )
-  }) %>%
-    bind_rows(.id = "effect")
-
-  pval <- as_data_frame(impacts_summary$pzmat) %>%
-    mutate(
-     var = rownames(impacts_summary$pzmat)
-    ) %>%
-    gather("effect", "p-val", -var)
-
-  left_join(coef, pval, by = c("effect", "var"))
-}
-```
-
-```{r obesity_impacts}
-obesity_impacts <- obesity_models %>%
-  lapply(impacts_extractor) %>%
-  bind_rows(.id = "model") %>%
-  transmute(
-    model, var, effect, 
-    output = paste(round(impact, 5), 
-                   gtools::stars.pval(`p-val`))
-  ) %>%
-  spread(model, output) 
-```
-
-# Physical Activity
-
-We cannot necessarily assume that the logsum parameter will be constructed the
-same way for physical activity as for obesity. (maybe?) But that's the assumption
-we'll make for now.
-
-```{r physact_estimation}
-pa_base_sdm <- update(obese_base_sdm, log_physact ~ .)
-pa_access_sdm <- update(access_sdm, log_physact ~ .)
-pa_tweets_sdm <- update(tweets_sdm, log_physact ~ .)
-```
-
-```{r physact_models}
-physact_models <- list(Base = pa_base_sdm, Access = pa_access_sdm, 
-                       `Access + Tweets` = pa_tweets_sdm)
-screenreg(physact_models, digits = 4)
-```
-
-```{r physact_impacts}
-physact_impacts <- physact_models %>%
-  lapply(impacts_extractor) %>%
-  bind_rows(.id = "model") %>%
-  transmute(
-    model, var, effect, 
-    output = paste(round(impact, 5), 
-                   gtools::stars.pval(`p-val`))
-  ) %>%
-  spread(model, output) 
-```
-
-# Results
-```{r impacts_table}
-list("Obesity" = obesity_impacts, "Physical Activity" = physact_impacts) %>%
-  bind_rows(.id = "Outcome") %>%
-  filter(var %in% c("access_ls", "tweets_ls")) %>%
-  dplyr::select(-var, -Base)  %>%
-  gather("Measure", "Impact", -Outcome, -effect) %>%
-  filter(!is.na(Impact))  %>%
-  spread(effect, Impact)
-
-```

### 2019-02-12 manuscript.Rnw
Complete initial draft of paper

@@
 
+\usepackage{url}
+
 \begin{document}
@@
 %% Front matter =========================
-\title{The impacts of park access on health outcomes: a spatial comprehensive approach}
+\title{The impacts of park access on health outcomes: A spatial comprehensive approach}
 
-\author{Nico Boyd\affilnum{1}, Gregory S. Macfarlane\affilnum{2}, and Kari Watkins\affilnum{1}}
+\author{Nico Boyd\affilnum{1}, Gregory S. Macfarlane\affilnum{2}, Kari Watkins\affilnum{1}, 
+and Professor WhogaveTweets\affilnum{3}}
 \runninghead{Boyd, et al.}
@@
 \affiliation{\affilnum{1}School of Civil and Environmental Engineering, Georgia Institute of Technology\\
-\affilnum{2}Civil and Environmental Engineering Department, Brigham Young University}
+\affilnum{2}Civil and Environmental Engineering Department, Brigham Young University\\
+\affilnum{3}Somewhere at Georgia Tech}
 
@@
 In this paper, we present a holistic and flexible measurement for park
-accessibility based in random utility theory. This measure can consider the continuous
+accessibility based in random utility choice theory. This measure can consider 
+the continuous
 distance to all parks in the region, weighted by the size of the park and other
@@
 activity participation and obesity rates in New York City, controlling for
-spatial correlation and dependence. We find a positive relationship where the
+spatial correlation and tract-level socioeconomic characteristics. 
+We find a positive relationship where the
 least park-accessible tracts have an expected physical activity participation
-rate 3.8 percent lower than the most accessible tracts, controlling for
-socioeconomic characteristics. We also find that increased physical activity
-rates lead to reduced aggregate obesity rates, but that when controlling for
-physical activity rates, accessibility to parks does not reduce overall accessibility.
-This suggests that to be an effective tool for addressing chronic diseases,
-communities must do more than simply provide parks, but also incentivize more
-citizens to use them.
+rate XX percent lower than the most accessible tracts, though this result is 
+only suggestive of correlation from a statistical perspective.
+However, we also find a strong and significant negative correlation between 
+increased park accessibility and decreased obesity rates, again controlling for
+spatial correlation, socioeconomic attributes, and physical activity rates.
+
+We also demonstrate how the choice-based accessibility metric could potentially 
+be expanded to account for other elements of a park's attractiveness by 
+introducing the number of tweets emanating from the park during 
+(WE NEED TO DESCRIBE THE PERIOD). The paper concludes with a discussion of
+opportunities for future research.
 
@@
 
-\begin{table*}[ht]
+\begin{table*}
   \caption{Commonly cited benefits of parks and greenspaces}
@@
 Metropolitan-level analyses do not capture the within-city variation in
-accessibility to parks that exist in many cities. Some cities with large amounts of total
-greenspace may nevertheless inequally distribute this space throughout the city,
-leading to areas with poor access. Conversely, a city with a smaller overall
-proportion of greenspace may give all of its citizens better access to sufficient
-greenspace to meet their needs or wants for physical activity. Considering access at the neighborhood level within a single city
-may also eliminate some regional or cultural fixed effects affecting
-metropolitan-level analyses.
+accessibility to parks that exist in many cities. Some cities with large amounts
+of total greenspace may nevertheless inequally distribute this space throughout
+the city, leading to areas with poor access. Conversely, a city with a smaller
+overall proportion of greenspace may give all of its citizens better access to
+sufficient greenspace to meet their needs or wants for physical activity.
+Considering access at the neighborhood level within a single city may also
+eliminate some regional or cultural fixed effects affecting metropolitan-level
+analyses.
 
@@
 
-To evaluate the impacts of parks and greenspaces on health outcomes at a
-sub-metropolitan level, one of the fundamental tasks involved is the selection
-of a measure of accessibility. In a comprehensive meta-analysis of published
-studies that measure active accessibility (accessibility using non-motorized
-travel modes including walking and cycling), \citet{Vale2016} identified four
-broad categories of studies based on the metric employed:
+When evaluating the impacts of parks and greenspaces on health outcomes at a
+sub-metropolitan level, one of the fundamental challenges researchers face is
+selecting a measure of accessibility. In a comprehensive meta-analysis of
+published studies that measure active accessibility (accessibility using
+non-motorized travel modes including walking and cycling), \citet{Vale2016}
+identified four broad categories of studies based on the metric employed:
 \begin{description}
@@
 setting, noting that ''ways to measure active accessibility are as varied as the
-number of scholars the measure them.'' For example, infrastructure-based distances
-can vary widely depending on which modes of travel infrastructure are considered.
+number of scholars the measure them.'' This is at least partly because
+each method has inherent strengths and weaknesses.
+Distance is an inherently continuous phenomenon, but software tools ease and enable
+buffer analyses where distance becomes a binary phenomon of somewhat arbitrary
+extent. \citet{Logan2017} note that these commonly used approaches can also lead
+researchers to overlook access-poor populations.
+Infrastructure-based distances can vary widely
+depending on which modes of travel are considered. a highway network may not
+adequately capture trails and bike paths.
 Similarly, gravity-based measures may not always account for surrounding land
-uses when considering the attractiveness or impedance of a destination.
-In evaluating methodologies of measuring access to urban services, including
-parks and greenspaces, \citet{Logan2017} note that existing and commonly used
-approaches "often simplify their measure of proximity by using large areal
-units and by imposing arbitrary distance thresholds," which often results in
-access-poor populations being overlooked.
+uses when considering the attractiveness or impedance of a destination; the
+findings of \citet{Pretty2005} and other suggest unpleasant pedestrian routes
+attenuate perceived travel costs.
+
+Though there is no shortage of possible accessibility metrics, in this paper
+we wish to present and apply a technique that is explicitly based in the choices
+of individuals as they consider how to access parks.
 
@@
 \subsection{Data}
-This study uses data available to the public from a variety of agencies and
-data providers.
 
@@
 
+This study uses data available to the public from a variety of federal and state
+data agencies. The datasets, as well as the analysis code, is available on GitHub
+at \url{https://github.com/boydnpb/park_access_new.git}
 The primary dataset is a geographic polygons shapefile of Census tracts in
@@
 features: baseball fields within Central Park, for example, are themselves
-unique greenspace. We elect to retain these independent features as they may
-generate additional activity.
+delineated green spaces. We elect to retain these independent features as they may
+generate additional physical activity participation aside from the containing 
+park.
 
@@
 \begin{equation}\label{eq:themodel}
- y_i = f(\vect{x}_i, CS_i, g(W, \vect{y} \vect{X}))
+ y_i = f(\vect{x}_i, CS_i, g(W, \vect{y}, \vect{X}))
 \end{equation}
 where $\vect{x}_i$ is a vector of tract-level sociodemographic attributes,
-$CS_i$ is a choice-based
-accessibility measure for the tract computed from Equation \ref{eq:cs} and
-$g(W, \vect{y}, X)$
-represents the impact of spatial neighborhood effects on the outcome.
+$CS_i$ is a choice-based accessibility measure for the tract computed from
+Equation \ref{eq:cs} and $g(W, \vect{y}, X)$ represents the impact of spatial
+neighborhood effects on the outcome.
 
@@
 common specifications \citep{Dubin1998}. In this study we specify that tracts
-with centroids located within 1.8 miles of each other are considered neighbors,
-and the strength of the relationship is the inverse of the distance between
-them:
+with population-weighted centroids located within 1.8 Euclidean miles of each 
+other are 
+considered neighbors, and the strength of the relationship is the inverse of 
+the distance between them:
 \begin{equation}\label{eq:weightsrule}
@@
 \end{equation}
-
 <<weights, warning = FALSE>>=
@@
 @
+This is a similar specification to then one \citet{tra_autoregression_paper} 
+used in a study of home prices in Atlanta. We selected this specification after 
+comparing overall model likelihoods between it and an adjacent (queen)
+tract specification.
 
@@
 estimated coefficients using size terms relevant to park acreage or amenities.
-Given this gap in the literature, we instead assert values for $\beta$ to generate
-the choice-based accessibility statistic used in this study.
+Given this gap in the literature, we instead assert plausible values for 
+$\beta$ to generate the choice-based accessibility statistic used in this study.
 
@@
 
-We initially estimated potential values of the $\beta$ coefficients by
-applying a limited-memory, box-constrained optimization algorithm maximizing the
-SDM model log-likehood \cite. The constraint required that the coefficient on
-size be positive and that on distance negative; all else equal, we assert people
-will prefer to use larger parks and parks that are nearer to their residence. This
-resulted in $\beta_d = -1.98$  and $\beta_s = 0.000$. Though the "most likely"
-parameter value, in this case it was unsatisfactory for two reasons: first,
-our prior assertion that park size is a valuable attribute of a park, and second
-the observation that in this specification tracts bordering on the corners of
-Central Park were measured with poor overall accessibility. we subsequently adjusted the $\beta_s$ value
-so that the
-
-to a value of $1.00$
-
 <<distances, warning=FALSE>>=
@@
 @
+
+The functional form of the accessibility statistic is adapted from 
+Equation \ref{eq:v_sd} by log-transforming first the distance (Euclidean, in miles) 
+between the park boundaries and population-weighted tract centroids and second the 
+park size (in acres).
+\begin{equation}\label{eq:v_logged}
+V_{ij} = \log(size_j)\lambda_s + \log(distance_{ij})\lambda_d
+\end{equation}
+We then calculate the consumer surplus $C$  (log-sum of the exponentiated 
+utilities) for each tract and standardize the result,
+\begin{equation}\label{eq:v_logged}
+C_i' = \frac{C_i - \bar{\vect{C}}}{sd(\vect{C})}
+\end{equation}
+
+We initially estimated potential values of the $\lambda$ coefficients by
+applying a limited-memory, box-constrained optimization algorithm 
+maximizing the model log-likehood.\footnote{Using the {\texttt nlm} function
+in R \citep{RCoreTeam2018}} The constraint required that the coefficient on
+size be positive and that on distance negative; all else equal, we assert people
+will prefer to use larger parks and parks that are nearer to their residence. This
+resulted in $\lambda_d = -1.98$  and $\lambda_s = 0.000$. Though the 
+``most likely''
+parameter value, in this case it was unsatisfactory for two reasons: first,
+our prior assertion that park size is a valuable attribute of a park, and second
+the observation that with these maximum likelihood values tracts bordering on 
+the corners of
+Central Park were measured with poor overall accessibility. We subsequently 
+adjusted the $\lambda_s$ value to $0.3$; the relative ratio of the coefficients 
+implies that all else equal, an individual would travel 6.6 times further to
+access a park twice as large.
 
@@
 @
+
+Figure \ref{fig:map} shows a map of New York City with each tract highlighted
+based on its relative accessibility score. The most continuous region of good park 
+access is in upper Manhattan and the Bronx, bracketed by Central Park and the
+Bronx River. Conversely, some of the poorest-accessibility areas are in 
+Brooklyn tracts not immediately adjacent to Prospect Park. Because the 
+accessibility statistic is normalized, the worst values are slightly below $-3$, 
+the best somehere above $3$.
 
@@
 \ref{subsec:accessibility}  added as a predictor. In both income specifications,
-the
-
-base model as
-well as models including logsum-based accessibility with and without Twitter
-activity. A few initial observations can be made from this table. First, adding
-the logsum measurement significantly improves the overall model fit, using a
-likelihood ratio test on the model log-likelihood statistics. Second, the
-coefficient estimates on the controlling variables remain virtually unchanged
-from the base model.
+the access logsum is positively correlated with physical activity rates; all else
+equal, people in tracts with higher accessibility to parks report higher rates of
+physical activity. The estimated value of this coefficient and its standard
+error is effectively the same with both representations of income, though in the
+log representation the estimated $t$-statistic technically passes the $0.1$ level.
+This is suggestive evidence only from a statistical perspective, but it deserves
+comment. Taking a compromise value of $0.10$ between the two models, the data
+suggest going from the least-accessible ($-3.0$) to the most-accessible ($3.0$) 
+tract implies a $0.6$ percentage-point rise in physical activity rates, though
+with a wide confidence interval around that estimate.
 
@@
                "Access logsum")
-texreg(pa_models, digits = 4, table = FALSE, single.row = TRUE,
+texreg(pa_models, digits = 4, table = FALSE,
        booktabs = TRUE, use.packages = FALSE,
-       custom.coef.names = termnames)
+       custom.coef.names = termnames,  stars = c(0.001, 0.01, 0.05, 0.1))
 @
@@
 obesity_models <- lapply(pa_models, function(model){
-  update(update(model, obesity ~ . ))
+  update(update(model, obesity ~ . + physact))
 })
@@
 @
+
 We now consider the impacts of a model where the dependent variable is
@@
 alongside the controlling variables and the accessibility logsum variable. 
+Table \ref{tab:obese_model} presents the estimated coefficients for both
+income specifications, with and without the accessibility logsum. As in the 
+physical activity models, the coefficients are largely significant and
+in the expected sign. The accessibility coefficient has practically the same 
+effect as in the physical activity model: moving from poor to excellent parks 
+access lowers the changes the expected obesity rate by around half a percentage
+point. In this case, however, the parameter $p$-value is lower than the 
+conventional $0.05$ level in each income representation.
 
@@
                "Access logsum")
-texreg(obesity_models, digits = 4, table = FALSE, single.row = TRUE,
+texreg(obesity_models, digits = 4, table = FALSE, 
        booktabs = TRUE, use.packages = FALSE,
-       custom.coef.names = termnames)
+       custom.coef.names = termnames,  stars = c(0.001, 0.01, 0.05, 0.1))
 @
@@
 
-\section{Discussion}
+\subsection{Additional Amenities}
 
-We readily acknowledge limitations in this study. To begin, it would be beneficial
-to have estimated the logsum coefficients from a high-quality survey rather than
-asserting them to match {\em a priori expectations of}
+As discussed when presenting the choice-based accessibility metric above,
+a key benefit it offers is the natural way in which park amenities beyond simply size
+can be accomodated. For example, the number of tweets emanating from a park
+might be seen as a measure of the park's popularity, or a proxy for its use. It 
+may be construed that parks with high popularity or use contain other amenities
+that encourage residents to use the park space, raising physical activity rates
+and lowering obesity rates beyond what would be expected with the park's size
+and proximity alone. It it reasonable to question whether park size and twitter
+activity are highly correlated, and thus would result in a double-counting of
+park size in the accessibility statistic. In our 
+data the number of tweets is positively but only loosely correlated  with park 
+size in our data ($\rho = \Sexpr{cor(parks$log_size, parks$log_tweets)}$), indicating
+that the Twitter data in fact could provide new information or proxy for the 
+desired amenities.
+
+To test this hypothesis, we can use the twitter data described earlier
+and suggested in Equation \ref{eq:v_sdt}. In this application, we again 
+log-transform all the components of utility:
+\begin{equation}\label{eq:v_logsdt}
+V_{ij} = \log(size_j)\lambda_s + \log(distance_{ij})\lambda_d + g(tweets_j)\lambda_t
+\end{equation}
+where $g(x)$ is a Yeo-Johnson transformation to preserve cases where $log(0)$
+would be otherwise undefined. As before, there is little information by which to
+determine the values of $\lambda$ in this utility specification; we use the
+previous values with the addition of $\lambda_t = 0.1$. The ratio between the 
+utility coefficients implies residents would travel almost twenty times as far or
+use a park three times smaller if it had twice the twitter activity, all else 
+equal.
+
+Figure \ref{fig:tweets_map} shows the percent change in the normalized log
+sum statistic after including the twitter infromation. There is not a recognizable
+overarching pattern to the changes, though tracts immediately surrounding 
+Prospect Park and Central Park gain even more relative accessibility, and tracts
+in midtown Manhattan, Queens, and southern Bronx tend to lose some.
+
+\begin{figure*}
+
+<<tweets_map, echo=FALSE, message=FALSE, warning=FALSE>>=
+library(ggmap)
+map <- ggmap(get_map(unname(st_bbox(tracts_sf)), source = "stamen", maptype = "toner-lite"),
+             extent = "device")
+map +
+  geom_sf(data = tracts_sf %>% 
+            mutate(diff = tweets_ls - access_ls, 
+                   pct_diff = diff / access_ls * 100), 
+          inherit.aes=FALSE,
+          aes(fill = cut(pct_diff, quantile(pct_diff, 
+                                            probs = seq(0, 1, by = 0.1)))),
+          alpha = 0.8, color = NA) +
+  scale_fill_brewer("Percent Change", palette = "PiYG")
+@
+  \caption{Percent change in relative accessibility when accounting for Twitter 
+  activity.}
+  \label{fig:tweets_map}
+\end{figure*}
+
+<<tweets_models>>=
+tweets_models <- list(pa_models[[3]], pa_models[[4]], 
+                      obesity_models[[3]], obesity_models[[4]]) %>%
+  lapply(function(model){
+    update(update(model, . ~ . -access_ls + tweets_ls))
+  })
+@
+
+Table \ref{tab:tweets_models} presents the estimated coefficients relating
+accessibility to obesity and physical activity rates, with and without the 
+inclusion of Twitter data. In all four cases, including twitter information in
+the accessibility statistic tempers the strength of the relationship with the
+dependent variable and modestly widens the standard error without substantively
+affecting the significance of the test statistics.  
+From this limited example,
+it appears Twitter activity does not enhance the ability of parks to induce 
+physical activity rates or lower obesity rates, though further investigation 
+of this and other park amenities would certainly be warranted.
+
+<<tab:tweets_models, results="asis">>=
+names(tweets_models) <- c("Physical Activity: High and Low Income", 
+                          "Physical Activity: log(Income)",
+                          "Obesity: High and Low Income", "Obesity: log(Income)")
+names(pa_models)[3:4] <- c("Physical Activity: High and Low Income", 
+                           "Physical Activity: log(Income)")
+names(obesity_models)[3:4] <- c("Obesity: High and Low Income", 
+                                "Obesity: log(Income)")
+lapply(c(tweets_models, pa_models, obesity_models), function(model){
+  s <- summary(model)
+  as_data_frame(s$Coef) %>%
+    mutate(term = rownames(s$Coef)) %>%
+    select(term, estimate = Estimate, std.error = `Std. Error`, 
+           statistic = `z value`, p.value = `Pr(>|z|)`)
+}) %>%
+  bind_rows(.id = "model") %>%
+  filter(term %in% c("tweets_ls", "access_ls")) %>%
+  separate(model, c("y", "income"), ": ") %>%
+  mutate(term = ifelse(term == "tweets_ls", 
+                       "Tweets, Size and Distance", "Size and Distance")) %>%
+  arrange(desc(y), income, term) %>%
+  select(income, term, estimate, std.error, p.value) %>%
+
+  kable(escape = FALSE, digits = 4,
+        table.envir = "table*", booktabs = TRUE,
+        caption = "Estimated Relationship between Accessiblity and Outcomes",
+        col.names = c("", "Accessibility Components", "Estimate", "Std. Error",
+                      "$p$-value")) %>%
+  kable_styling() %>%
+    group_rows("$y = $Physical Activity", 1, 4) %>%
+    group_rows("$y = $Obesity", 5, 8)
+@
+
+\section{Limitations and Future Research Direction}
+
+We readily acknowledge limitations in this study. As in any study conducted 
+with areal data, we at risk of falling victim to the ecological inference 
+fallacy, where aggregate statistics mask or contradict disaggregated or 
+individual-level trends. A large-sample survey of individuals in New York City,
+including measured physical activities and obesity would always be prefferable
+to the tract-level data used in this study. It would be also be
+preferrable to have obtained the $\lambda$ accessibility utility coefficients 
+from a high-quality survey of leisure destination choice rather than
+asserting them to match our prior expectations of what constitutes quality
+park access. An ideal survey to address the question would incorporate both sets
+of questions: physical activity and health data on one hand and park use (including
+which parks were used and how frequently) on the other. As no such dataset
+exists to our knowledge, this tract-level aggregate analysis with asserted 
+utility coefficients is the possibility that remains.
+
+This paper presents a holistic accessibility statistic that could, in theory,
+accommodate many attributes of the destination parks as well as the people who 
+might use them. As an illustration: the park-going population could be separated
+into at least four delineated clusters, each preferring different amenities of a 
+park: 
+\begin{itemize}
+ \item{runners and cyclists: long, interesting trail systems}
+ \item{sports players: soccer fields, basketball courts, or baseball diamonds}
+ \item{families with small children: water features and playgrounds}
+ \item{casual users: water features, gardens, performances, etc.}
+\end{itemize} 
+An analyst could then compute the accessibility logsum for each cluster with
+different utility values for each park's amenities, and obtain a measure of
+a neighborhood's accessibility to park features that its residents most care 
+about. In this paper, we proceed only incrementally beyond this theory by 
+adding Twitter activity as an an element of a park's attractiveness above and
+beyond its size. Exploring additional amenities or market segmentation strategies
+could provide a more definitive understanding of the relationship between park 
+accessibility and health outcomes.
+
+Finally, this study is primarily focused on the hypothesis that accessibility to
+parks encourages physical activity, which in turn reduces obesity. There are a 
+multitude of other hypotheses that might be proposed and tested with the basic
+methodology we have employed here, or competing explanations for the outcomes we
+have observed. It is distinctly possible, for instance, that individuals who wish
+to exercise regularly in parks choose to live near them; this could potentially
+explain why the observed effects of accessibility seem partially dependent on
+the representation of income in our econometric model. Controlling for this
+potential self-selection effect would be necessary to isolate the exogenous
+impacts of park access on obesity or other health outcomes. And regarding these
+other variables: this study did not consider potential relationships between park
+access and hospitalization rates, life expectancy, respiratory disease, 
+mental health, or any number of potential beneficial outcomes hypothesized 
+or explored in the existing literature. Exploring these connections and their
+underlying mechanisms should be a priority as city planners and urban architects
+attempt to improve the quality of life of urban residents in the future.
 

### 2019-02-13 manuscript.Rnw
First complete draft for team review.

@@
 \begin{abstract}
-This research identifies the correlation between size and access to urban parks
-and physical activity and obesity outcomes at the neighborhood level. Proximity
-to parks is associated with increased physical activity and reduced obesity, but
-little research has been conducted on the relationship between accessibility to
-parks and health outcomes. Using data for three urban areas, we created a new
-measure for access to parks called ‘park choice accessibility.’ Park choice
-accessibility uses a gravity model to interact distance to parks and the quality
-of those parks as defined by their size. A small park very close to a
-neighborhood can have a major impact, but a larger park at a similar distance
-may have an even larger impact. Similarly, a large park can be further away and
-still have an impact on health outcomes. Using spatial econometric analysis, we
-assess whether park choice accessibility is associated with increased physical
-activity or decreased prevalence of obesity at the neighborhood level. The
-analysis controls for socioeconomic covariates such as age, marital status,
-income, and educational attainment.
+Recent research has underscored the potential for public green spaces to 
+influence individual and societal health outcomes, but 
+empirical measurements of such influences have yielded mixed results to date, with
+particular disagreement surrounding how access to parks ought to be defined while
+controlling alternate explanations. In this paper we present a comprehensive 
+measure of park accessibility drawn from random utility choice theory, which avoids
+arbitrary assertions of proximity while incorporating potentially numerous 
+amenities and attributes of both the parks and the population. We apply this 
+choice-based accessibility measure to Census tract-level health and 
+socioeconomic data in New York City, paired with geographic parks data and  
+and geolocated data collected from the Twitter API. Controlling for the 
+socioeconomic variables and spatially correlated error terms, we show a positive
+but suggestive relationship between park access and physical activity rates, and
+a clear negative relationship between park access and obesity rates.
 \end{abstract}
@@
 slow development, including cardiovascular diseases, chronic respiratory
-diseases, diabetes, stroke, joint and bone diseases, and cancer \citep{WHO2014}
+diseases, diabetes, stroke, joint and bone diseases, and cancer.
 Recent statistics suggest that approximately 300,000 premature deaths each year
@@
 effective strategy for reducing and managing obesity and many of the
-aforementioned chronic diseases \citet{CDC2009, Durstine2013},
-\citet{Wolf2008} notes that “more than 50 percent of U.S. adults do not get enough
-physical activity to provide health benefits; 24 percent are not active at all
-in their leisure time. Activity decreases with age and sufficient activity is
-less common among women than men, and among those with lower incomes
-and less education” (p. 22). As the epidemiological transition from
+aforementioned chronic diseases \citep{CDC2009, Durstine2013},
+\citet{Wolf2008} observed that a large portion of U.S. adults do not participate
+in sufficient physical activity, a problem compounded by age and socioeconomic status. As the epidemiological transition from
 infectious to chronic diseases is now complete in the developed world,
@@
 activity, they are increasingly looking to public spaces like parks and
-greenways as key elements of the built environment that can support exercise
-\citep{Bedimo-Rung2005a}.
+greenways as key elements of the built environment that can support physical 
+activity \citep{Bedimo-Rung2005a}.
 
-In this paper, we present a holistic and flexible measurement for park
-accessibility based in random utility choice theory. This measure can consider 
-the continuous
-distance to all parks in the region, weighted by the size of the park and other
-amenities. We apply this measurement to study the link between park
+In spite of this interest and the numerous studies attempting to empirically 
+estimate the strength of such support, evidence to date has been mixed. 
+A major challenge researchers have faced is a proliferation of techniques for
+measuring access to green spaces, of which many are plausible but lack 
+analytical or theoretical basis. In this paper, we present a holistic 
+and flexible measurement for park accessibility based in random utility 
+choice theory. This choice-based accessibility measure compares the continuous
+distance to all parks in the region, weighted against the size of the park and 
+its other amenities. We apply this measurement to study the link between park
 accessibility and attractiveness and Census tract-level aggregate physical
@@
 least park-accessible tracts have an expected physical activity participation
-rate XX percent lower than the most accessible tracts, though this result is 
-only suggestive of correlation from a statistical perspective.
-However, we also find a strong and significant negative correlation between 
+rate half a percentage point lower than the most accessible tracts, though 
+this result is only suggestive of correlation from a statistical perspective.
+However, we also find a strong and significant correlation between 
 increased park accessibility and decreased obesity rates, again controlling for
@@
 be expanded to account for other elements of a park's attractiveness by 
-introducing the number of tweets emanating from the park during 
+introducing the number of tweets emanating from within a park during 
 (WE NEED TO DESCRIBE THE PERIOD). The paper concludes with a discussion of
@@
 between green space and physical activity is somewhat mixed.
-\citet{Wolf2008} report
-residents self-identify as more active and attaining a higher quality of life when
-a greater portion of the environment is greenspace.
-\citet{Larson2016}
-used self-reported scores on the Gallup-Healthways Wellbeing
-Index (citation) to evaluate the relationship between different areas of wellbeing,
+\citet{Wolf2008} reports that people self-identify as more active and attain a 
+higher quality of life when a greater portion of their environment is greenspace.
+\citet{Larson2016} similarly used self-reported scores on the Gallup-Healthways Wellbeing
+Index to evaluate the relationship between different areas of wellbeing,
 including physical health, and park quantity, quality, and accessibility in 44
@@
 activity and a significant, negative association between park density and
-obesity.
-Conversely, \citet{Richardson2012} examined the
+obesity. Conversely, \citet{Richardson2012} examined the
 relationship between urban greenspace and selected mortality rates, and though the authors
@@
 car accidents, they did find that all-cause mortality was significantly higher in
-greener cities.
-In a meta-analysis of 20 peer reviewed journal articles
+greener cities. In a meta-analysis of 20 peer reviewed journal articles
 exploring the relationship between parks and objectively measured physical
@@
 hypothesis that park access and use confer health benefits stemming from
-increased physical activity and attendant decreases in obesity, all of the
-studies -- except some included in the meta-analysis by \citet{Bancroft2015} -–
-were conducted at the level of the city or metropolitan statistical area.
+increased physical activity and attendant decreases in obesity, the large majority
+of extant research has been conducted at the level of the city or 
+metropolitan statistical area.
 Metropolitan-level analyses do not capture the within-city variation in
@@
 consensus on the most appropriate accessibility metric to use in a given
-setting, noting that ''ways to measure active accessibility are as varied as the
-number of scholars the measure them.'' This is at least partly because
+setting, noting that ``ways to measure active accessibility are as varied as the
+number of scholars the measure them.''(p. 227) This is at least partly because
 each method has inherent strengths and weaknesses.
@@
 Infrastructure-based distances can vary widely
-depending on which modes of travel are considered. a highway network may not
+depending on which modes of travel are considered: a highway network may not
 adequately capture trails and bike paths.
@@
 uses when considering the attractiveness or impedance of a destination; the
-findings of \citet{Pretty2005} and other suggest unpleasant pedestrian routes
+findings of \citet{Pretty2005} and others suggest unpleasant pedestrian routes
 attenuate perceived travel costs.
@@
 A key theoretical outcome of random utility choice models is that the
-consumer surplus, or the total weighted value of all alternatives of the choice set, can
-be obtained as the log-sum of the denominator of the choice probability equation
-\citep{Small1981}:
+consumer surplus to individual $i$, or the total weighted value of all 
+alternatives in individual $i$'s choice set, can
+be obtained as the logarithm of the sum in the denominator of the choice 
+probability equation \citep{Small1981}:
 \begin{equation}\label{eq:cs}
-CS_i = \ln\left({\sum_{p \in J}\exp(V_{ip})}\right)
+C_i = \ln\left({\sum_{p \in J}\exp(V_{ip})}\right)
 \end{equation}
-There are several advantages to a log-sum defined metric relative to
+There are several advantages to such a log-sum defined metric relative to
 buffer-based accessibility metrics more commonly found in the literature. First,
 all individuals are defined as having some access to all parks, rather than an
-arbitrary cutoff asserted by the researcher. This allows for the fact that some
-people are more or less sensitive to distances, and that distance is a
-continuous, non-binary phenomenon. Second, the random utility formulation
-allows the researcher to include any attribute of the park; in this case, we
-consider the size of the park as an element of accessibility. This suggests that
+arbitrary cutoff asserted by the researcher. 
+This allows for the fact that some people are more or less sensitive to 
+distances, and that distance is a continuous, non-binary phenomenon. 
+It defies reason to assume person living 1.1 miles 
+from a park has functionally different accessibility than someone living at 0.9 miles.
+Second, the random utility formulation
+allows the researcher to include any attribute of the park as part of its utility
+specification; in this case, we consider the size of the park weighted against its
+distance as an element of relative accessibility. This suggests that
 not all parks are equal, and that a large park such as Central Park on Manhattan
@@
 Note that this formulation is an extension of the gravity-based accessibility
-statistics, in the same way that the gravity model is a specific case of
+statistics, in the same way that the gravity-based trip distribution model model 
+is a specific case of
 a destination-choice model \citep{Daly1982}. But the extension is meaningful, and
@@
 
-In spite of its advantages, logsum-based accessibilities have not received
-as much application in the accessibility literature as distance-based or even
-gravity-based measures. They are commonly used, however, in alternatives analyses of
+In spite of its flexibility and basis in choice theory, accessibility measures 
+based in log-sums
+have not received much application in the accessibility literature compared with
+distance-based or even
+gravity-based measures. \citet{Zhang2011} present a metric with a similar functional
+form, though the authors arrive at it from a motivation other than destination 
+choice theory. Log-sums are commonly used, however, in alternatives analyses of
 transit infrastructure improvements \citep{DeJong2007}. A reason for this is
-likely that such alternatives analyses are usually conducted in the context of a travel
-demand model, where calibrated and multimodal logsums are readily available
-\citep{Geurs2010}.
+likely that a regional travel demand model is available to the analysts, thus making 
+calibrated and multimodal logsums are readily available \citep{Geurs2010}.
 
@@
 This study uses data available to the public from a variety of federal and state
-data agencies. The datasets, as well as the analysis code, is available on GitHub
-at \url{https://github.com/boydnpb/park_access_new.git}
+data agencies. The datasets, as well as the analysis code, are available on GitHub
+at \url{https://github.com/boydnpb/park_access_new.git}.
 The primary dataset is a geographic polygons shapefile of Census tracts in
@@
 After removing tracts with missing population information or outlying and unusual
-characteristics, we have Sexpr{nrow(tracts)} complete cases. Table
+characteristics, we have \Sexpr{nrow(tracts)} complete cases. Table
 \ref{tab:tractsdata} presents key descriptive statistics for these data.
@@
 
-In a destination choice framework, the tracts represent the "origins" and
-"destinations" are parks in New York City.
+In a destination choice framework, the tracts represent the ``origins'' and
+``destinations'' are parks in New York City.
 We retrieved a shapefile of public parks and greenspaces within New York City’s
@@
 \begin{equation}\label{eq:themodel}
- y_i = f(\vect{x}_i, CS_i, g(W, \vect{y}, \vect{X}))
+ y_i = f(\vect{x}_i, C_i, g(W, \vect{y}, \vect{X}))
 \end{equation}
 where $\vect{x}_i$ is a vector of tract-level sociodemographic attributes,
-$CS_i$ is a choice-based accessibility measure for the tract computed from
+$C_i$ is a choice-based accessibility measure for the tract computed from
 Equation \ref{eq:cs} and $g(W, \vect{y}, X)$ represents the impact of spatial
@@
 should be used because its estimates are unbiased. If the test cannot reject
-equivalents, the SEM is more efficient.
+equivalence, the SEM is more efficient.
 
@@
 @
-This is a similar specification to then one \citet{tra_autoregression_paper} 
-used in a study of home prices in Atlanta. We selected this specification after 
+This is a similar specification to that used by  \citet{tra_autoregression_paper} 
+in a study of home prices in Atlanta. We selected this specification after 
 comparing overall model likelihoods between it and an adjacent (queen)
@@
 the corners of
-Central Park were measured with poor overall accessibility. We subsequently 
+Central Park were measured with poor overall accessibility, a unacceptably 
+unintuitive result. We subsequently 
 adjusted the $\lambda_s$ value to $0.3$; the relative ratio of the coefficients 
@@
 accessibility statistic is normalized, the worst values are slightly below $-3$, 
-the best somehere above $3$.
+the best somewhere above $3$.
 
@@
           alpha = 0.8, color = NA) +
-  scale_fill_brewer("Relative Park Logsum", palette = "GnBu")
+  scale_fill_brewer("Normalized Log-sum", palette = "GnBu")
 @
-  \caption{Relative surplus-based accessibility in New York City}
+  \caption{Normalized choice-based park accessibility log-sum values in New York City}
   \label{fig:map}
@@
 The other two models presented in Table \ref{tab:pa_models} include
-the log-sum accessibility statistic described in Section
-\ref{subsec:accessibility}  added as a predictor. In both income specifications,
+the log-sum accessibility statistic described in above added as a predictor. 
+In both income specifications,
 the access logsum is positively correlated with physical activity rates; all else
@@
 alongside the controlling variables and the accessibility logsum variable. 
-Table \ref{tab:obese_model} presents the estimated coefficients for both
+Table \ref{tab:obese_models} presents the estimated coefficients for both
 income specifications, with and without the accessibility logsum. As in the 
@@
 and lowering obesity rates beyond what would be expected with the park's size
-and proximity alone. It it reasonable to question whether park size and twitter
+and proximity alone. It is reasonable to question whether park size and Twitter
 activity are highly correlated, and thus would result in a double-counting of
@@
 data the number of tweets is positively but only loosely correlated  with park 
-size in our data ($\rho = \Sexpr{cor(parks$log_size, parks$log_tweets)}$), indicating
-that the Twitter data in fact could provide new information or proxy for the 
+size in our data ($\rho = \Sexpr{round(cor(parks$log_size, parks$log_tweets), 3)}$), 
+indicating that the Twitter data in fact could provide new information or proxy for the 
 desired amenities.
@@
 where $g(x)$ is a Yeo-Johnson transformation to preserve cases where $log(0)$
-would be otherwise undefined. As before, there is little information by which to
+would be otherwise undefined \citep{Yeo2000}. As before, there is little 
+information by which to
 determine the values of $\lambda$ in this utility specification; we use the
@@
 utility coefficients implies residents would travel almost twenty times as far or
-use a park three times smaller if it had twice the twitter activity, all else 
+use a park three times smaller if it had twice the Twitter activity, all else 
 equal.
@@
 @
-  \caption{Percent change in relative accessibility when accounting for Twitter 
+  \caption{Percent change in relative accessibility to parks when accounting for Twitter 
   activity.}
@@
 
-Table \ref{tab:tweets_models} presents the estimated coefficients relating
+Table \ref{tab:tweets_coefs} presents the estimated coefficients relating
 accessibility to obesity and physical activity rates, with and without the 
@@
 affecting the significance of the test statistics.  
-From this limited example,
-it appears Twitter activity does not enhance the ability of parks to induce 
-physical activity rates or lower obesity rates, though further investigation 
-of this and other park amenities would certainly be warranted.
+From this limited example, we cannot say whether Twitter activity enhances the 
+ability of parks to induce physical activity rates or lower obesity rates, 
+though further investigation of this and other park amenities would certainly 
+be warranted.
 
-<<tab:tweets_models, results="asis">>=
+<<tweets_coefs, results="asis">>=
 names(tweets_models) <- c("Physical Activity: High and Low Income", 
@@
     mutate(term = rownames(s$Coef)) %>%
-    select(term, estimate = Estimate, std.error = `Std. Error`, 
+    dplyr::select(term, estimate = Estimate, std.error = `Std. Error`, 
            statistic = `z value`, p.value = `Pr(>|z|)`)
@@
   arrange(desc(y), income, term) %>%
-  select(income, term, estimate, std.error, p.value) %>%
+  dplyr::select(income, term, estimate, std.error, p.value) %>%
 
@@
   kable_styling() %>%
-    group_rows("$y = $Physical Activity", 1, 4) %>%
-    group_rows("$y = $Obesity", 5, 8)
+    group_rows("y = Physical Activity", 1, 4) %>%
+    group_rows("y = Obesity", 5, 8)
 @
@@
 to the tract-level data used in this study. It would be also be
-preferrable to have obtained the $\lambda$ accessibility utility coefficients 
+preferrable to have obtained the $\vect{\lambda}$ accessibility utility coefficients 
 from a high-quality survey of leisure destination choice rather than
@@
 
+As a travel impedance measure, we used the Euclidean distance between each tract's 
+population-weighted centroid and the nearest point on the edge of a park's border.
+Euclidean distances have well-rehearsed limitations regarding their fidelity with 
+the underlying infrastructure network, etc.. Network-based distances can also
+suffer from challenges when applied to multimodal problems; these challenges are
+exacerbated when the non-highway mode share is high, as is likely when considering
+access to parks. A better metric of travel impedance may be a mode choice model
+log-sum, which weights all travel alternatives against each other. A
+
 Finally, this study is primarily focused on the hypothesis that accessibility to

### 2019-02-21 manuscript.Rnw
Rebuild data
Instead of using the data file put together over several iterations, we're going to load up (almost) everything inside the document. This commit changes the tracts data frame from a geojson file to calls to the CDC and Census API.

@@
 library(texreg)
+library(RSocrata)
+library(tidycensus)
+
+options(tigris_use_cache = TRUE)
 opts_chunk$set(cache=TRUE, message = FALSE, echo = FALSE)
@@
 
-<<tracts>>=
-tracts <- geojson_read("data/nyc_tracts.geojson", what = "sp")
-
-# identify complete cases
-complete_tracts_index <- tracts@data %>%
-  tbl_df() %>%
-  complete.cases()
-tracts <- tracts[complete_tracts_index, ]
-
-tracts@data <-  tbl_df(tracts@data) %>%
-  dplyr::transmute(
+<<acs, results='hide'>>=
+# The study area includes the five boroughs of New York City
+counties <- c("081", "047", "061", "005", "085")
+names(counties) <- c("Queens", "Brooklyn", "Manhattan", "Bronx", "Staten Island")
+
+# We need to get lots of variables from the American Community Survey 5-year data.
+# This vector defines the variable names as well as the table/cell in the Census
+# API that each variable is drawn from. The comments help to identify which
+# table name and type they come from.
+variables <- c(
+  "population" = "B25008_001", # TOTAL POPULATION IN OCCUPIED HOUSING UNITS BY TENURE
+  "housing_units" = "B25001_001", # HOUSING UNITS
+  "households" = "B19001_001", #HOUSEHOLD INCOME IN THE PAST 12 MONTHS (IN 2017 INFLATION-ADJUSTED DOLLARS)
+  # Employment status
+  "men_fulltime" = "B23022_004",
+  # Estimate!!Total!!Male!!Worked in the past 12 months!!Usually worked 35 or more hours per week
+  "women_fulltime" = "B23022_028",
+  #Estimate!!Total!!Female!!Worked in the past 12 months!!Usually worked 35 or more hours per week
+  # LIVING ARRANGEMENTS OF ADULTS 18 YEARS AND OVER BY AGE
+  # everyone who
+  "adults" = "B09021_001",
+  "lives_alone" = "B09021_002",
+  "child_of_hhr" = "B09021_005",
+  "other_rel" = "B09021_006",
+  "other_nonrel" = "B09021_007",
+  # EDUCATIONAL ATTAINMENT BY EMPLOYMENT STATUS FOR THE POPULATION 25 TO 64 YEARS
+  "college" = "B23006_016",
+  # RACE
+  "black" = "B02001_003",
+  "asian" = "B02001_005",
+  "pacific" = "B02001_006",
+  "nativeam" = "B02001_004",
+  # HISPANIC OR LATINO ORIGIN BY SPECIFIC ORIGIN
+  # The number of hispanic individuals needs to be drawn from a different table.
+  "hispanic" = "B03001_003",
+  #MEDIAN HOUSEHOLD INCOME IN THE PAST 12 MONTHS (IN 2017 INFLATION-ADJUSTED DOLLARS)
+  "income" = "B19013_001",
+  #HOUSEHOLD INCOME IN THE PAST 12 MONTHS (IN 2017 INFLATION-ADJUSTED DOLLARS)
+  "inc_0010" = "B19001_002",  "inc_1015" = "B19001_003", "inc_1520" = "B19001_004",
+  "inc_2025" = "B19001_005", "inc_2530" = "B19001_006", "inc_3035" = "B19001_007",
+  "inc_125"  = "B19001_015", "inc_150"  = "B19001_016", "inc_200"  = "B19001_017",
+  #AGE BY RATIO OF INCOME TO POVERTY LEVEL IN THE PAST 12 MONTHS.
+  # we really only need the numbers of individuals in each age group, and the groups
+  # are more coarse here than in the sex by age table.
+  "age_0005" = "B17024_002", "age_0611" = "B17024_015", "age_1217" = "B17024_028",
+  "age_1824" = "B17024_041", "age_2534" = "B17024_054",
+  "age_3544" = "B17024_067", "age_4554" = "B17024_080", "age_5564" = "B17024_093",
+  "age_6574" = "B17024_106",
+  "age_75up" = "B17024_119"
+)
+
+acs <- get_acs(geography = "tract", variables = variables, year = 2017,
+               state = "NY", county = counties, geometry = TRUE) %>%
+  select(-moe) %>%
+  spread(variable, estimate) %>%
+  filter(adults > 0) %>%
+  mutate(area = as.numeric(st_area(geometry) * 1e-6)) %>%
+  # area is in m^2, change to km^2
+  transmute(
     geoid = GEOID,
@@
     ),
-    obesity = OBESITY,
-    physact = 100 - Phys_Act,
-    log_obesity = log(OBESITY),
-    log_physact = log(100 - Phys_Act),
-    density = Pop_Density,
-    fulltime = FulltimeWork, college = CollegeDeg,
-    single = Single_Percent,
-    black = PctBlack, asian = PctAsian, hispanic = PctHispanic,
-    other = PctNative + PctPacific,
-    Pct0to17, Pct18to29, Pct65plus,
-    income = NYC_Income,
-    lowincome = Income1 + Income2 + Income3,
-    highincome = Income8 + Income9 + Income10
+    population, households, housing_units, adults,
+    density = households / area,
+    income,
+    # many of the variables come in raw counts, but we want to consider
+    # them as shares of a relevant denominator.
+    lowincome    = 100 * (inc_0010 + inc_1015 + inc_1520 + inc_2530 +
+                            inc_3035) / households,
+    highincome   = 100 * (inc_125 + inc_150 + inc_200) / households,
+    fulltime     = 100 * (men_fulltime + women_fulltime) / adults,
+    college      = 100 * college / adults,
+    single       = 100 * (lives_alone + child_of_hhr + other_rel +
+                            other_nonrel) / adults,
+    youth        = 100 * (age_0005 + age_0611 + age_1217) / population,
+    young_adults = 100 * (age_1824 + age_2534) / population,
+    seniors      = 100 * (age_6574 + age_75up) / population,
+    black        = 100 * black / population,
+    asian        = 100 * asian / population,
+    hispanic     = 100 * hispanic / population,
+    other        = 100 * (nativeam + pacific) / population
   )
+
+@
+
+<<cdc500>>=
+# The CDC 500 cities project makes its data available through the Socrata API.
+# The url endpoints for each data object are different, though we have identified
+# them as below. We pre-filter the data to New York to save ourselves and the
+# API a bit of time.
+cdc_urls <- list(
+  "physact" = "https://chronicdata.cdc.gov/resource/axqz-rkkp.json?stateabbr=NY",
+  "obesity" = "https://chronicdata.cdc.gov/resource/bjvu-3y7d.json?stateabbr=NY"
+)
+
+# For each URL endpoint, get the data through Socrata (you will need your own
+# Socrata API key), filter to the FIPS codes we specified above, and trim
+# to only the data values we want to keep.
+cdc500 <- lapply(cdc_urls, function(url){
+  read.socrata(url, app_token = Sys.getenv("CDC_KEY"),
+               stringsAsFactors = FALSE) %>%
+    tbl_df() %>%
+    filter(tractfips %in% acs$geoid) %>%
+    transmute( geoid = tractfips, value = as.numeric(data_value ), year = year)
+}) %>%
+  bind_rows(.id = "measure") %>%
+  spread(measure, value) %>%
+  mutate(physact = 100 - physact)
+
+@
+
+<<tracts>>=
+tracts <- left_join(acs, cdc500, by = "geoid") %>%
+  filter(!is.na(obesity), !is.na(income))
 @
@@
 <<tractsdata, results="asis", message=FALSE, warning=FALSE, cache=FALSE>>=
+# This is an index pairing the variable names with the order we want them
+# to be shown in in the table.
 tracts_data_index <- data_frame(
   variable =  c(
-    "obesity", "physact", "income", "lowincome", "highincome",
+    "obesity", "physact", "income",
     "density", "fulltime", "college", "single",
-    "Pct0to17", "Pct18to29",  "Pct65plus", "black", "asian", "hispanic",
-    "other")
-  ) %>% mutate(order = row_number())
-
-tract_summary <- tracts@data %>%
-  dplyr::select(obesity:highincome) %>%
+    "youth", "young_adults",  "seniors", "black", "asian", "hispanic",
+    "other"),
+  source = c("CDC 500 Cities", "CDC500 Cities",
+             rep("ACS", 12))
+) %>%
+  mutate(order = row_number())
+
+# Get the tract SE / health data, calculate group statistics, and get it
+# in the right order
+tract_summary <- tracts %>%
+  st_set_geometry(NULL) %>%
+  tbl_df() %>%
+  dplyr::select(tracts_data_index$variable) %>%
   gather(variable, value) %>%
@@
     variable = c(
-      "Obesity Rate", "Physical Activity Rate",
-      "Median tract income",
-      "Share of low income",
-      "Share of high income",
-      "Density: population per square mile",
-      "Full-time workers", "College-educated",
-      "Single Adults", "Share of children (0-17)", "Share of young adults (18-29)",
-      "Share of seniors (65+)", "Black population share", "Asian population share",
-      "Hispanic population share", "Other Minorities")
+      "Obesity: Share of adults over 18 who are obese",
+      "Physical Activity: Share of adults over 18 who engage in some leisure-time physical activity",
+      "Income: Median tract income",
+      "Density: Households per square kilometer",
+      "Fulltime: Share of adults over 18 with full-time work",
+      "College: Share of adults over 24 with a college degree",
+      "Single: Share of adults over 18 living alone or in a non-partnership household",
+      "Youth: Share of population under 18",
+      "Young adults: Share of population between 18 and 34",
+      "Seniors: Share of population who are 65 and over",
+      "Black: Share of population who is black",
+      "Asian: Share of population who is Asian",
+      "Hispanic: Share of population who is Hispanic",
+      "Other: Share of population who belong to other minority groups")
   ) %>%
@@
   ) %>%
-  mutate( variable = c("Park Size in Acres", "Tweets Emanating from Park"))
+  mutate( 
+    variable = c("Size: Park size in acres", 
+                 "Tweets: Tweets emanating from park in September 2014"),
+    source = c("NYC", "Twitter API")
+    )
 
 bind_rows(tract_summary, park_summary) %>%
+  separate(variable, c("var", "desc"), sep = ": ") %>%
   kable(escape = FALSE, digits = 2,
@@
         caption = "Descriptive Statistics of Tract and Park Variables",
-        col.names = c(" ", "Minimum", "Median (IQR)", "Maximum")) %>%
+        col.names = c(" ", "Description", "Minimum", "Median (IQR)", "Maximum", "Source")) %>%
   kable_styling() %>%
-  group_rows(paste("Tract Variables, N =", nrow(tracts@data)), 1, 16) %>%
-  group_rows(paste("Park Variables, N =", nrow(parks)), 17, 18)
+  group_rows(paste("Tract Variables, N =", nrow(tracts)), 1, 14) %>%
+  group_rows(paste("Park Variables, N =", nrow(parks)), 15, 16)
 
@@
 <<weights, warning = FALSE>>=
-pop_weighted_centroids <- read_csv("data/CenPop2010_Mean_TR36.csv") %>%
-  mutate(geoid = str_c(STATEFP, COUNTYFP, TRACTCE))
-pop_weighted_centroids <- left_join(tracts@data, pop_weighted_centroids, by = "geoid") %>%
-  dplyr::select(geoid, LONGITUDE, LATITUDE) %>%
+# Instead of using the centroid of the tract geometry, we want to use the 
+# tract's population-weighted centroid.
+pop_weighted_centroids <- tracts %>% tbl_df() %>%
+  dplyr::select(geoid) %>%
+  left_join(
+    read_csv("data/CenPop2010_Mean_TR36.csv") %>%
+      transmute(geoid = str_c(STATEFP, COUNTYFP, TRACTCE), LATITUDE, LONGITUDE)
+  ) %>%
+  # convert to sf and project into the feet-based projection used in the
+  # parks dataset.
   st_as_sf(crs = 4326, coords = c("LONGITUDE", "LATITUDE")) %>%
@@
 <<logsum_betas>>=
-tracts@data <- tracts@data %>%
+# This is where we actually calculate the accessibility statistic, using the
+# distances matrix between tracts (in the rows) and parks (in the columns).
+tracts <- tracts %>%
   mutate(
-    access_ls = calculate_park_logsums(log(distances), parks$log_size,
-                                       betas = c(-1.98, 0.3)),
-    tweets_ls = calculate_park_logsums(log(distances), parks$log_size, parks$log_tweets,
+    # Log-sum based utilities, without and with Twitter data
+    access_ls = calculate_park_logsums(
+      log(distances), parks$log_size, betas = c(-1.98, 0.3)),
+    tweets_ls = calculate_park_logsums(
+      log(distances), parks$log_size, parks$log_tweets,
       betas = c(-1.98, 0.3, 0.1)),
-    access_ls = (access_ls - mean(access_ls) )/ sd(access_ls),
-    tweets_ls = (tweets_ls - mean(tweets_ls) )/ sd(tweets_ls)
+    # Normalize the logsum onto the Z-scale
+    access_ls = (access_ls - mean(access_ls) ) / sd(access_ls),
+    tweets_ls = (tweets_ls - mean(tweets_ls) ) / sd(tweets_ls)
   )
 
-tracts_sf <- tracts %>% st_as_sf() %>% st_transform(4326)
+# also calculate binary accessibility with a point-in-buffer analysis
+# on the population-weighted centroids and the parks
+parks_025_buffer <- st_buffer(parks, 0.25 * 5280)
+inside_025 <- pop_weighted_centroids %>%
+  st_join(parks_025_buffer %>% select(size),
+          st_within, largest = TRUE) %>%
+  filter(!is.na(size))
+
+tracts <- tracts %>%
+  mutate(access_025 = ifelse(geoid %in% inside_025$geoid, TRUE, FALSE))
+
+tracts_sp <- as(tracts, "Spatial")
 @
@@
 # this interactive map is much easier to create and explore
-pal <- colorQuantile("GnBu", tracts_sf$access_ls, n = 5)
-leaflet(tracts_sf) %>%
+pal <- colorQuantile("GnBu", tracts$access_ls, n = 5)
+bin <- colorFactor("Dark2", tracts$access_025)
+leaflet(tracts) %>%
   addProviderTiles(providers$CartoDB.Positron) %>%
@@
               label = ~as.character(round(tweets_ls,2))) %>%
-  addLayersControl(baseGroups = c("access", "tweets"))
+  addPolygons(group = "1/4 mile", color = ~ bin(access_025)) %>%
+  addLayersControl(baseGroups = c("access", "tweets", "1/4 mile"))
 @
@@
 library(ggmap)
-map <- ggmap(get_map(unname(st_bbox(tracts_sf)), source = "stamen", maptype = "toner-lite"),
+map <- ggmap(get_map(unname(st_bbox(tracts)), source = "stamen", maptype = "toner-lite"),
              extent = "device")
 map +
-  geom_sf(data = tracts_sf, inherit.aes=FALSE,
+  geom_sf(data = tracts, inherit.aes=FALSE,
           aes(fill = cut(access_ls, breaks = c(-Inf, -3, -2, -1, 0, 2, 3, Inf))),
@@
 base_formula <- formula(
-  ~ log(density) +
+  ~ log(density) + log(income) + 
     fulltime + college + single +
-    Pct0to17 + Pct18to29 + Pct65plus + # need to leave out a category for collinearity
+    youth + young_adults + seniors + # need to leave out a category for collinearity
     black + asian + hispanic + other)
-
 @
@@
 <<spatial selection>>=
-physact_base_sem <- errorsarlm(update(base_formula,
-                                      physact ~ . + highincome + lowincome),
-                             data = tracts@data, listw = W, zero.policy = TRUE)
-physact_base_sdm <- lagsarlm(update(base_formula,
-                                    physact ~ . + highincome + lowincome),
-                           data = tracts@data, listw = W, zero.policy = TRUE,
+physact_base_sem <- errorsarlm(update(base_formula, physact ~ .),
+                               data = tracts_sp, listw = W, zero.policy = TRUE)
+physact_base_sdm <- lagsarlm(update(base_formula, physact ~ .),
+                             data = tracts_sp, listw = W, zero.policy = TRUE,
                              type = "mixed")
-pa_base_logincome <- update(physact_base_sem, . ~ . +log(income) - highincome - lowincome)
 test_sdmsem <- lmtest::lrtest(physact_base_sem, physact_base_sdm)
@@
 pa_access_sem <- update(physact_base_sem, .~ . + access_ls)
-pa_access_logincome <- update(pa_access_sem, . ~ . + log(income) - highincome - lowincome)
+pa_quarter_sem <- update(physact_base_sem, .~ . + access_025)
+
 pa_models <- list("Base" = physact_base_sem,
-                  "Base: log(Income)" = pa_base_logincome,
-                  "Access" = pa_access_sem,
-                  "Access: log(Income)" = pa_access_logincome)
-
+                  "Access: logsum" = pa_access_sem,
+                  "Access: 1/4 mile" = pa_quarter_sem)
 @
@@
 <<pamodelchart, results="asis">>=
-termnames <- c("(Intercept)", "log(density)" ,
-               "Full-time workers", "College-educated",  "Single Adults",
-               "Share of children (0-17)", "Share of young adults (18-29)",
-               "Share of seniors (65+)", "Black population share",
+termnames <- c("(Intercept)", "log(density)", "$\\log($Income$)$",
+               "Fulltime", "College-educated",  "Single Adults",
+               "Youth (0-17)", "Young adults (18-34)",
+               "Seniors (65+)", "Black population share",
                "Asian population share",
                "Hispanic population share", "Other Minorities",
-               "Share high income", "Share low income",
-               "$\\lambda$", "$\\log($Income$)$",
-               "Access logsum")
+               "$\\lambda$", "Access logsum", "1/4 mile access")
+
 texreg(pa_models, digits = 4, table = FALSE,
        booktabs = TRUE, use.packages = FALSE,
-       custom.coef.names = termnames,  stars = c(0.001, 0.01, 0.05, 0.1))
+       custom.coef.names = termnames)
 @
@@
 <<obesemodelchart, results="asis">>=
-termnames <- c("(Intercept)", "log(density)" ,
-               "Full-time workers", "College-educated",  "Single Adults",
-               "Share of children (0-17)", "Share of young adults (18-29)",
-               "Share of seniors (65+)", "Black population share",
+termnames <- c("(Intercept)", "log(density)", "$\\log($Income$)$",
+               "Fulltime", "College-educated",  "Single Adults",
+               "Youth (0-17)", "Young adults (18-34)",
+               "Seniors (65+)", "Black population share",
                "Asian population share",
                "Hispanic population share", "Other Minorities",
-               "Share high income", "Share low income",
-               "$\\lambda$", "$\\log($Income$)$",
-               "Access logsum")
+               "Physical activity",
+               "$\\lambda$", "Access logsum", "1/4 mile access")
 texreg(obesity_models, digits = 4, table = FALSE,
        booktabs = TRUE, use.packages = FALSE,
-       custom.coef.names = termnames,  stars = c(0.001, 0.01, 0.05, 0.1))
+       custom.coef.names = termnames,  stars = c(0.001, 0.01, 0.05))
 @
@@
 library(ggmap)
-map <- ggmap(get_map(unname(st_bbox(tracts_sf)), source = "stamen", maptype = "toner-lite"),
+map <- ggmap(get_map(unname(st_bbox(tracts)), source = "stamen", maptype = "toner-lite"),
              extent = "device")
 map +
-  geom_sf(data = tracts_sf %>% 
+  geom_sf(data = tracts %>%
             mutate(diff = tweets_ls - access_ls,
@@
 <<tweets_models>>=
-tweets_models <- list(pa_models[[3]], pa_models[[4]], 
-                      obesity_models[[3]], obesity_models[[4]]) %>%
+tweets_models <- list(pa_models[[2]], obesity_models[[2]]) %>%
   lapply(function(model){
@@
 <<tweets_coefs, results="asis">>=
-names(tweets_models) <- c("Physical Activity: High and Low Income", 
-                          "Physical Activity: log(Income)",
-                          "Obesity: High and Low Income", "Obesity: log(Income)")
-names(pa_models)[3:4] <- c("Physical Activity: High and Low Income", 
-                           "Physical Activity: log(Income)")
-names(obesity_models)[3:4] <- c("Obesity: High and Low Income", 
-                                "Obesity: log(Income)")
-lapply(c(tweets_models, pa_models, obesity_models), function(model){
+access_models <- list("Physical Activity" = pa_models[[2]],
+                      "Physical Activity" = tweets_models[[1]],
+                      "Obesity" = obesity_models[[2]],
+                      "Obesity" = tweets_models[[2]] )
+
+lapply(access_models, function(model){
   s <- summary(model)
@@
 }) %>%
-  bind_rows(.id = "model") %>%
+  bind_rows(.id = "y") %>%
   filter(term %in% c("tweets_ls", "access_ls")) %>%
-  separate(model, c("y", "income"), ": ") %>%
-  mutate(term = ifelse(term == "tweets_ls", 
-                       "Tweets, Size and Distance", "Size and Distance")) %>%
-  arrange(desc(y), income, term) %>%
-  dplyr::select(income, term, estimate, std.error, p.value) %>%
-
+  mutate(
+    term = ifelse(term == "tweets_ls", "Tweets, Size and Distance", "Size and Distance")
+  ) %>%
+  arrange(desc(y), term) %>%
+  dplyr::select(y, term, estimate, std.error, p.value) %>%
   kable(escape = FALSE, digits = 4,
@@
         caption = "Estimated Relationship between Accessiblity and Outcomes",
-        col.names = c("", "Accessibility Components", "Estimate", "Std. Error",
-                      "$p$-value")) %>%
-  kable_styling() %>%
-    group_rows("y = Physical Activity", 1, 4) %>%
-    group_rows("y = Obesity", 5, 8)
+        col.names = c("Outcome Variable", "Accessibility Components", "Estimate", "Std. Error",
+                      "p-value")) %>%
+  kable_styling()
+
 @

### 2019-02-22 manuscript.Rnw
Improve data summary table

@@
 \usepackage{url}
+\usepackage{array}
+\usepackage{pdflscape}
+\usepackage{makecell}
 
@@
   separate(variable, c("var", "desc"), sep = ": ") %>%
+  mutate(
+    msd = str_replace(msd, " \\(", "\\\\\\\\ \\("),
+    msd = str_c("\\makecell{", msd, "}")
+  ) %>%
   kable(escape = FALSE, digits = 2,
@@
         col.names = c(" ", "Description", "Minimum", "Median (IQR)", "Maximum", "Source")) %>%
-  kable_styling() %>%
+  kable_styling(latex_options = "scale_down") %>%
+  column_spec(1, width = "0.5in") %>%
+  column_spec(2, width = "2in") %>%
   group_rows(paste("Tract Variables, N =", nrow(tracts)), 1, 14) %>%
@@
 
-tracts.dnn <- dnearneigh(gCentroid(as(pop_weighted_centroids, "Spatial"), byid = TRUE),
-                         0, 1.8 * 5280)
-dists <- nbdists(tracts.dnn, gCentroid(tracts, byid = TRUE))
+# determine which tracts are neighbors of each other. Need to be less than
+# 1.8 miles away from the centroids
+tracts.dnn <- dnearneigh(as(pop_weighted_centroids, "Spatial"), 0, 1.8 * 5280)
+# calculate the inverse distance to use as a weight on the neighbor connection
+dists <- nbdists(tracts.dnn, as(pop_weighted_centroids, "Spatial"))
 dists.inv <- lapply(dists, function(x) 1 / x)
+# get the spatial weights matrix and its trace to use in any impact analyses
 W <- nb2listw(neighbours = tracts.dnn, glist = dists.inv,
@@
 <<distances, warning=FALSE>>=
-# calculate centroids of tracts and parks
+# project the parks data into the same projection system as the tracts
 parks <- parks %>% st_transform(st_crs(pop_weighted_centroids))
 
-# distances from all tracts to all zones
+# distances from all tract centroids to all parks
 distances <- st_distance(pop_weighted_centroids, parks, byid = TRUE) %>%

### 2019-02-22 manuscript.Rnw
Revise literature review and data discussions

@@
 
-\author{Nico Boyd\affilnum{1}, Gregory S. Macfarlane\affilnum{2}, Kari Watkins\affilnum{1}, 
-and Professor WhogaveTweets\affilnum{3}}
+\author{Nico Boyd\affilnum{1}, Gregory S. Macfarlane\affilnum{2},
+John Taylor\affilnum{1}, and Kari Watkins\affilnum{1}}
 \runninghead{Boyd, et al.}
@@
 \affiliation{\affilnum{1}School of Civil and Environmental Engineering, Georgia Institute of Technology\\
-\affilnum{2}Civil and Environmental Engineering Department, Brigham Young University\\
-\affilnum{3}Somewhere at Georgia Tech}
+\affilnum{2}Civil and Environmental Engineering Department, Brigham Young University}
 
@@
 socioeconomic data in New York City, paired with geographic parks data and
-and geolocated data collected from the Twitter API. Controlling for the 
+geolocated position data collected from the Twitter API. Controlling for the
 socioeconomic variables and spatially correlated error terms, we show a positive
@@
 \maketitle
-<<setup, cache=FALSE, echo=FALSE, message=FALSE>>=
+<<setup, cache=FALSE, echo=FALSE, message=FALSE, warning=FALSE>>=
 library(tidyverse)
@@
 The United States is currently facing an epidemic of obesity and chronic
-diseases, which are non-communicable diseases of long duration and typically
-slow development, including cardiovascular diseases, chronic respiratory
+diseases, including cardiovascular diseases, chronic respiratory
 diseases, diabetes, stroke, joint and bone diseases, and cancer.
@@
 
-Until recently, most large-scale health promotion efforts focused on
+Historically, most large-scale health promotion efforts focused on
 individual-level interventions intended to educate people about healthy
 lifestyles and behaviors, touching on topics including diet and exercise.
-According to \citet{Coutts2008} however, this trend is shifting as professionals
-“have begun adopting an ecological paradigm, accepting that both individual and
-environmental determinants play a role in health behavior. This new, arguably
-revisited, public-health paradigm accounts not only for the compositional (who
-you are) but also for the contextual (where you are) influences on physical
-activity.” As professionals begin to operate from the assumption that the design
-and configuration of the built environment can facilitate or inhibit physical
-activity, they are increasingly looking to public spaces like parks and
-greenways as key elements of the built environment that can support physical 
-activity \citep{Bedimo-Rung2005a}.
+According to \citet{Duhl1999} however, the field of public health has
+moved toward a new paradigm that is grounded in a social model of health,
+treating health ``as an outcome of the effects of socioeconomic status, culture,
+environmental conditions, housing, employment and community influences.''
+Further, \citet{Coutts2008} maintains that awareness of the impacts of the built
+environment on health will increase connections between the practices of public
+health and urban planning. As professionals from both fields begin to operate
+from the assumption that the design and configuration of the built environment
+can facilitate or inhibit physical activity, they are increasingly looking to
+public spaces like parks and greenways as key elements of the built environment
+that can support physical activity \citep{Bedimo-Rung2005a}.
 
@@
 spatial correlation and tract-level socioeconomic characteristics.
-We find a positive relationship where the
-least park-accessible tracts have an expected physical activity participation
-rate half a percentage point lower than the most accessible tracts, though 
-this result is only suggestive of correlation from a statistical perspective.
+We find a positive relationship where the least park-accessible tracts have an
+expected physical activity participation rate half a percentage point lower than
+the most accessible tracts, though this result is only suggestive of correlation
+($p$-value $<0.1$ from a statistical) perspective.
 However, we also find a strong and significant correlation between
@@
 be expanded to account for other elements of a park's attractiveness by
-introducing the number of tweets emanating from within a park during 
-(WE NEED TO DESCRIBE THE PERIOD). The paper concludes with a discussion of
+introducing the number of geolocated tweets emanating from within each park during
+September 2014. The paper concludes with a discussion of
 opportunities for future research.
@@
   & Reduce aggression \citep{Kuo2001a}\\
-  & Enhance emotional and cognitive development\\
-  & Improve behavioral outcomes in youth \\
+  & Improve behavioral outcomes in youth \citep{Taylor1998, Madzia2018} \\
 \midrule
-Social Health & Enhanced community aesthetics\\
- & Reduce crime \citep{Kuo2013,Wolfe2012}\\
+Social Health & Reduce crime \citep{Wolfe2012}\\
  & Increase social interaction \citep{Sullivan2004}\\
 \midrule
-Economic Health & Provide ecosystem services\\
- & Increase residential property values and municipal property tax revenues\\
- & Attract more shoppers and increase economic activity to commercial districts \\
+Economic Health & Increase residential property values and municipal property tax revenues \citep{Donovan2010}\\
+ & Attract more shoppers and increase economic activity to commercial districts \citep{Schwab2009}\\
 \bottomrule
@@
 urban greenspace and mortality caused by lung cancer, diabetes, heart disease or
-car accidents, they did find that all-cause mortality was significantly higher in
+car accidents, they did find that all-cause mortality was oddly, {\em higher} in
 greener cities. In a meta-analysis of 20 peer reviewed journal articles
@@
 accessibility to parks that exist in many cities. Some cities with large amounts
-of total greenspace may nevertheless inequally distribute this space throughout
+of total greenspace may nevertheless unequally distribute this space throughout
 the city, leading to areas with poor access. Conversely, a city with a smaller
@@
 
-In a study of neighborhoods in New York City, \citet{Stark2014} found that a
+In a study of neighborhoods in New York City, \citet{Stark2014} used
+hierarchical linear models to explore relationships between the proportion of
+zip codes that was park space, the cleanliness of parks, and the body mass index
+of adult residents. The authors found that a
 higher proportion of park space in a neighborhood was associated with a lower
-BMI for its residents. \citet{Pretty2005} showed the presence of trees and other
-vegetation in outdoor environments is positively associated with physical
-activity; further, \citet{Suminski2005} showed that adding vegetation to
+BMI for its residents, and lower park cleanliness scores were associated with a
+higher BMI, suggesting that increasing the supply of clean, well-maintained
+parks could positively impact adult BMI.
+\citet{Suminski2005} showed that adding vegetation to
 existing sidewalks and trails increased local motivation to engage in physical
@@
 setting, noting that ``ways to measure active accessibility are as varied as the
-number of scholars the measure them.''(p. 227) This is at least partly because
+number of scholars that measure them.''(p. 227) This is at least partly because
 each method has inherent strengths and weaknesses.
 Distance is an inherently continuous phenomenon, but software tools ease and enable
-buffer analyses where distance becomes a binary phenomon of somewhat arbitrary
-extent. \citet{Logan2017} note that these commonly used approaches can also lead
+buffer analyses where distance becomes a binary phenomenon of somewhat arbitrary
+extent, as a cut-off distance is implied in the buffer distance assumption.
+\citet{Logan2017} note that these commonly used approaches can also lead
 researchers to overlook access-poor populations.
@@
 distances, and that distance is a continuous, non-binary phenomenon.
-It defies reason to assume person living 1.1 miles 
+It defies reason to assume a person living 1.1 miles
 from a park has functionally different accessibility than someone living at 0.9 miles.
@@
 distance as an element of relative accessibility. This suggests that
-not all parks are equal, and that a large park such as Central Park on Manhattan
+not all parks are equal, and that a large park such as Central Park in Manhattan
 may provide health and activity benefits over a much larger area than a smaller
@@
 Note that this formulation is an extension of the gravity-based accessibility
-statistics, in the same way that the gravity-based trip distribution model model 
+statistics, in the same way that the gravity-based trip distribution model
 is a specific case of
 a destination-choice model \citep{Daly1982}. But the extension is meaningful, and
-allows us to consider more attributes of the park or the indivdual beyond size
+allows us to consider more attributes of the park or the individual beyond size
 and impedance. For example, we can aggregate the number of tweets emanating from
@@
 likely that a regional travel demand model is available to the analysts, thus making
-calibrated and multimodal logsums are readily available \citep{Geurs2010}.
+calibrated and multimodal logsums readily available \citep{Geurs2010}.
 
 \section{Empirical Application}
-In this section we describe an experiment with data for New York City, where
-we compute a holistic logsum-based accessibility to parks metric and model the
+In this section, we describe an experiment with data for New York City, where
+we compute a holistic logsum-based accessibility to parks measure and model the
 relationship between this measure and physical activity rates, controlling for
 spatial effects and socioeconomic factors. We subsequently model the effect
-of this accessibility on obesity.
+of the accessibility metric on obesity rates, accounting for physical activity and
+the other controls. We then model the relationship between a modified logsum
+considering geolocated Twitter data and physical activity and obesity rates.
 
 \subsection{Data}
+
+This study uses data available to the public from a variety of federal and state
+data agencies. The datasets, as well as the analysis code, are available on GitHub
+at \url{https://github.com/boydnpb/park_access_new}.
 
@@
 
+The Centers for Disease Control and Prevention makes small-area estimates on
+key health indicators available through its 500 Cities data program. The models
+are estimated on Census sociodemographic data paired with national and local
+health surveys, and have been validated at multiple geographic levels
+\citep{Wang2018, Wang2017, Zhang2015}. We use two synthetic indicators as our
+outcome variables in this study: the share of adults in a Census tract who are
+obese, and the share of adults who participate in no leisure-time physical activity.
+To improve clarity in our interpretation, we use the complement of the second variable ---
+the share of tract adults who participate in {\em some} physical activity ---
+even if the amount is not sufficient to affect overall health. Both indicators
+are estimated for the year 2016.
+
 <<tracts>>=
@@
 
-This study uses data available to the public from a variety of federal and state
-data agencies. The datasets, as well as the analysis code, are available on GitHub
-at \url{https://github.com/boydnpb/park_access_new.git}.
-The primary dataset is a geographic polygons shapefile of Census tracts in
-New York City. We appended relevant sociodemographic data for each tract level
-from the American Community Survey 2011-2015 5-year estimates. Data on physical
-activity participation rate and obesity rates for each tract are available from
-the Centers for Disease Control and Prevention’s 500 Cities Project (WE NEED A CITATION).
-After removing tracts with missing population information or outlying and unusual
-characteristics, we have \Sexpr{nrow(tracts)} complete cases. Table
-\ref{tab:tractsdata} presents key descriptive statistics for these data.
+To the health data, we join data collected through the Census Bureau API via
+the {\texttt tidycensus} package for R \citep{Walker2019}. The primary dataset
+is a geographic polygons layer of Census tracts in the five boroughs of New York
+City. We append to each Census tract relevant sociodemographic data
+for each tract level from the American Community Survey 2013-2017 5-year
+estimates. For a small handful of tracts in our sample, Census supressed the
+median income estimate; these appear to be primarily wealthy tracts and in
+almost all cases the CDC estimates of obesity and physical activity are missing
+as well. After removing these tracts from the estimation dataset, we have
+\Sexpr{nrow(tracts)} complete cases. Table \ref{tab:tractsdata} presents key
+descriptive statistics for these data.
 
@@
 We retrieved a shapefile of public parks and greenspaces within New York City’s
-municipal boundaries and checked it for accuracy and relevance (WE NEED A CITATION).
-Upon inspection,
-we removed several facilities that do not qualify as publicly accessible green
-space, such as Yankee Stadium and its surrounding parking lots. We also removed
-parks of less than half an acre in size, as these appear to be predominantly
-vacant lots or parking strips rather than legitimate public green space. This
-leaves us with \Sexpr{nrow(parks)} parks. It is worth noting that the particular
-dataset classifies specific facilities within larger parks as independent
-features: baseball fields within Central Park, for example, are themselves
-delineated green spaces. We elect to retain these independent features as they may
-generate additional physical activity participation aside from the containing 
-park.
+municipal boundaries and checked it for accuracy and relevance \citep{nycparks}.
+Upon inspection, we removed several facilities that do not qualify as
+publicly accessible green space, such as Yankee Stadium and its surrounding
+parking lots. We also removed parks of less than half an acre in size, as these
+appear to be predominantly vacant lots or parking strips rather than legitimate
+public green space. This leaves us with \Sexpr{nrow(parks)} parks. It is worth
+noting that the particular dataset classifies specific facilities within larger
+parks as independent features: baseball fields within Central Park, for example,
+are themselves delineated green spaces. We elect to retain these independent
+features as they may generate additional physical activity participation aside
+from the containing park. Additionally, large cemeteries are not considered
+open space in these data, though people may use the space for physical activities
+such as walking or running. For each park we calculate its size in acres, and
+identify the number of geolocated tweets emanating from the park in September
+2014, as a proxy for the park's use or popularity.
 

### 2019-02-22 manuscript.Rnw
Revise literature review and physical activity model discussion

@@
 We find a positive relationship where the least park-accessible tracts have an
-expected physical activity participation rate half a percentage point lower than
-the most accessible tracts, though this result is only suggestive of correlation
-($p$-value $<0.1$ from a statistical) perspective.
+expected physical activity participation rate roughly a full percentage point
+lower than the most accessible tracts; a traditional quarter-mile buffer
+analysis estimates an effect twice as large but marginally less significant.
 However, we also find a strong and significant correlation between
-increased park accessibility and decreased obesity rates, again controlling for
-spatial correlation, socioeconomic attributes, and physical activity rates.
+the choice-based park accessibility metric and decreased obesity rates --- again
+controlling for spatial correlation, socioeconomic attributes, and physical
+activity rates. A traditional buffer analysis does not show significance.
 
 We also demonstrate how the choice-based accessibility metric could potentially
-be expanded to account for other elements of a park's attractiveness by
-introducing the number of geolocated tweets emanating from within each park during
-September 2014. The paper concludes with a discussion of
-opportunities for future research.
+be expanded to account for other measures of a park's attractiveness by
+introducing the number of geolocated tweets emanating from within each park
+during September 2014. The paper concludes with a discussion of opportunities
+for future research.
 
@@
 in a study of home prices in Atlanta. We selected this specification after
-comparing overall model likelihoods between it and an adjacent (queen)
-tract specification.
+comparing overall model likelihoods between it and specification where tracts
+sharing a boundary are considered neighbors (queen-adjacent).
 
@@
 will prefer to use larger parks and parks that are nearer to their residence. This
-resulted in $\lambda_d = -1.98$  and $\lambda_s = 0.000$. Though the 
-``most likely''
-parameter value, in this case it was unsatisfactory for two reasons: first,
-our prior assertion that park size is a valuable attribute of a park, and second
-the observation that with these maximum likelihood values tracts bordering on 
-the corners of
-Central Park were measured with poor overall accessibility, a unacceptably 
-unintuitive result. We subsequently 
-adjusted the $\lambda_s$ value to $0.3$; the relative ratio of the coefficients 
-implies that all else equal, an individual would travel 6.6 times further to
-access a park twice as large.
-
-<<logsum_betas>>=
+resulted in $\lambda_d = -1.98$  and $\lambda_s = 0.000$. Though the ``most
+likely'' parameter value, in this case it was unsatisfactory for two reasons:
+first, our prior assertion that park size is a valuable attribute of a park, and
+second the observation that with these maximum likelihood values tracts
+bordering on the corners of Central Park were measured with poor overall
+accessibility, an unacceptably unintuitive result. We subsequently adjusted the
+$\lambda_s$ value to $0.3$; the relative ratio of the coefficients implies that
+all else equal, an individual would travel 6.6 times further to access a park
+twice as large.
+
+<<logsum_betas, warning = FALSE>>=
 # This is where we actually calculate the accessibility statistic, using the
@@
 against the base covariates (without the accessibility logsum). A likelihood
-ratio test failed to reject that the SEM is equivalent
-($\chi^2 = \Sexpr{test_sdmsem$Chisq[2]}$ on \Sexpr{test_sdmsem$Df[2]} degrees of
-freedom), so we use the SEM specification only going forward.
-Table \ref{tab:pa_models} presents the estimated coefficients for the base
-model with income represented in two ways. In the first, the share of the tract
-population in
-the bottom three income deciles and the share of the population in the top three
-income deciles are each included as independent predictors. In the second, the
-logarithm of the mean tract income is included as a predictor.
-Notwithstanding the higher model likelihood of the deciles model, comparing
-coefficients across the models shows that the predictors change magnitude
-or significance only marginally, if at all. Overall, the base model
-predictors are significant and of intuitive direction. Tracts with higher shares
+ratio test failed to reject that the SDM is different from the SEM, so we use
+the SEM specification only going forward. Table \ref{tab:pa_models} presents the
+estimated coefficients for the base model with the socioecononomic controls and
+no representation of park accessibility. For the most part the coefficients are
+highly significant and of the expected sign. Tracts with higher shares
 of full-time workers, college-educated adults, young adults, and high-income
@@
 share of single adults, children, seniors, minorities of all types, and low
-income households all report lower rates of physical activity.
-
-The other two models presented in Table \ref{tab:pa_models} include
-the log-sum accessibility statistic described in above added as a predictor. 
-In both income specifications,
-the access logsum is positively correlated with physical activity rates; all else
-equal, people in tracts with higher accessibility to parks report higher rates of
-physical activity. The estimated value of this coefficient and its standard
-error is effectively the same with both representations of income, though in the
-log representation the estimated $t$-statistic technically passes the $0.1$ level.
-This is suggestive evidence only from a statistical perspective, but it deserves
-comment. Taking a compromise value of $0.10$ between the two models, the data
-suggest going from the least-accessible ($-3.0$) to the most-accessible ($3.0$) 
-tract implies a $0.6$ percentage-point rise in physical activity rates, though
-with a wide confidence interval around that estimate.
+income households all have lower modeled rates of physical activity.
+
+To this base model we add a measure of park accessibility considered in two
+ways: a logsum-based statistic as described earlier and a more typical quarter-mile
+binary access buffer. The estimated coefficients for these models are also
+given in Table \ref{tab:pa_models}. In both accessibility specifications,
+the access measure is positively correlated with physical activity rates; all else
+equal, people in tracts with higher accessibility to parks have higher estimated
+rates of physical activity. Though the estimated coefficient on the quarter-mile
+buffer appears to have a larger absolute value than the logsum-based statistic,
+this masks a key difference in the specification worth remembering.
+The logsum allow for variation in access unavailable in the buffer-based method.
+In our dataset almost every tract
+(\Sexpr{round(sum(tracts$access_025) / nrow(tracts) * 100, 0)} \%) is within
+a quarter-mile of {\em some} kind of park, but this park may not be large enough
+to support physical activities all individuals would like to participate in.
+Because tracts with only adequate park access are grouped with tracts having
+excellent park access, the overall effect is comparatively unclear. The logsum-based
+coefficient, on the other hand, suggest going from the least-accessible ($-3.0$)
+to the most-accessible ($3.0$) tract implies roughly a $1.1$\% percentage-point
+rise in estimated physical activity rates
 
@@
 screenreg(obesity_models, digits = 4, table = FALSE,
-          stars = c(0.001, 0.01, 0.05, 0.1) )
+          stars = c(0.001, 0.01, 0.05) )
 @

### 2019-02-27 manuscript.Rnw
Revise results discussion with only one income specification.

@@
 the {\em obesity} rate, and physical activity becomes an independent covariate
-alongside the controlling variables and the accessibility logsum variable.
-Table \ref{tab:obese_models} presents the estimated coefficients for both
-income specifications, with and without the accessibility logsum. As in the
-physical activity models, the coefficients are largely significant and
-in the expected sign. The accessibility coefficient has practically the same
-effect as in the physical activity model: moving from poor to excellent parks
-access lowers the changes the expected obesity rate by around half a percentage
-point. In this case, however, the parameter $p$-value is lower than the
-conventional $0.05$ level in each income representation.
+alongside the controlling variables and the accessibility metric.
+Table \ref{tab:obese_models} again presents the base model without
+any accessiblity measure, and models with accessibility captured through a
+logsum and through a 1/4 mile buffer. As in the
+physical activity models, the coefficients are typically significant and
+on the expected sign, with a few exceptions: in this case an increasing hispanic
+population share  and an increasing share of children have no significant
+relationship on obesity. Unlike in the physical activity models however,
+the inference on the accessibility statistic {\em does} change between
+specifications. The 1/4 mile buffer neither substantially improves the model
+fit nor is its term significant at any threshold, but the accessibility logsum
+indicates a significant negative correlation: moving from the least to most
+accessible tract decreases the expected obesity rate by approximately half a
+percentage point.
 
@@
 activity are highly correlated, and thus would result in a double-counting of
-park size in the accessibility statistic. In our 
-data the number of tweets is positively but only loosely correlated  with park 
-size in our data ($\rho = \Sexpr{round(cor(parks$log_size, parks$log_tweets), 3)}$), 
-indicating that the Twitter data in fact could provide new information or proxy for the 
-desired amenities.
-
-To test this hypothesis, we can use the twitter data described earlier
-and suggested in Equation \ref{eq:v_sdt}. In this application, we again 
+park size in the accessibility statistic. In our data the number of tweets is
+positively but only loosely correlated  with park
+size ($\rho = \Sexpr{round(cor(parks$log_size, parks$log_tweets), 3)}$),
+indicating that the Twitter data in fact could provide new information or proxy
+for the desired amenities.
+
+To test this hypothesis, we can use the twitter data described earlier and
+suggested in Equation \ref{eq:v_sdt}. In this application, we again
 log-transform all the components of utility:
@@
 would be otherwise undefined \citep{Yeo2000}. As before, there is little
-information by which to
-determine the values of $\lambda$ in this utility specification; we use the
-previous values with the addition of $\lambda_t = 0.1$. The ratio between the 
-utility coefficients implies residents would travel almost twenty times as far or
-use a park three times smaller if it had twice the Twitter activity, all else 
-equal.
+information by which to determine the values of $\lambda$ in this utility
+specification; we use the previous values with the addition of $\lambda_t =
+0.1$. The ratio between the utility coefficients implies residents would travel
+almost twenty times as far or use a park three times smaller if it had twice the
+Twitter activity, all else equal.
 
@@
 accessibility to obesity and physical activity rates, with and without the
-inclusion of Twitter data. In all four cases, including twitter information in
+inclusion of Twitter data. In both cases, including twitter information in
 the accessibility statistic tempers the strength of the relationship with the
 dependent variable and modestly widens the standard error without substantively
-affecting the significance of the test statistics.  
+affecting the significance of the test statistics or the interpretation.
 From this limited example, we cannot say whether Twitter activity enhances the

### 2019-02-28 manuscript.Rnw
Update with reviews from Nico and Kari

@@
 socioeconomic variables and spatially correlated error terms, we show a positive
-but suggestive relationship between park access and physical activity rates, and
-a clear negative relationship between park access and obesity rates.
+and significant relationship between park access and physical activity rates, and
+a clear negative relationship between park access and obesity rates. Applying
+a more traditional analysis considering 1/4 mile acess to parks yields mixed
+results.
 \end{abstract}
@@
 treating health ``as an outcome of the effects of socioeconomic status, culture,
-environmental conditions, housing, employment and community influences.''
+environmental conditions, housing, employment and community influences.'' (p. 7)
 Further, \citet{Coutts2008} maintains that awareness of the impacts of the built
@@
 distance to all parks in the region, weighted against the size of the park and
-its other amenities. We apply this measurement to study the link between park
+its other amenities.
+
+We apply this measurement to study the link between park
 accessibility and attractiveness and Census tract-level aggregate physical
@@
 lower than the most accessible tracts; a traditional quarter-mile buffer
-analysis estimates an effect twice as large but marginally less significant.
+analysis estimates an similar albeit marginally less significant effect.
 However, we also find a strong and significant correlation between
@@
 controlling for spatial correlation, socioeconomic attributes, and physical
-activity rates. A traditional buffer analysis does not show significance.
+activity rates. In this case a traditional buffer analysis does not show
+significance.
 
@@
 parks could positively impact adult BMI.
-\citet{Suminski2005} showed that adding vegetation to
-existing sidewalks and trails increased local motivation to engage in physical
-activity.
 
@@
 is a specific case of
-a destination-choice model \citep{Daly1982}. But the extension is meaningful, and
+a destination-choice model \citep{Daly1982}. But the extension is meaningful and
 allows us to consider more attributes of the park or the individual beyond size
@@
 In spite of its flexibility and basis in choice theory, accessibility measures
-based in log-sums
-have not received much application in the accessibility literature compared with
-distance-based or even
-gravity-based measures. \citet{Zhang2011} present a metric with a similar functional
-form, though the authors arrive at it from a motivation other than destination
-choice theory. Log-sums are commonly used, however, in alternatives analyses of
-transit infrastructure improvements \citep{DeJong2007}. A reason for this is
-likely that a regional travel demand model is available to the analysts, thus making
+based in log-sums have not received much application in the accessibility
+literature compared with distance-based or even gravity-based measures.
+\citet{Zhang2011} present a metric with a similar functional form, though the
+authors arrive at it from a motivation other than destination choice theory.
+Log-sums are commonly used, however, in alternatives analyses of transit
+infrastructure improvements \citep{DeJong2007}. A reason for this is likely that
+a regional travel demand model is available to the analysts, thus making
 calibrated and multimodal logsums readily available \citep{Geurs2010}.
@@
 The Centers for Disease Control and Prevention makes small-area estimates on
-key health indicators available through its 500 Cities data program. The models
+key health indicators available through its 500 Cities data program \citep{CDC5002016}. The models
 are estimated on Census sociodemographic data paired with national and local
@@
 To the health data, we join data collected through the Census Bureau API via
-the {\texttt tidycensus} package for R \citep{Walker2019}. The primary dataset
+the \texttt{tidycensus} package for R \citep{Walker2019}. The primary dataset
 is a geographic polygons layer of Census tracts in the five boroughs of New York
 City. We append to each Census tract relevant sociodemographic data
-for each tract level from the American Community Survey 2013-2017 5-year
+for each tract from the American Community Survey 2013-2017 5-year
 estimates. For a small handful of tracts in our sample, Census supressed the
@@
 buffer appears to have a larger absolute value than the logsum-based statistic,
-this masks a key difference in the specification worth remembering.
+this masks a key difference in the specification worth further comment.
 The logsum allow for variation in access unavailable in the buffer-based method.
@@
 excellent park access, the overall effect is comparatively unclear. The logsum-based
-coefficient, on the other hand, suggest going from the least-accessible ($-3.0$)
+coefficient, on the other hand, suggests going from the least-accessible ($-3.0$)
 to the most-accessible ($3.0$) tract implies roughly a $1.1$\% percentage-point
@@
 a key benefit it offers is the natural way in which park amenities beyond simply size
-can be accomodated. For example, the number of tweets emanating from a park
+can be accommodated. For example, the number of tweets emanating from a park
 might be seen as a measure of the park's popularity, or a proxy for its use. It
@@
 Figure \ref{fig:tweets_map} shows the percent change in the normalized log
-sum statistic after including the twitter infromation. There is not a recognizable
+sum statistic after including the twitter information. There is not a recognizable
 overarching pattern to the changes, though tracts immediately surrounding
@@
 From this limited example, we cannot say whether Twitter activity enhances the
-ability of parks to induce physical activity rates or lower obesity rates,
+ability of parks to increase physical activity rates or lower obesity rates,
 though further investigation of this and other park amenities would certainly
@@
 We readily acknowledge limitations in this study. As in any study conducted
-with areal data, we at risk of falling victim to the ecological inference 
+with areal data, we are at risk of falling victim to the ecological inference
 fallacy, where aggregate statistics mask or contradict disaggregated or
 individual-level trends. A large-sample survey of individuals in New York City,
-including measured physical activities and obesity would always be prefferable
-to the tract-level data used in this study. It would be also be
+including measured physical activities and obesity would always be preferable
+to the tract-level data used in this study. It would also be
 preferrable to have obtained the $\vect{\lambda}$ accessibility utility coefficients
@@
 about. In this paper, we proceed only incrementally beyond this theory by
-adding Twitter activity as an an element of a park's attractiveness above and
+adding Twitter activity as an element of a park's attractiveness above and
 beyond its size. Exploring additional amenities or market segmentation strategies
@@
 Euclidean distances have well-rehearsed limitations regarding their fidelity with
-the underlying infrastructure network, etc.. Network-based distances can also
+the underlying infrastructure network, etc. Network-based distances can also
 suffer from challenges when applied to multimodal problems; these challenges are
@@
 access to parks. A better metric of travel impedance may be a mode choice model
-log-sum, which weights all travel alternatives against each other. A
+logsum, which weights all travel alternatives against each other.
 
@@
 have observed. It is distinctly possible, for instance, that individuals who wish
-to exercise regularly in parks choose to live near them; this could potentially
-explain why the observed effects of accessibility seem partially dependent on
-the representation of income in our econometric model. Controlling for this
-potential self-selection effect would be necessary to isolate the exogenous
+to exercise regularly in parks choose to live near them. Given that the CDC models
+generating the obesity and physical activity estimates presumably include variables
+likely to influence such preferences (income, etc.), our investigation cannot
+isolate the preferences from the effect. Controlling for such a
+self-selection effect would be necessary to isolate the exogenous
 impacts of park access on obesity or other health outcomes. And regarding these
@@
 
+\section{Conclusion}
+
+Increased physical activity and decreased obesity rates are critical measures of
+improvement in public health. Although many have theorized the link between park
+accessibility and these metrics, previous literature has been mixed. Using New
+York City as a case study, we presented a holistic and flexible measurement for
+park accessibility that compares the continuous distance to all parks in the
+region, weighted against the size of the park and its other amenities. In terms
+of physical activity, we found a positive relationship where the least
+park-accessible tracts have an expected physical activity participation rate
+roughly one percentage point lower than the most accessible tracts. In terms of
+obesity, we also found a strong and significant correlation between increased
+park accessibility and decreased obesity rates, controlling for
+spatial correlation, socioeconomic attributes, and physical activity rates.
+
+In both cases a traditional, buffer-based analysis estimated a weaker relationship
+with greater uncertainty. Buffer analyses are relatively common in the literature,
+perhaps because of the widespread availability of GIS software. In spite of their
+widespread use, they are relatively limited in terms of both their theoretical
+underpinnings and their flexibility to accommodate attributes of parks beyond
+their proximity. And even proximity may not be adequately handled, as the buffer
+distance may be arbitrarily asserted by the researcher. Adopting choice-based
+accessibilities of the kind used in this study will allow researchers to
+encompass the full universe of park amenities in their accessibility analyses.
+This will in turn enable planners to consider how multiple attributes of a park
+--- from its location to its size to its amenities and beyond --- benefit the health
+of the community the park serves.
+
 \begin{funding}
-This project was funded in part via \ldots.
+This project was funded by the Speedwell Foundation.
 \end{funding}

### 2019-03-01 manuscript.Rnw
Add reference and clarify 500 Cities data

@@
 amenities and attributes of both the parks and the population. We apply this
-choice-based accessibility measure to Census tract-level health and
-socioeconomic data in New York City, paired with geographic parks data and
-geolocated position data collected from the Twitter API. Controlling for the
-socioeconomic variables and spatially correlated error terms, we show a positive
-and significant relationship between park access and physical activity rates, and
-a clear negative relationship between park access and obesity rates. Applying
-a more traditional analysis considering 1/4 mile acess to parks yields mixed
-results.
+choice-based accessibility measure to correlate Census tract-level obesity and
+physical activity rate estimates from the Centers for Disease Control and
+Prevention's 500 Cities project with tract-level American Community Survey
+socioeconomic data in New York City, paired with geographic parks data from New
+York City and geolocated tweet data collected from the Twitter API as a metric of
+a park's activity levels. Controlling for the socioeconomic variables and
+spatially correlated error terms, we show a positive and significant
+relationship between park access and physical activity rates, and a clear and
+significant negative relationship between park access and obesity rates.
+Applying a more traditional analysis considering 1/4 mile acess to parks yields
+mixed results and cannot accommodate features such as park size and twitter
+activity.
 \end{abstract}
@@
 
-Note that this formulation is an extension of the gravity-based accessibility
+Note that this formulation is an extension of gravity-based accessibility
 statistics, in the same way that the gravity-based trip distribution model
@@
 The Centers for Disease Control and Prevention makes small-area estimates on
-key health indicators available through its 500 Cities data program \citep{CDC5002016}. The models
-are estimated on Census sociodemographic data paired with national and local
-health surveys, and have been validated at multiple geographic levels
-\citep{Wang2018, Wang2017, Zhang2015}. We use two synthetic indicators as our
-outcome variables in this study: the share of adults in a Census tract who are
-obese, and the share of adults who participate in no leisure-time physical activity.
-To improve clarity in our interpretation, we use the complement of the second variable ---
-the share of tract adults who participate in {\em some} physical activity ---
-even if the amount is not sufficient to affect overall health. Both indicators
-are estimated for the year 2016.
+key health indicators available through its 500 Cities data program \citep{CDC5002016}.
+The indicators are multilevel aggregations and imputations of BRFSS responses
+\citep{Wang2018, Wang2017}, and have been recently used to study the tract-level
+link between gentrification and urban health \citep{Gibbons2018}. We use two
+indicators as our dependent variables in this study: the share of adults
+in a Census tract who are obese, and the share of adults who participate in no
+leisure-time physical
+activity. To improve clarity in our interpretation, we use the complement of the
+second variable --- the share of tract adults who participate in {\em some}
+physical activity --- even if the amount is not sufficient to affect overall
+health. Both indicators are estimated for the year 2016.
 

### 2019-03-04 manuscript.Rnw
Changes suggested by Kari and Nico, plus a couple of others.

@@
 significant negative relationship between park access and obesity rates.
-Applying a more traditional analysis considering 1/4 mile acess to parks yields
+Applying a more traditional analysis considering 1/4 mile access to parks yields
 mixed results and cannot accommodate features such as park size and twitter
@@
   kable_styling(latex_options = "scale_down") %>%
-  column_spec(1, width = "0.5in") %>%
   column_spec(2, width = "2in") %>%
@@
 \begin{align}
-{\boldsymbol y} &= X{\boldsymbol \beta} + {\boldsymbol u},  {\boldsymbol u} = \lambda W {\boldsymbol u} +
+{\boldsymbol y} &= X{\boldsymbol \beta} + {\boldsymbol u},  {\boldsymbol u} = \gamma W {\boldsymbol u} +
 {\boldsymbol \epsilon} \label{eq:SEM}\\
-{\boldsymbol y} &= X{\boldsymbol \beta} + (I-\lambda W)^{-1}{\boldsymbol \epsilon}\notag
+{\boldsymbol y} &= X{\boldsymbol \beta} + (I-\gamma W)^{-1}{\boldsymbol \epsilon}\notag
 \end{align}
@@
 with population-weighted centroids located within 1.8 Euclidean miles of each
-other are
-considered neighbors, and the strength of the relationship is the inverse of
-the distance between them:
+other are considered neighbors, and the strength of the relationship is the
+row-standardized (each row sums to 1) inverse of the distance between them:
 \begin{equation}\label{eq:weightsrule}
 W_{ij} = \begin{cases}
-  {\frac{1/d_{ij}} {\sum_{k=1}^n( 1/d_{ik})}} & \text{for $d_{ij} \leq 1.8$ miles}\\
+  {\frac{d_{ij}^{-1}} {\sum_{k=1}^n d_{ik}^{-1}}} & \text{for $d_{ij} \leq 1.8$ miles}\\
   0  & \text{for $d_{ij} > 1.8$ miles}
@@
 in a study of home prices in Atlanta. We selected this specification after
-comparing overall model likelihoods between it and specification where tracts
-sharing a boundary are considered neighbors (queen-adjacent).
+comparing overall model likelihoods between it and a specification where tracts
+sharing a boundary are considered neighbors.
 
@@
 
-We estimated SDM and SEM models regressing the physical activity rate against
+We estimated SDM and SEM models regressing the physical activity rate
 against the base covariates (without the accessibility logsum). A likelihood
@@
 this masks a key difference in the specification worth further comment.
-The logsum allow for variation in access unavailable in the buffer-based method.
+The logsum allows for variation in access unavailable in the buffer-based method.
 In our dataset almost every tract
@@
 coefficient, on the other hand, suggests going from the least-accessible ($-3.0$)
-to the most-accessible ($3.0$) tract implies roughly a $1.1$\% percentage-point
-rise in estimated physical activity rates
+to the most-accessible ($3.0$) tract implies roughly a $1.1$ percentage point
+rise in estimated physical activity rates.
 
@@
                "Hispanic population share", "Other Minorities",
-               "$\\lambda$", "Access logsum", "1/4 mile access")
+               "$\\gamma$: spatial correlation", "Access logsum", "1/4 mile access")
 
@@
        booktabs = TRUE, use.packages = FALSE,
-       custom.coef.names = termnames)
+       custom.coef.names = termnames,
+       custom.note = "%stars. Standard errors in parentheses.")
 @
@@
 physical activity models, the coefficients are typically significant and
-on the expected sign, with a few exceptions: in this case an increasing hispanic
+of the expected sign, with a few exceptions: in this case an increasing hispanic
 population share  and an increasing share of children have no significant
@@
                "Physical activity",
-               "$\\lambda$", "Access logsum", "1/4 mile access")
+               "$\\gamma$: spatial correlation", "Access logsum", "1/4 mile access")
 texreg(obesity_models, digits = 4, table = FALSE,
        booktabs = TRUE, use.packages = FALSE,
-       custom.coef.names = termnames,  stars = c(0.001, 0.01, 0.05))
+       custom.coef.names = termnames,
+       custom.note = "%stars. Standard errors in parentheses.")
 @

### 2019-03-05 manuscript.Rnw
Incorporate changes from John Taylor

@@
 %% Front matter =========================
-\title{The impacts of park access on health outcomes: A spatial comprehensive approach}
+\title{Modeling the impacts of park access on health outcomes: a choice-based accessibility approach}
 
 \author{Nico Boyd\affilnum{1}, Gregory S. Macfarlane\affilnum{2},
-John Taylor\affilnum{1}, and Kari Watkins\affilnum{1}}
+John E. Taylor\affilnum{1}, and Kari Watkins\affilnum{1}}
 \runninghead{Boyd, et al.}
@@
 lifestyles and behaviors, touching on topics including diet and exercise.
-According to \citet{Duhl1999} however, the field of public health has
+However, \citet{Duhl1999} claim the field of public health has
 moved toward a new paradigm that is grounded in a social model of health,
@@
 lower than the most accessible tracts; a traditional quarter-mile buffer
-analysis estimates an similar albeit marginally less significant effect.
+analysis estimates a similar albeit marginally less significant effect.
 However, we also find a strong and significant correlation between
@@
 and impedance. For example, we can aggregate the number of tweets emanating from
-a park as a measure of the park's attractiveness beyond its size, and the
+a park as a measure of the park's attractiveness beyond its size. The
 utility for the choice maker from Equation \ref{eq:v_sd} then becomes
@@
 \section{Empirical Application}
-In this section, we describe an experiment with data for New York City, where
+In this section, we develop a model with data for New York City, where
 we compute a holistic logsum-based accessibility to parks measure and model the
@@
 of the accessibility metric on obesity rates, accounting for physical activity and
-the other controls. We then model the relationship between a modified logsum
-considering geolocated Twitter data and physical activity and obesity rates.
+the other controls. Finally, we model the relationship between a modified logsum
+measure considering geolocated Twitter data and physical activity and obesity rates.
 
@@
 open space in these data, though people may use the space for physical activities
-such as walking or running. For each park we calculate its size in acres, and
-identify the number of geolocated tweets emanating from the park in September
-2014, as a proxy for the park's use or popularity.
+such as walking or running. For each park we calculate its size in acres. As a
+proxy for the park's use or popularity, we add a measure of Twitter activity. Using
+a Twitter application programming interface (API) implemented with the Python
+package Tweepy, we collected and stored tweets containing geotags of the exact
+geographic coordinates \citep{Wang2016}.  From this information, we were
+able to segregate tweets that were generated within a park in September 2014.
 
@@
 improvement in public health. Although many have theorized the link between park
-accessibility and these metrics, previous literature has been mixed. Using New
-York City as a case study, we presented a holistic and flexible measurement for
+accessibility and these metrics, previous literature has produced mixed
+findings, perhaps owing to the range of variables modeled and the coarse spatial
+scale of the analyses. Using New York City as a case study, we presented a
+holistic and flexible measurement for
 park accessibility that compares the continuous distance to all parks in the
@@
 their proximity. And even proximity may not be adequately handled, as the buffer
-distance may be arbitrarily asserted by the researcher. Adopting choice-based
-accessibilities of the kind used in this study will allow researchers to
-encompass the full universe of park amenities in their accessibility analyses.
+distance may be arbitrarily asserted by the researcher. The model we develop and
+apply in this paper extends buffer driven models to contribute a more
+comprehensive and flexible approach for measuring the impact of park access on
+health outcomes. Adopting choice-based accessibilities of the kind used in this
+study will allow researchers to
+encompass the full range of park amenities in their accessibility analyses.
 This will in turn enable planners to consider how multiple attributes of a park

### 2019-03-05 manuscript.Rnw
Incorporate John's suggested change to abstract

@@
 relationship between park access and physical activity rates, and a clear and
-significant negative relationship between park access and obesity rates.
-Applying a more traditional analysis considering 1/4 mile access to parks yields
-mixed results and cannot accommodate features such as park size and twitter
-activity.
+significant negative relationship between park access and obesity rates. In
+doing so, this research contributes a new, more comprehensive modeling approach
+for measuring the impact of park access on health, and may improve our
+understanding of the role  parks can serve in furthering public health
+objectives.
 \end{abstract}

### 2019-03-05 manuscript.Rnw
Change author order.

@@
 
-\author{Nico Boyd\affilnum{1}, Gregory S. Macfarlane\affilnum{2},
-John E. Taylor\affilnum{1}, and Kari Watkins\affilnum{1}}
+\author{Gregory S. Macfarlane\affilnum{1}, Nico Boyd\affilnum{2},
+John E. Taylor\affilnum{2}, and Kari Watkins\affilnum{2}}
 \runninghead{Boyd, et al.}
 
-\affiliation{\affilnum{1}School of Civil and Environmental Engineering, Georgia Institute of Technology\\
-\affilnum{2}Civil and Environmental Engineering Department, Brigham Young University}
+\affiliation{\affilnum{1}Civil and Environmental Engineering Department, Brigham Young University\\
+\affilnum{2}School of Civil and Environmental Engineering, Georgia Institute of Technology}
 

### 2019-03-22 manuscript.Rnw
Minor edits for submittal

@@
 John E. Taylor\affilnum{2}, and Kari Watkins\affilnum{2}}
-\runninghead{Boyd, et al.}
+\runninghead{Modeling health outcomes via choice-based accessibility to parks}
 
 \affiliation{\affilnum{1}Civil and Environmental Engineering Department, Brigham Young University\\
-\affilnum{2}School of Civil and Environmental Engineering, Georgia Institute of Technology}
-
-\corrauth{Gregory Macfarlane\\
-430 Engineering Building\\
-Brigham Young University\\
-Provo, UT 84602\\
-USA}
+\affilnum{2}Fehr \& Peers, Los Angeles, California\\
+\affilnum{3}School of Civil and Environmental Engineering, Georgia Institute of Technology}
+
+\corrauth{Gregory Macfarlane: 430 Engineering Building, Provo, UT 84602 USA}
+\email{gregmacfarlane@byu.edu}
+
 \begin{abstract}
@@
 \end{abstract}
+
+\keywords{accessibility, spatial modelling, green space}
 
@@
 
-<<map, echo=FALSE, message=FALSE, warning=FALSE>>=
+<<map, echo=FALSE, message=FALSE, warning=FALSE, dev = 'tikz', fig.height=4>>=
 library(ggmap)
@@
 @
-  \caption{Normalized choice-based park accessibility log-sum values in New York City}
+  \caption{Normalized choice-based park accessibility log-sum values in New York City.}
   \label{fig:map}
@@
 
-<<tweets_map, echo=FALSE, message=FALSE, warning=FALSE>>=
+<<tweets_map, echo=FALSE, message=FALSE, warning=FALSE, dev = 'tikz', fig.height=4>>=
 library(ggmap)

### 2019-06-06 manuscript.Rnw
Rebuilt as a generic manuscript after rejection from EPB.

@@
-\documentclass[Crown,sageh,times]{sagej}
-\def\journalname{Submitted to Environment and Planning B: Urban Analytics and City Science}
+\documentclass[12]{article}
 
@@
 \usepackage{url}
+\usepackage{natbib}
 \usepackage{array}
@@
 \usepackage{makecell}
+\usepackage{fullpage}
+\usepackage{booktabs}
+
+%% Front matter =========================
+\title{Modeling the impacts of park access on health outcomes: a choice-based 
+accessibility approach}
+
+\author{Gregory S. Macfarlane, Nico Boyd, John E. Taylor, and Kari Watkins}
 
 \begin{document}
-
-%% Front matter =========================
-\title{Modeling the impacts of park access on health outcomes: a choice-based accessibility approach}
-
-\author{Gregory S. Macfarlane\affilnum{1}, Nico Boyd\affilnum{2},
-John E. Taylor\affilnum{2}, and Kari Watkins\affilnum{2}}
-\runninghead{Modeling health outcomes via choice-based accessibility to parks}
-
-\affiliation{\affilnum{1}Civil and Environmental Engineering Department, Brigham Young University\\
-\affilnum{2}Fehr \& Peers, Los Angeles, California\\
-\affilnum{3}School of Civil and Environmental Engineering, Georgia Institute of Technology}
-
-\corrauth{Gregory Macfarlane: 430 Engineering Building, Provo, UT 84602 USA}
-\email{gregmacfarlane@byu.edu}
+\maketitle
 
@@
 
-\keywords{accessibility, spatial modelling, green space}
-
-\maketitle
 <<setup, cache=FALSE, echo=FALSE, message=FALSE, warning=FALSE>>=
@@
 \citet{Wolf2008} observed that a large portion of U.S. adults do not participate
-in sufficient physical activity, a problem compounded by age and socioeconomic status. As the epidemiological transition from
+in sufficient physical activity, a problem compounded by age and socioeconomic 
+status. As the epidemiological transition from
 infectious to chronic diseases is now complete in the developed world,
@@
 
-\begin{funding}
+\subsection*{Acknowledgements}
 This project was funded by the Speedwell Foundation.
-\end{funding}
 

### 2019-12-11 NYC_Twitter_SDM_models - Copy.Rmd
download data and refactor Rmd

@@
 ```{r read_tracts}
-tracts <- geojson_read("nyc_tracts.geojson", what = "sp")
-tracts_sf <- st_read("nyc_tracts.geojson", quiet = TRUE) %>%
+tracts <- geojson_read("data/nyc_tracts.geojson", what = "sp")
+tracts_sf <- st_read("data/nyc_tracts.geojson", quiet = TRUE) %>%
   st_transform(4326)
@@
 ```{r read_parks, echo=FALSE}
-parks <- geojson_read("nyc_parks.geojson", what = "sp")
-parks_sf <- st_read("nyc_parks.geojson") %>%
+parks <- geojson_read("data/nyc_parks.geojson", what = "sp")
+parks_sf <- st_read("data/nyc_parks.geojson") %>%
   st_transform(4326)
@@
 ```{r basemodels}
+# Time consuming.
 obese_base_lm <- lm(update(base_formula, OBESITY ~ .), data = tracts@data)
@@
 ```{r obesity, echo=FALSE, message=FALSE, warning=FALSE}
+# Time consuming.
 obese_ls1_sdm <- lagsarlm(update(parks_formula_1, OBESITY ~ .), 
@@
   ) %>%
-  spread(model, output) %>% View
+  spread(model, output)
 
@@
 
-# Estimate Physical Activity Models
+<!-- # Estimate Physical Activity Models -->
 
-```{r phy, echo=FALSE, message=FALSE, warning=FALSE}
+```{r phy, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
 physact_base_sdm <- lagsarlm(update(base_formula, Phys_Act ~ .), 
@@
 
-```{r physact_impacts}
+```{r physact_impacts, eval=FALSE, include=FALSE}
 physact_models %>%

### 2019-12-11 NYC_Twitter_SDM_models - Copy.Rmd
extract the function `impacts_extractor`

@@
 ```{r obese_sdm_age_impacts}
-impacts_extractor <- function(sdm) {
-
-  impacts_summary <- summary(impacts(sdm, tr=trMC, R=1000), zstats = TRUE)
-
-  coef <- list(
-    direct = impacts_summary$direct_sum,
-    indirect = impacts_summary$indirect_sum,
-    total = impacts_summary$total_sum
-  ) %>%
-  lapply(function(s) {
-    data_frame(
-      var = names(s$statistics[,1]),
-      impact = s$statistics[,1]
-    )
-  }) %>%
-    bind_rows(.id = "effect")
-
-  pval <- as_data_frame(impacts_summary$pzmat) %>%
-    mutate(
-     var = rownames(impacts_summary$pzmat)
-    ) %>%
-    gather("effect", "p-val", -var) %>%
-    mutate(effect = tolower(effect))
-
-  left_join(coef, pval, by = c("effect", "var"))
-}
 

### 2019-12-11 R/impacts_extractor.R

@@ -1,0 +1,23 @@
+impacts_extractor <- function(sdm) {
+  impacts_summary <-
+    summary(impacts(sdm, tr = trMC, R = 1000), zstats = TRUE)
+  
+  coef <- list(
+    direct = impacts_summary$direct_sum,
+    indirect = impacts_summary$indirect_sum,
+    total = impacts_summary$total_sum
+  ) %>%
+    lapply(function(s) {
+      data_frame(var = names(s$statistics[, 1]),
+                 impact = s$statistics[, 1])
+    }) %>%
+    bind_rows(.id = "effect")
+  
+  pval <- as_data_frame(impacts_summary$pzmat) %>%
+    mutate(var = rownames(impacts_summary$pzmat)) %>%
+    gather("effect", "p-val",-var) %>%
+    mutate(effect = tolower(effect))
+  
+  
+  left_join(coef, pval, by = c("effect", "var"))
+}

### 2019-12-11 NYC_Twitter_SDM_models - Copy.Rmd
extract modeling part.

@@
 ```
-
-# Obesity model estimation
-
-Create a spatial weights matrix. It's showing that we have three regions with
-no links, so we need to ensure that these are estimated in the model.
-
-```{r W, echo=FALSE}
-neighbors <- poly2nb(tracts, queen = TRUE)
-W <- nb2listw(neighbours = neighbors, zero.policy = TRUE)
-print(W, zero.policy = TRUE)
-trMC <- trW(as(W, "CsparseMatrix"), type="MC") # trace used in montecarlo impacts
-```
-
-```{r obese_lm, echo=FALSE, message=FALSE, include=FALSE}
-base_formula <- formula( 
-  ~ log(Pop_Density) + 
-    FulltimeWork + CollegeDeg + Single_Percent + 
-    Pct0to17 + Pct18to29 + Pct65plus + # need to leave out a category for collinearity  
-    PctBlack + PctAsian + PctOther + PctHispanic)
-
-parks_formula_1 <- update(base_formula, . ~ . + park_ls1 )
-parks_formula_2 <- update(base_formula, . ~ . + park_ls2 )
-parks_formula_4 <- update(base_formula, . ~ . + park_ls4 )
-parks_formula_5 <- update(base_formula, . ~ . + park_ls5 )
-```
-
-First, let's estimate the base model as an LM, SDM, and SEM model
-
-```{r basemodels}
-# Time consuming.
-obese_base_lm <- lm(update(base_formula, OBESITY ~ .), data = tracts@data)
-obese_base_sar <- lagsarlm(update(base_formula, OBESITY ~ .), 
-                           data = tracts@data, listw = W, zero.policy = TRUE)
-obese_base_sem <- errorsarlm(update(base_formula, OBESITY ~ .), 
-                             data = tracts@data, listw = W, zero.policy = TRUE)
-obese_base_sdm <- lagsarlm(update(base_formula, OBESITY ~ .), 
-                           data = tracts@data, listw = W, zero.policy = TRUE, 
-                           type = "mixed")
-```
-
-A likelihood ratio test reveals that the SEM is not preferred, so we use the SDM
-only going forward.
-
-```{r lrtest}
-lmtest::lrtest(obese_base_sem, obese_base_sdm)
-```
-
-```{r obesity, echo=FALSE, message=FALSE, warning=FALSE}
-# Time consuming.
-obese_ls1_sdm <- lagsarlm(update(parks_formula_1, OBESITY ~ .), 
-                          data = tracts@data,  listw = W, zero.policy = TRUE, type = "mixed")
-obese_ls2_sdm <- lagsarlm(update(parks_formula_2, OBESITY ~ .), 
-                          data = tracts@data,  listw = W, zero.policy = TRUE, type = "mixed")
-obese_ls4_sdm <- lagsarlm(update(parks_formula_4, OBESITY ~ .), 
-                          data = tracts@data,  listw = W, zero.policy = TRUE, type = "mixed")
-obese_ls5_sdm <- lagsarlm(update(parks_formula_5, OBESITY ~ .), 
-                          data = tracts@data,  listw = W, zero.policy = TRUE, type = "mixed")
-
-obesity_models <- list(obese_base_sdm, obese_ls1_sdm, obese_ls2_sdm,
-                       obese_ls4_sdm, obese_ls5_sdm)
-names(obesity_models) <- c("Base", ls_names)
-```
-
-Estimate spatial models for the `Phys_Act` variable. Model 1 does not include park access. Models 2 & 3 each use a different park access variable with a different combination of parameters:
-
-```{r}
-screenreg(obesity_models)
-```
-
-Estimate the impacts for `obese_sdm_age`.
-
-```{r obese_sdm_age_impacts}
-
-obesity_models %>%
-  lapply(impacts_extractor) %>%
-  bind_rows(.id = "model") %>%
-  transmute(
-    model, var, effect, 
-    output = paste(round(impact, 5), 
-                   gtools::stars.pval(`p-val`))
-  ) %>%
-  spread(model, output)
-
-```
-
-<!-- # Estimate Physical Activity Models -->
-
-```{r phy, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
-physact_base_sdm <- lagsarlm(update(base_formula, Phys_Act ~ .), 
-                             data = tracts@data,  listw = W, zero.policy = TRUE, 
-                             type = "mixed")
-physact_ls1_sdm <- lagsarlm(update(parks_formula_1, Phys_Act ~ .), 
-                            data = tracts@data,  listw = W, zero.policy = TRUE, 
-                            type = "mixed")
-physact_ls2_sdm <- lagsarlm(update(parks_formula_2, Phys_Act ~ .), 
-                            data = tracts@data,  listw = W, zero.policy = TRUE, 
-                            type = "mixed")
-physact_ls4_sdm <- lagsarlm(update(parks_formula_4, Phys_Act ~ .), 
-                            data = tracts@data,  listw = W, zero.policy = TRUE, 
-                            type = "mixed")
-physact_ls5_sdm <- lagsarlm(update(parks_formula_5, Phys_Act ~ .), 
-                            data = tracts@data,  listw = W, zero.policy = TRUE, 
-                            type = "mixed")
-
-physact_models <- list(physact_base_sdm, physact_ls1_sdm, physact_ls2_sdm,
-                       physact_ls4_sdm, physact_ls5_sdm)
-names(physact_models) <- c("Base", ls_names)
-```
-
-```{r physact_impacts, eval=FALSE, include=FALSE}
-physact_models %>%
-  lapply(impacts_extractor) %>%
-  bind_rows(.id = "model") %>%
-  transmute(
-    model, var, effect, 
-    output = paste(round(impact, 5), 
-                   gtools::stars.pval(`p-val`))
-  ) %>%
-  spread(model, output) %>% View
-```

### 2019-12-11 analysis/sdm.Rmd
extract modeling part.

# Obesity model estimation

Create a spatial weights matrix. It's showing that we have three regions with
no links, so we need to ensure that these are estimated in the model.

```{r W, echo=FALSE}
# tracts %>% class
# neighbors <- poly2nb(tracts, queen = TRUE)
# neighbors %>% write_rds("../output/neighbors.rds")
neighbors <- read_rds("../output/neighbors.rds")
neighbors %>% class
W <- nb2listw(neighbours = neighbors, zero.policy = TRUE)
print(W, zero.policy = TRUE)
trMC <- trW(as(W, "CsparseMatrix"), type="MC") # trace used in montecarlo impacts
```

```{r obese_lm, echo=FALSE, message=FALSE, include=FALSE}
base_formula <- formula( 
  ~ log(Pop_Density) + 
    FulltimeWork + CollegeDeg + Single_Percent + 
    Pct0to17 + Pct18to29 + Pct65plus + # need to leave out a category for collinearity  
    PctBlack + PctAsian + PctOther + PctHispanic)
# leave out y
parks_formula_1 <- update(base_formula, . ~ . + park_ls1 )
parks_formula_2 <- update(base_formula, . ~ . + park_ls2 )
parks_formula_4 <- update(base_formula, . ~ . + park_ls4 )
parks_formula_5 <- update(base_formula, . ~ . + park_ls5 )
```

First, let's estimate the base model as an LM, SDM, and SEM model

```{r}
# tracts@data %>% write_rds("../output/tracts_data.rds")
data <- read_rds("../output/tracts_data.rds")
data %>% head()
```

```{r basemodels}
# Time consuming.
obese_base_lm <- lm(update(base_formula, OBESITY ~ .), data = data)
obese_base_sar <- lagsarlm(update(base_formula, OBESITY ~ .), 
                           data = data, listw = W, zero.policy = TRUE)
obese_base_sem <- errorsarlm(update(base_formula, OBESITY ~ .), 
                             data = data, listw = W, zero.policy = TRUE)
obese_base_sdm <- lagsarlm(update(base_formula, OBESITY ~ .), 
                           data = data, listw = W, zero.policy = TRUE, 
                           type = "mixed")
```

A likelihood ratio test reveals that the SEM is not preferred, so we use the SDM
only going forward.

```{r lrtest}
lmtest::lrtest(obese_base_sem, obese_base_sdm)
```

```{r obesity, echo=FALSE, message=FALSE, warning=FALSE}
# Time consuming.
obese_ls1_sdm <- lagsarlm(update(parks_formula_1, OBESITY ~ .), 
                          data = data,  listw = W, zero.policy = TRUE, type = "mixed")
obese_ls2_sdm <- lagsarlm(update(parks_formula_2, OBESITY ~ .), 
                          data = data,  listw = W, zero.policy = TRUE, type = "mixed")
obese_ls4_sdm <- lagsarlm(update(parks_formula_4, OBESITY ~ .), 
                          data = data,  listw = W, zero.policy = TRUE, type = "mixed")
obese_ls5_sdm <- lagsarlm(update(parks_formula_5, OBESITY ~ .), 
                          data = data,  listw = W, zero.policy = TRUE, type = "mixed")

obesity_models <- list(obese_base_sdm, obese_ls1_sdm, obese_ls2_sdm,
                       obese_ls4_sdm, obese_ls5_sdm)
names(obesity_models) <- c("Base", ls_names)
```

Estimate spatial models for the `Phys_Act` variable. Model 1 does not include park access. Models 2 & 3 each use a different park access variable with a different combination of parameters:

```{r}
screenreg(obesity_models)
```

Estimate the impacts for `obese_sdm_age`.

```{r}
obesity_models %>% class
obesity_models
```

```{r obese_sdm_age_impacts}
devtools::load_all()
obesity_models %>%
  lapply(impacts_extractor) %>%
  bind_rows(.id = "model") %>%
  transmute(model, var, effect,
            output = paste(round(impact, 5),
                           gtools::stars.pval(`p-val`))) %>%
  spread(model, output)
```

<!-- # Estimate Physical Activity Models -->

```{r phy, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
physact_base_sdm <- lagsarlm(update(base_formula, Phys_Act ~ .), 
                             data = data,  listw = W, zero.policy = TRUE, 
                             type = "mixed")
physact_ls1_sdm <- lagsarlm(update(parks_formula_1, Phys_Act ~ .), 
                            data = data,  listw = W, zero.policy = TRUE, 
                            type = "mixed")
physact_ls2_sdm <- lagsarlm(update(parks_formula_2, Phys_Act ~ .), 
                            data = data,  listw = W, zero.policy = TRUE, 
                            type = "mixed")
physact_ls4_sdm <- lagsarlm(update(parks_formula_4, Phys_Act ~ .), 
                            data = data,  listw = W, zero.policy = TRUE, 
                            type = "mixed")
physact_ls5_sdm <- lagsarlm(update(parks_formula_5, Phys_Act ~ .), 
                            data = data,  listw = W, zero.policy = TRUE, 
                            type = "mixed")

physact_models <- list(physact_base_sdm, physact_ls1_sdm, physact_ls2_sdm,
                       physact_ls4_sdm, physact_ls5_sdm)
names(physact_models) <- c("Base", ls_names)
```

```{r physact_impacts, eval=FALSE, include=FALSE}
physact_models %>%
  lapply(impacts_extractor) %>%
  bind_rows(.id = "model") %>%
  transmute(
    model, var, effect, 
    output = paste(round(impact, 5), 
                   gtools::stars.pval(`p-val`))
  ) %>%
  spread(model, output) %>% View
```

### 2019-12-11 analysis/sdm.Rmd
render Rmd

@@
-# Obesity model estimation
+---
+output: github_document
+---
+
+<!-- # Obesity model estimation -->
 
@@
 no links, so we need to ensure that these are estimated in the model.
+
+```{r message=FALSE, warning=FALSE}
+library(tidyverse)
+library(geojsonio)
+library(tidyverse)
+library(maptools)
+library(rgdal)
+library(rgeos)
+library(spdep)
+library(texreg)
+library(leaflet)
+library(sf)
+library(VGAM)
+```
 
@@
 ```{r obesity, echo=FALSE, message=FALSE, warning=FALSE}
+ls_names <- c("Short, no tweets",  "Short, tweets",  
+              "Long, no tweets",  "Long, tweets")
 # Time consuming.
@@
 ```{r obese_sdm_age_impacts}
+# add2pkg::create_desc()
 devtools::load_all()

### 2019-12-16 analysis/sdm.Rmd
make the code visiable

@@
 
-```{r W, echo=FALSE}
+```{r W}
 # tracts %>% class
